### 2010-10-06 Text/paper.tex
Cleaned repo, added text with all the necessary pictures.

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[numbers]{natbib}
\usepackage{color}

\begin{document}
\section{Introduction}
The Taylor flow which has analytical solution for low Capillary numbers
\cite{bretherton} is the attractive benchmark for the diffusion interface
models. The film width needed to be resolved is proportional to $Ca^{2/3}$
which is for small capillary numbers can reach numbers from $0.001$ to $0.01$.
When it comes to the resolution of the film thickness the diffusion interface
should be resolved in such way as to have negligible effect on physics in the
film. The mass transfer which exists in the small film drops the bubble driven
pressure to proportionality $Ca^{2/3}$ \cite{kreutzer-pressure-drop} instead of
linear drop as in Poiesuille driven flows.

The goal of this paper is to give the answer as to how to choose parameters and
resolution to obtain the Bretherton/Taylor problem film thickness for the
diffusion interface model such as the free-energy model \cite{swift}.

\section{Benchmark}

There are a number of the numerical methods which were used for the Taylor flow
simulation. \citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillar.
\citet{kreutzer-pressure-drop} performed numerical simulation by the
finite volume method the circular capillary for the number of different
Reynolds and capillary numbers. \citeauthor{wong-films} in
\cite{wong-films,wong-pressure} studied three-dimensional bubbles in the
polygonal capillaries and showed the derivations of different shapes in the
slug crossections and menisci appearance.
\citet{heil-bretherton,ingham-plates} studied the air finger propagation to
the two-dimensional channel for the range of Reynolds numbers and capillary
numbers. \citet{giavedoni-numerical} performed crossvalidation of the
finite element solution developed by them with the already published results
across the different capillary numbers. The solution is available for the
circular and planar cases.

While the applicability of other methods are well established for the
simulation of the Bretherton/Taylor problem, the lattice Boltzmann
applicability and thorough parameters study to the best authors' knowledge is
not yet done. One should acknowledge the works of
\citet{pagonabarraga-fingers} on meniscus in thin films for the fingering
phenomena. \citet{sehgal-microchannel} performed the lattice Boltzmann
simulations of two-dimensional channel flow for larger capillary numbers where
the authors found the discrepancies with the classical Bretherton theory having
limitations to the low capillary numbers regime \cite{giavedoni-numerical}.

\section{Lattice Botlzmann benchmark}
The purpose of this section is to establish the benchmark and perform
parameters study for the classical Bretherton/Taylor problem. There are certain
challenges for the lattice Boltzmann method which can be summarized as follows:
\begin{itemize}
 \item The diffusive interface nature of the multiphase model should not
influence the film resolution.
 \item The spurious currents arising from the surface tension discretization
can influence the mass transfer in the thin film therefore comprimising the
overall solution.
 \item The wettability coefficient which is in principle can control the
dynamic angle and menisci appearance \cite{pagonabarraga-finger} should not
influence on the overall solution.
 \item The free-energy model examined in the work is one density/different
viscosities model. Classical Bretherton problem is the free-interface problem
where the film thickness is established by itself. The viscosities ratio should
be large as it is indicated in \cite{pagonabarraga-fingers}, however no studies
address this issue. Because the free-interface nature of the Bretherton/Taylor
problem one can have difficulties with imposing the pressure boundary
conditions which do require to impose two separate boundary conditions for two
distribution functions.
 \item A number of simulations are done across the circular shape of channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate the laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed for the curved boundaries
however the analysis of error which can be addressed to the boundary conditions
treatment becomes an issue. 
\end{itemize}
Given all the concerns and challenges the suggested Lattice Boltzmann framework
is the two-dimensional flow driven by gravity or by pressure boundary
conditions. The advantage of the gravity driven bubble is that it's easy to
implement with the periodic boundary conditions. However, the analytical
solution to this problem, to the best authors' knowledge, is not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and there
is also the additional pressure drop due to front and rear menisci.

The solution
for relatively long bubbles driven by pressure specified on the inlet and the
outlet coincides with the classical theory of Bretherton. However, to implement
the pressure boundary conditions for the binary liquid model requires the
specification of boundaries conditions for two populations, for the momentum and
for the advection-diffusion equation (which are quite different). To the best
authors' knowledge such the situation is not yet covered in the literature. We
address this situation in Section \ref{appendix-pressure}.

\section{The bubble motion in the planar case}
This section covers the results for the pressure and force driven bubbles. The
idea is to establish what parameters are needed, i.e. grid size, bubble length,
viscosity ratio,to obtain the proper behaviour of Bretherton law.

\subsection{The flow driven by the gravity}
\citet{wong-films} indicated that the film width for polygonal capillaries
changes from $Ca^{1/2}$ on distances $x\propto Ca^{-1}$ to $Ca^{2}{3}$ for
distances $x \gg Ca^{-1}$. In simulations the bubble length is taken around 5-10
diameters of the channel {\color{red} Why is that? Is it because of
curvatures?}.

Also, gravity or force driven bubble flow has periodical boundary conditions.
Therefore, the situation to be simulated is the infinite train of bubbles
coming one after another. Thus, to avoid the influence of bubbles on each other
one needs to take quite long gap between bubbles. We take this gap around 3-5
lengths of the bubble.

\subsubsection{The influence of initial film width. Right parameters}
It is important to show that the steady film width is independent on initial
conditions where the most crucial condition is the initialized bubble width. If
in steady state the bubble width addjust itself to reveal certain film width
therefore the simulations are consistent. This section is to validate it.

For the sake of simplicity we picked up the following parameters for the
simulation:
\begin{equation}
\begin{aligned}
&Ca=0.005\\
&Ca=\frac{U_{bubble} \mu_{slug}}{\gamma}=0.005.
\end{aligned}
\end{equation}
Some parameters are fixed with the model parameters, as surface tension:
\begin{equation}
\gamma=\sqrt{\frac{8 k A}{9}}=\sqrt{\frac{8 0.04 0.04}{9}}=0.037712
\end{equation}

The film width for such small Capillary numbers is given by equation
\cite{giavedoni-numerical}:
\begin{equation}
\frac{h_{\inf}}{H}=1.3375 Ca^{2/3}=0.03899
\end{equation}
where we assume that $2 H$ or $Ny$ in lattice units is the channel width. That
means that one needs to resolve the interface which is around $2$ percent of
whole channel width with for instance $4$ lattice units. Therefore the channel
width is $200$ lattice units. We take the bubble length as $5$ channel widths
and the distance between slugs we take as $2$ bubble lengths. This will give
right away the grid size as $200\mathrm{x}3000$. Next step is to calculate the
velocity. The velocity can be calculated from the Capillary number as:
\begin{equation}
U_{bubble}=0.005 \frac{\gamma}{\mu_{slug}}=0.005 \frac{0.037712}{0.66666}=2.82\,
10^{-4},
\end{equation}
where $\mu_{slug}=1/3 (2.5-0.5)=0.6666$.
This velocity implies the following pressure gradient calculated from the
Poisuielle flow solution:
\begin{equation}
\begin{aligned}
&U_{bubble}=\frac{H^2}{8\mu} \frac{\mathrm{d}P}{\mathrm{d}x}=\frac{Ny^2}{8/3
(\tau_{liq}-1/2)}\frac{\mathrm{d}P}{\mathrm{d}x}\\
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca
\end{aligned}
\end{equation}
Therefore the pressure gradient established is the following:
\begin{equation}
\begin{aligned}
\frac{\mathrm{d}P}{\mathrm{d}x}=3.77\, 10^{-8}.
\end{aligned}
\end{equation}

\subsubsection{The influence of initial film width. Easy parameters to check}
It is important to show that the steady film width is independent on initial
conditions where the most crucial condition is the initialized bubble width. If
in steady state the bubble width addjust itself to reveal certain film width
therefore the simulations are consistent. This section is to validate it.

The parameters in the previous subsection imply very long simulations. Even
Bretherton delivered his solution for the range of Capillary number less than
$0.005$. However, a lot of simulations devoted to the Bretherton problem
simulated extended it in the range of Capillary numbers greater than $0.005$
and for different Reynolds numbers. One can refer to figures
\ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from citations
\cite{giavedoni-numerical,heil-bretherton} which crossvalidated their
simulations with Bretherton and each other. As far as this section is intended
to show physics behavior the moderate for simulation parameters are taken which
can be easily and fast validated on a computer. 
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/giavedoni_planar.eps}
\caption{\citet{giavedoni-numerical} gathered results across the
literature for different Capillary numbers \label{fig:giavedoni:planar}}
\includegraphics[width=0.97\textwidth]{Figures/heil-planar.eps}
\caption{\citet{heil-bretherton} performed simulations for different ranges of
Capillary numbers and Reynolds numbers. \label{fig:heil:planar}}
\end{figure}

One can see from the Fig. \ref{fig:heil:planar} that the width for the
Capillary number equal to $0.05$ ranges from $0.12$ to $0.15$ while Reynolds
number ranges from $0$ to $300$. This capillary number has moderate film widths
to be resolved with certain accuracy. Let us recalculate all the necessary
parameters:
\begin{equation}
\begin{aligned}
&Ca=0.05\\
&Ca=\frac{U_{bubble} \mu_{liq}}{\gamma}=0.05.
\end{aligned}
\end{equation}
Some parameters are fixed with the model parameters, as surface tension:
\begin{equation}
\gamma=\sqrt{\frac{8 k A}{9}}=\sqrt{\frac{8 0.04 0.04}{9}}=0.037712
\end{equation}
The film width for such small Capillary numbers is taken from
\ref{fig:heil:planar} for Reynolds number $40$ equals to $0.12$. We want to
resolve $12\%$ with $6$ lattice units. Therefore the channel
width is $50$ lattice units. We take the bubble length as $5$ channel widths
and the distance between slugs we take as $2$ bubble lengths. This will give
right away the grid size as $50\mathrm{x}750$. Next step is to calculate the
velocity. The velocity can be calculated from the Capillary number as:
\begin{equation}
U_{bubble}=0.05 \frac{\gamma}{\mu_{liq}}=0.05 \frac{0.037712}{0.6666}=2.82\,
10^{-3},
\end{equation}
where $\mu_{liq}=1/3 (2.5-0.5)=0.6666$. This implies the viscosity ratio to be
equal $R_{\nu}=2/0.2=10$ if we take the relaxation parameter for gas as
$\tau_{gas}=0.7$.
This velocity implies the following pressure gradient calculated from the
Poisuielle flow solution:
\begin{equation}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca=6.003\,10^{-6}.
\end{aligned}
\end{equation}

Some pictures after $40000$ steps for the gravity driven flow and for different
initialization of bubble width as $Ny-12$, $Ny-20$, $Ny-28$.
\begin{figure}
\includegraphics*[bb=18 350 594 450,width=0.97\textwidth]
{Figures/force_width_6.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/force_width_10.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/force_width_14.eps}
\caption{The phase plots for different bubble widths as $Ny-12$,$Ny-20$ and
$Ny-28$.}
\end{figure}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_6.eps}
\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_10.eps}
\begin{center}
\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_14.eps}
\end{center}
\caption{The phase profiles for different initialization parameters as the
bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
\label{fig:phase:profiles:different:widths:force}}
\end{figure}

One can see from Figures \ref{fig:phase:profiles:different:widths:force} that
the film pattern in the beginning of the slug is almost independent of the
initial conditions and it ranges from 4 to 5 lattice boltzmann units, where the
is taken where the phase field has a value of zero. This preliminary result
justify that lattice Boltzmann method is able to predict the film thickness.

However, with the smaller initialization of the bubble one can see that the
front meniscus where the film thickness is validated accross is decreasing.
Therefore to be on the safe side one needs to initialize bubbles to be near the
analytical length.

We assumed that the interface thickness should be around $6$ lattice boltzmann
units, but we assume the calculations to be based on the $U_{slug}$ taken from
the simple Poisuille flow velocities. In reality the slug has slippage factor
with with the flow and velocity is slightly bigger than the liquid media. One
can take the velocity of the bubble from the simulations and recalculate all
the necessary lengths. Velocity profiles behind the bubble and in the front
meniscus can be seen in Fig. \ref{fig:velocity:profiles:different:widths:force}
and \ref{fig:velocity:profiles:different:widths:pressure} for the force driven
and pressure difference driven flows, correspondingly. As one can see average
velocity for the bubble and the slug larger than the calculated velocity which
should give larger the film. However, the film width for the planar case is
different and can have another coefficient in front of the proportionality to
$Ca^{2/3}$. Also, the phase gradient is taken as $0$ right now which can have
influence on the measurements as well.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_6.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_6.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_10.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_10.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_14.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_14.eps}\\
\caption{The velocity profiles in bulk and bubble for different initialization
parameters as the
bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
\label{fig:velocity:profiles:different:widths:force}}
\end{figure}

While the shapes of droplets are different for the force and pressure
difference driven flows the velocity profiles show that the force driven bubble
move faster than in the case of pressure difference driven flow:
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_6.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_10.eps}
\\
\begin{center}
\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_14.eps}\\
\end{center}
\caption{Comparison between velocity bubble
profiles.\label{fig:velocity:comparison:different:widths}}
\end{figure}

Note that in the case of pressure driven flows the termination bulk and bubble
velocity have more consistent behavior and the values are not different too
much. That is one of the reasons why in the case of the force driven flows the
width has a bit different values depending on the initialization procedure.

\subsection{Grid refinement}
To examine the interface width we performed grid refinement for both force and
pressure difference driven flows. To avoid the effects of rear and front
menisci the problem is initialized with the film width occupying $12$ percent
of the channel height. All other parameters are kept the same, that includes
keeping the same Capillary number and viscosities ratio. Therefore the
following parameters were utilized: $Ny=51,100,149,198$,corresponding to the
effective channels widths as $H_{eff}=49,98,147,196$, $Nx=736,1471,2206,2941=15
H_{eff}$. It is easy to check
that keeping all liquid viscosities the same yields the following to be
constant while doing the grid refinement
$H^2\frac{\mathrm{d}P}{\mathrm{d}x}=Ny^2\frac{\mathrm{d}P}{\mathrm{d}x}=Const$.
The results for the force driven flows velocity phase profiles are shown in
Fig. \ref{fig:grid:profiles}. The corresponding ratios to the effective channel
width (without bounce-back nodes) is as $0.0809472165313,
0.0708111118485, 0.0678483852412, 0.0664499716904$. One can see
that for the channel widths as $149,198$, the results are convergent. The
velocities in the center of the bubble are $0.004176469553,
0.00401004133, 0.003969909065, 0.00394134433$, which corresponds
to the $Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}=\frac{\frac{2}{3}\,
0.0045}{0.03771}=0.069$, which is slightly higher when for the assumed
profile. However, if we define Capillary number through the velocity of the
slug, then the Capillary number will have another value. The velocities of the
slug are $0.003411444831,0.003412693869,0.003420338754,0.003425757312$ and
Capillary number defined through these velocities is $0.0602$. Those numbers
are adequate if we assume that the results presented in
Fig.\ref{fig:giavedoni:planar} and Fig. \ref{fig:heil:planar} correspond to the
half of the channel but not the whole channel.

Let us calculate to which extent we need to refine the interface to obtain
convergent results. The interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. This interface width divided to the effective channel width
($5\xi/H_{eff}$) is as follows $0.102,0.051,0.034,0.025$ for the given
$H_{eff}$. However, the interface width needs to be takes as the ratio with the
film width which we want to resolve, this ratio $5\xi/H_{film}$ is as follows
$1.2605,  0.7205,  0.501,  0.3839$. In reality this number is
a bit overestimated as the interface doesn't fill the influence of the wall and
it maintains all the macroscopic parameters jumps. Therefore, one needs to
resolve the interface as $40\%$ of the expected film width.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_49.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_98.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_147.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_196.eps}\\
\caption{Grid refined profiles for the effictive channel widths
$H_{eff}=49,98,147,196$.\label{fig:grid:profiles}}
\end{figure}

{\color{red} The next step is to perform calculations for different Capillary
numbers to restore necessary characteristics as it's seen in Fig.
\ref{fig:giavedoni:planar} and \ref{fig:heil:planar}. Also note that grid
refinement is done for the gravity driven flows, which can have pretty
different characteristics in comparison with pressure simulations (unstable
for the grid refinement - need to adjust parameters)}.

\section{Capillary number region}
This section is devoted to the study of different Capillary number. We refer to
grid refinement section where we obtained that the interface width should be
at most $40$ percent of the interface width. For the benchmark results we
suggest the following Capillary numbers as:
\begin{equation}
0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0
\end{equation}
The corresponding approximate film widths from \cite{giavedoni-numerical} are as
follows (we take the half width as it's done in the grid refinement section):
$0.04,0.06,0.08,0.1,0.12,0.13,0.15$. $5$ lattice boltzmann units occupy
$40\%$ of the interface thickness. Therefore the grid size should be as
following $314\mathrm{x}4681$,$210\mathrm{x}3121$,
$158\mathrm{x}2341$, $127\mathrm{x}1876$, $106\mathrm{x}1561$,
$98\mathrm{x}1441$,$85\mathrm{x}1246$.

To utilize previous results to properly initialize simulations to run one need
to do some modifications. The result is that for the grid with $H_{eff}=196$ we
obtained the film width as $0.0685$ and velocity in the center of the bubble as
$0.0045$. The force gradient was $6\, 10^{-6} /16$. The predicted Capillary
number was $0.05$ and the actual one is $0.0795$. But we performed simulations
according the assumed value of $0.05$. Let us perform some correlations to
calculate how the system needs to be initialized. 
\begin{equation}
\begin{aligned}
&Ca_{lit} \propto U_{slug}\\
&U_{slug} \propto \frac{\mathrm{d}P}{\mathrm{d}x} Ny^2 
&Ca_{lit} \propto \frac{\mathrm{d}P}{\mathrm{d} x} Ny^2 \text{ or }\\
\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{Ny^2}
\end{aligned}
\end{equation}
Therefore the initialized consistent pressure gradients can be taken from
previous consistent simulations:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d} x}=6\,10^{-6}/16 \frac{Ca_{literature}}{0.05}
\frac{196^2}{Ny^2}=0.28812 \frac{Ca_{literature}}{Ny^2}
\end{equation}
Note that $Ny$ is calculated through the expected resolution of the interface
in comparison with the interface thickness which itself is based on
Capillary number simulations. In terms of time the steady state succesfull
simulation for the grid $196$ was obtained after $200,000$ steps. The physical
time is as follows:
\begin{equation}
\begin{aligned}
&\Delta t=\frac{U_{slug,LB} \Delta X}{U_{phys}} \propto \frac{1}{Ny}\\
&N_{iter}=200,000 \frac{Ny}{196}\approx 1020 Ny\\
&N_{output}=51 Ny\\
\end{aligned}
\end{equation}

Steady simulations were obtained for the following Capillary numbers
$0.03,0.05,0.08,0.1$ with the film thicknesses as
$0.0337433850478, 0.063192148642, 0.0799078589211, 0.0891674460655$ which are
quite close to the predicted one $0.04,0.06,0.08,0.1$. The thing is to identify
the Capillary numbers associated with those numbers. The velocity in the
centers are $0.00209258774,0.003826643811,0.006506324324,0.008333889077$. The
Capillary numbers associated with these velocities are $0.036992074529545796, 
0.067646144698590829,0.1150166512524845,0.14732373700007995$, which again a bit
overpredictive but consistent. The bulk liquid velocities for the simulation
are $0.001947507726,0.003364703007,0.005483194352,0.00687516499$. The capillary
numbers associated with those velocities are slightly less than with the bubble
are as follows:$0.034427397986694822, 0.059480107823211698,
0.096930097721574432, 0.12153689465513355$. We can define the average capillary
number defined through the bubble capillary and slug capillaries numbers as
$Ca_{aver}=\frac{Ca_{bubble}+Ca_{slug}}{2}$. The fig.
\ref{fig:capillary:simple} represents the thickness against the Capillary
number which mimics the curve behavior for the Capillary number dependance as
indicated in Fig. \ref{fig:giavedoni:planar}.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/capillary_thickness_simple.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/capillary_range.eps}
\caption{Thickness versus the capillary number for different scalings based on
the average capillary number. \label{fig:capillary:simple}}
\end{figure}

\subsection{The influence of the force}
\begin{description}
 \item 
On the grid $198\mathrm{x}2941$ and the initial film width as $20$ units
the
influence of the force is studied. $16$ different values of force are examined,
which have the following expression $\frac{0.000006}{16} i$, where $i=1 \dots
16$. The system is allowed to come to the equilibrium after $200,000$ steps. The
goal is to study the influence of the force on the velocity and therefore the
capillary number. Unfortunately in the first case only a few simulations are
stable, i.e.
$i=1,2$ is stable. The results for the interface thicknesses are
$0.0700122222668,0.0985152303286$ with the velocities associated as
$0.004490013864,0.009597299475$ and slug velocities as
$0.003861354387,0.006350100792$. Those numbers correspond to the bubble
capillary numbers as $0.079372981271400328, 0.16965788849626481$ and slug
capillary numbers as $0.068259746790303111, 0.11225498328103167$. The average
capillary numbers are $0.07381636,0.14095644$.

\item
On the grid $394\mathrm{x}5881$ and the initial film width as $20+4 i$,where
$i=1 \dots 5$ units the
influence of the force is studied. The force is scheduled as
$\frac{0.000006}{128}+i \frac{0.000006}{640}$.The intention is to cover the
calculation of the Capillary numbers in the range $0.03,0.035,0.04,0.045,0.05$. 
$5$ different values of force are examined. The system is allowed to come to the
equilibrium after $400,000-40,000 i$ steps. The associated thicknesses were as
it follows
$0.0488915027016,0.0538098132568,0.0587100833678,0.0615470184196,
0.0661331899123$. Bubble velocities are
$0.002592856205,0.003077834398,0.003562975124,0.004034533823,0.004470657603$.
Slug velocities are
$0.002337542358,0.002752687109,0.003158125555,0.003552413499,0.003926980914$.
Bubble Capillary numbers are $0.04583566,  0.05440894,  0.0629851 , 
0.07132116, 0.07903081$; slug Capillary numbers are $0.0413223 ,  0.04866109, 
0.0558283, 0.06279839, 0.06941987$. The average capillary numbers are
$0.04357898, 0.05153502, 0.0594067, 0.06705977, 0.07422534$.
\end{description}

All the range of available Capillary numbers are shown in Fig.
\ref{fig:capillary:simple}. The figure looks consistent with different
scalings and provide good results.

\subsection{The influence of the wall gradient}
The influence of the wall gradient. We examined the wall gradient influence on
the standard grid $198\mathrm{x}2941$ and the initial film width as $20$
lattice boltzmann units. We examined $11$ different values for the wall
gradient ranging from $-1$ to $1$. The results are consistent among all the
wall gradients given that the interface is resolved in the right way. For the
given wall gradients the values of interface thicknesses are
$0.0709130255276, 0.0708903443305, 0.0708661222722,
0.0708369188112, 0.0707962325316, 0.0707266326981, 0.0705839443249,
0.0702489349415, 0.0697553433989$. For the wall gradients $0.8$ and $1.0$ the
values are NAN. The negative values of the phase gradient are preferable, as
far as the phase of the liquid adjacent to the wall has the value of $1$. That
means that the phase values near the wall are above $1$. In case of positive
gradients the values near the wall are below $1$ and the gradient profile can
fuse with the values of the gas phase as $-1$, creating the wall stickyness,
unphysical one.

\subsection{The flow driven by the pressure difference}
The details of imposing pressure boundaries for the binary liquid model is
given in Section \ref{appendix-pressure}.

\subsubsection{The influence of initial bubble width. Easy parameters to check}
Here we impose absolutely the same parameters as for gravity driven flow.
However instead of pressure gradient one needs to put different densities at
the inlet and the outlet which can be easily calculated from the pressure
gradient as follows:
\begin{equation}
\begin{aligned}
&\rho_{inlet}=1+3 Nx \frac{\mathrm{d}P}{\mathrm{d}x}=3\, 750\, 6\,
10^{-6}=1.0135\\
&\rho_{outlet}=1.
\end{aligned}
\end{equation}

Some pictures after $40000$ steps for the pressure driven flows for different
initialization of bubble width as $Ny-12$, $Ny-20$, $Ny-28$.
\begin{figure}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/pressure_width_6.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/pressure_width_10.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/pressure_width_14.eps}
\caption{The phase plots for different bubble widths as $Ny-12$,$Ny-20$ and
$Ny-28$.}
\end{figure}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/plane_pressure_width_6.eps}
\includegraphics[width=0.47\textwidth]{Figures/plane_pressure_width_10.eps}
\begin{center}
\includegraphics[width=0.47\textwidth]{Figures/plane_pressure_width_14.eps}
\end{center}
\caption{The phase profiles for different initialization parameters as the
bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
\label{fig:phase:profiles:different:widths:pressure}}
\end{figure}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_pressure_width_6.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_pressure_width_6.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_pressure_width_10.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_pressure_width_10.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_pressure_width_14.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_pressure_width_14.eps}\\
\caption{The velocity profiles in bulk and bubble for different initialization
parameters as the
bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
\label{fig:velocity:profiles:different:widths:pressure}}
\end{figure}

\section{Appendix I. Pressure boundaries conditions}
\label{appendix-pressure}
Bulk pressure for the binary liquid model is a function not only of density
but the phase field as well. Thus, one of the possible ways to fix the pressure
boundary condition is to fix density and phase at the inleat and at the outlet.
The pressure difference is introduced through the difference in densities
keeping phases at the inlet and at the outlet the same. One can introduce the
pressure difference through the phase, however it's unphysical since the phase
field is governed by passive advection-diffusion equation, where velocity is
given from the hydrodynamic simulations.

We examined many different pressure boundaries conditions, including Zou-He BC
\cite{zouhe-boundary,harting-pressure}, equilibrium conditions for
the phase field, etc. However, the stable situation was obtained only with the
boundary conditions by \citeauthor{guo-pressure} \cite{guo-pressure,
guo-boundary-conditions}. The idea of the boundary conditions is simple: the
coming populations to domain are decomposed to the equilibrium and
non-equilibrium part:
\begin{equation}
f_i=f_i^{eq}(\rho_{wall},\bm{u}_{wall})+f_i^{neq}.
\end{equation}
The incoming population after the collision is changed due to collision:
\begin{equation}
f_i=f_i^{eq}(\rho_{wall},\bm{u}_{wall})+f_i^{neq} (1-\omega).
\end{equation}
The method is called extrapolation as the wall macroscopic parameters are
extrapolated from fluid using first or second accuracy numerical stencils. If
the pressure boundary condition is specified then it specifies $\rho_{wall}$ as
well extrapolating the wall velocity $\bm{u}_{wall}$ from nearest neighbours.
The same goes for the velocity boundary conditions, where velocity
$\bm{u}_{wall}$ is specified and $\rho_{wall}$ is extrapolated from the fluid.

It is pretty straightforward to obtain non-equilibrium contribution from the
fluid, however the equilibrium for the wall side needs special treatment. If
one wants to take the hydrodynamic equilibrium, the numerical scheme blows
up. Therefore, one needs to construct the equilibrium function as it's given
for the binary liquid model. The binary liquid model equilibrium contains local
density, local phase and phase laplacians and gradients. The pressure boundary
conditions specify the local phase and the local density. Therefore, we
extrapolate not only velocity to construct the equilibrium, but phase laplacian
and gradients. This situation produces stable simulations to be validated in
the future. 
\bibliographystyle{plainnat}
\bibliography{paper}

\end{document}

### 2010-10-06 Text/paper.tex
Added small introduction.

@@
 
+\title{The binary liquid model for gravity-driven microchannel simulations}
+\date{\today}
+
 \begin{document}
+
+\maketitle
+
+\begin{abstract}
+The motion towards exploration of nano- and microscale phenomena brings new
+challenges to simulation tools. The continuous physics approach is limited on
+those scales, therefore such methods as Molecular Dynamics (MD) or Lattice
+Boltzmann Method (LBM) become appropriate. This work has aim to
+reproduce the gravity driven microchannel flow of free surface liquid with the
+help of the binary liquid lattice Boltzmann method. In the low capillary number
+$Ca$ region the flow is thoroughly described by Bretherton \cite{bretherton},
+where the deposition film thickness is proportional to $Ca^{2/3}$. The binary
+liquid model is continuous interface multiphase model. Therefore, the issue of
+how to properly resolve the interface to obtain proper correlations holds. In
+the work we address the binary liquid parameters, as mobility $M$, surface
+tension parameter $k$, the free energy constant $A$, grid resolution, to
+properly simulate microchannel flow. The work is valuable for industry and
+researchers willing to simulate Bretherton phenomena.
+\end{abstract}
+
 \section{Introduction}
-The Taylor flow which has analytical solution for low Capillary numbers
-\cite{bretherton} is the attractive benchmark for the diffusion interface
-models. The film width needed to be resolved is proportional to $Ca^{2/3}$
-which is for small capillary numbers can reach numbers from $0.001$ to $0.01$.
-When it comes to the resolution of the film thickness the diffusion interface
-should be resolved in such way as to have negligible effect on physics in the
-film. The mass transfer which exists in the small film drops the bubble driven
-pressure to proportionality $Ca^{2/3}$ \cite{kreutzer-pressure-drop} instead of
-linear drop as in Poiesuille driven flows.
+Taylor/Bretherton \cite{bretherton} flow describes flow of long bubbles in
+microchannels. It was found that the front meniscus film width is proportional
+to $Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
+of works which simulate the Bretherton problem with the boundary value method
+\cite{ingham-plates,heil-planar}, the simulation of Bretherton problem is of
+difficulty for diffusion interface methods, where the interface is spread over
+the few grid nodes. In this case the question of proper film width resolution
+arises. While the discontinuous interface models are good to simulate
+non-complicated geometries phenomena, the continuous interface methods are more
+flexible to resolve problem evolving coalescence or droplet breakup. This work
+concentrates on the resolving parameters range of the binary liquid lattice
+Boltzmann method to validate it across the Bretherton problem.
 
-The goal of this paper is to give the answer as to how to choose parameters and
+Lattice Boltzmann method has emerged as successful method to simulate various
+phenomena as hydrodynamics \cite{yu}, thermal-flows
+\cite{karlin-minimalmodels}, micro-flows \cite{ansumali-small-knudsen},
+ferrofluids \cite{kuzmin-aniso}, and multiphase flows
+\cite{swift,Shan-chen:extended}. Based on kinetic nature and as a particle
+method LBM tackles relatively easy complex geometries and the incorporation of
+physics phenomena on the microscopic level, as multiphase models. Most
+multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
+the interface in the continuous way. Such representation as it was pointed out
+brings issues with the film thickness resolution versus the interface
+resolution. In other words, when it comes to the resolution of the film
+thickness the diffusion interface should be resolved in such way as to have
+negligible effect on physics in the film.
+
+This work examines the binary liquids free-energy LBM model \cite{swift}. The
+model simulates binary liquids within the assumption of the uniform overall
+density. Therefore, the question of how to simulate the Bretherton problem
+where two phases are completely of different densities (liquid and gas) is of
+interest. However, the motion of a liquid slug and a bubble in a capillary is
+governed by the capillary number. Therefore, the viscosities ratio not the
+densities ratio is the critical parameter.
+
+The goal of this paper is to give the answer as to how to choose related to the
+binary liquid model parameters, viscosities ratio and grid
 resolution to obtain the Bretherton/Taylor problem film thickness for the
-diffusion interface model such as the free-energy model \cite{swift}.
+diffusion interface model. The paper is organized as follows. We briefly
+explain the binary liquid lattice Boltzmann model and summarize literature
+review for the Bretherton problem. Then the parameters involved in the
+simulation are examined and presented in the results section. The paper is
+concluded with main findings.
 

### 2010-10-08 Text/paper.tex
Added introduction about Bretherton problem and LBM description.

@@
 of works which simulate the Bretherton problem with the boundary value method
-\cite{ingham-plates,heil-planar}, the simulation of Bretherton problem is of
+\cite{ingham-plates,heil-bretherton}, the simulation of Bretherton problem is of
 difficulty for diffusion interface methods, where the interface is spread over
@@
 
-\section{Benchmark}
+\section{Bretherton bubble flow}
+Bretherton in his significant work \cite{bretherton} studied a long bubble
+moving in a tube filled with liquid. It was found that the deposition film is
+proportional to $Ca^{2/3}$ in the range of small capillary numbers. Later on
+it was figured \cite{wong-films,wong-pressure} that the film deposition
+proportinal to $Ca^{2/3}$ only near the front meniscus and the film width is
+variated through the bubble length for not infinite bubbles. Numerical
+simulations \cite{giavedoni-numerical} and experimental studies
+\cite{kreutzer-pressure-drop} show as well the deviation of rule $Ca^{2/3}$ for
+capillary numbers more than $0.003$ and Reynolds number dependance. To be able
+to predict consistently the flow pattern for capillaries in different range of
+the parameters researchers validate their codes with the small capillary
+numbers Bretherton problem.
 
-There are a number of the numerical methods which were used for the Taylor flow
+There are a number of the numerical methods which
+were used for the Taylor/Bretherton flow
 simulation. \citet{vanbaten-circular} studied the mass transfer and film
@@
 finite element solution developed by them with the already published results
-across the different capillary numbers. The solution is available for the
-circular and planar cases.
+across the different capillary numbers. The solution is available for circular
+and planar cases.
 
@@
 applicability and thorough parameters study to the best authors' knowledge is
-not yet done. One should acknowledge the works of
+not yet done. Especially that concerns the diffuse interface multiphase
+models, where one needs to resolve the film accordingly. One should acknowledge
+the works of
 \citet{pagonabarraga-fingers} on meniscus in thin films for the fingering
@@
 
+While to resolve interface on a fine level for capillary numbers less than
+$0.003$ is computationally extensive even for a two-dimensional case. For
+instance for capillary number $0.01$ to properly resolve the interface one
+needs around 22 GB memory requirement. Therefore, this work is based on the
+comparison with the already established results which were in their case
+validated over the Bretherton bubble flow. The work presents recipes for the
+initialization and the optimal parameter range for the binary liquid LBM model,
+which is described in the next section.
+
+\section{Lattice Botlzmann binary liquid model}
+The lattice Boltzmann equation operates on the rectangular grid representing the
+physical domain. It utilizes the
+probability distribution functions containing the information about the
+macroscopic variables, such as density,
+momentum. LBE consists from two parts: local collision and propagation operated
+from one node to another in the certain
+direction specified by the velocity set.  The lattice Boltzmann implementation
+in a standard view representing the
+collision and propagation steps:
+\begin{equation}
+\label{standard:implementation}
+\begin{aligned}
+&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
+F_i,&&\text{ collision step}\\
+&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
+\end{aligned}
+\end{equation}
+where $F_i$ is the external force population. The binary liquid LBM model is
+based on the free-energy functional \cite{swift,landau} and operates with two
+sets of populations one to track pressure and velocities fields, another one is
+for the phase field. The model to be utilized is two-dimensional nine-velocity
+D2Q9 model.
+
+In terms of D2Q9 the equilibrium populations are as \cite{pooley-contact}:
+\begin{equation}
+\begin{aligned}
+&f_i^{eq}=w_i 
+\biggl(3
+p_0 - k \phi \Delta \phi
++\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
+\beta}}{2 c_s^4}\biggr), i=1\div8\\
+&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
+&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i=1\div8 \\
+&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
+\end{aligned}
+\end{equation}
+where $\Gamma$ is the mobility parameter; chemical potential
+$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
+tension; $A$ is the parameter of the free-energy model; bulk pressure
+$p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. D2Q9 related parameters are
+$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
+\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and
+$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
+the sound speed parameter $c_s^2=\dfrac{1}{3}$. Weights related to the
+inclusion of the surface tension coefficient are as follows:
+$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
+$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
+$w^{xy}_{7-8}=-1/4$. Note that the system restores the macroscopic
+equations as:
+\begin{equation}
+\begin{aligned}
+&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
+&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
+-\partial_{\alpha}P_{\alpha \beta} +
+\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
++ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
+&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
+\end{aligned}
+\label{binary:fluid:system}
+\end{equation}
+where the viscosity $\nu=c_s^2 (\tau-1/2)$ and mobility parameter
+$M=\Gamma(\tau_{\phi}-1/2)$. The system allows the separation of the liquid
+phase with $\phi=1$ and so-called gas phase with $\phi=-1$. Note that the
+spatial variation of the relaxation parameter depends on the relaxation
+parameter of $\tau_{gas}$ and $\tau_{liq}$ linearly as
+$\tau=\tau_{gas}+\frac{\phi+1}{2}\tau_{liq}$.
+
 \section{Lattice Botlzmann benchmark}
-The purpose of this section is to establish the benchmark and perform
-parameters study for the classical Bretherton/Taylor problem. There are certain
-challenges for the lattice Boltzmann method which can be summarized as follows:
+To properly simulate and to compare simulations results with the already
+performed simulations one needs to design a benchmark by addressing the
+following challenges for the lattice Boltzmann method:
 \begin{itemize}

### 2010-10-08 Text/paper.tex
Modified sections about dimensionalization and initialization. Need to come out with nice tests.

@@
 parameter of $\tau_{gas}$ and $\tau_{liq}$ linearly as
-$\tau=\tau_{gas}+\frac{\phi+1}{2}\tau_{liq}$.
+$\tau=\tau_{gas}+\frac{\phi+1}{2}\tau_{liq}$. Note that the surface tension in
+the framework of binary liquid model as $\sqrt{8 k A}{9}$.
 
@@
 To properly simulate and to compare simulations results with the already
-performed simulations one needs to design a benchmark by addressing the
-following challenges for the lattice Boltzmann method:
+performed simulations one needs to design a numerical benchmark by addressing
+the following challenges for the lattice Boltzmann method:
 \begin{itemize}
  \item The diffusive interface nature of the multiphase model should not
-influence the film resolution.
+influence the film width results. This is obviously achieved only for certain
+resolutions of the diffuse interface in comparison with the film width.
  \item The spurious currents arising from the surface tension discretization
 can influence the mass transfer in the thin film therefore comprimising the
-overall solution.
+overall solution. Therefore the bubble should be long enough to decrease the
+influence of the spurious currents.
  \item The wettability coefficient which is in principle can control the
@@
 be large as it is indicated in \cite{pagonabarraga-fingers}, however no studies
-address this issue. Because the free-interface nature of the Bretherton/Taylor
-problem one can have difficulties with imposing the pressure boundary
-conditions which do require to impose two separate boundary conditions for two
-distribution functions.
- \item A number of simulations are done across the circular shape of channel,
+address this issue.
+  \item The Bretherton problem addresses the bubbles flow driven by pressure
+difference but not body force. However, it's relatively difficult to impose the
+pressure difference in the framework of the binary liquid model because two
+boundary conditions with the associated gradients and laplacians are required.
+The pressure boundary conditions for the binary liquid model are currently
+under development. Preliminary results show that the differences between
+pressure difference flows and body force driven flows are small for the film
+deposition thickness, the interest of this study. Thus, the work's focus is at
+the body force driven flows because of the better simulation stability and
+realization along with the periodic boundary conditions.
+ \item Most simulations are done across the circular shape of channel,
 which is quite difficult to address in terms of lattice Boltzmann free energy
@@
 however the analysis of error which can be addressed to the boundary conditions
-treatment becomes an issue. 
+treatment becomes an issue. Therefore the benchmark is for an axis-symmetric
+case. 
 \end{itemize}
 Given all the concerns and challenges the suggested Lattice Boltzmann framework
-is the two-dimensional flow driven by gravity or by pressure boundary
-conditions. The advantage of the gravity driven bubble is that it's easy to
-implement with the periodic boundary conditions. However, the analytical
+is the two-dimensional flow driven by a body force. However, the analytical
 solution to this problem, to the best authors' knowledge, is not yet known.
@@
 validate the Bretherton film thickness, the latter should be used with caution
-since the velocity of the bubble is different from the fluid velocity and there
-is also the additional pressure drop due to front and rear menisci.
+since the velocity of the bubble is different from the fluid velocity and a
+bubble shape
+is different in front and rear menisci. \citet{wong-films} indicated that the
+film width for polygonal capillaries
+changes from $Ca^{1/2}$ on distances $x\propto Ca^{-1}$ to $Ca^{2/3}$ for
+distances $x \gg Ca^{-1}$. Therefore, for a bubble to have a formed shape in
+the  front meniscus we picked up the bubble length as $5$ channel heights. The
+channel length is taken to minimize the influence of one bubble on another,
+because of the periodicity of boundary conditions. For the simulations the
+channel length is taken as $3$ bubble lenghts giving all together $15$ channel
+heights. Figure \ref{fig:benchmark:sketch} represents the benchmark sketch.
+\begin{figure}
+%\includegraphics[width=0.97\textwidth]{}
+\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
+\label{fig:benchmark:sketch}}
+\end{figure}
 
-The solution
-for relatively long bubbles driven by pressure specified on the inlet and the
-outlet coincides with the classical theory of Bretherton. However, to implement
-the pressure boundary conditions for the binary liquid model requires the
-specification of boundaries conditions for two populations, for the momentum and
-for the advection-diffusion equation (which are quite different). To the best
-authors' knowledge such the situation is not yet covered in the literature. We
-address this situation in Section \ref{appendix-pressure}.
-
-\section{The bubble motion in the planar case}
-This section covers the results for the pressure and force driven bubbles. The
-idea is to establish what parameters are needed, i.e. grid size, bubble length,
-viscosity ratio,to obtain the proper behaviour of Bretherton law.
-
-\subsection{The flow driven by the gravity}
-\citet{wong-films} indicated that the film width for polygonal capillaries
-changes from $Ca^{1/2}$ on distances $x\propto Ca^{-1}$ to $Ca^{2}{3}$ for
-distances $x \gg Ca^{-1}$. In simulations the bubble length is taken around 5-10
-diameters of the channel {\color{red} Why is that? Is it because of
-curvatures?}.
-
-Also, gravity or force driven bubble flow has periodical boundary conditions.
-Therefore, the situation to be simulated is the infinite train of bubbles
-coming one after another. Thus, to avoid the influence of bubbles on each other
-one needs to take quite long gap between bubbles. We take this gap around 3-5
-lengths of the bubble.
-
-\subsubsection{The influence of initial film width. Right parameters}
-It is important to show that the steady film width is independent on initial
-conditions where the most crucial condition is the initialized bubble width. If
-in steady state the bubble width addjust itself to reveal certain film width
-therefore the simulations are consistent. This section is to validate it.
-
-For the sake of simplicity we picked up the following parameters for the
-simulation:
+\section{Results}
+\subsection{The nondimensionalization and initialization procedure}
+This section is to introduce the procedure of nondimensionalization and
+initialization techniques to be able to predict parameters necessary for
+simulation. The capillary number governs the interface thickness and defined as:
+\begin{equation}
+Ca=\frac{U_{bubble} \mu_{liq}}{\gamma},
+\end{equation}
+where $\mu_{liq}$ is the liquid viscosity, $\gamma$ is the surface tension. The
+bubble velocity $U_{bubble}$ is taken in the center of the bubble where the film
+thickness is examined. 
+While parameters as $\mu_{liq}$, $\gamma$ and capillary number $Ca$ are given
+or defined through the model parameters, the velocity $U_{bubble}$ is not
+defined. In the simulation we approximately match the velocity with the
+Poiseuille profile center velocit, which is given by the equation:
 \begin{equation}
 \begin{aligned}
-&Ca=0.005\\
-&Ca=\frac{U_{bubble} \mu_{slug}}{\gamma}=0.005.
+U_{bubble} \approx U_{liq}=\frac{H^2}{8
+\mu_{liq}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
+U_{bubble}\approx
+\frac{{N_y}^2}{\frac{8}{6}(\tau_{liq}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
+x } ,
 \end{aligned}
 \end{equation}
-Some parameters are fixed with the model parameters, as surface tension:
+By taking the Poiseuille profile one can estimate the body force through the
+capillary number and grid resolution:
 \begin{equation}
-\gamma=\sqrt{\frac{8 k A}{9}}=\sqrt{\frac{8 0.04 0.04}{9}}=0.037712
+\begin{aligned}
+&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{N_y}^2}\gamma Ca.
+\end{aligned}
 \end{equation}
+In reality we iterate the body force value
+until the desired velocity and the capillary number as well are achieved.
 
-The film width for such small Capillary numbers is given by equation
-\cite{giavedoni-numerical}:
+After one resolves necessary parameters to run a simulation, one needs to
+initialize the macroscopic fields and populations.
+The velocity is initialized with $0$ everywhere. Populations are initialized
+with the equilbrium values for the binary liquid model including all the phase
+gradients and laplacians. The liquid phase is initialized with $1$ and the
+gas phase is initialized with $-1$. Though the initialization film thickness
+is not involved in the final result for the front meniscus, Fig.
+\ref{fig:different:initialization:widths}, we keep the initialized film
+thickness as close as possible to the already obtained numerical film
+thicknesses \cite{giavedoni-numerical}. Otherwise, the region where the
+Bretherton phenomena occurs is dillute, which is obviously seen for the
+simulation initialized with the film thickness as $Ny-28$ {\color{red} Michal,
+can you please prepare the simulations with the different initialization. Or
+just delete different thicknesses initialization}.
+\begin{figure}
+\includegraphics*[bb=18 350 594 450,width=0.97\textwidth]
+{Figures/force_width_6.eps}
+\includegraphics*[bb=18 350 594 450,
+width=0.97\textwidth]{Figures/force_width_10.eps}
+\includegraphics*[bb=18 350 594 450,
+width=0.97\textwidth]{Figures/force_width_14.eps}
+\caption{\color{red} The phase plots for different bubble widths as
+$Ny-12$,$Ny-20$ and
+$Ny-28$. Michal, could you please prepare nice pictures for
+simulations. Indicate as well the
+grid and time steps\label{fig:different:initialization:widths}}
+\end{figure}
+
+Here are two different examples presented that can quantitatively show the
+initialization procedure for quite different capillary numbers. In these
+examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
+That explicitly implies the surface tension to be $\sqrt{\frac{8 k
+A}{9}}=0.0377$ and the characteristic length of the interface
+$4\sqrt{\frac{k}{A}}=4$ lattice boltzmann units. For the sake of simplicity and
+stability the relaxations parameter for liquid and gas phases are taken as
+$\tau_{liq}=2.5$ and $\tau_{gas}=0.7$. That gives the gas-liquid viscosity ratio
+as $10$.  
+\begin{description}
+ \item[I \bm{$Ca=0.005$}]
+  The film width for such small Capillary numbers \cite{bretherton} is
+given by equation:
   \begin{equation}
-\frac{h_{\inf}}{H}=1.3375 Ca^{2/3}=0.03899
+  \frac{h_{\infty}}{H}=1.3375 Ca^{2/3}=0.03899,
   \end{equation}
-where we assume that $2 H$ or $Ny$ in lattice units is the channel width. That
-means that one needs to resolve the interface which is around $2$ percent of
-whole channel width with for instance $4$ lattice units. Therefore the channel
-width is $200$ lattice units. We take the bubble length as $5$ channel widths
-and the distance between slugs we take as $2$ bubble lengths. This will give
-right away the grid size as $200\mathrm{x}3000$. Next step is to calculate the
-velocity. The velocity can be calculated from the Capillary number as:
+  where we assume that $2 H$ or $Ny$ in lattice units is the channel width.
+Therefore one needs to resolve the interface which is around $2$ percent of
+  whole channel width with for instance $8$ lattice units, where $4$ lattice
+units occupies the interface width for given. Note that the interface occupies
+$50$ percent of the interface.  Therefore the channel
+  width is $400$ lattice units. We take the bubble length as $5$ channel widths
+  and the distance between bubble is as $2$ bubble lengths. This will give
+  right away the grid size as $400\mathrm{x}6000$. In terms of computer memory
+requirement the calculations with the double precision accuracy demand:
 \begin{equation}
-U_{bubble}=0.005 \frac{\gamma}{\mu_{slug}}=0.005 \frac{0.037712}{0.66666}=2.82\,
-10^{-4},
+\begin{aligned}
+&2\text{ set of populations }\mathrm{x}\,\, 8\,\, \text{bytes per population
+}\mathrm{x}\,\, 9 \text{ number of
+populations}\\
+&\mathrm{x}\,400\,\mathrm{x}\,6000\approx 345 \mathrm{MB}. 
+\end{aligned}
+\end{equation}
+Even with this naive assumption for the $50$ percent resolution of the film
+widht in comparison with the interface width, one can get the memory
+requirement as $345 MB$.
+
+After resolving the interface one can  calculate the
+  velocity from the capillary number as:
+  \begin{equation}
+  U_{bubble}=0.005 \frac{\gamma}{\mu_{slug}}=0.005
+\frac{0.0377}{0.66666}=2.82\,10^{-4},
   \end{equation}
@@
   \begin{aligned}
-\frac{\mathrm{d}P}{\mathrm{d}x}=3.77\, 10^{-8}.
+  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42\, 10^{-9}.
   \end{aligned}
   \end{equation}
+  Note that in the case of large grids  height resolution is directly
+involved in	
+the physical time step as:
+\begin{equation}
+\Delta t =\frac{U_{LBM}}{U_{phys}} \Delta x ,
+\end{equation}
+where for example the real velocities in the physical world can reach
+approximately $1$ m/s and physical channel height upto $1$ mm. In this case one
+can easily approximate the physical time step:
+\begin{equation}
+\Delta t = 7.05 \, 10^{-10} \,\mathrm{s}
+\end{equation}
+which is a small time step. The typical simulation needs approximately
+$100,000$ multiplication time steps to come to the steady state solution.
 
-\subsubsection{The influence of initial film width. Easy parameters to check}
-It is important to show that the steady film width is independent on initial
-conditions where the most crucial condition is the initialized bubble width. If
-in steady state the bubble width addjust itself to reveal certain film width
-therefore the simulations are consistent. This section is to validate it.
-
-The parameters in the previous subsection imply very long simulations. Even
-Bretherton delivered his solution for the range of Capillary number less than
-$0.005$. However, a lot of simulations devoted to the Bretherton problem
-simulated extended it in the range of Capillary numbers greater than $0.005$
+ \item[II $\bm{Ca=0.05}$] 
+   The parameters in the previous subsection imply very long
+simulations. Bretherton delivered his solution for the range of Capillary number
+less than
+  $0.005$. However, a lot of simulations devoted to the Bretherton/Taylor
+flow
+  simulated and extended it in the range of Capillary numbers greater than
+$0.005$
   and for different Reynolds numbers. One can refer to figures
@@
   \cite{giavedoni-numerical,heil-bretherton} which crossvalidated their
-simulations with Bretherton and each other. As far as this section is intended
-to show physics behavior the moderate for simulation parameters are taken which
+  simulations with Bretherton and each other. Even though LBM can simulate the
+flows with small capillary numbers, i.e. less than $0.005$, but waiting time
+for the simulation to be accomplished is relatively long. Therefore, we focus
+our work on the simulations of \citet{giavedoni-numerical} which crossvalidated
+results not only for small capillary numbers but for large capillary numbers as
+well. As far as the work intention is
+  to show proper physics behavior the moderate parameters are
+taken which
   can be easily and fast validated on a computer. 
+  One can see from the Fig. \ref{fig:heil:planar} that the width for the
+  Capillary number equal to $0.05$ ranges from $0.12$ to $0.15$ while Reynolds
+  number ranges from $0$ to $300$. This capillary number has moderate film
+widths
+  to be resolved with certain accuracy.
+
+  The film width for such small Capillary numbers is taken from
+  \ref{fig:heil:planar} for Reynolds number $40$ equals to $0.12$. We want to
+  resolve $6\%$ of whole channel width with $12$ lattice units, where $4$
+lattice units are related to the interface resolution.  Therefore the channel
+  width is $200$ lattice units. We take the bubble length as $5$ channel widths
+  and the distance between slugs we take as $2$ bubble lengths. This will give
+  right away the grid size as $200\mathrm{x}3000$. The velocity can be
+calculated from the Capillary number as:
+  \begin{equation}
+  U_{bubble}=0.05 \frac{\gamma}{\mu_{liq}}=0.05 \frac{0.037712}{0.6666}=2.82\,
+  10^{-3}.
+  \end{equation}
+  The velocity implies the following pressure gradient calculated from the
+  Poisuielle flow solution:
+  \begin{equation}
+  \begin{aligned}
+  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca=3.77\,10^{-7}.
+  \end{aligned}
+  \end{equation}
+  The initialization approach is to take values from the already performed
+numerical simulation and to calculate the grid numbers from them matching the
+body force gradient through the solution of Poiseuille flow.
+
+\end{description}
 \begin{figure}
@@
 \end{figure}
+Note that we base the initialization techniques through correlations for the
+capillary number and the body force. However, in the binary liquid framework
+those correlations give only the approximation. First of all the body force
+which is based on the Poiesuille velocity profile for the liquid phase is the
+biggest assumption. In reality the less viscous bubble will have a slippage
+factor and its velocity is larger than the surrounding liquid velocity. By
+plugging in the obtained velocity back to the capillary number equation the
+larger capillary number can be obtained. While we present the initialization
+tips in reality one needs to take the bubble velocity from the simulations and
+recalculate all the necessary characteristics.
 
-One can see from the Fig. \ref{fig:heil:planar} that the width for the
-Capillary number equal to $0.05$ ranges from $0.12$ to $0.15$ while Reynolds
-number ranges from $0$ to $300$. This capillary number has moderate film widths
-to be resolved with certain accuracy. Let us recalculate all the necessary
-parameters:
-\begin{equation}
-\begin{aligned}
-&Ca=0.05\\
-&Ca=\frac{U_{bubble} \mu_{liq}}{\gamma}=0.05.
-\end{aligned}
-\end{equation}
-Some parameters are fixed with the model parameters, as surface tension:
-\begin{equation}
-\gamma=\sqrt{\frac{8 k A}{9}}=\sqrt{\frac{8 0.04 0.04}{9}}=0.037712
-\end{equation}
-The film width for such small Capillary numbers is taken from
-\ref{fig:heil:planar} for Reynolds number $40$ equals to $0.12$. We want to
-resolve $12\%$ with $6$ lattice units. Therefore the channel
-width is $50$ lattice units. We take the bubble length as $5$ channel widths
-and the distance between slugs we take as $2$ bubble lengths. This will give
-right away the grid size as $50\mathrm{x}750$. Next step is to calculate the
-velocity. The velocity can be calculated from the Capillary number as:
-\begin{equation}
-U_{bubble}=0.05 \frac{\gamma}{\mu_{liq}}=0.05 \frac{0.037712}{0.6666}=2.82\,
-10^{-3},
-\end{equation}
-where $\mu_{liq}=1/3 (2.5-0.5)=0.6666$. This implies the viscosity ratio to be
-equal $R_{\nu}=2/0.2=10$ if we take the relaxation parameter for gas as
-$\tau_{gas}=0.7$.
-This velocity implies the following pressure gradient calculated from the
-Poisuielle flow solution:
-\begin{equation}
-\begin{aligned}
-&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca=6.003\,10^{-6}.
-\end{aligned}
-\end{equation}
+\subsubsection{The influence of initial film width. Easy parameters to check}
 
-Some pictures after $40000$ steps for the gravity driven flow and for different
-initialization of bubble width as $Ny-12$, $Ny-20$, $Ny-28$.
-\begin{figure}
-\includegraphics*[bb=18 350 594 450,width=0.97\textwidth]
-{Figures/force_width_6.eps}
-\includegraphics*[bb=18 350 594 450,
-width=0.97\textwidth]{Figures/force_width_10.eps}
-\includegraphics*[bb=18 350 594 450,
-width=0.97\textwidth]{Figures/force_width_14.eps}
-\caption{The phase plots for different bubble widths as $Ny-12$,$Ny-20$ and
-$Ny-28$.}
-\end{figure}
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_6.eps}
-\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_10.eps}
-\begin{center}
-\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_14.eps}
-\end{center}
-\caption{The phase profiles for different initialization parameters as the
-bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
-\label{fig:phase:profiles:different:widths:force}}
-\end{figure}
-
-One can see from Figures \ref{fig:phase:profiles:different:widths:force} that
-the film pattern in the beginning of the slug is almost independent of the
-initial conditions and it ranges from 4 to 5 lattice boltzmann units, where the
-is taken where the phase field has a value of zero. This preliminary result
-justify that lattice Boltzmann method is able to predict the film thickness.
-
-However, with the smaller initialization of the bubble one can see that the
-front meniscus where the film thickness is validated accross is decreasing.
-Therefore to be on the safe side one needs to initialize bubbles to be near the
-analytical length.
-
-We assumed that the interface thickness should be around $6$ lattice boltzmann
-units, but we assume the calculations to be based on the $U_{slug}$ taken from
-the simple Poisuille flow velocities. In reality the slug has slippage factor
-with with the flow and velocity is slightly bigger than the liquid media. One
-can take the velocity of the bubble from the simulations and recalculate all
-the necessary lengths. Velocity profiles behind the bubble and in the front
-meniscus can be seen in Fig. \ref{fig:velocity:profiles:different:widths:force}
+ Velocity profiles behind the bubble and in the front
+  meniscus can be seen in Fig.
+\ref{fig:velocity:profiles:different:widths:force}
   and \ref{fig:velocity:profiles:different:widths:pressure} for the force driven
@@
   \end{figure}
-
-While the shapes of droplets are different for the force and pressure
-difference driven flows the velocity profiles show that the force driven bubble
-move faster than in the case of pressure difference driven flow:
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_6.eps}
-\hfill
-\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_10.eps}
-\\
-\begin{center}
-\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_14.eps}\\
-\end{center}
-\caption{Comparison between velocity bubble
-profiles.\label{fig:velocity:comparison:different:widths}}
-\end{figure}
-
-Note that in the case of pressure driven flows the termination bulk and bubble
-velocity have more consistent behavior and the values are not different too
-much. That is one of the reasons why in the case of the force driven flows the
-width has a bit different values depending on the initialization procedure.
 

### 2010-10-13 Text/paper.tex
More work on the paper.

@@
 \begin{equation}
+\label{poiseuille:velocity:center}
 \begin{aligned}
@@
 A}{9}}=0.0377$ and the characteristic length of the interface
-$4\sqrt{\frac{k}{A}}=4$ lattice boltzmann units. For the sake of simplicity and
+$5\sqrt{\frac{k}{A}}=5$ lattice boltzmann units. For the sake of simplicity and
 stability the relaxations parameter for liquid and gas phases are taken as
@@
 Therefore one needs to resolve the interface which is around $2$ percent of
-  whole channel width with for instance $8$ lattice units, where $4$ lattice
+  whole channel width with for instance $8$ lattice units, where $5$ lattice
 units occupies the interface width for given. Note that the interface occupies
-$50$ percent of the interface.  Therefore the channel
+$62.5$ percent of the interface.  Therefore the channel
   width is $400$ lattice units. We take the bubble length as $5$ channel widths
-  and the distance between bubble is as $2$ bubble lengths. This will give
-  right away the grid size as $400\mathrm{x}6000$. In terms of computer memory
+  and the distance between bubble is as $2$ bubble lengths. It gives the
+grid size as $400\mathrm{x}6000$. In terms of computer memory
 requirement the calculations with the double precision accuracy demand:
@@
   \ref{fig:heil:planar} for Reynolds number $40$ equals to $0.12$. We want to
-  resolve $6\%$ of whole channel width with $12$ lattice units, where $4$
+  resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
 lattice units are related to the interface resolution.  Therefore the channel
   width is $200$ lattice units. We take the bubble length as $5$ channel widths
-  and the distance between slugs we take as $2$ bubble lengths. This will give
-  right away the grid size as $200\mathrm{x}3000$. The velocity can be
+  and the distance between slugs we take as $2$ bubble lengths. That gives
+the grid size as $200\mathrm{x}3000$. The velocity can be
 calculated from the Capillary number as:
@@
 larger capillary number can be obtained. While we present the initialization
-tips in reality one needs to take the bubble velocity from the simulations and
-recalculate all the necessary characteristics.
-
-\subsubsection{The influence of initial film width. Easy parameters to check}
-
- Velocity profiles behind the bubble and in the front
-  meniscus can be seen in Fig.
-\ref{fig:velocity:profiles:different:widths:force}
-  and \ref{fig:velocity:profiles:different:widths:pressure} for the force driven
-  and pressure difference driven flows, correspondingly. As one can see average
-  velocity for the bubble and the slug larger than the calculated velocity which
-  should give larger the film. However, the film width for the planar case is
-  different and can have another coefficient in front of the proportionality to
-  $Ca^{2/3}$. Also, the phase gradient is taken as $0$ right now which can have
-  influence on the measurements as well.
-  \begin{figure}
-  \includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_6.eps}
-  \hfill
-  \includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_6.eps}\\
-  \includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_10.eps}
-  \hfill
-  \includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_10.eps}\\
-  \includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_14.eps}
-  \hfill
-  \includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_14.eps}\\
-  \caption{The velocity profiles in bulk and bubble for different initialization
-  parameters as the
-  bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
-  \label{fig:velocity:profiles:different:widths:force}}
-  \end{figure}
+tips, in reality one needs to take the bubble velocity from the simulations and
+recalculate all the necessary characteristics or in other words adjust the
+forcing to obtain the desired velocity/capillary number.
 
 \subsection{Grid refinement}
-To examine the interface width we performed grid refinement for both force and
-pressure difference driven flows. To avoid the effects of rear and front
-menisci the problem is initialized with the film width occupying $12$ percent
-of the channel height. All other parameters are kept the same, that includes
-keeping the same Capillary number and viscosities ratio. Therefore the
-following parameters were utilized: $Ny=51,100,149,198$,corresponding to the
-effective channels widths as $H_{eff}=49,98,147,196$, $Nx=736,1471,2206,2941=15
+To properly estimate the interface resolution one needs to perform the
+convergence study dependance on the grid resolution. All parameters are fixed,
+including a bubble velocity, the capillary number and change the grid
+resolution. Our goal is to know exactly the ratio of the interface width to the
+film thickness where results are not dependant on the grid resolution.
+
+The example what we want to simulate is the capillary number
+$0.05$. As it was pointed out the film thickness in that case is $6$ percent.
+Note that in the case of half-way bounce-back walls which are used in the
+simulations one needs to calculate the film thickness as: 
+\begin{equation}
+\delta=(\phi_0-0.5)/(N_y-2),
+\end{equation}
+where $\phi_0$ is the grid coordinate where phase field equals to $0$, $N_y-2$
+is related to the effective channel height. Note that it's a big
+simplifications to impose the boundary in the middle beetween the bounce-back
+node and the fluid node. In the case of $\delta=0.06$ the initial grid number
+is $N_y=102$, which gives the horizontal grid number as $N_x-1=15(N_y-2)=1500$.
+The bubble is initialized as rectangular box with coordinates
+$y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
+$\phi_{bubble}=-1$. All other nodes are initialized with the phase field
+$\phi=1$. The force gradient can be estimated through the Poiseuille
+profile formula, Eq.\ref{poiseuille:velocity:center}, and it equals
+$1.508\,10^{-6}$.
+
+After choosing the reference parameters, the grid refinement needs to keep the
+same the same macroscopic parameters as capillary number and viscosities ratio.
+Therefore the simulations were performed with grid
+numbers $Ny=102,127,152,177,202,227$, corresponding to the
+effective channels widths as $H_{eff}=100,125,150,175$,
+$Nx=1501,1876,2251,2625,3000,3375=15
 H_{eff}$. It is easy to check
-that keeping all liquid viscosities the same yields the following to be
+that keeping all liquid viscosities the same yields the following quantity to
+be a
 constant while doing the grid refinement
-$H^2\frac{\mathrm{d}P}{\mathrm{d}x}=Ny^2\frac{\mathrm{d}P}{\mathrm{d}x}=Const$.
+$H_{eff}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
+x } =Const$.
+{\color{red} Michal, please add some results here and description to it as well.
 The results for the force driven flows velocity phase profiles are shown in
@@
 slug are $0.003411444831,0.003412693869,0.003420338754,0.003425757312$ and
-Capillary number defined through these velocities is $0.0602$. Those numbers
-are adequate if we assume that the results presented in
-Fig.\ref{fig:giavedoni:planar} and Fig. \ref{fig:heil:planar} correspond to the
-half of the channel but not the whole channel.
-
+Capillary number defined through these velocities is $0.0602$. 
+}
 Let us calculate to which extent we need to refine the interface to obtain
@@
 $\xi=\sqrt{k/A}=1$. This interface width divided to the effective channel width
-($5\xi/H_{eff}$) is as follows $0.102,0.051,0.034,0.025$ for the given
-$H_{eff}$. However, the interface width needs to be takes as the ratio with the
-film width which we want to resolve, this ratio $5\xi/H_{film}$ is as follows
-$1.2605,  0.7205,  0.501,  0.3839$. In reality this number is
-a bit overestimated as the interface doesn't fill the influence of the wall and
-it maintains all the macroscopic parameters jumps. Therefore, one needs to
-resolve the interface as $40\%$ of the expected film width.
+($5\xi/H_{eff}$) is as follows {\color{red} Please add new values here
+$0.102,0.051,0.034,0.025$ } for the given
+$H_{eff}$. Let us calculate the interface width ratio to the film thickness as
+$5\xi/H_{film}$ is as follows
+{\color{red} Please add here new values $1.2605,  0.7205,  0.501,  0.3839$}.
+Therefore, one needs to
+resolve the interface as {\color{red} Please give number here $40\%$} of the
+expected film width for simulations to be converged.
+{\color{red} If we need this pictures???
 \begin{figure}
@@
 \includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_196.eps}\\
-\caption{Grid refined profiles for the effictive channel widths
+\caption{Do we need these pictures? Grid refined profiles for the effictive
+channel widths
 $H_{eff}=49,98,147,196$.\label{fig:grid:profiles}}
 \end{figure}
+}
+
+\section{Capillary number region}
 
@@
 
-\section{Capillary number region}
 This section is devoted to the study of different Capillary number. We refer to

### 2010-10-14 Text/paper.tex
Specified all the necessary benchmarks.

@@
 which is a small time step. The typical simulation needs approximately
-$100,000$ multiplication time steps to come to the steady state solution.
+{\color{red} $100,000$} time steps to come to the steady state solution.
 
@@
 Therefore the simulations were performed with grid
-numbers $Ny=102,127,152,177,202,227$, corresponding to the
-effective channels widths as $H_{eff}=100,125,150,175$,
-$Nx=1501,1876,2251,2625,3000,3375=15
-H_{eff}$. It is easy to check
+numbers $N_y=102,127,152,177,202,227$, corresponding to the
+effective channels widths as $H_{eff}=100,125,150,175,200,225$,
+$N_x=1501,1876,2251,2626,3001,3376=15
+H_{eff}+1$. It is easy to check
 that keeping all liquid viscosities the same yields the following quantity to
@@
 
+\subsection{The influence of the wall gradient}
+Once the procedure of how to resolve the interface gradient is known, we want
+to examine the influence of the phase gradient, i.e. the wettability of the
+wall, to the film thickness. We examined the
+wall gradient influence on
+the standard grid $202\mathrm{x}3001$ and the initial film width as
+{\color{red} Take the value from previous calculations $20$}
+lattice boltzmann units. We examined $11$ different values for the wall
+gradient ranging from $-1$ to $1$. The results are consistent among all the
+wall gradients given that the interface is resolved in the right way. For the
+given wall gradients the values of interface thicknesses are
+{\color{red} Change these numbers $0.0709130255276, 0.0708903443305,
+0.0708661222722,
+0.0708369188112, 0.0707962325316, 0.0707266326981, 0.0705839443249,
+0.0702489349415, 0.0697553433989$}. For the wall gradients $0.8$ and $1.0$ the
+values are NAN. The negative values of the phase gradient are preferable, as
+far as the phase of the liquid adjacent to the wall has the value of $1$. That
+means that the phase values near the wall are above $1$. In case of positive
+gradients the values near the wall are below $1$ and the gradient profile can
+fuse with the values of the gas phase as $-1$, creating the wall stickyness,
+unphysical one. {\color{red} Michal, if you want to add some pictures go ahead
+with it}.
+
+\section{The influence of different viscosities}
+The results obtained earlier are taken for the liquiqs viscosity ratio $10$.
+We performed the same simulations for the liquids viscosity ratio as $20$.
+Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. 
+{\color{red} The results showed that blah-blah-blah - I guess will be the same}.
+
 \section{Capillary number region}
+The purpose of this section is to validate the correlations of
+\citet{giavedoni-numerical}, Fig. \ref{fig:giavedoni:planar}, and
+\citet{heil-bretherton}, Fig. \ref{fig:heil:planar}, for the range of capillary
+numbers. Because of limiting computational factors ({\color{red} What is the
+least capillary number which is OK?}) we cannot go to the range of small
+capillary numbers. However, all the other methods were validated across the
+small capilary numbers and extended for the range of capillary numbers from
+$0.005$ to $1.0$. Our goal is to compare the film thickness over the capillary
+range from $0.03$ to $1.0$.
 
-{\color{red} The next step is to perform calculations for different Capillary
-numbers to restore necessary characteristics as it's seen in Fig.
-\ref{fig:giavedoni:planar} and \ref{fig:heil:planar}. Also note that grid
-refinement is done for the gravity driven flows, which can have pretty
-different characteristics in comparison with pressure simulations (unstable
-for the grid refinement - need to adjust parameters)}.
-
-This section is devoted to the study of different Capillary number. We refer to
+To perform numerical simulations we refer to
 grid refinement section where we obtained that the interface width should be
-at most $40$ percent of the interface width. For the benchmark results we
+at most {\color{red} $40$} percent of the interface width. For the benchmark
+results we
 suggest the following Capillary numbers as:
 \begin{equation}
-0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0
+0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0\, .
 \end{equation}
 The corresponding approximate film widths from \cite{giavedoni-numerical} are as
-follows (we take the half width as it's done in the grid refinement section):
+follows:
 $0.04,0.06,0.08,0.1,0.12,0.13,0.15$. $5$ lattice boltzmann units occupy
-$40\%$ of the interface thickness. Therefore the grid size should be as
-following $314\mathrm{x}4681$,$210\mathrm{x}3121$,
+{\color{red} Could be another number $40\%$} of the interface thickness.
+Therefore the grid size should be as
+following {\color{red} Michal, different numbers would be here.
+$314\mathrm{x}4681$,$210\mathrm{x}3121$,
 $158\mathrm{x}2341$, $127\mathrm{x}1876$, $106\mathrm{x}1561$,
-$98\mathrm{x}1441$,$85\mathrm{x}1246$.
+$98\mathrm{x}1441$,$85\mathrm{x}1246$ Michal, for the larger capillary
+numbers such small grid is not
+appropriate because of the instabilities.
+}.
 
+{\color{red} Probably we don't need the explanations about initialization here,
+just to present results}
 To utilize previous results to properly initialize simulations to run one need
-to do some modifications. The result is that for the grid with $H_{eff}=196$ we
-obtained the film width as $0.0685$ and velocity in the center of the bubble as
-$0.0045$. The force gradient was $6\, 10^{-6} /16$. The predicted Capillary
-number was $0.05$ and the actual one is $0.0795$. But we performed simulations
+to do some modifications. The result is that for the grid with $H_{eff}=200$ we
+obtained the film width as {\color{red} Will be another number $0.0685$} and
+velocity in the center of the bubble as
+{\color{red} Will be another number $0.0045$}. The force gradient was
+{\color{red} Probably the same number $6\, 10^{-6} /16$}. The predicted
+Capillary
+number was $0.05$ and the actual one based on the center velocity calculation is
+{\color{red} Will be another number $0.0795$}. However, we performed simulations
 according the assumed value of $0.05$. Let us perform some correlations to
@@
 &Ca_{lit} \propto U_{slug}\\
-&U_{slug} \propto \frac{\mathrm{d}P}{\mathrm{d}x} Ny^2 
-&Ca_{lit} \propto \frac{\mathrm{d}P}{\mathrm{d} x} Ny^2 \text{ or }\\
-\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{Ny^2}
+&U_{slug} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\ 
+&Ca_{lit} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
+&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{Ny^2},
 \end{aligned}
 \end{equation}
-Therefore the initialized consistent pressure gradients can be taken from
+where the index $lit$ stands for the presented in the literature in
+\cite{giavedoni-numerical,heil-bretherton}. Therefore the initialized consistent
+pressure gradients can be taken from
 previous consistent simulations:
@@
 \frac{\mathrm{d}P}{\mathrm{d} x}=6\,10^{-6}/16 \frac{Ca_{literature}}{0.05}
-\frac{196^2}{Ny^2}=0.28812 \frac{Ca_{literature}}{Ny^2}
+\frac{200^2}{N_y^2}=0.3 \frac{Ca_{literature}}{N_y^2}
 \end{equation}
-Note that $Ny$ is calculated through the expected resolution of the interface
+Note that $N_y$ is calculated through the expected resolution of the interface
 in comparison with the interface thickness which itself is based on
-Capillary number simulations. In terms of time the steady state succesfull
-simulation for the grid $196$ was obtained after $200,000$ steps. The physical
+Capillary number simulations. In terms of time the steady state successful
+simulation for the grid $202\mathrm{x}3001$ was obtained after $200,000$ steps.
+The physical
 time is as follows:
@@
 \begin{aligned}
-&\Delta t=\frac{U_{slug,LB} \Delta X}{U_{phys}} \propto \frac{1}{Ny}\\
-&N_{iter}=200,000 \frac{Ny}{196}\approx 1020 Ny\\
-&N_{output}=51 Ny\\
+&\Delta t=\frac{U_{slug,LB} \Delta X}{U_{phys}} \propto \frac{1}{N_y-2}\\
+&N_{iter}=200,000 \frac{N_y-2}{200}\approx 1000 N_y\\
+%&N_{output}=51 Ny\\
 \end{aligned}
@@
 
-Steady simulations were obtained for the following Capillary numbers
-$0.03,0.05,0.08,0.1$ with the film thicknesses as
-$0.0337433850478, 0.063192148642, 0.0799078589211, 0.0891674460655$ which are
+{\color{red} That's why we need better grid for larger Capillary numbers. Steady
+simulations were obtained for the following Capillary numbers
+$0.03,0.05,0.08,0.1$  }with the film thicknesses as
+{\color{red} $0.0337433850478, 0.063192148642, 0.0799078589211,
+0.0891674460655$} which are
 quite close to the predicted one $0.04,0.06,0.08,0.1$. The thing is to identify
@@
 0.067646144698590829,0.1150166512524845,0.14732373700007995$, which again a bit
-overpredictive but consistent. The bulk liquid velocities for the simulation
-are $0.001947507726,0.003364703007,0.005483194352,0.00687516499$. The capillary
-numbers associated with those velocities are slightly less than with the bubble
-are as follows:$0.034427397986694822, 0.059480107823211698,
-0.096930097721574432, 0.12153689465513355$. We can define the average capillary
-number defined through the bubble capillary and slug capillaries numbers as
-$Ca_{aver}=\frac{Ca_{bubble}+Ca_{slug}}{2}$. The fig.
+overpredictive but consistent. The fig.
 \ref{fig:capillary:simple} represents the thickness against the Capillary
@@
 \includegraphics[width=0.47\textwidth]{Figures/capillary_range.eps}
-\caption{Thickness versus the capillary number for different scalings based on
+\caption{{\color{red} This one to eliminate - just present here
+comparison between our simulations in the longer range of capillary
+numbers and Giavedoni}Thickness versus the capillary
+number for different scalings based on
 the average capillary number. \label{fig:capillary:simple}}
@@
 \subsection{The influence of the force}
+
+{\color{red} Michal if you find it useful - use this section. My suggestion is
+to eliminate it.}
 \begin{description}
@@
 $0.0700122222668,0.0985152303286$ with the velocities associated as
-$0.004490013864,0.009597299475$ and slug velocities as
-$0.003861354387,0.006350100792$. Those numbers correspond to the bubble
-capillary numbers as $0.079372981271400328, 0.16965788849626481$ and slug
-capillary numbers as $0.068259746790303111, 0.11225498328103167$. The average
+$0.004490013864,0.009597299475$. Those numbers correspond to the bubble
+capillary numbers as $0.079372981271400328, 0.16965788849626481$. The average
 capillary numbers are $0.07381636,0.14095644$.
@@
 $0.002592856205,0.003077834398,0.003562975124,0.004034533823,0.004470657603$.
-Slug velocities are
-$0.002337542358,0.002752687109,0.003158125555,0.003552413499,0.003926980914$.
 Bubble Capillary numbers are $0.04583566,  0.05440894,  0.0629851 , 
-0.07132116, 0.07903081$; slug Capillary numbers are $0.0413223 ,  0.04866109, 
-0.0558283, 0.06279839, 0.06941987$. The average capillary numbers are
-$0.04357898, 0.05153502, 0.0594067, 0.06705977, 0.07422534$.
+0.07132116, 0.07903081$.
 \end{description}
 
-All the range of available Capillary numbers are shown in Fig.
-\ref{fig:capillary:simple}. The figure looks consistent with different
-scalings and provide good results.
+\section{Conclusion}
+The work presents the numerical tips and the benchmark for the
+Bretherton/Taylor problem using the binary liquid lattice Botlzmann method. The
+results show consistency with the already obtained numerical correlations using
+other methods. Though results are specific for the binary liquid lattice
+Botlzmann method but numerical tips and correlations can be used for any
+continuous interface methods. The work is valuable for people simulating
+microchannel flows.
 
-\subsection{The influence of the wall gradient}
-The influence of the wall gradient. We examined the wall gradient influence on
-the standard grid $198\mathrm{x}2941$ and the initial film width as $20$
-lattice boltzmann units. We examined $11$ different values for the wall
-gradient ranging from $-1$ to $1$. The results are consistent among all the
-wall gradients given that the interface is resolved in the right way. For the
-given wall gradients the values of interface thicknesses are
-$0.0709130255276, 0.0708903443305, 0.0708661222722,
-0.0708369188112, 0.0707962325316, 0.0707266326981, 0.0705839443249,
-0.0702489349415, 0.0697553433989$. For the wall gradients $0.8$ and $1.0$ the
-values are NAN. The negative values of the phase gradient are preferable, as
-far as the phase of the liquid adjacent to the wall has the value of $1$. That
-means that the phase values near the wall are above $1$. In case of positive
-gradients the values near the wall are below $1$ and the gradient profile can
-fuse with the values of the gas phase as $-1$, creating the wall stickyness,
-unphysical one.
-
-\subsection{The flow driven by the pressure difference}
-The details of imposing pressure boundaries for the binary liquid model is
-given in Section \ref{appendix-pressure}.
-
-\subsubsection{The influence of initial bubble width. Easy parameters to check}
-Here we impose absolutely the same parameters as for gravity driven flow.
-However instead of pressure gradient one needs to put different densities at
-the inlet and the outlet which can be easily calculated from the pressure
-gradient as follows:
-\begin{equation}
-\begin{aligned}
-&\rho_{inlet}=1+3 Nx \frac{\mathrm{d}P}{\mathrm{d}x}=3\, 750\, 6\,
-10^{-6}=1.0135\\
-&\rho_{outlet}=1.
-\end{aligned}
-\end{equation}
-
-Some pictures after $40000$ steps for the pressure driven flows for different
-initialization of bubble width as $Ny-12$, $Ny-20$, $Ny-28$.
-\begin{figure}
-\includegraphics*[bb=18 350 594 450,
-width=0.97\textwidth]{Figures/pressure_width_6.eps}
-\includegraphics*[bb=18 350 594 450,
-width=0.97\textwidth]{Figures/pressure_width_10.eps}
-\includegraphics*[bb=18 350 594 450,
-width=0.97\textwidth]{Figures/pressure_width_14.eps}
-\caption{The phase plots for different bubble widths as $Ny-12$,$Ny-20$ and
-$Ny-28$.}
-\end{figure}
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/plane_pressure_width_6.eps}
-\includegraphics[width=0.47\textwidth]{Figures/plane_pressure_width_10.eps}
-\begin{center}
-\includegraphics[width=0.47\textwidth]{Figures/plane_pressure_width_14.eps}
-\end{center}
-\caption{The phase profiles for different initialization parameters as the
-bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
-\label{fig:phase:profiles:different:widths:pressure}}
-\end{figure}
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_pressure_width_6.eps}
-\hfill
-\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_pressure_width_6.eps}\\
-\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_pressure_width_10.eps}
-\hfill
-\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_pressure_width_10.eps}\\
-\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_pressure_width_14.eps}
-\hfill
-\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_pressure_width_14.eps}\\
-\caption{The velocity profiles in bulk and bubble for different initialization
-parameters as the
-bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
-\label{fig:velocity:profiles:different:widths:pressure}}
-\end{figure}
-
-\section{Appendix I. Pressure boundaries conditions}
-\label{appendix-pressure}
-Bulk pressure for the binary liquid model is a function not only of density
-but the phase field as well. Thus, one of the possible ways to fix the pressure
-boundary condition is to fix density and phase at the inleat and at the outlet.
-The pressure difference is introduced through the difference in densities
-keeping phases at the inlet and at the outlet the same. One can introduce the
-pressure difference through the phase, however it's unphysical since the phase
-field is governed by passive advection-diffusion equation, where velocity is
-given from the hydrodynamic simulations.
-
-We examined many different pressure boundaries conditions, including Zou-He BC
-\cite{zouhe-boundary,harting-pressure}, equilibrium conditions for
-the phase field, etc. However, the stable situation was obtained only with the
-boundary conditions by \citeauthor{guo-pressure} \cite{guo-pressure,
-guo-boundary-conditions}. The idea of the boundary conditions is simple: the
-coming populations to domain are decomposed to the equilibrium and
-non-equilibrium part:
-\begin{equation}
-f_i=f_i^{eq}(\rho_{wall},\bm{u}_{wall})+f_i^{neq}.
-\end{equation}
-The incoming population after the collision is changed due to collision:
-\begin{equation}
-f_i=f_i^{eq}(\rho_{wall},\bm{u}_{wall})+f_i^{neq} (1-\omega).
-\end{equation}
-The method is called extrapolation as the wall macroscopic parameters are
-extrapolated from fluid using first or second accuracy numerical stencils. If
-the pressure boundary condition is specified then it specifies $\rho_{wall}$ as
-well extrapolating the wall velocity $\bm{u}_{wall}$ from nearest neighbours.
-The same goes for the velocity boundary conditions, where velocity
-$\bm{u}_{wall}$ is specified and $\rho_{wall}$ is extrapolated from the fluid.
-
-It is pretty straightforward to obtain non-equilibrium contribution from the
-fluid, however the equilibrium for the wall side needs special treatment. If
-one wants to take the hydrodynamic equilibrium, the numerical scheme blows
-up. Therefore, one needs to construct the equilibrium function as it's given
-for the binary liquid model. The binary liquid model equilibrium contains local
-density, local phase and phase laplacians and gradients. The pressure boundary
-conditions specify the local phase and the local density. Therefore, we
-extrapolate not only velocity to construct the equilibrium, but phase laplacian
-and gradients. This situation produces stable simulations to be validated in
-the future. 
 \bibliographystyle{plainnat}

### 2010-10-21 Sailfish/Grid/binary_microchannel.py
Started the sailfish script for the Grid refinement. Can pass arguments from the command line directly.

@@ -1,0 +1,69 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import sys
+import optget
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+# TODO: [(736, 51), (1471, 100), (2206, 149), (2941, 198)]
+
+class GeoFE(geo.LBMGeo2D):
+
+    def define_nodes(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(hy <= 1, hy >= self.lat_ny-2)
+        self.set_geo(wall_map, self.NODE_WALL)
+ 
+    def init_fields(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        width = 6
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(
+                       np.logical_and(hx >= 1.0*(self.lat_nx-1)/3.0,
+                                      hx <= 2.0*(self.lat_nx-1)/3.0),
+                       np.logical_and(hy >= width + 1 , hy <= self.lat_ny-width-2))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_seperation_2d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 736,
+                    'lat_ny': 51+2, 'grid': 'D2Q9',
+                    'kappa': 0.04, 'Gamma': 1.0, 'A': 0.04,
+                    'scr_scale': 2,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'periodic_x': True, 'periodic_y': True}
+        settings.update(defaults)
+
+        lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=[], defaults=settings)
+
+        self.add_body_force((6.0e-6, 0.0), grid=0, accel=False)
+
+        # Use the fluid velocity in the relaxation of the order parameter field,
+        # and the molecular velocity in the relaxation of the density field.
+        self.use_force_for_eq(None, 0)
+        self.use_force_for_eq(0, 1)
+
+
+
+def main(argv=None):
+    try:                                
+        opts, args = getopt.getopt(argv, "nx:ny:", ["--nx=", "--ny="]) 
[…]

### 2010-10-21 Sailfish/Grid/grid.py

@@ -1,0 +1,89 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+ 
+def Run_Simulations():
+    print os.getcwd()
+    #os.mkdir("temp")
+    for i in range(1, 5):
+        dir_temp=str(49*i+2)
+        os.mkdir(dir_temp)
+        os.chdir(dir_temp)
+        os.mkdir("Pressure")
+        os.mkdir("Force")
+        subprocess.Popen('cp ../../binary_simple.cpp Force/', shell=True)
+        subprocess.Popen('cp ../../binary_guo_pressure_bb_walls.cpp Pressure/', shell=True)
+        os.chdir("Force")
+        os.mkdir("tmp2")
+        cmd="sed -e 's/const int NY=[0-9]*;/const int NY="+str(49*i+2)+";/' -e 's/const int NX=[0-9]*;/const int NX="+str(15*49*i+1)+";/'"
+        #print cmd
+        sed=subprocess.Popen(cmd+" binary_simple.cpp > binary_simple_out.cpp", shell=True)
+        sed.wait()
+        print os.getcwd()
+        comp=subprocess.Popen("g++ -O3 binary_simple_out.cpp -o main.out")
+        comp.wait()
+        subprocess.Popen("./main.out "+str(6*i)+" "+str(i)+" &")
+        os.chdir("..")
+        os.chdir("Pressure")
+        os.mkdir("tmp")
+        sed=subprocess.Popen(cmd+" binary_guo_pressure_bb_walls.cpp > binary_guo_pressure_bb_walls_out.cpp",  shell=True)
+        sed.wait()
+        comp=subprocess.Popen("g++ -O3 binary_guo_pressure_bb_walls_out.cpp -o main.out")
+        comp.wait()
+        subprocess.Popen("./main.out "+str(6*i)+" "+str(i)+" &")
+        os.chdir("..")
+        os.chdir("..")
+
+def Get_Zero(prof, factor):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    print (zero-0.5)/(49*factor)
+
+def Analyze_Simulations():
+    print os.getcwd()
+    
+    for i in range(1, 5):
[…]

### 2010-10-21 Sailfish/Grid/binary_microchannel.py
Revised binary_microchannel file. Something is wrong.

@@ -5,7 +5,7 @@ from sailfish import lbm
 from sailfish import geo
 
 import sys
-import optget
+import getopt
 import optparse
 from optparse import OptionGroup, OptionParser, OptionValueError
 
@@ -30,7 +30,7 @@ class GeoFE(geo.LBMGeo2D):
         self.sim.rho[:] = 1.0
 
 class FESim(lbm.BinaryFluidFreeEnergy):
-    filename = 'fe_seperation_2d'
+    filename = 'fe_separation_2d'
 
     def __init__(self, geo_class, defaults={}):
         settings = {'verbose': True, 'lat_nx': 736,
@@ -43,26 +43,37 @@ class FESim(lbm.BinaryFluidFreeEnergy):
 
         lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=[], defaults=settings)
 
-        self.add_body_force((6.0e-6, 0.0), grid=0, accel=False)
+        #self.add_body_force((6.0e-6, 0.0), grid=0, accel=False)
+        self.add_body_force((force, 0.0), grid=0, accel=False)
 
         # Use the fluid velocity in the relaxation of the order parameter field,
         # and the molecular velocity in the relaxation of the density field.
         self.use_force_for_eq(None, 0)
         self.use_force_for_eq(0, 1)
+        print "Force=",force
 
 
 
 def main(argv=None):
     try:                                
-        opts, args = getopt.getopt(argv, "nx:ny:", ["--nx=", "--ny="]) 
+        opts, args = getopt.getopt(argv, '', ['force='])
+        print opts
+        print args
     except getopt.GetoptError:           
         usage()                          
         sys.exit(2)                     
-
-
+    
+    global force
+    flag=False
+    for opt, arg in opts:
+        if opt == "--force":
+            force=arg
+            flag=True
+    if not flag:
+        force=6e-6
 
 if __name__ == "__main__":
-    #main(sys.argv[1:])
+    main(sys.argv[1:])
 
     sim = FESim(GeoFE)
     sim.run()

### 2010-10-21 Sailfish/Grid/grid.py

@@ -6,17 +6,17 @@ import numpy
  
 def Run_Simulations():
     print os.getcwd()
+    ny=[102,127,152,177,202,227]
+    nx=[1501,1876,152,177,202,227]
+    force_init=6e-6/4;
     #os.mkdir("temp")
-    for i in range(1, 5):
-        dir_temp=str(49*i+2)
+    for i in range(0, len(ny)):
+        dir_temp=str(ny[i])
         os.mkdir(dir_temp)
+        subprocess.call(['cp','binary_microchannel.py',dir_temp+"/"])
         os.chdir(dir_temp)
-        os.mkdir("Pressure")
-        os.mkdir("Force")
-        subprocess.Popen('cp ../../binary_simple.cpp Force/', shell=True)
-        subprocess.Popen('cp ../../binary_guo_pressure_bb_walls.cpp Pressure/', shell=True)
-        os.chdir("Force")
-        os.mkdir("tmp2")
+        subprocess.call(['./binary_microchannel.py','--lat_nx='+str(nx[i]),'--lat_ny='+str(ny[i])],'--bc_wall=halfbb','--force=')
+        
         cmd="sed -e 's/const int NY=[0-9]*;/const int NY="+str(49*i+2)+";/' -e 's/const int NX=[0-9]*;/const int NX="+str(15*49*i+1)+";/'"
         #print cmd
         sed=subprocess.Popen(cmd+" binary_simple.cpp > binary_simple_out.cpp", shell=True)
@@ -84,6 +84,7 @@ def Analyze_Velocities():
 
 
 if __name__=="__main__":
-    Analyze_Simulations()    
+    #Analyze_Simulations()    
     #Analyze_Velocities()
+    Run_Simulations()
     pylab.show()

### 2010-10-21 Text/paper.tex
Added the initialization results to the text.

@@
 thickness as close as possible to the already obtained numerical film
-thicknesses \cite{giavedoni-numerical}. Otherwise, the region where the
-Bretherton phenomena occurs is dillute, which is obviously seen for the
-simulation initialized with the film thickness as $Ny-28$ {\color{red} Michal,
-can you please prepare the simulations with the different initialization. Or
-just delete different thicknesses initialization}.
+thicknesses \cite{giavedoni-numerical}. Otherwise, the simulation time for the
+bubble to obtain a steady-state form increases.
 \begin{figure}
-\includegraphics*[bb=18 350 594 450,width=0.97\textwidth]
-{Figures/force_width_6.eps}
-\includegraphics*[bb=18 350 594 450,
-width=0.97\textwidth]{Figures/force_width_10.eps}
-\includegraphics*[bb=18 350 594 450,
-width=0.97\textwidth]{Figures/force_width_14.eps}
-\caption{\color{red} The phase plots for different bubble widths as
-$Ny-12$,$Ny-20$ and
-$Ny-28$. Michal, could you please prepare nice pictures for
-simulations. Indicate as well the
-grid and time steps\label{fig:different:initialization:widths}}
+\includegraphics*[bb=-20 360 620 430,width=0.97\textwidth]
+{Figures/Init/initbegin12.eps}
+\includegraphics*[bb=-20 360 620 430,
+width=0.97\textwidth]{Figures/Init/initbegin16.eps}
+\includegraphics*[bb=-20 360 620 430,
+width=0.97\textwidth]{Figures/Init/initbegin20.eps}
+\includegraphics*[bb=-20 360 620
+430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
+\includegraphics*[bb=-20 360 620
+430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
+\includegraphics*[bb=-20 360 620
+430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
+\caption{The phase plots for different bubble widths as
+$H_{eff}-12$,$H_{eff}-16$ and
+$H_{eff}-20$. The grid for the simulation is $102\mathrm{x}1501$. The phase
+profiles were obtained after $200000$ time steps. One can see that even the
+system is initialized with different bubble volumes the system comes to the
+shape, where the film thickness is conserved. For the given parameters the
+interface thickness is $0.0615$, $0.0611$, $0.0605$.
+\label{fig:different:initialization:widths}}
 \end{figure}

### 2010-10-21 Sailfish/Grid/bmc.py

@@ -1,0 +1,59 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo2D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(hy < self.wall_width, hy >= self.lat_ny-self.wall_width)
+        self.set_geo(wall_map, self.NODE_WALL)
+
+    def init_fields(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(
+                       np.logical_and(hx >= (self.lat_nx-1)/3.0,
+                                      hx <= 2*(self.lat_nx-1)/3.0),
+                       np.logical_and(hy >= width + self.wall_width, hy <= self.lat_ny-width-1-self.wall_width))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_seperation_2d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 736,
+                    'lat_ny': 51+10, 'grid': 'D2Q9',
+                    'kappa': 0.04, 'Gamma': 1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'bc_wall': 'halfbb',
+                    'periodic_x': True, 'periodic_y': True}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=0.0))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=5))
+
+        lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=opts, defaults=settings)
+
+        self.add_body_force((self.options.force, 0.0), grid=0, accel=False)
+
+        # Use the fluid velocity in the relaxation of the order parameter field,
+        # and the molecular velocity in the relaxation of the density field.
[…]

### 2010-10-21 Sailfish/Init/binary_microchannel.py

@@ -20,21 +20,21 @@ class GeoFE(geo.LBMGeo2D):
  
     def init_fields(self):
         hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
-        width = 6
+        width = self.sim.options.init_width
 
         self.sim.phi[:] = 1.0
         self.sim.phi[np.logical_and(
                        np.logical_and(hx >= 1.0*(self.lat_nx-1)/3.0,
                                       hx <= 2.0*(self.lat_nx-1)/3.0),
-                       np.logical_and(hy >= width + 1 , hy <= self.lat_ny-width-2))] = -1.0
+                       np.logical_and(hy >= width + 2 , hy <= self.lat_ny-width-3))] = -1.0
         self.sim.rho[:] = 1.0
 
 class FESim(lbm.BinaryFluidFreeEnergy):
     filename = 'fe_separation_2d'
 
     def __init__(self, geo_class, defaults={}):
-        settings = {'verbose': True, 'lat_nx': 736,
-                    'lat_ny': 51+2, 'grid': 'D2Q9',
+        settings = {'verbose': True, 'lat_nx': 1501,
+                    'lat_ny': 102+2, 'grid': 'D2Q9',
                     'kappa': 0.04, 'Gamma': 1.0, 'A': 0.04,
                     'scr_scale': 2,
                     'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
@@ -43,43 +43,16 @@ class FESim(lbm.BinaryFluidFreeEnergy):
 
         lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=[], defaults=settings)
 
-        #self.add_body_force((6.0e-6, 0.0), grid=0, accel=False)
-        self.add_body_force((force, 0.0), grid=0, accel=False)
+        self.add_body_force((6.0e-6/4.0, 0.0), grid=0, accel=False)
+        #self.add_body_force((force, 0.0), grid=0, accel=False)
 
         # Use the fluid velocity in the relaxation of the order parameter field,
         # and the molecular velocity in the relaxation of the density field.
         self.use_force_for_eq(None, 0)
         self.use_force_for_eq(0, 1)
-        print "Force=",force
 
-
-
-def pre_main(argv=None):
-    try:                                
-        opts, args = getopt.getopt(argv, '', ['force='])
-        print opts
-        print args
-    except getopt.GetoptError:           
-        usage()                          
[…]

### 2010-10-21 Sailfish/Init/init.py

@@ -1,0 +1,76 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+ 
+def Run_Simulations():
+    print os.getcwd()
+    width=[12,14,16,18,20]
+    force_init=6e-6/4;
+    #os.mkdir("temp")
+    for i in range(0, len(width)):
+        dir_temp=str(width[i])
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','binary_microchannel.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        subprocess.call(['./binary_microchannel.py','--bc_wall=halfbb','--init_width='+dir_temp,'--every=20000','--max_iters=200000','--output=init'])
+        os.chdir("..")
+  
+def Get_Zero(prof):
+    zero=0
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    print (zero-1.5)/(len(prof)-4)
+
+def Analyze_Simulations():
+    print os.getcwd()
+    width=[12,14,16,18,20]
+    for i in range(0, len(width)):
+        dir_temp="Results/"+str(width[i])
+        os.chdir(dir_temp)
+        name="init180000.npz"
+        array=numpy.load(name)
+        fig=pylab.figure(figsize=(11,1.5))
+        pylab.imshow(array['phi'])
+        pylab.yticks([0,25,50,75,100])
+        pylab.savefig("../../initfinish"+str(width[i])+".eps",format="EPS",dpi=70)
+        
+        fig_init=pylab.figure(figsize=(11,1.5))
+        array_init=numpy.load("init000000.npz")
+        pylab.imshow(array_init['phi'])
+        pylab.yticks([0,25,50,75,100])
+        pylab.savefig("../../initbegin"+str(width[i])+".eps",format="EPS",dpi=70)
+        
+        #pylab.plot(array[:, 520*i])
+        #pylab.savefig("grid_phase_prof_"+str(49*i)+".eps", dpi=300)
+        Get_Zero(array['phi'][:, 1300])
+        #extrapolator=UnivariateSpline(array[0:(49*i+2)/2, 600*i], numpy.arange(0, (49*i+2)/2),  k=2)
+        #print extrapolator(0)
+        os.chdir("../..")
+        
+
[…]

### 2010-10-21 Sailfish/Init/refinement.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+for a in `seq 1 4`
+do
+  echo $a
+  b=49*$a
+  echo $b
+  #mkdir Omega${a*49+2}
+  #cp main.out Omega$a/
+  #cp athena_palettes/LeeLoo.txt Omega$a/
+  
+  #cd Omega$a
+ 
+  #for b in `seq 20 2 30`
+  #do
+  #  mkdir R$b
+  #  mkdir R$b/tmp
+  #  mkdir R$b/athena_palettes
+
+  #  cp LeeLoo.txt R$b/athena_palettes/
+  #  cp main.out R$b/
+  #  cd R$b
+  #  ./main.out $a $b
+  #  rm main.out
+  #  rm -rf athena_palettes
+  #  cd ..
+  #done
+    
+  #rm main.out
+  #rm LeeLoo.txt
+  #cd ..
+done
+exit 0

### 2010-10-21 Sailfish/Init/test.py

@@ -1,0 +1,24 @@
+#!/usr/bin/python
+import getopt
+import sys
+
+def main(argv=None):
+    try:                                
+        opts, args = getopt.getopt(argv, '', ['force='])
+    except getopt.GetoptError:           
+        usage()                          
+        sys.exit(2)                     
+    
+    global force
+    flag=False
+    for opt, arg in opts:
+        if opt == "--force":
+            force=arg
+            flag=True
+    if not flag:
+        force=6e-6
+    
+    print force
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2010-10-22 Text/paper.tex
Added results for the wall gradient. Need to add some numbers.

@@
 
-Here are two different examples presented that can quantitatively show the
+Here two different examples are presented that can quantitatively show the
 initialization procedure for quite different capillary numbers. In these
@@
 \subsection{The influence of the wall gradient}
-Once the procedure of how to resolve the interface gradient is known, we want
-to examine the influence of the phase gradient, i.e. the wettability of the
-wall, to the film thickness. We examined the
-wall gradient influence on
-the standard grid $202\mathrm{x}3001$ and the initial film width as
-{\color{red} Take the value from previous calculations $20$}
-lattice boltzmann units. We examined $11$ different values for the wall
+The results should not depend on the wall wettability. Wettability is defined
+through the phase gradient \cite{pooley-contact}, $\partial_n \phi$. This
+section is to examine the wall gradient influence on simulations. We take large
+enough resolution for simulations to be consistent. Simulation domain was
+chosen as $177\mathrm{x}2626$ and the initial film width as
+$12$ lattice boltzmann units, which correspond to the predicted film thickness.
+We examined $11$ different values for the wall
 gradient ranging from $-1$ to $1$. The results are consistent among all the
@@
 0.0702489349415, 0.0697553433989$}. For the wall gradients $0.8$ and $1.0$ the
-values are NAN. The negative values of the phase gradient are preferable, as
+simulations are unstable. One can see the simulated profiles in Fig.
+\ref{fig:gradients:profiles} The negative values of the phase gradient are
+preferable, as
 far as the phase of the liquid adjacent to the wall has the value of $1$. That
@@
 with it}.
-
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
+\caption{Phase profiles for different wall gradients. One can see if the width
+is properly resolved then the wettability of the wall doesn't play any role
+affecting the film thickness.
+\label{fig:gradients:profiles}
+}
+\end{figure}
 \section{The influence of different viscosities}

### 2010-10-22 Sailfish/Wall_Gradient/binary_microchannel.py

@@ -1,0 +1,60 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import sys
+import getopt
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+# TODO: [(736, 51), (1471, 100), (2206, 149), (2941, 198)]
+
+class GeoFE(geo.LBMGeo2D):
+
+    def define_nodes(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(hy <= 1, hy >= self.lat_ny-2)
+        self.set_geo(wall_map, self.NODE_WALL)
+ 
+    def init_fields(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        #width = int(7*1.75)
+        width=12
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(
+                       np.logical_and(hx >= 1.0*(self.lat_nx-1)/3.0,
+                                      hx <= 2.0*(self.lat_nx-1)/3.0),
+                       np.logical_and(hy >= width + 2 , hy <= self.lat_ny-width-3))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_separation_2d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 2626,
+                    'lat_ny': 177+2, 'grid': 'D2Q9',
+                    'kappa': 0.04, 'Gamma': 1.0, 'A': 0.04,
+                    'scr_scale': 2,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'periodic_x': True, 'periodic_y': True}
+        settings.update(defaults)
+
+        lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=[], defaults=settings)
+
+        self.add_body_force((6.0e-6/(4.0*1.75*1.75), 0.0), grid=0, accel=False)
+        #self.add_body_force((force, 0.0), grid=0, accel=False)
+
+        # Use the fluid velocity in the relaxation of the order parameter field,
+        # and the molecular velocity in the relaxation of the density field.
+        self.use_force_for_eq(None, 0)
[…]

### 2010-10-22 Sailfish/Wall_Gradient/gradient.py

@@ -1,0 +1,85 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+ 
+def Run_Simulations():
+    print os.getcwd()
+    gradient=numpy.arange(-1.0,1.2,0.2)
+    str_gradient=['-10','-8','-6','-4','-2','0','2','4','6','8','10']
+    for i in range(0, len(gradient)):
+        dir_temp="Grad"+str_gradient[i]
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','binary_microchannel.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        subprocess.call(['./binary_microchannel.py','--bc_wall=halfbb','--phase_grad='+str(gradient[i]),'--every=20000','--max_iters=200000','--output=grad','--batch'])
+        os.chdir("..")
+  
+def Get_Zero(prof):
+    zero=0
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    print (zero-1.5)/(len(prof)-4)
+
+def Analyze_Simulations():
+    print os.getcwd()
+    gradient=[-1.0,-0.8,-0.6,-0.4,-0.2,0.0,0.2,0.4,0.6,0.8,1.0]
+    str_gradient=['-10','-8','-6','-4','-2','0','2','4','6','8','10']
+    style=["bo","r+","g--","k-.","c<-.","y>","bs","g^","r^-"]
+    fig=pylab.figure()
+    for i in range(0, len(str_gradient)-2):
+        dir_temp="Results/Grad"+str_gradient[i]
+        os.chdir(dir_temp)
+        name="grad180000.npz"
+        array=numpy.load(name)
+        #fig=pylab.figure(figsize=(11,1.5))
+        #pylab.imshow(array['phi'])
+        pylab.plot(array['phi'][0:22,2000],style[i],linewidth=3,markersize=10)
+        #pylab.yticks([0,25,50,75,100])
+        #pylab.savefig("../../initfinish"+str(width[i])+".eps",format="EPS",dpi=70)
+        
+        #fig_init=pylab.figure(figsize=(11,1.5))
+        #array_init=numpy.load("init000000.npz")
+        #pylab.imshow(array_init['phi'])
[…]

### 2010-10-22 Text/paper.tex
Finished the wall gradient stuff. Moving to the grid refining.

@@
 which is a small time step. The typical simulation needs approximately
-{\color{red} $100,000$} time steps to come to the steady state solution.
+$100000-300000$ time steps to come to the steady state solution giving the
+physical time simulation around $10^{-5}--10^{-4}\,\mathrm{s}$, which is
+probably not enough to find desired correlations.
 
@@
 given wall gradients the values of interface thicknesses are
-{\color{red} Change these numbers $0.0709130255276, 0.0708903443305,
-0.0708661222722,
-0.0708369188112, 0.0707962325316, 0.0707266326981, 0.0705839443249,
-0.0702489349415, 0.0697553433989$}. For the wall gradients $0.8$ and $1.0$ the
+$0.0633$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0633$,$0.0632$,
+$0.0631$. The results are different in 4th digit. For the wall gradients $0.8$
+and $1.0$ the
 simulations are unstable. One can see the simulated profiles in Fig.
@@
 far as the phase of the liquid adjacent to the wall has the value of $1$. That
-means that the phase values near the wall are above $1$. In case of positive
+means that the phase values near the wall are above $1$ and do not interact
+with the bubble which has phase value $-1$.  In case of positive
 gradients the values near the wall are below $1$ and the gradient profile can
 fuse with the values of the gas phase as $-1$, creating the wall stickyness,
-unphysical one. {\color{red} Michal, if you want to add some pictures go ahead
-with it}.
+unphysical one.
 \begin{figure}

### 2010-10-22 Sailfish/Wall_Gradient/gradient.py

@@ -27,7 +27,7 @@ def Analyze_Simulations():
     print os.getcwd()
     gradient=[-1.0,-0.8,-0.6,-0.4,-0.2,0.0,0.2,0.4,0.6,0.8,1.0]
     str_gradient=['-10','-8','-6','-4','-2','0','2','4','6','8','10']
-    style=["bo","r+","g--","k-.","c<-.","y>","bs","g^","r^-"]
+    style=["bo","rH","g--","k-.","c<-.","y>","bs","g^","rD"]
     fig=pylab.figure()
     for i in range(0, len(str_gradient)-2):
         dir_temp="Results/Grad"+str_gradient[i]
@@ -48,9 +48,10 @@ def Analyze_Simulations():
         
         #pylab.plot(array[:, 520*i])
         #pylab.savefig("grid_phase_prof_"+str(49*i)+".eps", dpi=300)
-        #Get_Zero(array['phi'][:, 1300])
+        Get_Zero(array['phi'][:, 2000])
         #extrapolator=UnivariateSpline(array[0:(49*i+2)/2, 600*i], numpy.arange(0, (49*i+2)/2),  k=2)
         #print extrapolator(0)
+        
         os.chdir("../..")
     fig.subplots_adjust(left=0.15)
     pylab.xlabel(r'''$x$''',fontsize=30)

### 2010-10-22 Text/paper.tex
Added grid results to the text.

@@
 is related to the effective channel height. Note that it's a big
-simplifications to impose the boundary in the middle beetween the bounce-back
-node and the fluid node. In the case of $\delta=0.06$ the initial grid number
+simplification to impose the boundary in the middle beetween the bounce-back
+node and the fluid node which give rise to the viscosity dependent location of
+the wall \cite{ginzburg-multireflection}. In the case of $\delta=0.06$ the
+initial grid number
 is $N_y=102$, which gives the horizontal grid number as $N_x-1=15(N_y-2)=1500$.
-The bubble is initialized as rectangular box with coordinates
+The bubble is initialized as a rectangular box with coordinates
 $y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
@@
 $H_{eff}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
-x } =Const$.
-{\color{red} Michal, please add some results here and description to it as well.
-The results for the force driven flows velocity phase profiles are shown in
-Fig. \ref{fig:grid:profiles}. The corresponding ratios to the effective channel
-width (without bounce-back nodes) is as $0.0809472165313,
-0.0708111118485, 0.0678483852412, 0.0664499716904$. One can see
-that for the channel widths as $149,198$, the results are convergent. The
-velocities in the center of the bubble are $0.004176469553,
-0.00401004133, 0.003969909065, 0.00394134433$, which corresponds
-to the $Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}=\frac{\frac{2}{3}\,
-0.0045}{0.03771}=0.069$, which is slightly higher when for the assumed
-profile. However, if we define Capillary number through the velocity of the
-slug, then the Capillary number will have another value. The velocities of the
-slug are $0.003411444831,0.003412693869,0.003420338754,0.003425757312$ and
-Capillary number defined through these velocities is $0.0602$. 
-}
-Let us calculate to which extent we need to refine the interface to obtain
-convergent results. The interface itself occupies approximately $5 \xi$, where
-$\xi=\sqrt{k/A}=1$. This interface width divided to the effective channel width
-($5\xi/H_{eff}$) is as follows {\color{red} Please add new values here
-$0.102,0.051,0.034,0.025$ } for the given
-$H_{eff}$. Let us calculate the interface width ratio to the film thickness as
-$5\xi/H_{film}$ is as follows
-{\color{red} Please add here new values $1.2605,  0.7205,  0.501,  0.3839$}.
-Therefore, one needs to
-resolve the interface as {\color{red} Please give number here $40\%$} of the
-expected film width for simulations to be converged.
-{\color{red} If we need this pictures???
+x } = Const$. We performed a number of simulations for the prescribed grids.
+The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
+see the visual results convergence for $H_{eff}\geq 175$. The calculated widths
+are $0.0606$,$0.0618$,$0.0617$,$0.0629$,$0.0624$,$0.0624$. We accept the
+convergent results for grids larger than $H_{eff}=175$. However, one can see
+that imposing proper initialization technique, large enough time and the wall
+wettability reveal results which are different only in 3rd digit even for
+underresolved film widths. To calculate how good the interface thickness is
+resolved the interface thickness to the film width thickness is calculated.The
+interface itself occupies approximately $5 \xi$, where
+$\xi=\sqrt{k/A}=1$. Let us calculate the interface width ratio to the film
+thickness as
+$5\xi/H_{film}$ is as follows $0.824$, $0.646$,  $0.539$, $0.453$,
+$0.400$, $0.355$. Therefore, one needs to
+resolve the interface as $40-50$ percent of the
+expected film width for simulations to be converged. We further examine the
+velocities in the center of the bubble to calculate the Capillary number. The
+calculated numbers are $0.0041$, $0.0041$,$0.0040$,$0.0040$,$0.0041$,$0.0039$.
+This corresponds to the average capillary number as $0.07$, which is slightly
+larger than the expected capillary number. The corresponding error can be
+attributed to the density and viscosity ratio which is much larger in the
+Bretherton problem.
 \begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_49.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_98.eps}\\
-\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_147.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_196.eps}\\
-\caption{Do we need these pictures? Grid refined profiles for the effictive
+\includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
+\caption{Grid refined profiles for the effictive
 channel widths
-$H_{eff}=49,98,147,196$.\label{fig:grid:profiles}}
+$H_{eff}=100,125,150,175,200,225$.One can see
+convergent results even for the underresolved film
+thickness.\label{fig:grid:profiles}}
 \end{figure}
-}
 

### 2010-10-22 Sailfish/Grid/bmc.py

@@ -31,8 +31,8 @@ class FESim(lbm.BinaryFluidFreeEnergy):
     filename = 'fe_seperation_2d'
 
     def __init__(self, geo_class, defaults={}):
-        settings = {'verbose': True, 'lat_nx': 736,
-                    'lat_ny': 51+10, 'grid': 'D2Q9',
+        settings = {'verbose': True, 'lat_nx': 1501,
+                    'lat_ny': 102+2, 'grid': 'D2Q9',
                     'kappa': 0.04, 'Gamma': 1.0, 'A': 0.04,
                     'scr_scale': 1,
                     'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,

### 2010-10-22 Sailfish/Grid/grid.py

@@ -7,84 +7,97 @@ import numpy
 def Run_Simulations():
     print os.getcwd()
     ny=[102,127,152,177,202,227]
-    nx=[1501,1876,152,177,202,227]
+    nx=[1501,1876,2251,2626,3001,3376]
     force_init=6e-6/4;
+    init_width=6
     #os.mkdir("temp")
     for i in range(0, len(ny)):
         dir_temp=str(ny[i])
         os.mkdir(dir_temp)
-        subprocess.call(['cp','binary_microchannel.py',dir_temp+"/"])
+        subprocess.call(['cp','bmc.py',dir_temp+"/"])
         os.chdir(dir_temp)
-        subprocess.call(['./binary_microchannel.py','--lat_nx='+str(nx[i]),'--lat_ny='+str(ny[i])],'--bc_wall=halfbb','--force=')
+        ratio=float(ny[i]-2)/100.0
+        width=int(init_width*ratio)
+        subprocess.call(['./bmc.py','--lat_nx='+str(nx[i]),'--lat_ny='+str(ny[i]),'--bc_wall=halfbb',\
+            '--force='+str(force_init/(ratio*ratio)),'--batch','--every=50000','--max_iters='+str(int(200000*ratio)+1),'--output=grid','--iwidth='+str(width)])
         
-        cmd="sed -e 's/const int NY=[0-9]*;/const int NY="+str(49*i+2)+";/' -e 's/const int NX=[0-9]*;/const int NX="+str(15*49*i+1)+";/'"
-        #print cmd
-        sed=subprocess.Popen(cmd+" binary_simple.cpp > binary_simple_out.cpp", shell=True)
-        sed.wait()
-        print os.getcwd()
-        comp=subprocess.Popen("g++ -O3 binary_simple_out.cpp -o main.out")
-        comp.wait()
-        subprocess.Popen("./main.out "+str(6*i)+" "+str(i)+" &")
-        os.chdir("..")
-        os.chdir("Pressure")
-        os.mkdir("tmp")
-        sed=subprocess.Popen(cmd+" binary_guo_pressure_bb_walls.cpp > binary_guo_pressure_bb_walls_out.cpp",  shell=True)
-        sed.wait()
-        comp=subprocess.Popen("g++ -O3 binary_guo_pressure_bb_walls_out.cpp -o main.out")
-        comp.wait()
-        subprocess.Popen("./main.out "+str(6*i)+" "+str(i)+" &")
[…]

### 2010-10-22 Text/paper.tex
Added capillary number script. Need to put some results in.

@@
 \citet{heil-bretherton}, Fig. \ref{fig:heil:planar}, for the range of capillary
-numbers. Because of limiting computational factors ({\color{red} What is the
-least capillary number which is OK?}) we cannot go to the range of small
+numbers. Because of limiting computational factors as memory and the
+performance we cannot go to the range of small
 capillary numbers. However, all the other methods were validated across the
@@
 $0.005$ to $1.0$. Our goal is to compare the film thickness over the capillary
-range from $0.03$ to $1.0$.
-
-To perform numerical simulations we refer to
-grid refinement section where we obtained that the interface width should be
-at most {\color{red} $40$} percent of the interface width. For the benchmark
+range from $0.03$ to $1.0$.  For the
+benchmark
 results we
@@
 follows:
-$0.04,0.06,0.08,0.1,0.12,0.13,0.15$. $5$ lattice boltzmann units occupy
-{\color{red} Could be another number $40\%$} of the interface thickness.
-Therefore the grid size should be as
-following {\color{red} Michal, different numbers would be here.
-$314\mathrm{x}4681$,$210\mathrm{x}3121$,
-$158\mathrm{x}2341$, $127\mathrm{x}1876$, $106\mathrm{x}1561$,
-$98\mathrm{x}1441$,$85\mathrm{x}1246$ Michal, for the larger capillary
-numbers such small grid is not
-appropriate because of the instabilities.
-}.
-
-{\color{red} Probably we don't need the explanations about initialization here,
-just to present results}
-To utilize previous results to properly initialize simulations to run one need
-to do some modifications. The result is that for the grid with $H_{eff}=200$ we
-obtained the film width as {\color{red} Will be another number $0.0685$} and
-velocity in the center of the bubble as
-{\color{red} Will be another number $0.0045$}. The force gradient was
-{\color{red} Probably the same number $6\, 10^{-6} /16$}. The predicted
-Capillary
-number was $0.05$ and the actual one based on the center velocity calculation is
-{\color{red} Will be another number $0.0795$}. However, we performed simulations
-according the assumed value of $0.05$. Let us perform some correlations to
-calculate how the system needs to be initialized. 
+\begin{equation}
+0.04,0.06,0.08,0.1,0.12,0.13,0.15\,.
+\end{equation}
+To perform numerical simulations we refer
+to
+grid refinement section where we obtained that the interface width should be
+at most $50$ percent of the interface width. Therefore the grid was chosen as
+$202 \mathrm{x} 3001$. $5$ lattice boltzmann units do not occupy more than $60$
+percent of the effective width. Because the simulation get unstable with
+smaller grids and larger gradients, all the capillary number simulations were
+performed on the grid chosen. To properly initialize the force gradient the
+proportionality law was utilized. For instance, with the forcing
+$6\,10^{-6}/16$ was chosen for the grid to obtain predicted capillary number as
+$0.005$. Therefore, one can obtain the initial forcing for other grids using
+simple correlations: 
 \begin{equation}
 \begin{aligned}
-&Ca_{lit} \propto U_{slug}\\
-&U_{slug} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\ 
+&Ca_{lit} \propto U_{bubble}\\
+&U_{bubble} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\ 
 &Ca_{lit} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
-&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{Ny^2},
+&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{N_y^2},
 \end{aligned}
 \end{equation}
-where the index $lit$ stands for the presented in the literature in
-\cite{giavedoni-numerical,heil-bretherton}. Therefore the initialized consistent
-pressure gradients can be taken from
-previous consistent simulations:
+where the index $lit$ stands for the predicted capillary number
+\cite{giavedoni-numerical,heil-bretherton}. The grid number $Ny$ is not
+involved as because the simulations are conducted on the same grid.
+Therefore,the pressure gradient can be obtained through the capillary numbers
+ratio:
 \begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d} x}=6\,10^{-6}/16 \frac{Ca_{literature}}{0.05}
-\frac{200^2}{N_y^2}=0.3 \frac{Ca_{literature}}{N_y^2}
+\frac{\mathrm{d}P}{\mathrm{d} x}=6\,10^{-6}/16 \frac{Ca_{lit}}{0.05}
 \end{equation}
-Note that $N_y$ is calculated through the expected resolution of the interface
-in comparison with the interface thickness which itself is based on
-Capillary number simulations. In terms of time the steady state successful
-simulation for the grid $202\mathrm{x}3001$ was obtained after $200,000$ steps.
+The film width is initialized through the capillary numbers ratio as well:
+\begin{equation*}
+w=12 \frac{Ca_{lit}}{0.05}
+\end{equation*}
+The results were obtained after $200,000$ steps.
 The physical

### 2010-10-22 Sailfish/Capillary/capillary.py

@@ -1,0 +1,104 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+ 
+def Run_Simulations():
+    print os.getcwd()
+    capillary=[0.03,0.05,0.08,1.0,0.2,0.4,0.6,0.8,1.0]
+    capillary_str=["3","5","8","10","20","40","60","80","100"]
+    width=[0.04,0.06,0.08,0.1,0.12,0.13,0.15]
+    force_init=6e-6/16;
+    init_width=12
+    #os.mkdir("temp")
+    for i in range(0, len(capillary)):
+        dir_temp=capillary_str[i]
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','bmc.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        ratio=capillary[i]/0.05
+        force=force_init*ratio
+        width_value=int(width[i]*200)
+        subprocess.call(['./bmc.py','--bc_wall=halfbb','--force='+str(force),'--batch','--every=50000','--max_iters=200000','--output=capillary','--iwidth='+str(width_value)])
+        
+        os.chdir("..")
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    print (zero-0.5)/(len(prof)-2)
+
+def Analyze_Simulations():
+    print os.getcwd()
+    ny=[102,127,152,177,202,227]
+    nx=[1501,1876,2251,2626,3001,3376]
+    style=["bo","rH","c<","y>","bs","g^"]
+    color=["b","r","c","y","b","g"]
+    style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    fig=pylab.figure()
+
+
+    for i in range(0, len(ny)):
+        dir_temp="Results/"+str(ny[i])
+        ratio=float(ny[i]-2)/100.0
+        os.chdir(dir_temp)
+        #os.chdir("Force")
+        #pylab.figure()
+        name="grid"+str(200000+i*50000)+".npz"
+        array=numpy.load(name)
+        prof=array['phi'][:,int(1400*ratio)]
[…]

### 2010-10-22 Sailfish/Grid_Viscous/bmc.py

@@ -1,0 +1,59 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo2D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(hy < self.wall_width, hy >= self.lat_ny-self.wall_width)
+        self.set_geo(wall_map, self.NODE_WALL)
+
+    def init_fields(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(
+                       np.logical_and(hx >= (self.lat_nx-1)/3.0,
+                                      hx <= 2*(self.lat_nx-1)/3.0),
+                       np.logical_and(hy >= width + self.wall_width, hy <= self.lat_ny-width-1-self.wall_width))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_seperation_2d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 1501,
+                    'lat_ny': 102+2, 'grid': 'D2Q9',
+                    'kappa': 0.04, 'Gamma': 1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 4.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'bc_wall': 'halfbb',
+                    'periodic_x': True, 'periodic_y': True}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=0.0))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=5))
+
+        lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=opts, defaults=settings)
+
+        self.add_body_force((self.options.force, 0.0), grid=0, accel=False)
+
+        # Use the fluid velocity in the relaxation of the order parameter field,
[…]

### 2010-10-22 Sailfish/Grid_Viscous/grid.py

@@ -1,0 +1,103 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+ 
+def Run_Simulations():
+    print os.getcwd()
+    ny=[102,127,152,177,202,227]
+    nx=[1501,1876,2251,2626,3001,3376]
+    force_init=6e-6/4;
+    init_width=6
+    #os.mkdir("temp")
+    for i in range(0, len(ny)):
+        dir_temp=str(ny[i])
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','bmc.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        ratio=float(ny[i]-2)/100.0
+        width=int(init_width*ratio)
+        subprocess.call(['./bmc.py','--lat_nx='+str(nx[i]),'--lat_ny='+str(ny[i]),'--bc_wall=halfbb',\
+            '--force='+str(force_init/(ratio*ratio)),'--batch','--every=50000','--max_iters='+str(int(200000*ratio)+1),'--output=grid','--iwidth='+str(width)])
+        
+        os.chdir("..")
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    print (zero-1.5)/(len(prof)-4)
+
+def Analyze_Simulations():
+    print os.getcwd()
+    ny=[102,127,152,177,202,227]
+    nx=[1501,1876,2251,2626,3001,3376]
+    style=["bo","rH","c<","y>","bs","g^"]
+    color=["b","r","c","y","b","g"]
+    style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    fig=pylab.figure()
+
+
+    for i in range(0, len(ny)):
+        dir_temp="Results/"+str(ny[i])
+        ratio=float(ny[i]-2)/100.0
+        os.chdir(dir_temp)
+        #os.chdir("Force")
+        #pylab.figure()
+        name="grid"+str(200000+i*50000)+".npz"
+        array=numpy.load(name)
+        prof=array['phi'][:,int(1400*ratio)]
+        x=numpy.arange(0.0,float(ny[i]))/ratio
[…]

### 2010-10-24 results-michal/run_refinement.py
Add results from the grid refinement run.

@@ -1,0 +1,13 @@
+#!/usr/bin/python
+
+Ca = 0.05
+gamma = 0.03771
+
+for y in range(100, 250, 25):
+    ny = y + 2
+    nx = y * 15
+
+    iwidth = int(0.06 * y)
+    force = Ca * 8.0 * gamma / y**2 
+
+    print "./bmc.py --lat_nx=%d --lat_ny=%d --iwidth=%d --force=%e --output=ca0.05_cw%d_ --every=100000 --max_iters=1000000" % (nx, ny, iwidth, force, y)

### 2010-10-25 Sailfish/Capillary/capillary.py
Added Capillary number data and comparison.

@@ -3,12 +3,13 @@ import os
 import subprocess
 import pylab
 import numpy
+import math
  
 def Run_Simulations():
     print os.getcwd()
-    capillary=[0.03,0.05,0.08,1.0,0.2,0.4,0.6,0.8,1.0]
+    capillary=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0]
     capillary_str=["3","5","8","10","20","40","60","80","100"]
-    width=[0.04,0.06,0.08,0.1,0.12,0.13,0.15]
+    width=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16,0.17]
     force_init=6e-6/16;
     init_width=12
     #os.mkdir("temp")
@@ -20,7 +21,7 @@ def Run_Simulations():
         ratio=capillary[i]/0.05
         force=force_init*ratio
         width_value=int(width[i]*200)
-        subprocess.call(['./bmc.py','--bc_wall=halfbb','--force='+str(force),'--batch','--every=50000','--max_iters=200000','--output=capillary','--iwidth='+str(width_value)])
+        subprocess.call(['./bmc.py','--bc_wall=halfbb','--force='+str(force),'--batch','--every=50000','--max_iters=200001','--output=capillary','--iwidth='+str(width_value)])
         
         os.chdir("..")
 
@@ -31,49 +32,79 @@ def Get_Zero(prof):
     for counter in range(0, len(prof)/2):
         if prof[counter]>=0 and prof[counter+1]<0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
-    print (zero-0.5)/(len(prof)-2)
+    return (zero-0.5)/(len(prof)-2)
 
 def Analyze_Simulations():
+    from numpy import genfromtxt
     print os.getcwd()
-    ny=[102,127,152,177,202,227]
-    nx=[1501,1876,2251,2626,3001,3376]
-    style=["bo","rH","c<","y>","bs","g^"]
-    color=["b","r","c","y","b","g"]
-    style_diff=["b-","r:","c-.","y--","b^","g<"]
+    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0]
+    capillary_str=["3","5","8","10","20","40","60","80","100"]
+    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16,0.17]
+
[…]

### 2010-10-25 Sailfish/Capillary/capillary.py
Added proper Giavedoni numbers.

@@ -90,7 +90,7 @@ def Analyze_Simulations():
     print capillaries
     print velocities
     pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
-    pylab.loglog(giavedoni[:,0],giavedoni[:,1],"bD-",linewidth=3,markersize=10)
+    pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
     pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
     pylab.xlim(0.02,1.1)
     numpy.savetxt("capillary.dat",zip(capillaries,widths))

### 2010-10-25 Text/paper.tex
Added the capillary number text. Next iteration is the viscous capillary number simulations.

@@
 \end{figure}
-\section{The influence of different viscosities}
-The results obtained earlier are taken for the liquiqs viscosity ratio $10$.
-We performed the same simulations for the liquids viscosity ratio as $20$.
-Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. 
-{\color{red} The results showed that blah-blah-blah - I guess will be the same}.
 
@@
 small capilary numbers and extended for the range of capillary numbers from
-$0.005$ to $1.0$. Our goal is to compare the film thickness over the capillary
-range from $0.03$ to $1.0$.  For the
+$0.005$ to $10.0$. Our goal is to compare the film thickness over the capillary
+range from $0.03$ to $0.8$, which is a computationally reasonable task.  For the
 benchmark
@@
 \begin{equation}
-0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0\, .
+0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8\, .
 \end{equation}
@@
 where the index $lit$ stands for the predicted capillary number
-\cite{giavedoni-numerical,heil-bretherton}. The grid number $Ny$ is not
+\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
 involved as because the simulations are conducted on the same grid.
@@
 The results were obtained after $200,000$ steps.
-The physical
-time is as follows:
-\begin{equation}
-\begin{aligned}
-&\Delta t=\frac{U_{slug,LB} \Delta X}{U_{phys}} \propto \frac{1}{N_y-2}\\
-&N_{iter}=200,000 \frac{N_y-2}{200}\approx 1000 N_y\\
-%&N_{output}=51 Ny\\
-\end{aligned}
-\end{equation}
-
-{\color{red} That's why we need better grid for larger Capillary numbers. Steady
-simulations were obtained for the following Capillary numbers
-$0.03,0.05,0.08,0.1$  }with the film thicknesses as
-{\color{red} $0.0337433850478, 0.063192148642, 0.0799078589211,
-0.0891674460655$} which are
-quite close to the predicted one $0.04,0.06,0.08,0.1$. The thing is to identify
-the Capillary numbers associated with those numbers. The velocity in the
-centers are $0.00209258774,0.003826643811,0.006506324324,0.008333889077$. The
-Capillary numbers associated with these velocities are $0.036992074529545796, 
-0.067646144698590829,0.1150166512524845,0.14732373700007995$, which again a bit
-overpredictive but consistent. The fig.
-\ref{fig:capillary:simple} represents the thickness against the Capillary
-number which mimics the curve behavior for the Capillary number dependance as
-indicated in Fig. \ref{fig:giavedoni:planar}.
+Film thicknesses were obtained as $0.042$,$0.066$, $0.085$, $0.077$, $0.123$, $0.153$, $0.164$, $0.173$ with the bubble center velocities as $0.0022$, $0.0041$, $0.0068$, $0.0055$, $0.0185$, $0.0388$, $0.0592$, $0.0782$.
+Based on the obtained center velocities the capillary numbers
+are formed as $0.039$,$0.072$,$0.120$,$0.098$,$0.328$,$0.686$,$1.046$,$1.383$.
+Those capillary numbers are overpredicted but the obtained lengths are overpredicted as well due to the improper body force imposement.
+We extracted data from \cite{giavedoni-numerical,heil-bretherton} and compared it with the simulation. The results are shown in Fig. \ref{fig:capillary:comparison}.
 \begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/capillary_thickness_simple.eps}
-\hfill
-\includegraphics[width=0.47\textwidth]{Figures/capillary_range.eps}
-\caption{{\color{red} This one to eliminate - just present here
-comparison between our simulations in the longer range of capillary
-numbers and Giavedoni}Thickness versus the capillary
-number for different scalings based on
-the average capillary number. \label{fig:capillary:simple}}
+\includegraphics[width=0.97\textwidth]{Figures/Capillary/capillaries_comparison.eps}
+\caption{The comparison between simulation results and results from \cite{giavedoni-numerical} and \cite{heil-bretherton}.
+One can see the reasonable agreement.
+The film thickness against the capillary number range is presented.\label{fig:capillary:comparison}}
 \end{figure}
 
-\subsection{The influence of the force}
-
-{\color{red} Michal if you find it useful - use this section. My suggestion is
-to eliminate it.}
-\begin{description}
- \item 
-On the grid $198\mathrm{x}2941$ and the initial film width as $20$ units
-the
-influence of the force is studied. $16$ different values of force are examined,
-which have the following expression $\frac{0.000006}{16} i$, where $i=1 \dots
-16$. The system is allowed to come to the equilibrium after $200,000$ steps. The
-goal is to study the influence of the force on the velocity and therefore the
-capillary number. Unfortunately in the first case only a few simulations are
-stable, i.e.
-$i=1,2$ is stable. The results for the interface thicknesses are
-$0.0700122222668,0.0985152303286$ with the velocities associated as
-$0.004490013864,0.009597299475$. Those numbers correspond to the bubble
-capillary numbers as $0.079372981271400328, 0.16965788849626481$. The average
-capillary numbers are $0.07381636,0.14095644$.
-
-\item
-On the grid $394\mathrm{x}5881$ and the initial film width as $20+4 i$,where
-$i=1 \dots 5$ units the
-influence of the force is studied. The force is scheduled as
-$\frac{0.000006}{128}+i \frac{0.000006}{640}$.The intention is to cover the
-calculation of the Capillary numbers in the range $0.03,0.035,0.04,0.045,0.05$. 
-$5$ different values of force are examined. The system is allowed to come to the
-equilibrium after $400,000-40,000 i$ steps. The associated thicknesses were as
-it follows
-$0.0488915027016,0.0538098132568,0.0587100833678,0.0615470184196,
-0.0661331899123$. Bubble velocities are
-$0.002592856205,0.003077834398,0.003562975124,0.004034533823,0.004470657603$.
-Bubble Capillary numbers are $0.04583566,  0.05440894,  0.0629851 , 
-0.07132116, 0.07903081$.
-\end{description}
+\section{The influence of different viscosities}
+The results obtained earlier are taken for the liquiqs viscosity ratio $10$.
+We performed the same simulations for the liquids viscosity ratio as $20$.
+Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. 
+{\color{red} The results showed that blah-blah-blah - I guess will be the same}.
 

### 2010-10-25 Sailfish/Capillary_Viscous/capillary.py

@@ -132,7 +132,7 @@ def Analyze_Velocities():
 
 
 if __name__=="__main__":
-    Analyze_Simulations()    
+    #Analyze_Simulations()    
     #Analyze_Velocities()
-    #Run_Simulations()
+    Run_Simulations()
     pylab.show()

### 2010-10-25 Text/paper.tex
Added different viscosities results. Waiting for the grid refinement, but not sure if it's needed to be included.

@@
 We performed the same simulations for the liquids viscosity ratio as $20$.
-Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. 
-{\color{red} The results showed that blah-blah-blah - I guess will be the same}.
+Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. We
+haven't found significant differences with the results obtained earlier. The
+comparison for the liquid thicknesses is presented on Fig.
+\ref{fig:capillary:viscous}. Therefore one can validate the assumption that in
+the low capillary flow regime the density ratio is not involved. As well the
+gas-liquid viscosity ratio of the order $10$ is enough to obtain results
+consistent with the Bretherton/Taylor problem. 
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/Capillary_Viscous/capillaries_viscous.eps}
+\caption{The widths versus capillary number comparison for
+$\frac{\mu_{liq}}{\mu_{gas}}=10$ and $\frac{\mu_{liq}}{\mu_{gas}}=20$. The
+results show good consistency and agreement.\label{fig:capillary:viscous}}
+\end{figure}
 

### 2010-10-25 Sailfish/Capillary_Viscous/capillary.py

@@ -41,8 +41,9 @@ def Analyze_Simulations():
     capillary_str=["3","5","8","10","20","40","60","80","100"]
     width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16,0.17]
 
-    exam=[2000,2200,2600,2500,1100,950,600,350,350]
-    good=[0,1,2,3,4,5,6,7]
+    #exam=[2000,2200,2600,2500,1100,950,600,350,350] Old values
+    exam=[2000,2100,2200,2400,100,1600,200,1600,350]
+    good=[0,1,2,3,4,5,6,7,8]
     #style=["bo","rH","c<","y>","bs","g^"]
     #color=["b","r","c","y","b","g"]
     #style_diff=["b-","r:","c-.","y--","b^","g<"]
@@ -52,7 +53,10 @@ def Analyze_Simulations():
     velocities=[]
     
     #read the giavedoni data (not precise though)
-    giavedoni=genfromtxt("planarcasesolution.csv",delimiter=',',dtype=float)[1:]
+    #giavedoni=genfromtxt("planarcasesolution.csv",delimiter=',',dtype=float)[1:]
+    
+    #read old data
+    capillary_not_viscous=numpy.loadtxt("../Capillary/capillary.dat")
 
 
     for i in range(0, len(capillary_theor)):
@@ -80,20 +84,19 @@ def Analyze_Simulations():
         #pylab.plot(array[:, 520*i])
         #pylab.savefig("grid_phase_prof_"+str(49*i)+".eps", dpi=300)
         #Get_Zero(prof)
-        #extrapolator=UnivariateSpline(array[0:(49*i+2)/2, 600*i], numpy.arange(0, (49*i+2)/2),  k=2)
-        #print extrapolator(0)
 
         os.chdir("../..")
     
     fig=pylab.figure()
-    capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
+    capillaries=numpy.array(velocities)*(4.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
     print "Widths=",widths
     print "Capillaries=",capillaries
     print "Velocities",velocities
     
     pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
[…]

### 2010-10-25 results-michal/plot-chan-width.py
Add a script to plot the phase profile as a function of the channel width.

@@ -1,0 +1,33 @@
+#!/usr/bin/python
+
+import matplotlib
+matplotlib.use('cairo')
+
+import glob
+import numpy as np
+
+from pylab import *
+
+for x in sorted(glob.glob('ca0.05_cw???_1000000.npz')):
+    print x
+    a = np.load(x)
+    width = a['phi'].shape[0] - 2
+    mid = width/2
+
+    x1 = np.min(np.where(a['phi'][mid,:] < 0.0))
+    x2 = np.max(np.where(a['phi'][mid,:] < 0.0))
+
+    if (x2 - x1 > a['phi'].shape[1] / 2):
+        x = x2
+    else:
+        x = (x2 + x1) / 2
+
+    xdat = np.linspace(0, 1.0, width+1)
+    xmax = np.sum(xdat < 0.15)
+
+    plot(xdat[:xmax], a['phi'][:xmax,x], '-', label='%d' % width)
+
+legend()
+savefig('chan-width.pdf')
+
+

### 2010-10-25 Text/paper.tex
Minor style fixes in the abstract.

@@
 \date{\today}
-
+\newcommand{\Ca}{\mathrm{Ca}}
+\newcommand{\todo}[1]{{\color{red}#1}}
 \begin{document}
@@
 challenges to simulation tools. The continuous physics approach is limited on
-those scales, therefore such methods as Molecular Dynamics (MD) or Lattice
-Boltzmann Method (LBM) become appropriate. This work has aim to
-reproduce the gravity driven microchannel flow of free surface liquid with the
-help of the binary liquid lattice Boltzmann method. In the low capillary number
-$Ca$ region the flow is thoroughly described by Bretherton \cite{bretherton},
-where the deposition film thickness is proportional to $Ca^{2/3}$. The binary
-liquid model is continuous interface multiphase model. Therefore, the issue of
-how to properly resolve the interface to obtain proper correlations holds. In
-the work we address the binary liquid parameters, as mobility $M$, surface
-tension parameter $k$, the free energy constant $A$, grid resolution, to
-properly simulate microchannel flow. The work is valuable for industry and
-researchers willing to simulate Bretherton phenomena.
+those scales, and methods such as Molecular Dynamics (MD) or the Lattice
+Boltzmann Method (LBM) become more appropriate. This work aims to
+reproduce the gravity driven microchannel flow of a free surface liquid with the
+help of the free-energy binary liquid lattice Boltzmann method. In the low capillary number
+$\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
+where the deposition film thickness is proportional to $\Ca^{2/3}$. The binary
+liquid model we use is a continuous-interface multiphase model, which presents
+the question of how to properly resolve the interface to obtain proper correlations \todo{between what?}.
+In this work we investigate how the binary liquid model parameters and the grid resolution
+need to be adjusted to properly simulate microchannel flows.  The presented results
+should be of interest for industrial applications and to researchers simulating
+Bretherton phenomena.
 \end{abstract}

### 2010-11-04 Text/paper.tex
Adding results for the bubble length. Need to add qualitative pictures from Kreutzer and Yang.

@@
 
-\section{Capillary number region}
+\subsection{Capillary number region}
 The purpose of this section is to validate the correlations of
@@
 
-\section{The influence of different viscosities}
+\subsection{The film variation over the bubble}
+Though this section
+
+\subsection{The influence of different viscosities}
 The results obtained earlier are taken for the liquiqs viscosity ratio $10$.

### 2010-11-04 Sailfish/Capillary/capillary.py

@@ -109,6 +109,92 @@ def Analyze_Simulations():
     #pylab.xlim(xmax=15)
     pylab.savefig("capillaries_comparison.eps",format="EPS",dpi=300)
 
+def Analyze_Bubble():
+    from numpy import genfromtxt
+    print os.getcwd()
+    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8]
+    capillary_str=["3","5","8","10","20","40","60","80"]
+    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16]
+
+    exam=[2000,2200,2600,2500,1100,950,600,350]
+    good=[0,1,2,3,4,5,6,7]
+    good_value=[1,2,3]
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    low=[1509,1864,1747]
+    high=[2554,2900,2724]
+    widths=[]
+    velocities=[]
+    pylab.figure()
+    for counter,value in enumerate(good_value):
+        dir_temp="Results/"+capillary_str[value]
+        #ratio=float(ny[i]-2)/100.0
+        os.chdir(dir_temp)
+        #os.chdir("Force")
+        #pylab.figure()
+        name="capillary200000.npz"
+        array=numpy.load(name)
+        thicknesses=[]
+        for coor in range(low[counter],high[counter]):
+            prof=array['phi'][:,coor]
+            thicknesses.append(Get_Zero(prof))
+        #prof=array['phi'][:,exam[i]]
+        #x=numpy.arange(0.0,float(ny[i]))/ratio
+        
+        #pylab.imshow(array['phi'])
+    
+        #if i in good:
+            #pylab.plot(prof)
+        #    widths.append(Get_Zero(prof))
+        #    vel=array['v'][0]
+        #    vel_prof=vel[:,exam[i]]
+        #    velocities.append(vel_prof[len(vel_prof)/2])
+        
+        #pylab.plot(x[0:20],prof[0:20],color[i],linewidth=3)
+        #pylab.plot(x,prof,style_diff[i],markersize=10,linewidth=3)
+        #pylab.figure()
+        #pylab.plot(array[:, 520*i])
[…]

### 2010-11-04 Text/paper.tex
Finalized additional results.

@@
 \end{figure}
+Another question which arises is the influence of the wall gradient on the bubble velocity and the
+capillary number accordingly. The shear stress controlled by the wall phase gradient changes the
+effective velocity near the wall, thus changing the bubble velocity as well.
+One can see that if the film thickness is resolved properly the influence of the wall gradient can
+be neglected. Fig. \ref{fig:velocities:wall:gradient} shows the center bubble velocity for different
+gradients. The average relative error around the mean velocity is $0.1\%$. 
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/Wall/velocities_grad_profiles.eps}
+\caption{The center bubble velocity versus the phase gradient. One can neglect the influence of
+the wall once the film is properly resolved.
+\label{fig:velocities:wall:gradient}
+}
+\end{figure}
 
@@
 \subsection{The film variation over the bubble}
-Though this section
+This section is intended to show the qualitative variation of the film thickness over the bubble.
+Though there are simulations for the bubble length simulations for the two-dimensional case, due to
+the presentation of data it is hard to extract numerical data. One can see bubble lengths
+variations for the two-dimensional simulations for the range of capillary numbers in Fig.
+\ref{fig:sehgal:bubble:length}.
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_length.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_sehgal.eps}\\
+\caption{The qualitative comparison for the film thickness across
+the bubble length. Right (courtesy of \citet{sehgal-microchannel}),
+left (present simulations).\label{fig:sehgal:bubble:length}}
+\end{figure}
+One can see the qualitative agreement, i.e. the thickness increases towards the front meniscus and
+rapidly decreases towards the rear meniscus. This shape sometimes is called as a "bullet" shape.
 

### 2010-11-04 Sailfish/Capillary/capillary.py

@@ -112,23 +112,24 @@ def Analyze_Simulations():
 def Analyze_Bubble():
     from numpy import genfromtxt
     print os.getcwd()
-    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8]
-    capillary_str=["3","5","8","10","20","40","60","80"]
+    capillary_theor=numpy.array([0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8])
+    capillary_str=numpy.array(["3","5","8","10","20","40","60","80"])
     width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16]
 
     exam=[2000,2200,2600,2500,1100,950,600,350]
     good=[0,1,2,3,4,5,6,7]
-    good_value=[1,2,3]
-    #style=["bo","rH","c<","y>","bs","g^"]
+    good_value=[1,2,4]
+    style=["b-.","r--","c."]
+    labels=[r'''$Ca='''+ca+r'''$''' for ca in numpy.array(capillary_theor[good_value],dtype=str)]
     #color=["b","r","c","y","b","g"]
     #style_diff=["b-","r:","c-.","y--","b^","g<"]
     
     #fig=pylab.figure()
-    low=[1509,1864,1747]
-    high=[2554,2900,2724]
+    low=[1509,1865,381]
+    high=[2554,2900,1427]
     widths=[]
     velocities=[]
-    pylab.figure()
+    fig=pylab.figure()
     for counter,value in enumerate(good_value):
         dir_temp="Results/"+capillary_str[value]
         #ratio=float(ny[i]-2)/100.0
@@ -163,36 +164,24 @@ def Analyze_Bubble():
         #print extrapolator(0)
         
         #pylab.figure()
-        pylab.plot(thicknesses)
-        pylab.ylim(ymin=0.0,ymax=0.5)
-
-
+        pylab.plot(thicknesses,style[counter],linewidth=3)
 
         os.chdir("../..")
     
-    #fig=pylab.figure()
-    #capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
-    #print "Widths=",widths
-    #print "Capillaries=",capillaries
-    #print "Velocities",velocities
-    
[…]

### 2010-11-04 Sailfish/Wall_Gradient/gradient.py

@@ -63,24 +63,51 @@ def Analyze_Simulations():
 
 def Analyze_Velocities():
     print os.getcwd()
-    
-    for i in range(1, 5):
-        dir_temp="Proper Grid/"+str(49*i+2)
+    gradient=[-1.0,-0.8,-0.6,-0.4,-0.2,0.0,0.2,0.4,0.6,0.8,1.0]
+    str_gradient=['-10','-8','-6','-4','-2','0','2','4','6','8','10']
+    style=["bo","rH","g--","k-.","c<-.","y>","bs","g^","rD"]
+    fig=pylab.figure()
+    velocities=[]
+    for i in range(0, len(str_gradient)-2):
+        dir_temp="Results/Grad"+str_gradient[i]
         os.chdir(dir_temp)
-        #os.chdir("Force")
-        name="velocity"+(1-(i*40000)/100000)*"0"+str(40000*i)+".dat"
-        array=numpy.loadtxt(name)
-        prof=array[:,520*i]
-        pylab.figure()
-        pylab.imshow(array)
-        pylab.figure()
-        pylab.plot(array[:, 520*i])
-        print prof[len(prof)/2]
+        name="grad180000.npz"
+        array=numpy.load(name)
+        ny=array['phi'].shape[0]
+        nx=array['phi'].shape[1]
+        #fig=pylab.figure(figsize=(11,1.5))
+        #pylab.imshow(array['phi'])
+        #pylab.plot(array['phi'][0:22,2000],style[i],linewidth=3,markersize=10)
+        #pylab.yticks([0,25,50,75,100])
+        #pylab.savefig("../../initfinish"+str(width[i])+".eps",format="EPS",dpi=70)
+        velocities.append(array['v'][0][ny/2,2000])
+        #fig_init=pylab.figure(figsize=(11,1.5))
+        #array_init=numpy.load("init000000.npz")
+        #pylab.imshow(array_init['phi'])
+        #pylab.yticks([0,25,50,75,100])
+        #pylab.savefig("../../initbegin"+str(width[i])+".eps",format="EPS",dpi=70)
+        
+        #pylab.plot(array[:, 520*i])
+        #pylab.savefig("grid_phase_prof_"+str(49*i)+".eps", dpi=300)
+        #Get_Zero(array['phi'][:, 2000])
[…]

### 2010-11-08 Text/paper.tex
Multiple style and notation fixes in the whole paper.

@@
 $\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
-where the deposition film thickness is proportional to $\Ca^{2/3}$. The binary
+with the deposition film thickness proportional to $\Ca^{2/3}$. The binary
 liquid model we use is a continuous-interface multiphase model, which presents
@@
 \section{Introduction}
-Taylor/Bretherton \cite{bretherton} flow describes flow of long bubbles in
+The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
 microchannels. It was found that the front meniscus film width is proportional
-to $Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
+to $\Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
 of works which simulate the Bretherton problem with the boundary value method
-\cite{ingham-plates,heil-bretherton}, the simulation of Bretherton problem is of
-difficulty for diffusion interface methods, where the interface is spread over
-the few grid nodes. In this case the question of proper film width resolution
-arises. While the discontinuous interface models are good to simulate
-non-complicated geometries phenomena, the continuous interface methods are more
-flexible to resolve problem evolving coalescence or droplet breakup. This work
-concentrates on the resolving parameters range of the binary liquid lattice
-Boltzmann method to validate it across the Bretherton problem.
+\cite{ingham-plates,heil-bretherton}, its simulation using diffusive interface methods
+is difficult.  If the interfaec is spread over several grid nodes, the question of
+proper film width resolution arises. While the discontinuous interface models
+are good at simulating pheonomena in simple geometries, the continuous interface
+methods are more flexible and capable of resolving problems involving coalescence
+or droplet breakup. This work
+concentrates on resolving the parameter range of the binary liquid lattice
+Boltzmann method to validate it the Bretherton regime.
 
-Lattice Boltzmann method has emerged as successful method to simulate various
-phenomena as hydrodynamics \cite{yu}, thermal-flows
-\cite{karlin-minimalmodels}, micro-flows \cite{ansumali-small-knudsen},
+The lattice Boltzmann method has emerged as a successful method to simulate
+a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
+\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
 ferrofluids \cite{kuzmin-aniso}, and multiphase flows
-\cite{swift,Shan-chen:extended}. Based on kinetic nature and as a particle
-method LBM tackles relatively easy complex geometries and the incorporation of
-physics phenomena on the microscopic level, as multiphase models. Most
+\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
+method easily tackles complex geometries and the incorporation of
+physical phenomena on the microscopic level, as is the case in multiphase models. Most
 multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
-the interface in the continuous way. Such representation as it was pointed out
+the interface in a continuous way. Such a representation
 brings issues with the film thickness resolution versus the interface
 resolution. In other words, when it comes to the resolution of the film
-thickness the diffusion interface should be resolved in such way as to have
-negligible effect on physics in the film.
+thickness the diffusion \todo{diffusive?} interface should be resolved in such a way as to have
+negligible effect on the physics of the film.
 
-This work examines the binary liquids free-energy LBM model \cite{swift}. The
-model simulates binary liquids within the assumption of the uniform overall
+This work examines the binary liquid free-energy LB model \cite{swift}. The
+model simulates two liquids with the assumption of uniform overall
 density. Therefore, the question of how to simulate the Bretherton problem
-where two phases are completely of different densities (liquid and gas) is of
-interest. However, the motion of a liquid slug and a bubble in a capillary is
-governed by the capillary number. Therefore, the viscosities ratio not the
-densities ratio is the critical parameter.
+where the two phases have two completely different densities (i.e. the two fluids
+are liquid and gas) is of interest. However, in the case of the motion of a
+liquid slug and a bubble in a capillary the governing factor is the capillary number
+and the ratio of the viscosities, not of the densities, is the critical parameter.
 
-The goal of this paper is to give the answer as to how to choose related to the
-binary liquid model parameters, viscosities ratio and grid
-resolution to obtain the Bretherton/Taylor problem film thickness for the
-diffusion interface model. The paper is organized as follows. We briefly
-explain the binary liquid lattice Boltzmann model and summarize literature
-review for the Bretherton problem. Then the parameters involved in the
+The goal of this paper is to give an answer as to how to choose the
+binary liquid model parameters, the viscosities ratio and the grid
+resolution, in order to obtain correct Bretherton/Taylor film thickness.
+The paper is organized as follows.  First, we briefly
+explain the binary liquid lattice Boltzmann model and summarize the literature
+for the Bretherton problem. Then, the parameters involved in the
 simulation are examined and presented in the results section. The paper is
-concluded with main findings.
+concluded with a summary of the main findings.
 
@@
 Bretherton in his significant work \cite{bretherton} studied a long bubble
-moving in a tube filled with liquid. It was found that the deposition film is
-proportional to $Ca^{2/3}$ in the range of small capillary numbers. Later on
-it was figured \cite{wong-films,wong-pressure} that the film deposition
-proportinal to $Ca^{2/3}$ only near the front meniscus and the film width is
-variated through the bubble length for not infinite bubbles. Numerical
+moving in a tube filled with liquid. It was found that the deposition film width
+is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
+it was realized \cite{wong-films,wong-pressure} that the film deposition
+is proportional to $\Ca^{2/3}$ only near the front meniscus and that the film width
+varies throughout the bubble length for bubbles of finite length. Numerical
 simulations \cite{giavedoni-numerical} and experimental studies
-\cite{kreutzer-pressure-drop} show as well the deviation of rule $Ca^{2/3}$ for
-capillary numbers more than $0.003$ and Reynolds number dependance. To be able
-to predict consistently the flow pattern for capillaries in different range of
-the parameters researchers validate their codes with the small capillary
-numbers Bretherton problem.
+\cite{kreutzer-pressure-drop} showed as well a Reynolds number dependence, and
+a deviation from the $\Ca^{2/3}$ rule for capillary numbers larger than $0.003$.
+To be able to predict consistently the flow pattern for capillaries in
+different ranges of parameters, simulations are developed and their codes
+are typically validated with the small capillary numbers Bretherton problem.
 
-There are a number of the numerical methods which
-were used for the Taylor/Bretherton flow
-simulation. \citet{vanbaten-circular} studied the mass transfer and film
-thickness for rising bubbles in a circular capillar.
-\citet{kreutzer-pressure-drop} performed numerical simulation by the
-finite volume method the circular capillary for the number of different
+There are a number of numerical methods which
+were used for the simulation of the Taylor/Bretherton flow.
+\citet{vanbaten-circular} studied the mass transfer and film
+thickness for rising bubbles in a circular capillary.
+\citet{kreutzer-pressure-drop} used the finite volume method to perform
+simulations of a circular capillary for a number of different
 Reynolds and capillary numbers. \citeauthor{wong-films} in
-\cite{wong-films,wong-pressure} studied three-dimensional bubbles in the
+\cite{wong-films,wong-pressure} studied three-dimensional bubbles in
 polygonal capillaries and showed the derivations of different shapes in the
 slug crossections and menisci appearance.
-\citet{heil-bretherton,ingham-plates} studied the air finger propagation to
-the two-dimensional channel for the range of Reynolds numbers and capillary
+\citet{heil-bretherton,ingham-plates} studied air finger propagation in
+a two-dimensional channel for a range of Reynolds and capillary
 numbers. \citet{giavedoni-numerical} performed crossvalidation of the
-finite element solution developed by them with the already published results
-across the different capillary numbers. The solution is available for circular
-and planar cases.
+finite element solution developed by them with previously published results.
+The solution is available for the circular and planar cases.
 
-While the applicability of other methods are well established for the
-simulation of the Bretherton/Taylor problem, the lattice Boltzmann
-applicability and thorough parameters study to the best authors' knowledge is
-not yet done. Especially that concerns the diffuse interface multiphase
-models, where one needs to resolve the film accordingly. One should acknowledge
-the works of
-\citet{pagonabarraga-fingers} on meniscus in thin films for the fingering
-phenomena. \citet{sehgal-microchannel} performed the lattice Boltzmann
-simulations of two-dimensional channel flow for larger capillary numbers where
-the authors found the discrepancies with the classical Bretherton theory having
-limitations to the low capillary numbers regime \cite{giavedoni-numerical}.
+While the applicability of other methods is well established for the
+simulation of the Bretherton/Taylor problem, this is not the case for
+the lattice Boltzmann method. A thorough parameter study for this approach
+has not yet been done to the best of the authors' knowledge.
 
-While to resolve interface on a fine level for capillary numbers less than
-$0.003$ is computationally extensive even for a two-dimensional case. For
-instance for capillary number $0.01$ to properly resolve the interface one
-needs around 22 GB memory requirement. Therefore, this work is based on the
-comparison with the already established results which were in their case
-validated over the Bretherton bubble flow. The work presents recipes for the
-initialization and the optimal parameter range for the binary liquid LBM model,
-which is described in the next section.
+One should acknowledge the works of \citet{pagonabarraga-fingers} on menisci
+in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
+simulations of two-dimensional channel flows for larger capillary numbers, and
+the authors found discrepancies with the classical Bretherton theory, which
+is limited to the low capillary number regime \cite{giavedoni-numerical}.
 
-\section{Lattice Botlzmann binary liquid model}
-The lattice Boltzmann equation operates on the rectangular grid representing the
-physical domain. It utilizes the
-probability distribution functions containing the information about the
-macroscopic variables, such as density,
-momentum. LBE consists from two parts: local collision and propagation operated
-from one node to another in the certain
-direction specified by the velocity set.  The lattice Boltzmann implementation
-in a standard view representing the
-collision and propagation steps:
+To resolve the interface on a fine level for capillary numbers less than
+$0.003$ is computationally intensive even in the two-dimensional case
+(the amount of memory necessary to perform the simulation is on the order
+of tens or hundreds of gigabytes).  This work is therefore based on a
+comparison with the already established results, which were themselves
+validated for the Bretherton bubble flow.  In this paper, we present
+recipes for the initialization of the simulations and discuss optimal
+parameter ranges for the binary liquid LB model.
+
+\section{Lattice Boltzmann binary liquid model}
+The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
+physical domain. It utilizes
+probability distribution functions (also known as particle populations)
+containing information about
+macroscopic variables, such as fluid density and momentum. LBE consists of
+two parts: a local collision step, and a propagation step which transports
+information from one node to another in certain
+directions specified by the discretized velocity set.
+The LBE is typically implementated as follows:
 \begin{equation}
@@
 \end{equation}
-where $F_i$ is the external force population. The binary liquid LBM model is
-based on the free-energy functional \cite{swift,landau} and operates with two
-sets of populations one to track pressure and velocities fields, another one is
-for the phase field. The model to be utilized is two-dimensional nine-velocity
-D2Q9 model.
+where $F_i$ is the external force population. The binary fluid LB model is
+based on a free-energy functional \cite{swift,landau}, and operates with two
+sets of populations: one to track the pressure and velocity fields, and the
+other to represent the phase field.
 
-In terms of D2Q9 the equilibrium populations are as \cite{pooley-contact}:
+The model we use in this work is a two-dimensional nine-velocity (D2Q9) model,
+with equilibrium populations \cite{pooley-contact}:
 \begin{equation}
@@
 &f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
-&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
-\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i=1\div8 \\
+&g_i^{eq}=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
 &g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
@@
 \end{equation}
-where $\Gamma$ is the mobility parameter; chemical potential
+where $\Gamma$ is the mobility parameter; the chemical potential
 $\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
-tension; $A$ is the parameter of the free-energy model; bulk pressure
-$p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. D2Q9 related parameters are
+tension; $A$ is a parameter of the free-energy model  The bulk pressure
+is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
+Parameters speficic to the D2Q9 grid are the weights
 $w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
-\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and
+\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
 $Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
-the sound speed parameter $c_s^2=\dfrac{1}{3}$. Weights related to the
-inclusion of the surface tension coefficient are as follows:
+the sound speed parameter $c_s^2=1/3$.  The weights related to the
+inclusion of the surface tension coefficient into the equations are as follows:
 $w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
 $w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
-$w^{xy}_{7-8}=-1/4$. Note that the system restores the macroscopic
-equations as:
+$w^{xy}_{7-8}=-1/4$. The system can be shown to restore the macroscopic
+fluid equations as:
 \begin{equation}
@@
 \end{equation}
-where the viscosity $\nu=c_s^2 (\tau-1/2)$ and mobility parameter
-$M=\Gamma(\tau_{\phi}-1/2)$. The system allows the separation of the liquid
-phase with $\phi=1$ and so-called gas phase with $\phi=-1$. Note that the
-spatial variation of the relaxation parameter depends on the relaxation
-parameter of $\tau_{gas}$ and $\tau_{liq}$ linearly as
-$\tau=\tau_{gas}+\frac{\phi+1}{2}\tau_{liq}$. Note that the surface tension in
-the framework of binary liquid model as $\sqrt{8 k A}{9}$.
+where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
+$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter. The system allows the separation of the liquid
+phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
+relaxation time is taken to depend linearly on the relaxation
+times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
+$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
+The surface tension in the framework of the binary liquid model is $\sqrt{8 k A}{9}$ \todo{is this right?  should the 9 be in the
+denominator?  is not, perhaps put before the root?}.
 
-\section{Lattice Botlzmann benchmark}
-To properly simulate and to compare simulations results with the already
-performed simulations one needs to design a numerical benchmark by addressing
+\section{Lattice Boltzmann benchmark}
+To properly simulate and to compare simulation results with the data
+published in the literature, one needs to design a numerical benchmark addressing
 the following challenges for the lattice Boltzmann method:
 \begin{itemize}
- \item The diffusive interface nature of the multiphase model should not
+ \item The diffusive nature of the interface in the multiphase model should not
 influence the film width results. This is obviously achieved only for certain
-resolutions of the diffuse interface in comparison with the film width.
- \item The spurious currents arising from the surface tension discretization
-can influence the mass transfer in the thin film therefore comprimising the
-overall solution. Therefore the bubble should be long enough to decrease the
-influence of the spurious currents.
- \item The wettability coefficient which is in principle can control the
-dynamic angle and menisci appearance \cite{pagonabarraga-finger} should not
-influence on the overall solution.
- \item The free-energy model examined in the work is one density/different
-viscosities model. Classical Bretherton problem is the free-interface problem
-where the film thickness is established by itself. The viscosities ratio should
-be large as it is indicated in \cite{pagonabarraga-fingers}, however no studies
-address this issue.
-  \item The Bretherton problem addresses the bubbles flow driven by pressure
-difference but not body force. However, it's relatively difficult to impose the
-pressure difference in the framework of the binary liquid model because two
-boundary conditions with the associated gradients and laplacians are required.
-The pressure boundary conditions for the binary liquid model are currently
-under development. Preliminary results show that the differences between
-pressure difference flows and body force driven flows are small for the film
-deposition thickness, the interest of this study. Thus, the work's focus is at
-the body force driven flows because of the better simulation stability and
-realization along with the periodic boundary conditions.
- \item Most simulations are done across the circular shape of channel,
+resolutions of the interface in comparison with the film width.
+ \item Spurious currents arising from surface tension discretization
+can influence mass transfer in the thin film, thereby comoromising the
+overall solution. The bubble should therefore be long enough to decrease the
+influence of spurious currents.
+ \item The wettability coefficient which in principle can control the
+dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
+influence the overall solution.
+ \item The free-energy model examined in the work is single density, two
+viscosities model. The classical Bretherton problem is a free-interface problem
+where the film thickness is established by itself. The viscosity ratio should
+be large, as indicated in \cite{pagonabarraga-fingers}, however no studies
+address this issue. \todo{what's the conclusion of the impact of this issue to our paper then?}
+  \item The Bretherton problem addresses bubble flow driven by a pressure
+difference, not a body force.  It is however relatively difficult to impose a
+pressure difference in the framework of the binary liquid model, as two
+simultaneous boundary conditions with the associated gradients and laplacians are required.
+This kind of pressure boundary conditions for the binary liquid model is currently
+under development, and preliminary results show that the differences between
+pressure difference flows and flows driven by a body force are small as far as the film
+deposition thickness is concerned.  In this work we limit ourselves to the study
+of body force driven flows, because of their simplicity and better numerical stability.
+ \item Most simulations are done for a circular channel,
 which is quite difficult to address in terms of lattice Boltzmann free energy
-framework where one needs to calculate the laplacians near the boundaries.
+framework where one needs to calculate Laplacians near the boundaries.
 Although one can apply certain finite difference stencils
-\cite{arnold-boundary,hunt-boundary} developed for the curved boundaries
-however the analysis of error which can be addressed to the boundary conditions
-treatment becomes an issue. Therefore the benchmark is for an axis-symmetric
-case. 
+\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
+the analysis of error introduced by the boundary conditions becomes an issue.
+Our benchmark is therefore done is for an axis-symmetric case. \todo{since this paper
+is about a 2d flow, which we already said in the intro, what is the point of discussing
+curve boundaries which are only an issue in 3d?}
 \end{itemize}
-Given all the concerns and challenges the suggested Lattice Boltzmann framework
-is the two-dimensional flow driven by a body force. However, the analytical
-solution to this problem, to the best authors' knowledge, is not yet known.
+Given all the concerns and challenges, the suggested lattice Boltzmann framework
+is a two-dimensional flow driven by a body force.  An analytical
+solution to this problem is to the best of our knowledge not yet known.
 Though some authors \cite{sehgal-microchannel} use the gravity driven model to
 validate the Bretherton film thickness, the latter should be used with caution
-since the velocity of the bubble is different from the fluid velocity and a
+since the velocity of the bubble is different from the fluid velocity and the
 bubble shape
-is different in front and rear menisci. \citet{wong-films} indicated that the
+is different in the front and rear menisci. \citet{wong-films} indicated that the
 film width for polygonal capillaries
-changes from $Ca^{1/2}$ on distances $x\propto Ca^{-1}$ to $Ca^{2/3}$ for
-distances $x \gg Ca^{-1}$. Therefore, for a bubble to have a formed shape in
-the  front meniscus we picked up the bubble length as $5$ channel heights. The
-channel length is taken to minimize the influence of one bubble on another,
-because of the periodicity of boundary conditions. For the simulations the
-channel length is taken as $3$ bubble lenghts giving all together $15$ channel
-heights. Figure \ref{fig:benchmark:sketch} represents the benchmark sketch.
+changes from $\Ca^{1/2}$ on distances $x\propto \Ca^{-1}$ to $\Ca^{2/3}$ for
+distances $x \gg \Ca^{-1}$. Therefore, for a bubble to have a formed shape in
+the  front meniscus we chose the bubble length to be $5$ channel heights. The
+channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
+to minimize the influence of one bubble on another,
+because of the periodicity of boundary conditions.
+Figure \ref{fig:benchmark:sketch} represents a sketch of the geometry
+used in the benchmark.
 \begin{figure}
@@
 \subsection{The nondimensionalization and initialization procedure}
-This section is to introduce the procedure of nondimensionalization and
-initialization techniques to be able to predict parameters necessary for
-simulation. The capillary number governs the interface thickness and defined as:
+We now intrduce the procedure of nondimensionalization, as well as
+the initialization techniques we used to be able to predict parameters necessary for
+simulation. The capillary number governs the interface thickness and is defined as:
 \begin{equation}
-Ca=\frac{U_{bubble} \mu_{liq}}{\gamma},
+Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
 \end{equation}
-where $\mu_{liq}$ is the liquid viscosity, $\gamma$ is the surface tension. The
-bubble velocity $U_{bubble}$ is taken in the center of the bubble where the film
+where $\mu_{\mathrm{liq}}$ is the liquid viscosity, and $\gamma$ is the surface tension. The
+bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
 thickness is examined.
-While parameters as $\mu_{liq}$, $\gamma$ and capillary number $Ca$ are given
-or defined through the model parameters, the velocity $U_{bubble}$ is not
-defined. In the simulation we approximately match the velocity with the
-Poiseuille profile center velocit, which is given by the equation:
+While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
+or defined through the model parameters, the velocity $U_{\mathrm{bubble}}$ is not.
+In the simulation we approximately match the velocity with the
+Poiseuille profile center velocity, which is given by the equation:
 \begin{equation}
 \begin{aligned}
-U_{bubble} \approx U_{liq}=\frac{H^2}{8
-\mu_{liq}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
-U_{bubble}\approx
-\frac{{N_y}^2}{\frac{8}{6}(\tau_{liq}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
+U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H^2}{8
+\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
+U_{\mathrm{bubble}}\approx
+\frac{{N_y}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
 x } ,
@@
 \end{equation}
-In reality we iterate the body force value
-until the desired velocity and the capillary number as well are achieved.
+In practice, we iterate adjusting the body force value until both the desired velocity and the
+capillary numbe are achieved.
 
-After one resolves necessary parameters to run a simulation, one needs to
-initialize the macroscopic fields and populations.
+After one calculates the parameters necessary to run a simulation, one needs to
+initialize the macroscopic fields and particle populations.
 The velocity is initialized with $0$ everywhere. Populations are initialized
-with the equilbrium values for the binary liquid model including all the phase
-gradients and laplacians. The liquid phase is initialized with $1$ and the
-gas phase is initialized with $-1$. Though the initialization film thickness
-is not involved in the final result for the front meniscus, Fig.
-\ref{fig:different:initialization:widths}, we keep the initialized film
+with the equilibrium values for the binary liquid model, including all the phase
+gradients and Laplacians. The liquid phase is initialized with $\phi = 1$ and the
+gas phase is initialized with $\phi = -1$. Though the initialization film thickness
+is not involved in the final result for the front meniscus (see Fig.
+\ref{fig:different:initialization:widths}) we keep the initialized film
 thickness as close as possible to the already obtained numerical film
-thicknesses \cite{giavedoni-numerical}. Otherwise, the simulation time for the
-bubble to obtain a steady-state form increases.
+thickness values \cite{giavedoni-numerical}. This is done to minimize
+the simulation time, as starting far away from the steady-state requies
+additional time for the relaxation.
 \begin{figure}
@@
 \caption{The phase plots for different bubble widths as
-$H_{eff}-12$,$H_{eff}-16$ and
-$H_{eff}-20$. The grid for the simulation is $102\mathrm{x}1501$. The phase
-profiles were obtained after $200000$ time steps. One can see that even the
-system is initialized with different bubble volumes the system comes to the
-shape, where the film thickness is conserved. For the given parameters the
+$H_{\mathrm{eff}}-12$,$H_{\mathrm{eff}}-16$ and
+$H_{\mathrm{eff}}-20$. The grid for the simulation is $102 \times 1501$. The phase
+profiles were obtained after $2\cdot10^5$ time steps. One can see that even if the
+system is initialized with different bubble volumes, the bubbles always relaxate
+to a shape where the film thickness is conserved. For the given parameters the
 interface thickness is $0.0615$, $0.0611$, $0.0605$.
@@
 
-Here two different examples are presented that can quantitatively show the
+Here, two different examples are presented that can quantitatively show the
 initialization procedure for quite different capillary numbers. In these
@@
 $5\sqrt{\frac{k}{A}}=5$ lattice boltzmann units. For the sake of simplicity and
-stability the relaxations parameter for liquid and gas phases are taken as
-$\tau_{liq}=2.5$ and $\tau_{gas}=0.7$. That gives the gas-liquid viscosity ratio
-as $10$.  
+stability the relaxation times for liquid and gas phases are taken as
+$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which corresponds
+to a gas-liquid viscosity ratio of $10$.
 \begin{description}
@@
   The film width for such small Capillary numbers \cite{bretherton} is
-given by equation:
+given by the equation:
   \begin{equation}
-  \frac{h_{\infty}}{H}=1.3375 Ca^{2/3}=0.03899,
+  \frac{h_{\infty}}{H}=1.3375 \Ca^{2/3}=0.03899,
   \end{equation}
-  where we assume that $2 H$ or $Ny$ in lattice units is the channel width.
-Therefore one needs to resolve the interface which is around $2$ percent of
-  whole channel width with for instance $8$ lattice units, where $5$ lattice
+  where we assume that $2 H$ or $N_y$ is the channel width in lattice units.
+
+\todo{Therefore one needs to resolve the interface which is around $2$ percent of
+the  whole channel width with, for instance $8$ lattice units, where $5$ lattice
 units occupies the interface width for given. Note that the interface occupies
-$62.5$ percent of the interface.  Therefore the channel
-  width is $400$ lattice units. We take the bubble length as $5$ channel widths
-  and the distance between bubble is as $2$ bubble lengths. It gives the
-grid size as $400\mathrm{x}6000$. In terms of computer memory
-requirement the calculations with the double precision accuracy demand:
+$62.5$ percent of the interface. -- these two sentences need to be rewritten for clarity}  Therefore the channel
+  width is $400$ lattice units.
+
+  We take the bubble length as $5$ channel widths
+  and the distance between bubbles is $2$ bubble lengths. This gives a
+grid size of $400 \times 6000$. In terms of computer memory
+requirements, the calculations with double precision accuracy demand:
 \begin{equation}
 \begin{aligned}
-&2\text{ set of populations }\mathrm{x}\,\, 8\,\, \text{bytes per population
-}\mathrm{x}\,\, 9 \text{ number of
+&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
+} \times \,\, 9 \text{ number of
 populations}\\
-&\mathrm{x}\,400\,\mathrm{x}\,6000\approx 345 \mathrm{MB}. 
+&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
 \end{aligned}
@@
 Even with this naive assumption for the $50$ percent resolution of the film
-widht in comparison with the interface width, one can get the memory
-requirement as $345 MB$.
+width in comparison with the interface width, one can get the memory
+requirement of $345 \mathrm{MB}$.
 
@@
   \begin{equation}
-  U_{bubble}=0.005 \frac{\gamma}{\mu_{slug}}=0.005
-\frac{0.0377}{0.66666}=2.82\,10^{-4},
+  U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{slug}}}=0.005
+\frac{0.0377}{0.66666}=2.82 \cdot10^{-4},
   \end{equation}
-  where $\mu_{slug}=1/3 (2.5-0.5)=0.6666$.
+  where $\mu_{\mathrm{slug}}=1/3 (2.5-0.5)=0.6666$.
   This velocity implies the following pressure gradient calculated from the
-  Poisuielle flow solution:
+  Poiseuille flow solution: \todo{maybe add a ref, to this eq. from the previous section?}
   \begin{equation}
   \begin{aligned}
-  &U_{bubble}=\frac{H^2}{8\mu} \frac{\mathrm{d}P}{\mathrm{d}x}=\frac{Ny^2}{8/3
-  (\tau_{liq}-1/2)}\frac{\mathrm{d}P}{\mathrm{d}x}\\
-  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca
+  &U_{\mathrm{bubble}}=\frac{H^2}{8\mu} \frac{\mathrm{d}P}{\mathrm{d}x}=\frac{N_y^2}{8/3
+  (\tau_{\mathrm{liq}}-1/2)}\frac{\mathrm{d}P}{\mathrm{d}x}\\
+  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca
   \end{aligned}
   \end{equation}
-  Therefore the pressure gradient established is the following:
+  The pressure gradient is then:
   \begin{equation}
   \begin{aligned}
-  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42\, 10^{-9}.
+  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
   \end{aligned}
@@
   Note that in the case of large grids  height resolution is directly
-involved in	
-the physical time step as:
+involved in calculation of the physical time step :since:
 \begin{equation}
-\Delta t =\frac{U_{LBM}}{U_{phys}} \Delta x ,
+\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
 \end{equation}
 where for example the real velocities in the physical world can reach
-approximately $1$ m/s and physical channel height upto $1$ mm. In this case one
+approximately $1$ m/s and physical channel height up to $1$ mm. In this case one
 can easily approximate the physical time step:
 \begin{equation}
-\Delta t = 7.05 \, 10^{-10} \,\mathrm{s}
+\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
 \end{equation}
-which is a small time step. The typical simulation needs approximately
+which is quote small. A typical simulation needs approximately
 $100000-300000$ time steps to come to the steady state solution giving the
-physical time simulation around $10^{-5}--10^{-4}\,\mathrm{s}$, which is
-probably not enough to find desired correlations.
+physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
+probably not enough to find the desired correlations. \todo{correlations between what?}
 
@@
   $0.005$. However, a lot of simulations devoted to the Bretherton/Taylor
-flow
-  simulated and extended it in the range of Capillary numbers greater than
-$0.005$
-  and for different Reynolds numbers. One can refer to figures
-  \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from citations
-  \cite{giavedoni-numerical,heil-bretherton} which crossvalidated their
-  simulations with Bretherton and each other. Even though LBM can simulate the
-flows with small capillary numbers, i.e. less than $0.005$, but waiting time
-for the simulation to be accomplished is relatively long. Therefore, we focus
-our work on the simulations of \citet{giavedoni-numerical} which crossvalidated
-results not only for small capillary numbers but for large capillary numbers as
-well. As far as the work intention is
-  to show proper physics behavior the moderate parameters are
-taken which
-  can be easily and fast validated on a computer. 
-  One can see from the Fig. \ref{fig:heil:planar} that the width for the
-  Capillary number equal to $0.05$ ranges from $0.12$ to $0.15$ while Reynolds
-  number ranges from $0$ to $300$. This capillary number has moderate film
-widths
-  to be resolved with certain accuracy.
+flow extended it to the range of Capillary numbers greater than
+$0.005$, and to different Reynolds numbers. One can refer to figures
+  \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
+  \citet{giavedoni-numerical,heil-bretherton}, who crossvalidated their
+  simulations with Bretherton and each other. Even though the LBM can simulate
+flows with small capillary numbers, i.e. less than $0.005$, the simulation time
+is relatively long. Therefore, we focus our work on the simulations of
+\citet{giavedoni-numerical} who crossvalidated
+results for all ranges of capillary numbers.  Our intention is
+  to show proper physics behavior for moderate parameters
+which
+  can be easily and quickly validated on a computer.
+  One can see from Fig. \ref{fig:heil:planar} that the width for the
+  Capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
+  number ranges from $0$ to $300$.
 
   The film width for such small Capillary numbers is taken from
-  \ref{fig:heil:planar} for Reynolds number $40$ equals to $0.12$. We want to
+  \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
   resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
@@
   width is $200$ lattice units. We take the bubble length as $5$ channel widths
-  and the distance between slugs we take as $2$ bubble lengths. That gives
-the grid size as $200\mathrm{x}3000$. The velocity can be
+  with distance between slugs being $2$ bubble lengths \todo{this I think is the 3rd
+  time this particular geometry is mentioned in this paper, perhaps just skip it in this place?}. That gives
+the grid size of $200 \times 3000$. The velocity can be
 calculated from the Capillary number as:
   \begin{equation}
-  U_{bubble}=0.05 \frac{\gamma}{\mu_{liq}}=0.05 \frac{0.037712}{0.6666}=2.82\,
+  U_{\mathrm{bubble}}=0.05 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.05 \frac{0.037712}{0.6666}=2.82\cdot
   10^{-3}.
@@
   The velocity implies the following pressure gradient calculated from the
-  Poisuielle flow solution:
+  Poiseuille flow solution:
   \begin{equation}
   \begin{aligned}
-  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca=3.77\,10^{-7}.
+  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca=3.77\cdot10^{-7}.
   \end{aligned}
   \end{equation}
-  The initialization approach is to take values from the already performed
-numerical simulation and to calculate the grid numbers from them matching the
+  Our approach to the initial conditions is  to take values from the previously performed
+numerical simulations and to calculate the grid sizes from them, matching the
 body force gradient through the solution of Poiseuille flow.
@@
 \end{figure}
-Note that we base the initialization techniques through correlations for the
+Note that we base the initialization techniques on correlations for the
 capillary number and the body force. However, in the binary liquid framework
-those correlations give only the approximation. First of all the body force
-which is based on the Poiesuille velocity profile for the liquid phase is the
-biggest assumption. In reality the less viscous bubble will have a slippage
+those correlations are only approximate. The biggest assumption lies
+in calculating the body force from the Poiseuille velocity profile.
+In reality, the less viscous bubble will have a slippage
 factor and its velocity is larger than the surrounding liquid velocity. By
@@
 larger capillary number can be obtained. While we present the initialization
-tips, in reality one needs to take the bubble velocity from the simulations and
+tips, in practice one needs to take the bubble velocity from the simulations and
 recalculate all the necessary characteristics or in other words adjust the
@@
 \subsection{Grid refinement}
-To properly estimate the interface resolution one needs to perform the
-convergence study dependance on the grid resolution. All parameters are fixed,
-including a bubble velocity, the capillary number and change the grid
-resolution. Our goal is to know exactly the ratio of the interface width to the
+To properly estimate the interface resolution one needs to study the convergence
+of the results of a function of the grid resolution. To do that, all parameters are held fixed,
+including the bubble velocity, the capillary number, an only the grid
+resolution is changed. Our goal is to know exactly the ratio of the interface width to the
 film thickness where results are not dependant on the grid resolution.
 
-The example what we want to simulate is the capillary number
-$0.05$. As it was pointed out the film thickness in that case is $6$ percent.
-Note that in the case of half-way bounce-back walls which are used in the
+We illustrate the procedure in more detail for the capillary number
+$0.05$. As it was pointed out earlier, the film thickness in that case is $6$ percent.
+Note that in the case of the half-way bounce-back walls \todo{the bounce-back
+proceduce was not described earlier in this paper} which are used in the
 simulations one needs to calculate the film thickness as:
@@
 \end{equation}
-where $\phi_0$ is the grid coordinate where phase field equals to $0$, $N_y-2$
-is related to the effective channel height. Note that it's a big
+where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
+is related to the effective channel height. Note that it is a big
 simplification to impose the boundary in the middle beetween the bounce-back
@@
 the wall \cite{ginzburg-multireflection}. In the case of $\delta=0.06$ the
-initial grid number
-is $N_y=102$, which gives the horizontal grid number as $N_x-1=15(N_y-2)=1500$.
+initial grid size is $N_y=102$, which gives the horizontal grid size as $N_x-1=15(N_y-2)=1500$.
 The bubble is initialized as a rectangular box with coordinates
 $y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
-$\phi_{bubble}=-1$. All other nodes are initialized with the phase field
+$\phi_{\mathrm{bubble}}=-1$. All other nodes are initialized with the phase field
 $\phi=1$. The force gradient can be estimated through the Poiseuille
-profile formula, Eq.\ref{poiseuille:velocity:center}, and it equals
-$1.508\,10^{-6}$.
+profile formula, Eq.~\ref{poiseuille:velocity:center}, and it equals
+$1.508 \cdot 10^{-6}$.
 
-After choosing the reference parameters, the grid refinement needs to keep the
-same the same macroscopic parameters as capillary number and viscosities ratio.
-Therefore the simulations were performed with grid
-numbers $N_y=102,127,152,177,202,227$, corresponding to the
-effective channels widths as $H_{eff}=100,125,150,175,200,225$,
+After choosing the reference parameters, the grid refinement procedure
+needs to keep the macroscopic parameters constant.  Therefore the simulations
+were performed with grid numbers $N_y=102,127,152,177,202,227$, corresponding to the
+effective channels widths  $H_{\mathrm{eff}}=100,125,150,175,200,225$,
 $N_x=1501,1876,2251,2626,3001,3376=15
-H_{eff}+1$. It is easy to check
+H_{\mathrm{eff}}+1$. It is easy to check
 that keeping all liquid viscosities the same yields the following quantity to
-be a
-constant while doing the grid refinement
-$H_{eff}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
-x } = Const$. We performed a number of simulations for the prescribed grids.
+be independent of grid size
+$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
+x } = \mathrm{const}$. We performed a number of simulations for the prescribed grids.
 The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
-see the visual results convergence for $H_{eff}\geq 175$. The calculated widths
+see the visual results converge for $H_{\mathrm{eff}}\geq 175$. The calculated widths
 are $0.0606$,$0.0618$,$0.0617$,$0.0629$,$0.0624$,$0.0624$. We accept the
-convergent results for grids larger than $H_{eff}=175$. However, one can see
-that imposing proper initialization technique, large enough time and the wall
-wettability reveal results which are different only in 3rd digit even for
-underresolved film widths. To calculate how good the interface thickness is
-resolved the interface thickness to the film width thickness is calculated.The
+convergent results for grids larger than $H_{\mathrm{eff}}=175$. However, one can see
+that with proper initialization techniques, large enough time and the wall
+wettability, results are different only in the 3rd digit even for
+underresolved film widths. To calculate how well the interface thickness is
+resolved, the ratio of the interface thickness to the film width thickness is calculated. The
 interface itself occupies approximately $5 \xi$, where
@@
 thickness as
-$5\xi/H_{film}$ is as follows $0.824$, $0.646$,  $0.539$, $0.453$,
+$5\xi/H_{\mathrm{film}}$: $0.824$, $0.646$,  $0.539$, $0.453$,
 $0.400$, $0.355$. Therefore, one needs to
 resolve the interface as $40-50$ percent of the
-expected film width for simulations to be converged. We further examine the
+expected film width for simulations to be convergent. We further examine the
 velocities in the center of the bubble to calculate the Capillary number. The
 calculated numbers are $0.0041$, $0.0041$,$0.0040$,$0.0040$,$0.0041$,$0.0039$.
-This corresponds to the average capillary number as $0.07$, which is slightly
+This corresponds to the average capillary number of $0.07$, which is slightly
 larger than the expected capillary number. The corresponding error can be
-attributed to the density and viscosity ratio which is much larger in the
+attributed to the density and viscosity ratios which are much larger in the
 Bretherton problem.
@@
 \includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
-\caption{Grid refined profiles for the effictive
+\caption{Grid-refined profiles for the effictive
 channel widths
-$H_{eff}=100,125,150,175,200,225$.One can see
+$H_{\mathrm{eff}}=100,125,150,175,200,225$.One can see
 convergent results even for the underresolved film
@@
 \subsection{The influence of the wall gradient}
-The results should not depend on the wall wettability. Wettability is defined
-through the phase gradient \cite{pooley-contact}, $\partial_n \phi$. This
-section is to examine the wall gradient influence on simulations. We take large
-enough resolution for simulations to be consistent. Simulation domain was
-chosen as $177\mathrm{x}2626$ and the initial film width as
-$12$ lattice boltzmann units, which correspond to the predicted film thickness.
+The simulation results should not depend on the wall wettability. Wettability is defined
+through the phase gradient \cite{pooley-contact}, $\partial_n \phi$.  We took a large
+enough grid resolution in order for simulations to be consistent -- the grid size was
+$177 \times 2626$ and the initial film width was
+$12$ lattice boltzmann units, which corresponds to the predicted film thickness.
 We examined $11$ different values for the wall
@@
 wall gradients given that the interface is resolved in the right way. For the
-given wall gradients the values of interface thicknesses are
+given wall gradients the values of interface thickness values are
 $0.0633$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0633$,$0.0632$,
@@
 simulations are unstable. One can see the simulated profiles in Fig.
-\ref{fig:gradients:profiles} The negative values of the phase gradient are
-preferable, as
-far as the phase of the liquid adjacent to the wall has the value of $1$. That
-means that the phase values near the wall are above $1$ and do not interact
+\ref{fig:gradients:profiles}. Negative values of the phase gradient are
+preferable, since the phase of the liquid adjacent to the wall has the value of $1$.
+The phase values near the wall are above $1$ and do not interact
 with the bubble which has phase value $-1$.  In case of positive
-gradients the values near the wall are below $1$ and the gradient profile can
-fuse with the values of the gas phase as $-1$, creating the wall stickyness,
-unphysical one.
+gradients the values near the wall are below $1$, and the gradient profile can
+fuse with the values of the gas phase as, giving rise to an unphysical wall stickyness.
 \begin{figure}
 \includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
-\caption{Phase profiles for different wall gradients. One can see if the width
-is properly resolved then the wettability of the wall doesn't play any role
-affecting the film thickness.
+\caption{Phase profiles for different wall gradients. One can see that if the width
+is properly resolved then the wettability of the wall does affect the film thickness.
 \label{fig:gradients:profiles}
@@
 Another question which arises is the influence of the wall gradient on the bubble velocity and the
-capillary number accordingly. The shear stress controlled by the wall phase gradient changes the
+corresponding capillary number. The shear stress controlled by the wall phase gradient changes the
 effective velocity near the wall, thus changing the bubble velocity as well.
@@
 \citet{giavedoni-numerical}, Fig. \ref{fig:giavedoni:planar}, and
-\citet{heil-bretherton}, Fig. \ref{fig:heil:planar}, for the range of capillary
-numbers. Because of limiting computational factors as memory and the
-performance we cannot go to the range of small
-capillary numbers. However, all the other methods were validated across the
-small capilary numbers and extended for the range of capillary numbers from
-$0.005$ to $10.0$. Our goal is to compare the film thickness over the capillary
+\citet{heil-bretherton}, Fig. \ref{fig:heil:planar}, for a range of capillary
+numbers. Because of limited computational resources, we skip the
+small capillary numbers and compare the film thickness over the capillary number
 range from $0.03$ to $0.8$, which is a computationally reasonable task.  For the
@@
 results we
-suggest the following Capillary numbers as:
+suggest the following Capillary numbers:
 \begin{equation}
@@
 To perform numerical simulations we refer
-to
-grid refinement section where we obtained that the interface width should be
-at most $50$ percent of the interface width. Therefore the grid was chosen as
-$202 \mathrm{x} 3001$. $5$ lattice boltzmann units do not occupy more than $60$
+to the grid refinement section where we established that the interface width should be
+at most $50$ percent of the interface width.  We therefore choose the grid to be
+$202 \times 3001$. $5$ lattice Boltzmann units do not occupy more than $60$
 percent of the effective width. Because the simulation get unstable with
 smaller grids and larger gradients, all the capillary number simulations were
-performed on the grid chosen. To properly initialize the force gradient the
-proportionality law was utilized. For instance, with the forcing
-$6\,10^{-6}/16$ was chosen for the grid to obtain predicted capillary number as
-$0.005$. Therefore, one can obtain the initial forcing for other grids using
+performed on the chosen grid. To properly initialize the force gradient, the
+proportionality law was utilized. The forcing
+$6 \cdot 10^{-6}/16$ was chosen to obtain the predicted capillary
+number $0.005$. The forcing for other grids can be obtained using
 simple correlations:
@@
 \begin{aligned}
-&Ca_{lit} \propto U_{bubble}\\
-&U_{bubble} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\ 
-&Ca_{lit} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
-&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{N_y^2},
+&Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
+&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\
+&Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
+&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{N_y^2},
 \end{aligned}
 \end{equation}
-where the index $lit$ stands for the predicted capillary number
+where the index ,,lit'' stands for the predicted capillary number
 \cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
-involved as because the simulations are conducted on the same grid.
-Therefore,the pressure gradient can be obtained through the capillary numbers
+involved because all simulations are conducted on the same grid.
+The pressure gradient can be obtained through the capillary numbers
 ratio:
 \begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d} x}=6\,10^{-6}/16 \frac{Ca_{lit}}{0.05}
+\frac{\mathrm{d}P}{\mathrm{d} x}=6 \cdot 10^{-6}/16 \frac{Ca_{\mathrm{lit}}}{0.05}
 \end{equation}
@@
 \begin{equation*}
-w=12 \frac{Ca_{lit}}{0.05}
+w=12 \frac{Ca_{\mathrm{lit}}}{0.05}
 \end{equation*}
-The results were obtained after $200,000$ steps.
-Film thicknesses were obtained as $0.042$,$0.066$, $0.085$, $0.077$, $0.123$, $0.153$, $0.164$, $0.173$ with the bubble center velocities as $0.0022$, $0.0041$, $0.0068$, $0.0055$, $0.0185$, $0.0388$, $0.0592$, $0.0782$.
-Based on the obtained center velocities the capillary numbers
-are formed as $0.039$,$0.072$,$0.120$,$0.098$,$0.328$,$0.686$,$1.046$,$1.383$.
-Those capillary numbers are overpredicted but the obtained lengths are overpredicted as well due to the improper body force imposement.
-We extracted data from \cite{giavedoni-numerical,heil-bretherton} and compared it with the simulation. The results are shown in Fig. \ref{fig:capillary:comparison}.
+The results were obtained after $2\cdot10^5$ steps. The following film thicknesses were
+calculated: $0.042$,$0.066$, $0.085$, $0.077$, $0.123$, $0.153$, $0.164$, $0.173$ with
+the corresponding bubble center velocities $0.0022$, $0.0041$, $0.0068$, $0.0055$,
+$0.0185$, $0.0388$, $0.0592$, $0.0782$. Based on the obtained center velocities,
+the capillary numbers are calculated as
+$0.039$,$0.072$,$0.120$,$0.098$,$0.328$,$0.686$,$1.046$,$1.383$. These
+capillary numbers are overpredicted but the obtained lengths are overpredicted
+as well due to the improper body force imposement \todo{what is improper about? do you refer
+to the fact that the Poiseuille should not be sued here?}.  We extracted data from
+\cite{giavedoni-numerical,heil-bretherton} and compared it the results of our
+simulations, which is shown in Fig. \ref{fig:capillary:comparison}.
 \begin{figure}
 \includegraphics[width=0.97\textwidth]{Figures/Capillary/capillaries_comparison.eps}
-\caption{The comparison between simulation results and results from \cite{giavedoni-numerical} and \cite{heil-bretherton}.
-One can see the reasonable agreement.
-The film thickness against the capillary number range is presented.\label{fig:capillary:comparison}}
+\caption{A comparison between simulation results and results from
+\cite{giavedoni-numerical} and \cite{heil-bretherton}. One can see a
+reasonable agreement. The plots depict the film thickness as a function of the
+the capillary number.\label{fig:capillary:comparison}}
 \end{figure}
@@
 \subsection{The film variation over the bubble}
-This section is intended to show the qualitative variation of the film thickness over the bubble.
-Though there are simulations for the bubble length simulations for the two-dimensional case, due to
-the presentation of data it is hard to extract numerical data. One can see bubble lengths
+
+Next, we investigated the qualitative variation of the film thickness over the bubble length.
+Though there are simulations for the bubble length for the two-dimensional case, due to
+the presentation of data it is hard to extract numerical values from the plots. One can see bubble lengths
 variations for the two-dimensional simulations for the range of capillary numbers in Fig.
@@
 \end{figure}
-One can see the qualitative agreement, i.e. the thickness increases towards the front meniscus and
+One can see a qualitative agreement, i.e. the thickness increases towards the front meniscus and
 rapidly decreases towards the rear meniscus. This shape sometimes is called as a "bullet" shape.
@@
 \subsection{The influence of different viscosities}
-The results obtained earlier are taken for the liquiqs viscosity ratio $10$.
-We performed the same simulations for the liquids viscosity ratio as $20$.
-Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. We
-haven't found significant differences with the results obtained earlier. The
-comparison for the liquid thicknesses is presented on Fig.
-\ref{fig:capillary:viscous}. Therefore one can validate the assumption that in
-the low capillary flow regime the density ratio is not involved. As well the
-gas-liquid viscosity ratio of the order $10$ is enough to obtain results
+The results obtained earlier are taken for the liquid viscosity ratio $10$.
+We performed the same simulations for the liquid viscosity ratio of $20$.
+The relaxation parameters were taken as $\tau_{\mathrm{liquid}}=4.5$ and $\tau_{\mathrm{bubble}}=0.7$. We
+did not find any significant differences with the results obtained earlier. The
+comparison for the liquid thicknesses is presented in Fig.
+\ref{fig:capillary:viscous}. The results validate our assumption that in
+the low capillary flow regime the density ratio does not affect the film thickness,
+and that the gas-liquid viscosity ratio of the order $10$ is enough to obtain results
 consistent with the Bretherton/Taylor problem.
@@
 \includegraphics[width=0.97\textwidth]{Figures/Capillary_Viscous/capillaries_viscous.eps}
-\caption{The widths versus capillary number comparison for
-$\frac{\mu_{liq}}{\mu_{gas}}=10$ and $\frac{\mu_{liq}}{\mu_{gas}}=20$. The
-results show good consistency and agreement.\label{fig:capillary:viscous}}
+\caption{The film widths versus capillary number for
+$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=20$. The
+results show good consistency and mutual agreement.\label{fig:capillary:viscous}}
 \end{figure}
@@
 \section{Conclusion}
-The work presents the numerical tips and the benchmark for the
+This work presents numerical tips and a benchmark for the
 Bretherton/Taylor problem using the binary liquid lattice Botlzmann method. The
-results show consistency with the already obtained numerical correlations using
-other methods. Though results are specific for the binary liquid lattice
-Botlzmann method but numerical tips and correlations can be used for any
-continuous interface methods. The work is valuable for people simulating
-microchannel flows.
+results show consistency with the previously published numerical results obtained
+using other methods. Though our results are specific to the binary liquid lattice
+Botlzmann method, the numerical hints and procedures can be used for any
+continuous interface method.  We hope this work will prove to be valuable for
+people simulating microchannel flows.
 

### 2010-11-08 Text/paper.tex
Answered Michal's comments. Need to forward it to the supervisor.

@@
 liquid model we use is a continuous-interface multiphase model, which presents
-the question of how to properly resolve the interface to obtain proper correlations \todo{between what?}.
+the question of how to properly resolve the interface in comparison with the film thickness to
+obtain Bretherton correlations.
 In this work we investigate how the binary liquid model parameters and the grid resolution
@@
 resolution. In other words, when it comes to the resolution of the film
-thickness the diffusion \todo{diffusive?} interface should be resolved in such a way as to have
+thickness the diffusive interface should be resolved in such a way as to have
 negligible effect on the physics of the film.
@@
 $\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
-The surface tension in the framework of the binary liquid model is $\sqrt{8 k A}{9}$ \todo{is this right?  should the 9 be in the
-denominator?  is not, perhaps put before the root?}.
+The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k A}{9}}$.
 
@@
 influence the overall solution.
- \item The free-energy model examined in the work is single density, two
+ \item The free-energy model examined in the work is a single density, two
 viscosities model. The classical Bretherton problem is a free-interface problem
-where the film thickness is established by itself. The viscosity ratio should
-be large, as indicated in \cite{pagonabarraga-fingers}, however no studies
-address this issue. \todo{what's the conclusion of the impact of this issue to our paper then?}
+where the film thickness is established by itself. Thus, the viscosity and density ratios for the
+Bretherton problem should
+be large. However, in terms of the lattice Boltzmann free-energy framework those issues are not
+addressed and will be studied in the current work.
   \item The Bretherton problem addresses bubble flow driven by a pressure
@@
 the analysis of error introduced by the boundary conditions becomes an issue.
-Our benchmark is therefore done is for an axis-symmetric case. \todo{since this paper
-is about a 2d flow, which we already said in the intro, what is the point of discussing
-curve boundaries which are only an issue in 3d?}
+Our benchmark is therefore done is for a plane-symmetric case in the two-dimensional case. We do
+not address the axis-symmetric two-dimensional case since it implies the derivation of additional
+mass and force terms for the square based LBM grid \cite{halliday-circular}.
 \end{itemize}
@@
   \end{equation}
-  where we assume that $2 H$ or $N_y$ is the channel width in lattice units.
-
-\todo{Therefore one needs to resolve the interface which is around $2$ percent of
-the  whole channel width with, for instance $8$ lattice units, where $5$ lattice
-units occupies the interface width for given. Note that the interface occupies
-$62.5$ percent of the interface. -- these two sentences need to be rewritten for clarity}  Therefore the channel
-  width is $400$ lattice units.
+  where we assume that $2 H$ or $N_y$ is the channel width in lattice units. Thus, the film
+thickness occupies around $2$ percent of the whole channel height. Let us impose that $8$ lattice
+units are required to resolve  $2$ percent of the channel. Note that $5$ out of $8$ lattice units
+($62.5$ percent) are used to resolve the phase interface. Therefore the channel width is $400$
+lattice units.
 
@@
 \end{equation}
-Even with this naive assumption for the $50$ percent resolution of the film
+Even with this naive assumption for the $62.5$ percent resolution of the film
 width in comparison with the interface width, one can get the memory
@@
   This velocity implies the following pressure gradient calculated from the
-  Poiseuille flow solution: \todo{maybe add a ref, to this eq. from the previous section?}
-  \begin{equation}
-  \begin{aligned}
-  &U_{\mathrm{bubble}}=\frac{H^2}{8\mu} \frac{\mathrm{d}P}{\mathrm{d}x}=\frac{N_y^2}{8/3
-  (\tau_{\mathrm{liq}}-1/2)}\frac{\mathrm{d}P}{\mathrm{d}x}\\
-  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca
-  \end{aligned}
-  \end{equation}
-  The pressure gradient is then:
+  Poiseuille flow solution, Eq. \ref{poiseuille:velocity:center}: 
   \begin{equation}
@@
 physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
-probably not enough to find the desired correlations. \todo{correlations between what?}
+probably not enough to find the desired physics correlations taken place on macroscopic time level.
 
@@
 lattice units are related to the interface resolution.  Therefore the channel
-  width is $200$ lattice units. We take the bubble length as $5$ channel widths
-  with distance between slugs being $2$ bubble lengths \todo{this I think is the 3rd
-  time this particular geometry is mentioned in this paper, perhaps just skip it in this place?}. That gives
-the grid size of $200 \times 3000$. The velocity can be
+  width is $200$ lattice units. Therefore using the assumptions for the benchmark 
+the grid size is $200 \times 3000$. The velocity can be
 calculated from the Capillary number as:
@@
 $0.05$. As it was pointed out earlier, the film thickness in that case is $6$ percent.
-Note that in the case of the half-way bounce-back walls \todo{the bounce-back
-proceduce was not described earlier in this paper} which are used in the
+Note that in the case of the half-way bounce-back walls \cite{yu} which are used in the
 simulations one needs to calculate the film thickness as:
@@
 $0.039$,$0.072$,$0.120$,$0.098$,$0.328$,$0.686$,$1.046$,$1.383$. These
-capillary numbers are overpredicted but the obtained lengths are overpredicted
-as well due to the improper body force imposement \todo{what is improper about? do you refer
-to the fact that the Poiseuille should not be sued here?}.  We extracted data from
+capillary numbers are overpredicted but the obtained widths are overpredicted
+as well due to the improper body force imposement, because the assumption of Poiseuille flow is not
+valid. If one needs to obtain the desired capillary number, the shooting method for the forcing is
+necessary starting as the Poiseuille pressure gradient force as the initial condition. We extracted
+data from
 \cite{giavedoni-numerical,heil-bretherton} and compared it the results of our
-simulations, which is shown in Fig. \ref{fig:capillary:comparison}.
+simulations, which are shown in Fig. \ref{fig:capillary:comparison}.
 \begin{figure}

### 2010-11-22 Text/paper.tex
Added half modifications.

@@
 \usepackage{bm}
-\usepackage[numbers]{natbib}
+\usepackage[numbers,sort&compress]{natbib}
 \usepackage{color}
@@
 \begin{abstract}
-The motion towards exploration of nano- and microscale phenomena brings new
-challenges to simulation tools. The continuous physics approach is limited on
-those scales, and methods such as Molecular Dynamics (MD) or the Lattice
-Boltzmann Method (LBM) become more appropriate. This work aims to
-reproduce the gravity driven microchannel flow of a free surface liquid with the
-help of the free-energy binary liquid lattice Boltzmann method. In the low capillary number
-$\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
-with the deposition film thickness proportional to $\Ca^{2/3}$. The binary
-liquid model we use is a continuous-interface multiphase model, which presents
-the question of how to properly resolve the interface in comparison with the film thickness to
-obtain Bretherton correlations.
-In this work we investigate how the binary liquid model parameters and the grid resolution
-need to be adjusted to properly simulate microchannel flows.  The presented results
-should be of interest for industrial applications and to researchers simulating
-Bretherton phenomena.
+This work shows that the free-energy binary liquid lattice Boltzmann method is able to
+reproduce the gravity driven microchannel flow of an air bubble through liquid. In the low capillary
+number
+$\Ca$ regime (defined in Eq. \ref{capillary:number:definition}), the flow is thoroughly described by
+Bretherton \cite{bretherton},
+with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
+from the Bretherton correlations for $Ca\geq 0.005$, studied experimentally and numerically. This
+work aims to examine the lattice Boltzmann binary liquid model feasibility to obtain valid results
+for an air bubble flow in liquid for a range of $Ca$ number. The binary
+liquid model is a continuous-interface multiphase model, which raises
+the question of how to properly resolve the interface in comparison with the film thickness.
+It was found that the film interface needs to be as $50$
+percent in comparison with the film thickness for simulation to give convergent results. While
+the convergency criteria is fulfilled the results are of good comparison with the
+numerical simulations of \citet{giavedoni-numerical} in the range of $Ca$ from $0.01$ to $1$. 
 \end{abstract}
@@
 microchannels. It was found that the front meniscus film width is proportional
-to $\Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
+to $\Ca^{2/3}$ for capillary numbers less than $0.003$, where the capillary number is defined as:
+\begin{equation}
+\label{capillary:number:definition}
+Ca=\frac{\mu_{liq} U_{bubble}}{\gamma},
+\end{equation}
+where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is a propagation bubble velocity, and
+$\gamma$ is the surface tension between air and liquid.
+While there are number
 of works which simulate the Bretherton problem with the boundary value method
-\cite{ingham-plates,heil-bretherton}, its simulation using diffusive interface methods
-is difficult.  If the interfaec is spread over several grid nodes, the question of
-proper film width resolution arises. While the discontinuous interface models
-are good at simulating pheonomena in simple geometries, the continuous interface
-methods are more flexible and capable of resolving problems involving coalescence
-or droplet breakup. This work
+\cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
+involving complex geometries, free interface motion or for problems involving coalescence and/or
+droplet breakup. The continuous interface models are more flexible for those kind of simulations.
+However, if the interface is spread over several grid nodes, the question of
+proper film width resolution in comparison with the interface arises.  This work
 concentrates on resolving the parameter range of the binary liquid lattice
-Boltzmann method to validate it the Bretherton regime.
+Boltzmann method to validate it in the range of capillary numbers.
 
@@
 \cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
-method easily tackles complex geometries and the incorporation of
-physical phenomena on the microscopic level, as is the case in multiphase models. Most
+method easily tackles complex geometries andallows for incorporation of
+physical phenomena on the microscopic level, as in the case in multiphase models. Most
 multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
 the interface in a continuous way. Such a representation
-brings issues with the film thickness resolution versus the interface
+brings issues with film thickness resolution versus interface
 resolution. In other words, when it comes to the resolution of the film
@@
 model simulates two liquids with the assumption of uniform overall
-density. Therefore, the question of how to simulate the Bretherton problem
-where the two phases have two completely different densities (i.e. the two fluids
-are liquid and gas) is of interest. However, in the case of the motion of a
-liquid slug and a bubble in a capillary the governing factor is the capillary number
-and the ratio of the viscosities, not of the densities, is the critical parameter.
+density. While the classical Bretherton problem is stated for air and liquid, which are of largely
+different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
+neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
+negligible Reynolds number effect on the film thickness for the range of Reynolds numbers.
+Therefore, the governing factor for microchannel flows is not the densities ratio, but the
+viscosities ratio. Moreover, the results based on the uniform density, as in the case of the LBM
+binary liquid model, are of good comparison with other simulations
+\cite{giavedoni-numerical,heil-bretherton} taking into the account inertia effects.
 
@@
 binary liquid model parameters, the viscosities ratio and the grid
-resolution, in order to obtain correct Bretherton/Taylor film thickness.
+resolution, in order to obtain correct film thicknesses.
+
 The paper is organized as follows.  First, we briefly
@@
 \section{Bretherton bubble flow}
-Bretherton in his significant work \cite{bretherton} studied a long bubble
+Bretherton \cite{bretherton} studied a long bubble
 moving in a tube filled with liquid. It was found that the deposition film width
@@
 it was realized \cite{wong-films,wong-pressure} that the film deposition
-is proportional to $\Ca^{2/3}$ only near the front meniscus and that the film width
-varies throughout the bubble length for bubbles of finite length. Numerical
+is proportional to $\Ca^{2/3}$ only in the specific region behind the front meniscus and
+that the film width
+varies over the bubble length for bubbles of finite length. Numerical
 simulations \cite{giavedoni-numerical} and experimental studies
@@
 To be able to predict consistently the flow pattern for capillaries in
-different ranges of parameters, simulations are developed and their codes
-are typically validated with the small capillary numbers Bretherton problem.
+different ranges of parameters, simulations have been developed and are typically validated with the
+small capillary numbers Bretherton problem.
 
@@
 \citet{vanbaten-circular} studied the mass transfer and film
-thickness for rising bubbles in a circular capillary.
-\citet{kreutzer-pressure-drop} used the finite volume method to perform
+thickness for rising bubbles in a circular capillary using the finite volume method.
+\citet{kreutzer-pressure-drop} used the finite volume method as well to perform
 simulations of a circular capillary for a number of different
-Reynolds and capillary numbers. \citeauthor{wong-films} in
-\cite{wong-films,wong-pressure} studied three-dimensional bubbles in
+Reynolds and capillary numbers. \citeauthor{wong-films} \cite{wong-films,wong-pressure} studied
+three-dimensional bubbles in
 polygonal capillaries and showed the derivations of different shapes in the
@@
 a two-dimensional channel for a range of Reynolds and capillary
-numbers. \citet{giavedoni-numerical} performed crossvalidation of the
+numbers using finite element method. \citet{giavedoni-numerical} performed crossvalidation of the
 finite element solution developed by them with previously published results.
-The solution is available for the circular and planar cases.
+The solution is available for circular and planar cases.
 
@@
 information from one node to another in certain
-directions specified by the discretized velocity set.
+directions specified by the discrete velocity set.
 The LBE is typically implementated as follows:
@@
 \begin{aligned}
-&f_i^{eq}=w_i 
+&f_i^{eq}&&=w_i 
 \biggl(3
@@
 +\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
-\beta}}{2 c_s^4}\biggr), i=1\div8\\
-&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
-&g_i^{eq}=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
+\beta}}{2 c_s^4}\biggr)\\
+&&&+k\bigl(w_i^{xx} (\partial_x \phi)^2+w_i^{yy} (\partial_y \phi)^2 +w_i^{xy} \partial_x
+\phi \partial_y \phi \bigr), i=1\div8\\
+&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
+&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
 \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
-&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
+&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
 \end{aligned}
@@
 -\partial_{\alpha}P_{\alpha \beta} +
-\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
-+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
-&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
+\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
+&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
 \end{aligned}
@@
 phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
-relaxation time is taken to depend linearly on the relaxation
+relaxation time is taken as linearly dependant on the relaxation
 times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
@@
 influence the film width results. This is obviously achieved only for certain
-resolutions of the interface in comparison with the film width.
- \item Spurious currents arising from surface tension discretization
-can influence mass transfer in the thin film, thereby comoromising the
-overall solution. The bubble should therefore be long enough to decrease the
-influence of spurious currents.
+resolutions of the interface in comparison with the film width. The same applies to spurious
+currents arising from surface tension discretization which
+can influence mass transfer in the thin film, thereby compromising the
+overall solution. Therefore the diffusive interface should be of nengligible influence on the film
+thickness.
  \item The wettability coefficient which in principle can control the
@@
  \item The free-energy model examined in the work is a single density, two
-viscosities model. The classical Bretherton problem is a free-interface problem
+viscosities model. The classical Bretherton problem is a free-interface problem between liquid and
+air
 where the film thickness is established by itself. Thus, the viscosity and density ratios for the
+classical
 Bretherton problem should
-be large. However, in terms of the lattice Boltzmann free-energy framework those issues are not
-addressed and will be studied in the current work.
-  \item The Bretherton problem addresses bubble flow driven by a pressure
-difference, not a body force.  It is however relatively difficult to impose a
-pressure difference in the framework of the binary liquid model, as two
-simultaneous boundary conditions with the associated gradients and laplacians are required.
+be large. However, in terms of the lattice Boltzmann binary liquid framework only the viscosities
+ratio can be addressed. The results show that the results are consistent with other works and
+inertia effects can be neglected.
+  \item The Bretherton problem addresses a bubble flow driven by a pressure
+difference, not a body force. However, to the best authors knowledge the pressure boundaries
+conditions are still not yet developed for the binary liquid model, because it involves the
+treatment of associated gradients and laplacians on the boundary.
 This kind of pressure boundary conditions for the binary liquid model is currently
@@
 pressure difference flows and flows driven by a body force are small as far as the film
-deposition thickness is concerned.  In this work we limit ourselves to the study
-of body force driven flows, because of their simplicity and better numerical stability.
- \item Most simulations are done for a circular channel,
+deposition thickness is of interest.  In this work we limit ourselves to the study
+of body force driven flows, because of their simplicity and better numerical stability. The case of
+body force driven flows implies the periodic boundary conditions. As soon as the periodic boundary
+conditions are applied, not one bubble but whole bubble train is simulated. In this case one needs
+to insure that bubbles do not influence each other making large distance between them.
+ \item Most simulations are done for a channel with circular crosssection,
 which is quite difficult to address in terms of lattice Boltzmann free energy
-framework where one needs to calculate Laplacians near the boundaries.
+framework where one needs to calculate the Laplacian and gradients near boundaries.
 Although one can apply certain finite difference stencils
 \cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
-the analysis of error introduced by the boundary conditions becomes an issue.
-Our benchmark is therefore done is for a plane-symmetric case in the two-dimensional case. We do
-not address the axis-symmetric two-dimensional case since it implies the derivation of additional
-mass and force terms for the square based LBM grid \cite{halliday-circular}.
+the analysis of error introduced by the boundary conditions becomes an issue. Though the circular
+axis-symmetric case can be simulated on rectangular grid by introducing specific mass and force
+terms in the continuity and Navier-Stokes equation \cite{halliday-circular}, the lattice Boltzmann
+incorporation of force and mass terms in the process of development and validation. Therefore
+the benchmark is chosen as a plane-symmetric two-dimensional case.
+
 \end{itemize}
@@
 is a two-dimensional flow driven by a body force.  An analytical
-solution to this problem is to the best of our knowledge not yet known.
+solution to this problem is to the best of our knowledge not yet known for the range of capillary
+numbers.
 Though some authors \cite{sehgal-microchannel} use the gravity driven model to
@@
 bubble shape
-is different in the front and rear menisci. \citet{wong-films} indicated that the
-film width for polygonal capillaries
-changes from $\Ca^{1/2}$ on distances $x\propto \Ca^{-1}$ to $\Ca^{2/3}$ for
-distances $x \gg \Ca^{-1}$. Therefore, for a bubble to have a formed shape in
-the  front meniscus we chose the bubble length to be $5$ channel heights. The
+is different in the front and rear menisci depending on Bond number $Bo$. \citet{wong-films}
+analytically described the film thickness variation along a bubble varying from $\Ca^{1/2}$ to
+$\Ca^{2/3}$. The bubble should be long enough to have the film thickness proportional to
+$Ca^{2/3}$. Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
+to be sufficient to give consistent with theory results. The
 channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
@@
 because of the periodicity of boundary conditions.
-Figure \ref{fig:benchmark:sketch} represents a sketch of the geometry
+Figure \ref{fig:benchmark:sketch} is a sketch of the geometry
 used in the benchmark.
@@
 \subsection{The nondimensionalization and initialization procedure}
-We now intrduce the procedure of nondimensionalization, as well as
-the initialization techniques we used to be able to predict parameters necessary for
+We now introduce the procedure of nondimensionalization, as well as
+the initialization techniques we used to be able to determine parameters necessary for
 simulation. The capillary number governs the interface thickness and is defined as:
@@
 While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
-or defined through the model parameters, the velocity $U_{\mathrm{bubble}}$ is not.
-In the simulation we approximately match the velocity with the
+or defined through the model parameters, the bubble velocity $U_{\mathrm{bubble}}$ is not given
+explicitly.
+In the simulation we approximately match and initialize the velocity with the
 Poiseuille profile center velocity, which is given by the equation:
@@
 \begin{aligned}
-U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H^2}{8
+U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H_{\mathrm{eff}}^2}{8
 \mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
 U_{\mathrm{bubble}}\approx
-\frac{{N_y}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
+\frac{{(N_y-2)}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
 x } ,
@@
 \end{equation}
+where $N_y$ is the grid number in the vertical direction. As it will be discussed further, the
+effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes.
 By taking the Poiseuille profile one can estimate the body force through the
@@
 \begin{aligned}
-&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{N_y}^2}\gamma Ca.
+&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma Ca.
 \end{aligned}
 \end{equation}
-In practice, we iterate adjusting the body force value until both the desired velocity and the
-capillary numbe are achieved.
+The body force (\ref{poiseuille:velocity:center}) applied to the problem will cause the bubble
+velocity different from the implied by the capillary number because of the multiphase flow. That's
+why to achieve the 
+desired velocity and the
+capillary number the iteration procedure to find the appropriate body force needs to be done.
+However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
+starting approximation point.
 
@@
 gradients and Laplacians. The liquid phase is initialized with $\phi = 1$ and the
-gas phase is initialized with $\phi = -1$. Though the initialization film thickness
+gas phase is initialized with $\phi = -1$. Though the initial film thickness
 is not involved in the final result for the front meniscus (see Fig.
@@
 thickness values \cite{giavedoni-numerical}. This is done to minimize
-the simulation time, as starting far away from the steady-state requies
+the convergence simulation time, as starting far away from the steady-state requies
 additional time for the relaxation.
 \begin{figure}
-\includegraphics*[bb=-20 360 620 430,width=0.97\textwidth]
+\includegraphics*[bb=-15 360 630 430,width=0.97\textwidth]
 {Figures/Init/initbegin12.eps}
-\includegraphics*[bb=-20 360 620 430,
+\includegraphics*[bb=-15 360 630
+430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
+\includegraphics*[bb=-15 360 630 430,
 width=0.97\textwidth]{Figures/Init/initbegin16.eps}
-\includegraphics*[bb=-20 360 620 430,
+\includegraphics*[bb=-15 360 630
+430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
+\includegraphics*[bb=-15 360 630 430,
 width=0.97\textwidth]{Figures/Init/initbegin20.eps}
-\includegraphics*[bb=-20 360 620
-430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
-\includegraphics*[bb=-20 360 620
-430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
-\includegraphics*[bb=-20 360 620
+\includegraphics*[bb=-15 360 630
 430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
 \caption{The phase plots for different bubble widths as
-$H_{\mathrm{eff}}-12$,$H_{\mathrm{eff}}-16$ and
-$H_{\mathrm{eff}}-20$. The grid for the simulation is $102 \times 1501$. The phase
+$H_{\mathrm{eff}}-12$, $H_{\mathrm{eff}}-16$ and
+$H_{\mathrm{eff}}-20$. The grid for the simulation is $102 \times 1501$. The results are rescaled
+on $N_y$.The phase
 profiles were obtained after $2\cdot10^5$ time steps. One can see that even if the
-system is initialized with different bubble volumes, the bubbles always relaxate
+system is initialized with different bubble volumes, the bubbles always relax
 to a shape where the film thickness is conserved. For the given parameters the
-interface thickness is $0.0615$, $0.0611$, $0.0605$.
+interface thickness rescaled at the effective channel width $H_{\mathrm{eff}}$  is
+$0.0701$,$0.0697$,$0.0692$ measured at the dimensionless coordinate $x=13$.
 \label{fig:different:initialization:widths}}
@@
 
-Here, two different examples are presented that can quantitatively show the
+Here, two different examples are presented that quantitatively show the
 initialization procedure for quite different capillary numbers. In these
@@
 $\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which corresponds
-to a gas-liquid viscosity ratio of $10$.
+to a gas-liquid kinematic viscosity ratio of $10$.
 \begin{description}
@@
   \begin{equation}
-  \frac{h_{\infty}}{H}=1.3375 \Ca^{2/3}=0.03899,
+  \frac{h_{\infty}}{H_{\mathrm{eff}}}=0.6687 \Ca^{2/3}=0.0194.
   \end{equation}
-  where we assume that $2 H$ or $N_y$ is the channel width in lattice units. Thus, the film
+Thus, the film
 thickness occupies around $2$ percent of the whole channel height. Let us impose that $8$ lattice
-units are required to resolve  $2$ percent of the channel. Note that $5$ out of $8$ lattice units
+units are required to resolve  $2$ percent of the channel. Note that $5$ out of $8$ these lattice
+units
 ($62.5$ percent) are used to resolve the phase interface. Therefore the channel width is $400$
@@
 \end{equation}
-Even with this naive assumption for the $62.5$ percent resolution of the film
-width in comparison with the interface width, one can get the memory
+Even with this naive assumption for the $62.5$ percent resolution of the interface width of the film
+width, one can get the memory
 requirement of $345 \mathrm{MB}$.
@@
   U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{slug}}}=0.005
-\frac{0.0377}{0.66666}=2.82 \cdot10^{-4},
+\frac{0.0377}{0.6666}=2.82 \cdot10^{-4},
   \end{equation}
@@
   \end{equation}
-  Note that in the case of large grids  height resolution is directly
-involved in calculation of the physical time step :since:
+Note the velocity is directly
+involved in calculation of the physical time step. Small velocites imply longer running simulations:
 \begin{equation}
@@
 approximately $1$ m/s and physical channel height up to $1$ mm. In this case one
-can easily approximate the physical time step:
+can easily approximate the physical time step during one lattice Boltzmann iteration:
 \begin{equation}
@@
 \end{equation}
-which is quote small. A typical simulation needs approximately
-$100000-300000$ time steps to come to the steady state solution giving the
+which is small. A typical simulation runs approximately
+$100000-300000$ time steps due to computational resources. That gives 
 physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
-probably not enough to find the desired physics correlations taken place on macroscopic time level.
+probably not enough for physical system to come to the pseudo steady state.
 
  \item[II $\bm{Ca=0.05}$] 
-   The parameters in the previous subsection imply very long
-simulations. Bretherton delivered his solution for the range of Capillary number
+   The parameters in the previous subsection imply long
+simulations for system to come to equilibrium and this number increase with decrease of the
+capillary number. Bretherton obtained the analytical solution for the range of low capillary numbers
 less than
-  $0.005$. However, a lot of simulations devoted to the Bretherton/Taylor
-flow extended it to the range of Capillary numbers greater than
-$0.005$, and to different Reynolds numbers. One can refer to figures
+  $0.005$. However, a lot of simulations extended it to the range of Capillary numbers greater than
+$0.005$, and to higher Reynolds numbers. One can refer to figures
   \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
@@
 \caption{\citet{giavedoni-numerical} gathered results across the
-literature for different Capillary numbers \label{fig:giavedoni:planar}}
+literature for different Capillary numbers. Note that results are scaled on a half-width of the
+channel. \label{fig:giavedoni:planar}}
 \includegraphics[width=0.97\textwidth]{Figures/heil-planar.eps}
 \caption{\citet{heil-bretherton} performed simulations for different ranges of
-Capillary numbers and Reynolds numbers. \label{fig:heil:planar}}
+Capillary numbers and Reynolds numbers. Note that results are scaled on a half-width of the channel.
+\label{fig:heil:planar}}
 \end{figure}
@@
 capillary number and the body force. However, in the binary liquid framework
-those correlations are only approximate. The biggest assumption lies
+those correlations are only approximate. The stronggest assumption lies
 in calculating the body force from the Poiseuille velocity profile.
@@
 factor and its velocity is larger than the surrounding liquid velocity. By
-plugging in the obtained velocity back to the capillary number equation the
+substituting the obtained velocity back to the capillary number equation the
 larger capillary number can be obtained. While we present the initialization
@@
 To properly estimate the interface resolution one needs to study the convergence
-of the results of a function of the grid resolution. To do that, all parameters are held fixed,
-including the bubble velocity, the capillary number, an only the grid
-resolution is changed. Our goal is to know exactly the ratio of the interface width to the
-film thickness where results are not dependant on the grid resolution.
+of the results as a function of the grid resolution. To do that, all parameters are held fixed,
+including the bubble velocity, the capillary number, and only the grid
+resolution is changed. Our goal is to determine the ratio of the interface width to 
+film thickness where results are not dependent on the grid resolution.
 
-We illustrate the procedure in more detail for the capillary number
-$0.05$. As it was pointed out earlier, the film thickness in that case is $6$ percent.
+We illustrate the procedure for the capillary number
+$0.05$. As was pointed out earlier, the film thickness in that case is $6$ percent of the effective
+channel $H_{\mathrm{eff}}$.
 Note that in the case of the half-way bounce-back walls \cite{yu} which are used in the
@@
 where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
-is related to the effective channel height. Note that it is a big
+is related to the effective channel height. That can be cleared out by the following explanations.
+If the grid number in the vertical direction is $N_y$ then one has $N_y-1$ regions which represent
+domain. The effective wall location is in the middle between bounce-back and fluid nodes giving
+overall $N_y-2$ representing the channel width. Note that it is a big
 simplification to impose the boundary in the middle beetween the bounce-back
 node and the fluid node which give rise to the viscosity dependent location of
-the wall \cite{ginzburg-multireflection}. In the case of $\delta=0.06$ the
+the wall \cite{ginzburg-multireflection}, however the effective location of the wall for the
+multiphase models to the best authors' knowledge is not yet derived. In the case of $\delta=0.06$
+the
 initial grid size is $N_y=102$, which gives the horizontal grid size as $N_x-1=15(N_y-2)=1500$.
@@
 profile formula, Eq.~\ref{poiseuille:velocity:center}, and it equals
-$1.508 \cdot 10^{-6}$.
+$1.508 \cdot 10^{-6}$ lattice units.
 
@@
 \end{figure}
+%Here will be a consolidated table
+\begin{table}
+
+\end{table}
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{unsrtnat}
 \bibliography{paper}

### 2010-11-22 Sailfish/Init/init.py

@@ -22,7 +22,7 @@ def Get_Zero(prof):
     for counter in range(0, len(prof)/2):
         if prof[counter]>=0 and prof[counter+1]<0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
-    print (zero-1.5)/(len(prof)-4)
+    print (zero-0.5)/(len(prof)-2)
 
 def Analyze_Simulations():
     print os.getcwd()
@@ -33,14 +33,19 @@ def Analyze_Simulations():
         name="init180000.npz"
         array=numpy.load(name)
         fig=pylab.figure(figsize=(11,1.5))
-        pylab.imshow(array['phi'])
-        pylab.yticks([0,25,50,75,100])
+        pylab.imshow(array['phi'],cmap="gray",extent=[0.0,15.0,0.0,1.0])
+        pylab.yticks([0.0,0.5,1.0])
+        #pylab.axis([0.0,15.0,0.0,1.0])
         pylab.savefig("../../initfinish"+str(width[i])+".eps",format="EPS",dpi=70)
         
         fig_init=pylab.figure(figsize=(11,1.5))
         array_init=numpy.load("init000000.npz")
-        pylab.imshow(array_init['phi'])
-        pylab.yticks([0,25,50,75,100])
+        pylab.imshow(array_init['phi'],cmap="gray",extent=[0.0,15.0,0.0,1.0])
+        #pylab.axis([0.0,15.0,0.0,1.0])
+        #pylab.xlim(0.0,15.0)
+        #pylab.ylim(0.0,1.0)
+
+        pylab.yticks([0.0,0.5,1.0])
         pylab.savefig("../../initbegin"+str(width[i])+".eps",format="EPS",dpi=70)
         
         #pylab.plot(array[:, 520*i])

### 2010-11-24 Text/paper.tex
Revised all text. Need to wait Sasha's sketch.

@@
 \usepackage{color}
+\usepackage{tabularx}
 
@@
 the
-initial grid size is $N_y=102$, which gives the horizontal grid size as $N_x-1=15(N_y-2)=1500$.
+initial grid size is $N_y=102$, which gives the horizontal grid size as $N_x=15(N_y-2)=1500$.
 The bubble is initialized as a rectangular box with coordinates
@@
 After choosing the reference parameters, the grid refinement procedure
-needs to keep the macroscopic parameters constant.  Therefore the simulations
-were performed with grid numbers $N_y=102,127,152,177,202,227$, corresponding to the
-effective channels widths  $H_{\mathrm{eff}}=100,125,150,175,200,225$,
-$N_x=1501,1876,2251,2626,3001,3376=15
-H_{\mathrm{eff}}+1$. It is easy to check
-that keeping all liquid viscosities the same yields the following quantity to
+needs to keep the macroscopic parameters constant.  It is easy to check
+that it yields the following quantity to
 be independent of grid size
 $H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
-x } = \mathrm{const}$. We performed a number of simulations for the prescribed grids.
+x } = \mathrm{const}$. We performed a number of simulations for the prescribed grids. The
+simulations in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
+velocity $U_{bubble}$ are summarized in Table \ref{table:parameters:grid:refinement}.
 The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
-see the visual results converge for $H_{\mathrm{eff}}\geq 175$. The calculated widths
-are $0.0606$,$0.0618$,$0.0617$,$0.0629$,$0.0624$,$0.0624$. We accept the
-convergent results for grids larger than $H_{\mathrm{eff}}=175$. However, one can see
+see the visual results converge for $H_{\mathrm{eff}}\geq 175$. However, one can see
 that with proper initialization techniques, large enough time and the wall
@@
 interface itself occupies approximately $5 \xi$, where
-$\xi=\sqrt{k/A}=1$. Let us calculate the interface width ratio to the film
-thickness as
-$5\xi/H_{\mathrm{film}}$: $0.824$, $0.646$,  $0.539$, $0.453$,
-$0.400$, $0.355$. Therefore, one needs to
-resolve the interface as $40-50$ percent of the
+$\xi=\sqrt{k/A}=1$. The ratio of the interface width to the film thickness $5\xi/H_{\mathrm{film}}$
+ is shown in Table \ref{table:parameters:grid:refinement}. Based on results one can conclude that
+ the interface needs to be resolved as $40-50$ percent of the
 expected film width for simulations to be convergent. We further examine the
-velocities in the center of the bubble to calculate the Capillary number. The
-calculated numbers are $0.0041$, $0.0041$,$0.0040$,$0.0040$,$0.0041$,$0.0039$.
-This corresponds to the average capillary number of $0.07$, which is slightly
+velocities in the center of the bubble to calculate the Capillary number. One can see from Table
+\ref{table:parameters:grid:refinement} that the bubble velocities are pretty consistent and the
+calculated capillary number corresponding to those values is $0.07$, which is slightly
 larger than the expected capillary number. The corresponding error can be
 attributed to the density and viscosity ratios which are much larger in the
-Bretherton problem.
+Bretherton problem and Poiseuille profile body force initialization.
 \begin{figure}
 \includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
-\caption{Grid-refined profiles for the effictive
+\caption{Grid-refined profiles for the effective
 channel widths
-$H_{\mathrm{eff}}=100,125,150,175,200,225$.One can see
+$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_{eff}$ and $\delta=0$
+corresponds to the wall location. The profiles were taken at $x=14$ nondimensional coordinate.
+Capillary number is $Ca=0.05$. Capillary number obtained from simulations $Ca=0.07$. One can
+see
 convergent results even for the underresolved film
@@
 \begin{table}
-
+\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|}
+\hline
+$N_x$&$N_y$&$\delta$&$U_{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_{iter}$\\
+\hline
+$1500$&$102$&$0.0694$&$0.0041$&$0.824$&$200000$\\
+\hline
+$1875$&$127$&$0.0688$&$0.0041$&$0.646$&$250000$\\
+\hline
+$2250$&$152$&$0.0676$&$0.0040$&$0.539$&$300000$\\
+\hline
+$2625$&$177$&$0.0679$&$0.0040$&$0.453$&$350000$\\
+\hline
+$3000$&$202$&$0.0668$&$0.0041$&$0.400$&$400000$\\
+\hline
+$3375$&$227$&$0.0663$&$0.0039$&$0.355$&$450000$\\
+\hline
+\end{tabularx}
+\caption{The parameters and results for studying grid resolution simulations. Domain simulated is
+of size $N_x\, \mathrm{x}\, N_y$. $U_{bubble}$ is center
+bubble velocity measured at the scaled to $H_eff$ $x=14.0$.
+$5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of the film in
+lattice Boltzmann units. $Ca=0.05$ was taken to initialize the simulation.
+\label{table:parameters:grid:refinement}}
 \end{table}
@@
 We examined $11$ different values for the wall
-gradient ranging from $-1$ to $1$. The results are consistent among all the
-wall gradients given that the interface is resolved in the right way. For the
-given wall gradients the values of interface thickness values are
-$0.0633$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0633$,$0.0632$,
-$0.0631$. The results are different in 4th digit. For the wall gradients $0.8$
+gradient ranging from $-1$ to $1$. The parameters are summarized in Table
+\ref{table:parameters:wall:gradient}. The results are consistent among all the
+wall gradients given that the interface is resolved in the right way. For 
+given wall gradients the values of interface thickness values are of $0.2\%$ relative accuracy.  For
+the wall gradients $0.8$
 and $1.0$ the
 simulations are unstable. One can see the simulated profiles in Fig.
-\ref{fig:gradients:profiles}. Negative values of the phase gradient are
+\ref{fig:gradients:profiles}. The latter shows that the negative wall gradient values cause 
+unphysical phase values more than $1.0$ near the wall. One should attribute them to the numerical
+imposement of wall gradients. Therefore, they do not spoil numerical scheme as whole. Note that
+negative values of the phase gradient are
 preferable, since the phase of the liquid adjacent to the wall has the value of $1$.
@@
 gradients the values near the wall are below $1$, and the gradient profile can
-fuse with the values of the gas phase as, giving rise to an unphysical wall stickyness.
+fuse with the values of the gas phase. This case is attributed when to the slug flow when gas
+touches the wall.
 \begin{figure}
@@
 \caption{Phase profiles for different wall gradients. One can see that if the width
-is properly resolved then the wettability of the wall does affect the film thickness.
+is properly resolved then the wettability of the wall does affect the film thickness. The
+simulations were conducted for $200,000$ iterations. The measured profile was taken at the
+nondimensional coordinate scaled to $H_{\mathrm{eff}}$ as $x=11.42$. $Ca=0.05$ was taken to
+initialize simulation.
 \label{fig:gradients:profiles}
@@
 \end{figure}
+
 Another question which arises is the influence of the wall gradient on the bubble velocity and the
 corresponding capillary number. The shear stress controlled by the wall phase gradient changes the
-effective velocity near the wall, thus changing the bubble velocity as well.
+effective viscosity near the wall, thus changing the bubble velocity as well. If viscosity ratio is
+high enough then the free surface motion of the bubble should not be dependent on the shear stress
+near the wall.
 One can see that if the film thickness is resolved properly the influence of the wall gradient can
-be neglected. Fig. \ref{fig:velocities:wall:gradient} shows the center bubble velocity for different
-gradients. The average relative error around the mean velocity is $0.1\%$.
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/Wall/velocities_grad_profiles.eps}
-\caption{The center bubble velocity versus the phase gradient. One can neglect the influence of
-the wall once the film is properly resolved.
-\label{fig:velocities:wall:gradient}
-}
-\end{figure}
+be neglected. The center bubble velocities are presented in Table
+\ref{table:parameters:wall:gradient}. 
+%Fig. \ref{fig:velocities:wall:gradient} shows the center bubble velocity for different
+%gradients. 
+The average relative error around the mean velocity is $0.1\%$.
+% Uncomment if you need this picture
+% \begin{figure}
+% \includegraphics[width=0.97\textwidth]{Figures/Wall/velocities_grad_profiles.eps}
+% \caption{The center bubble velocity versus the phase gradient. One can neglect the influence of
+% the wall once the film is properly resolved.
+% \label{fig:velocities:wall:gradient}
+% }
+% \end{figure}
+\begin{table}
+\begin{tabularx}{\textwidth}{|X|X X X X X X X X X X X|}
+\hline
+$\scriptstyle \partial_n \phi$& $\scriptstyle -1.0$& $\scriptstyle -0.8$&
+$\scriptstyle -0.6$&$\scriptstyle -0.4$&$\scriptstyle -0.2$&$\scriptstyle
+0.0$&$\scriptstyle 0.2$&$\scriptstyle 0.4$&$\scriptstyle 0.6$&$\scriptstyle 0.8$&$\scriptstyle
+1.0$\\
+\hline
+$\scriptstyle \delta$& $\scriptstyle 0.0633$& $\scriptstyle 0.0634$& $\scriptstyle 0.0634$&
+$\scriptstyle 0.0634$& $\scriptstyle 0.0634$& $\scriptstyle 0.0634$& $\scriptstyle 0.0633$&
+$\scriptstyle 0.0632$& $\scriptstyle 0.0631$ &$\scriptstyle \mathrm{N/A}$&$\scriptstyle
+\mathrm{N/A}$\\
+\hline
+$\scriptstyle U_{b}$ &$\scriptstyle 0.0041$& $\scriptstyle 0.0042$& $\scriptstyle 0.0042$
+&$\scriptstyle 0.0042$ & $\scriptstyle 0.0041$& $\scriptstyle 0.0041$ & $\scriptstyle 0.0042$ &
+$\scriptstyle 0.0042$ & $\scriptstyle 0.0042$ & $\scriptstyle \mathrm{N/A}$ &$\scriptstyle
+\mathrm{N/A}$\\
+\hline
+\end{tabularx}
+\caption{The parameters and results for studying wall gradient simulations. $U_{b}$ stands for the
+center bubble velocity. $\delta$ is the interface thickness. $\partial_n \phi$ is the phase
+gradient near the wall responsible for hydrophilic and hydrophobic behavior. The results are
+calculated at nondimensional $x=11.42$ after $200000$ time iterations. $Ca=0.05$ was taken to
+initialize simulation.
+\label{table:parameters:wall:gradient}}
+\end{table}
 
@@
 small capillary numbers and compare the film thickness over the capillary number
-range from $0.03$ to $0.8$, which is a computationally reasonable task.  For the
-benchmark
-results we
-suggest the following Capillary numbers:
-\begin{equation}
-0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8\, .
-\end{equation}
-The corresponding approximate film widths from \cite{giavedoni-numerical} are as
-follows:
-\begin{equation}
-0.04,0.06,0.08,0.1,0.12,0.13,0.15\,.
-\end{equation}
+range from $0.03$ to $0.8$, which is a computationally reasonable task.  
 To perform numerical simulations we refer
@@
 $6 \cdot 10^{-6}/16$ was chosen to obtain the predicted capillary
-number $0.005$. The forcing for other grids can be obtained using
+number $0.05$. The forcing for other grids can be obtained using
 simple correlations:
@@
 \end{equation}
-where the index ,,lit'' stands for the predicted capillary number
+where the subscript ,,lit'' stands for the predicted capillary number
 \cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
@@
 \end{equation*}
-The results were obtained after $2\cdot10^5$ steps. The following film thicknesses were
-calculated: $0.042$,$0.066$, $0.085$, $0.077$, $0.123$, $0.153$, $0.164$, $0.173$ with
-the corresponding bubble center velocities $0.0022$, $0.0041$, $0.0068$, $0.0055$,
-$0.0185$, $0.0388$, $0.0592$, $0.0782$. Based on the obtained center velocities,
-the capillary numbers are calculated as
-$0.039$,$0.072$,$0.120$,$0.098$,$0.328$,$0.686$,$1.046$,$1.383$. These
+The results were obtained after $2\cdot10^5$ steps. Simulation results are presented in Table
+\ref{table:parameters:capillary:number}. One can see that the calculated
 capillary numbers are overpredicted but the obtained widths are overpredicted
-as well due to the improper body force imposement, because the assumption of Poiseuille flow is not
-valid. If one needs to obtain the desired capillary number, the shooting method for the forcing is
+as well due to the improper body force imposement, because of the Poiseuille flow assumption. If one
+needs to obtain the desired capillary number, the shooting method for the forcing is
 necessary starting as the Poiseuille pressure gradient force as the initial condition. We extracted
-data from
-\cite{giavedoni-numerical,heil-bretherton} and compared it the results of our
-simulations, which are shown in Fig. \ref{fig:capillary:comparison}.
+results of 
+\citet{giavedoni-numerical} and \citet{heil-bretherton} and compared them with our results, shown
+in Fig. \ref{fig:capillary:comparison}.
 \begin{figure}
 \includegraphics[width=0.97\textwidth]{Figures/Capillary/capillaries_comparison.eps}
-\caption{A comparison between simulation results and results from
-\cite{giavedoni-numerical} and \cite{heil-bretherton}. One can see a
+\caption{A comparison between simulation results and results of
+\citet{giavedoni-numerical} and \citet{heil-bretherton}. One can see a
 reasonable agreement. The plots depict the film thickness as a function of the
@@
 \end{figure}
+
+\begin{table}
+\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
+\hline
+$Ca_{\mathrm{lit}}$&$\delta_{\mathrm{lit}}$&$x$&$\delta$&$U_b$&$Ca$\\
+\hline
+$0.03$&$0.04$&$10.0$&$0.042$&$0.0022$&$0.039$\\
+\hline
+$0.05$&$0.06$&$11.0$&$0.066$&$0.0041$&$0.072$\\
+\hline
+$0.08$&$0.08$&$13.0$&$0.085$&$0.0068$&$0.120$\\
+\hline
+$0.1$&$0.1$&$12.5$&$0.077$&$0.0055$&$0.098$\\
+\hline
+$0.2$&$0.12$&$5.5$&$0.123$&$0.0185$&$0.328$\\
+\hline
+$0.4$&$0.13$&$4.7$&$0.153$&$0.0388$&$0.686$\\
+\hline
+$0.6$&$0.15$&$3.0$&$0.164$&$0.0592$&$1.046$\\
+\hline
+$0.8$&$0.16$&$1.7$&$0.173$&$0.0782$&$1.383$\\
+\hline
+\end{tabularx}
+\caption{The parameters and results for studying capillary number region simulations.
+$\delta_{\mathrm{lit}}$ is the film thickness with corresponding $Ca_{\mathrm{lit}}$ taken from
+literature. $\delta_{sim}$ is the simulation film thickness with corresponding $Ca$. $U_b$ is the
+center bubble velocity taken at location $x$. $Ca$ is based on the measured center bubble velocity
+$U_{b}$.
+\label{table:parameters:capillary:number}}
+\end{table}
 
@@
 \begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_length.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_sehgal.eps}\\
+\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_sehgal.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_length.eps}\\
 \caption{The qualitative comparison for the film thickness across
-the bubble length. Right (courtesy of \citet{sehgal-microchannel}),
-left (present simulations).\label{fig:sehgal:bubble:length}}
+the bubble length. Left (courtesy of \citet{sehgal-microchannel}),
+right (present simulations). $x$ is scaled to $H_{\mathrm{eff}}$
+and is taken in the flow direction.\label{fig:sehgal:bubble:length}}
 \end{figure}
 One can see a qualitative agreement, i.e. the thickness increases towards the front meniscus and
-rapidly decreases towards the rear meniscus. This shape sometimes is called as a "bullet" shape.
+rapidly decreases towards the rear meniscus. This shape sometimes is referred as a "bullet" shape.
 
@@
 This work presents numerical tips and a benchmark for the
-Bretherton/Taylor problem using the binary liquid lattice Botlzmann method. The
+Bretherton/Taylor problem using the binary liquid lattice Botlzmann method. To obtain correlations
+we specifically designed the benchmark. The bubble was chosen long enough for the film thickness to
+stabilize. Also, we used periodic boundary conditions for scheme robustness. Thus, the bubble
+train was simulated instead of one bubble motion. To avoid the influence of other bubbles
+distance between bubbles were chosen as $3$ bubble sizes. The
 results show consistency with the previously published numerical results obtained
-using other methods. Though our results are specific to the binary liquid lattice
+using other methods in terms of capillary number region, shape of the bubble. It comes with surprise
+but with large enough viscosities ratio the results are independent on inertia. This work supports
+this assumption as well. Also, we examined the influence of grid resolution on the results. We found
+that if the phase interface is resolved as $50$ percent of the simulated film thickness, the results
+are convergent. Though  our results are specific to the binary liquid lattice
 Botlzmann method, the numerical hints and procedures can be used for any
@@
 people simulating microchannel flows.
-
 \bibliographystyle{unsrtnat}

### 2010-11-24 Sailfish/Capillary/capillary.py

@@ -37,11 +37,11 @@ def Get_Zero(prof):
 def Analyze_Simulations():
     from numpy import genfromtxt
     print os.getcwd()
-    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0]
-    capillary_str=["3","5","8","10","20","40","60","80","100"]
-    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16,0.17]
+    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8]
+    capillary_str=["3","5","8","10","20","40","60","80"]
+    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16]
 
-    exam=[2000,2200,2600,2500,1100,950,600,350,350]
+    exam=[2000,2200,2600,2500,1100,950,600,350]
     good=[0,1,2,3,4,5,6,7]
     #style=["bo","rH","c<","y>","bs","g^"]
     #color=["b","r","c","y","b","g"]
@@ -91,10 +91,11 @@ def Analyze_Simulations():
     print "Capillaries=",capillaries
     print "Velocities",velocities
     
-    pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
     pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
     pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
-    pylab.xlim(0.02,1.1)
+    pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
+
+    pylab.xlim(0.02,1.5)
     pylab.ylim(ymin=0.01)
     numpy.savetxt("capillary.dat",zip(capillaries,widths))
     
@@ -105,7 +106,7 @@ def Analyze_Simulations():
     pylab.ylabel(r'''$\delta$''',fontsize=30)
     
     #labels=[r'''$H_{eff}='''+str(value-2)+r'''$''' for value in ny]
-    pylab.legend(["Simulations","Giavedoni","Heil"],loc=4)
+    pylab.legend(["Giavedoni","Heil","Simulations"],loc=4)
     #pylab.xlim(xmax=15)
     pylab.savefig("capillaries_comparison.eps",format="EPS",dpi=300)
 
@@ -164,14 +165,21 @@ def Analyze_Bubble():
         #print extrapolator(0)
         
[…]

### 2010-11-24 Sailfish/Capillary_Viscous/capillary.py

@@ -97,7 +97,7 @@ def Analyze_Simulations():
     pylab.loglog(capillary_not_viscous[:,0],capillary_not_viscous[:,1],"bD-",linewidth=3,markersize=10)
     #pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
     #pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
-    pylab.xlim(0.02,1.1)
+    pylab.xlim(0.02,2.5)
     pylab.ylim(ymin=0.01)
     numpy.savetxt("capillary.dat",zip(capillaries,widths))
     

### 2010-11-24 Sailfish/Grid/grid.py

@@ -30,7 +30,7 @@ def Get_Zero(prof):
     for counter in range(0, len(prof)/2):
         if prof[counter]>=0 and prof[counter+1]<0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
-    print (zero-1.5)/(len(prof)-4)
+    print (zero-0.5)/(len(prof)-2)
 
 def Analyze_Simulations():
     print os.getcwd()
@@ -52,10 +52,13 @@ def Analyze_Simulations():
         name="grid"+str(200000+i*50000)+".npz"
         array=numpy.load(name)
         prof=array['phi'][:,int(1400*ratio)]
+        prof_domain=prof[1:len(prof)-1]
+        delta_x=1.0/float(ny[i]-2)
+        x_domain=numpy.arange(0.5*delta_x,1.0,delta_x)
         x=numpy.arange(0.0,float(ny[i]))/ratio
         #pylab.imshow(array['phi'])
         #pylab.plot(x[0:20],prof[0:20],color[i],linewidth=3)
-        pylab.plot(x,prof,style_diff[i],markersize=10,linewidth=3)
+        pylab.plot(x_domain,prof_domain,style_diff[i],markersize=10,linewidth=3)
         #pylab.figure()
         #pylab.plot(array[:, 520*i])
         #pylab.savefig("grid_phase_prof_"+str(49*i)+".eps", dpi=300)
@@ -67,11 +70,11 @@ def Analyze_Simulations():
     fig.subplots_adjust(left=0.15,bottom=0.15)  
     pylab.xticks(fontsize=20)
     pylab.yticks(fontsize=20)
-    pylab.xlabel(r'''$x$''',fontsize=30)
+    pylab.xlabel(r'''$\delta$''',fontsize=30)
     pylab.ylabel(r'''$\phi$''',fontsize=30)
-    labels=[r'''$H_{eff}='''+str(value-2)+r'''$''' for value in ny]
+    labels=[r'''$H_{\mathrm{eff}}='''+str(value-2)+r'''$''' for value in ny]
     pylab.legend(labels)
-    pylab.xlim(xmax=15)
+    pylab.xlim(xmax=0.15)
     pylab.savefig("norm_grid_profs.eps",format="EPS",dpi=300)
 
 
@@ -97,7 +100,7 @@ def Analyze_Velocities():
 
 
 if __name__=="__main__":
[…]

### 2010-11-24 Sailfish/Wall_Gradient/gradient.py

@@ -21,7 +21,7 @@ def Get_Zero(prof):
     for counter in range(0, len(prof)/2):
         if prof[counter]>=0 and prof[counter+1]<0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
-    print (zero-1.5)/(len(prof)-4)
+    print (zero-0.5)/(len(prof)-2.0)
 
 def Analyze_Simulations():
     print os.getcwd()
@@ -36,7 +36,11 @@ def Analyze_Simulations():
         array=numpy.load(name)
         #fig=pylab.figure(figsize=(11,1.5))
         #pylab.imshow(array['phi'])
-        pylab.plot(array['phi'][0:22,2000],style[i],linewidth=3,markersize=10)
+        prof=array['phi'][:,2000]
+        prof_domain=prof[1:len(prof)-1]
+        x_delta=1.0/len(prof_domain)
+        x=numpy.arange(0.5*x_delta,1.0,x_delta)
+        pylab.plot(x,prof_domain,style[i],linewidth=3,markersize=10)
         #pylab.yticks([0,25,50,75,100])
         #pylab.savefig("../../initfinish"+str(width[i])+".eps",format="EPS",dpi=70)
         
@@ -53,8 +57,9 @@ def Analyze_Simulations():
         #print extrapolator(0)
         
         os.chdir("../..")
-    fig.subplots_adjust(left=0.15)
-    pylab.xlabel(r'''$x$''',fontsize=30)
+    fig.subplots_adjust(left=0.15,bottom=0.15)
+    pylab.xlim(xmax=0.141)
+    pylab.xlabel(r'''$\delta$''',fontsize=30)
     pylab.ylabel(r'''$\phi$''',fontsize=30)
     pylab.legend([r'''$\partial_n\phi=$'''+r'''$'''+str(x)+r'''$''' for x in gradient[:-2]])
     pylab.xticks(fontsize=20)
@@ -103,6 +108,7 @@ def Analyze_Velocities():
     pylab.xticks(fontsize=20)
     pylab.yticks(fontsize=20)
     #pylab.savefig("velocities_grad_profiles.eps",format="EPS",dpi=300)
+    print velocities
     print numpy.std(velocities)/numpy.mean(velocities)
 
 

### 2011-01-04 Text/paper.tex
Finalized my revisions. Sasha sent the figure.
Started doing modifications, but found that didn't merge previous ones.

@@
 This work shows that the free-energy binary liquid lattice Boltzmann method is able to
-reproduce the gravity driven microchannel flow of an air bubble through liquid. In the low capillary
-number
-$\Ca$ regime (defined in Eq. \ref{capillary:number:definition}), the flow is thoroughly described by
+reproduce the gravity driven microchannel flow of a gas bubble through liquid. In the regime of low
+capillary
+numbers
+$\Ca$ (defined in Eq. \ref{capillary:number:definition}), the flow is thoroughly described by
 Bretherton \cite{bretherton},
 with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
-from the Bretherton correlations for $Ca\geq 0.005$, studied experimentally and numerically. This
+from the Bretherton correlation for $Ca\geq 0.005$, studied experimentally and numerically. This
 work aims to examine the lattice Boltzmann binary liquid model feasibility to obtain valid results
-for an air bubble flow in liquid for a range of $Ca$ number. The binary
-liquid model is a continuous-interface multiphase model, which raises
-the question of how to properly resolve the interface in comparison with the film thickness.
+for a bubble flow in liquid for a range of $Ca$ number. The binary
+liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
 It was found that the film interface needs to be as $50$
@@
 The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
-microchannels. It was found that the front meniscus film width is proportional
+microchannels. It was found that the front meniscus film thickness is proportional
 to $\Ca^{2/3}$ for capillary numbers less than $0.003$, where the capillary number is defined as:
@@
 where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is a propagation bubble velocity, and
-$\gamma$ is the surface tension between air and liquid.
+$\gamma$ is the surface tension between gas and liquid.
 While there are number
@@
 However, if the interface is spread over several grid nodes, the question of
-proper film width resolution in comparison with the interface arises.  This work
+proper film thickness resolution in comparison with the interface arises.  This work
 concentrates on resolving the parameter range of the binary liquid lattice
@@
 model simulates two liquids with the assumption of uniform overall
-density. While the classical Bretherton problem is stated for air and liquid, which are of largely
+density. While the classical Bretherton problem is stated for gas and liquid, which are of largely
 different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
@@
 Bretherton \cite{bretherton} studied a long bubble
-moving in a tube filled with liquid. It was found that the deposition film width
+moving in a tube filled with liquid. It was found that the deposition film thickness
 is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
@@
 is proportional to $\Ca^{2/3}$ only in the specific region behind the front meniscus and
-that the film width
+that the film thickness
 varies over the bubble length for bubbles of finite length. Numerical
@@
 slug crossections and menisci appearance.
-\citet{heil-bretherton,ingham-plates} studied air finger propagation in
+\citet{heil-bretherton,ingham-plates} studied gas finger propagation in
 a two-dimensional channel for a range of Reynolds and capillary
@@
  \item The diffusive nature of the interface in the multiphase model should not
-influence the film width results. This is obviously achieved only for certain
-resolutions of the interface in comparison with the film width. The same applies to spurious
+influence the film thickness results. This is obviously achieved only for certain
+resolutions of the interface in comparison with the film thickness. The same applies to spurious
 currents arising from surface tension discretization which
@@
 viscosities model. The classical Bretherton problem is a free-interface problem between liquid and
-air
+gas
 where the film thickness is established by itself. Thus, the viscosity and density ratios for the
@@
 \end{equation}
-Even with this naive assumption for the $62.5$ percent resolution of the interface width of the film
-width, one can get the memory
+Even with this naive assumption for the $62.5$ percent resolution of the interface thickness of the
+film thickness, one can get the memory
 requirement of $345 \mathrm{MB}$.
@@
   can be easily and quickly validated on a computer.
-  One can see from Fig. \ref{fig:heil:planar} that the width for the
+  One can see from Fig. \ref{fig:heil:planar} that the film thickness for the
   Capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
@@
 
-  The film width for such small Capillary numbers is taken from
+  The film thickness for such small Capillary numbers is taken from
   \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
@@
 including the bubble velocity, the capillary number, and only the grid
-resolution is changed. Our goal is to determine the ratio of the interface width to 
+resolution is changed. Our goal is to determine the ratio of the interface thickness to 
 film thickness where results are not dependent on the grid resolution.
@@
 wettability, results are different only in the 3rd digit even for
-underresolved film widths. To calculate how well the interface thickness is
-resolved, the ratio of the interface thickness to the film width thickness is calculated. The
+underresolved film thicknesses. To calculate how well the interface thickness is
+resolved, the ratio of the interface thickness to the film thickness is calculated. The
 interface itself occupies approximately $5 \xi$, where
-$\xi=\sqrt{k/A}=1$. The ratio of the interface width to the film thickness $5\xi/H_{\mathrm{film}}$
+$\xi=\sqrt{k/A}=1$. The ratio of the interface thickness to the film thickness
+$5\xi/H_{\mathrm{film}}$
  is shown in Table \ref{table:parameters:grid:refinement}. Based on results one can conclude that
  the interface needs to be resolved as $40-50$ percent of the
-expected film width for simulations to be convergent. We further examine the
+expected film thickness for simulations to be convergent. We further examine the
 velocities in the center of the bubble to calculate the Capillary number. One can see from Table
@@
 enough grid resolution in order for simulations to be consistent -- the grid size was
-$177 \times 2626$ and the initial film width was
+$177 \times 2626$ and the initial film thickness was
 $12$ lattice boltzmann units, which corresponds to the predicted film thickness.
@@
 \includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
-\caption{Phase profiles for different wall gradients. One can see that if the width
+\caption{Phase profiles for different wall gradients. One can see that if the thickness
 is properly resolved then the wettability of the wall does affect the film thickness. The
@@
 To perform numerical simulations we refer
-to the grid refinement section where we established that the interface width should be
-at most $50$ percent of the interface width.  We therefore choose the grid to be
+to the grid refinement section where we established that the interface thickness should be
+at most $50$ percent of the interface thickness.  We therefore choose the grid to be
 $202 \times 3001$. $5$ lattice Boltzmann units do not occupy more than $60$
-percent of the effective width. Because the simulation get unstable with
+percent of the effective thickness. Because the simulation get unstable with
 smaller grids and larger gradients, all the capillary number simulations were
@@
 \end{equation}
-The film width is initialized through the capillary numbers ratio as well:
+The film thickness is initialized through the capillary numbers ratio as well:
 \begin{equation*}
@@
 \ref{table:parameters:capillary:number}. One can see that the calculated
-capillary numbers are overpredicted but the obtained widths are overpredicted
+capillary numbers are overpredicted but the obtained thicknesses are overpredicted
 as well due to the improper body force imposement, because of the Poiseuille flow assumption. If one
@@
 \includegraphics[width=0.97\textwidth]{Figures/Capillary_Viscous/capillaries_viscous.eps}
-\caption{The film widths versus capillary number for
+\caption{The film thicknesses versus capillary number for
 $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=20$. The

### 2011-01-04 Text/paper.tex
Prepared according to the Chemical Engineering Journal. Michal, please indicate your address and do the last glimpse.

@@
-\documentclass{article}
+%% This is file `elsarticle-template-1-num.tex',
+%%
+%% Copyright 2009 Elsevier Ltd
+%%
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%%
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%%
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%%
+%% Template article for Elsevier's document class `elsarticle'
+%% with numbered style bibliographic references
+%%
+%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
+%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
+%%
+\documentclass[preprint,12pt]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn]{elsarticle}
+%% \documentclass[final,5p,times]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+
+%% if you use PostScript figures in your article
+%% use the graphics package for simple commands
+%% \usepackage{graphics}
+%% or use the graphicx package for more complicated commands
+\usepackage{graphicx}
+%% or use the epsfig package if you prefer to use the old commands
+%% \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
 \usepackage{amsmath,amssymb}
-\usepackage{graphicx}
 \usepackage{bm}
-\usepackage[numbers,sort&compress]{natbib}
+%\usepackage[numbers,sort&compress]{natbib}
 \usepackage{color}
@@
 
-\title{The binary liquid model for gravity-driven microchannel simulations}
-\date{\today}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers after \end{frontmatter}.
+\usepackage{lineno}
+
+%% natbib.sty is loaded by default. However, natbib options can be
+%% provided with \biboptions{...} command. Following options are
+%% valid:
+
+%%   round  -  round parentheses are used (default)
+%%   square -  square brackets are used   [option]
+%%   curly  -  curly braces are used      {option}
+%%   angle  -  angle brackets are used    <option>
+%%   semicolon  -  multiple citations separated by semi-colon
+%%   colon  - same as semicolon, an earlier confusion
+%%   comma  -  separated by comma
+%%   numbers-  selects numerical citations
+%%   super  -  numerical citations as superscripts
+%%   sort   -  sorts multiple citations according to order in ref. list
+%%   sort&compress   -  like sort, but also compresses numerical citations
+%%   compress - compresses without sorting
+%%
+%% \biboptions{comma,round}
+
+% \biboptions{}
+
+\journal{Chemical Engineering Journal}
+
 \newcommand{\Ca}{\mathrm{Ca}}
 \newcommand{\todo}[1]{{\color{red}#1}}
+
 \begin{document}
 
-\maketitle
+\begin{frontmatter}
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+\title{The binary liquid model for gravity-driven microchannel simulations}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+\author[uofa]{A.~Kuzmin\corref{cor1}}
+\cortext[cor1]{Corresponding author}
+\ead{shurik.kuzmin@gmail.com}
+\author[google]{M.~Januszewski}
+\ead{michalj@gmail.com}
+\author[uofa]{J.~Derksen}
+\ead{kuzmin@ualberta.ca}
+\author[schlum]{D.~Eskin}
+\ead{deskin@slb.com}
+\author[schlum]{F.~Mostowfi}
+\ead{fmostowfi@slb.com}
+\address[google]{Google, Zurich, Switzerland}
+\address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
+116 St, Edmonton, Alberta, T6G
+2V4 Canada}
+\address[schlum]{DBR Schlumberger Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
+Canada}
 
@@
 liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
-It was found that the film interface needs to be as $50$
-percent in comparison with the film thickness for simulation to give convergent results. While
-the convergency criteria is fulfilled the results are of good comparison with the
+It was found that the interface thickness needs to be $50$
+percent of the film thickness for simulation to give convergent results. While
+the convergence criteria is fulfilled the results are in a good agreement with the
 numerical simulations of \citet{giavedoni-numerical} in the range of $Ca$ from $0.01$ to $1$. 
 \end{abstract}
+
+\begin{keyword}
+Bretherton problem \sep Microchannel simulation \sep Multiphase flow \sep Lattice Boltzmann method
+\sep Binary liquid model \sep Flow between plates \sep Gravity driven
+%% keywords here, in the form: keyword \sep keyword
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+\end{keyword}
+
+\end{frontmatter}
+
+%%
+%% Start line numbering here if you want
+%%
+\linenumbers
+
+%% main text
+
+%% The Appendices part is started with the command \appendix;
+%% appendix sections are then done as normal sections
+%% \appendix
+
+%% \section{}
+%% \label{}
+
+%% References
+%%
+%% Following citation commands can be used in the body text:
+%% Usage of \cite is as follows:
+%%   \cite{key}          ==>>  [#]
+%%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
+%%   \citet{key}         ==>>  Author [#]
 
@@
 microchannels. It was found that the front meniscus film thickness is proportional
-to $\Ca^{2/3}$ for capillary numbers less than $0.003$, where the capillary number is defined as:
+to $\Ca^{2/3}$ for capillary numbers smaller than $0.003$, where the capillary number is defined as:
 \begin{equation}
@@
 \end{equation}
-where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is a propagation bubble velocity, and
+where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is the bubble velocity, and
 $\gamma$ is the surface tension between gas and liquid.
+
 While there are number
-of works which simulate the Bretherton problem with the boundary value method
+of works which simulate the Bretherton problem by the boundary value approach
 \cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
-involving complex geometries, free interface motion or for problems involving coalescence and/or
+involving complex geometries, free interface motion, or for problems involving coalescence and/or
 droplet breakup. The continuous interface models are more flexible for those kind of simulations.
 However, if the interface is spread over several grid nodes, the question of
-proper film thickness resolution in comparison with the interface arises.  This work
-concentrates on resolving the parameter range of the binary liquid lattice
+proper film resolution in comparison with the interface resolution arises.  This work is focused on
+resolving the parameter range of the binary liquid lattice
 Boltzmann method to validate it in the range of capillary numbers.
@@
 \cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
-method easily tackles complex geometries andallows for incorporation of
-physical phenomena on the microscopic level, as in the case in multiphase models. Most
+method easily tackles complex geometries and allows incorporation of
+physical phenomena on the microscopic level, as in the case of multiphase models. Most
 multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
-the interface in a continuous way. Such a representation
-brings issues with film thickness resolution versus interface
-resolution. In other words, when it comes to the resolution of the film
-thickness the diffusive interface should be resolved in such a way as to have
+the interface using continuous interface methods where the interfaces spans over several grid nodes.
+Such a representation
+brings issues of the film thickness resolution versus interface
+resolution. Thus, the diffusive interface should be resolved in such a way as to have
 negligible effect on the physics of the film.
@@
 model simulates two liquids with the assumption of uniform overall
-density. While the classical Bretherton problem is stated for gas and liquid, which are of largely
+density. While the classical Bretherton problem is stated for gas and liquid, which are of
+significantly
 different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
 neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
-negligible Reynolds number effect on the film thickness for the range of Reynolds numbers.
-Therefore, the governing factor for microchannel flows is not the densities ratio, but the
-viscosities ratio. Moreover, the results based on the uniform density, as in the case of the LBM
-binary liquid model, are of good comparison with other simulations
+negligible Reynolds number effect on the film thickness for the relatively wide range of Reynolds
+numbers.
+Therefore, the major governing factor for microchannel flows is not the density ratio, but the
+viscosity ratio. Moreover, the results based on the uniform density, as in the case of the LBM
+binary liquid model, are good compared with other simulations
 \cite{giavedoni-numerical,heil-bretherton} taking into the account inertia effects.
 
-The goal of this paper is to give an answer as to how to choose the
-binary liquid model parameters, the viscosities ratio and the grid
-resolution, in order to obtain correct film thicknesses.
+The goal of this paper is to understand how to choose the
+binary liquid model parameters, the viscosity ratio and the grid
+resolution to obtain correct film thicknesses.
 
 The paper is organized as follows.  First, we briefly
-explain the binary liquid lattice Boltzmann model and summarize the literature
+explain the binary liquid lattice Boltzmann model and review the literature
 for the Bretherton problem. Then, the parameters involved in the
@@
 is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
-it was realized \cite{wong-films,wong-pressure} that the film deposition
-is proportional to $\Ca^{2/3}$ only in the specific region behind the front meniscus and
+it was realized \cite{wong-films,wong-pressure} that the film thickness
+is proportional to $\Ca^{2/3}$ only in the certain region behind the front meniscus and
 that the film thickness
@@
 simulations \cite{giavedoni-numerical} and experimental studies
-\cite{kreutzer-pressure-drop} showed as well a Reynolds number dependence, and
-a deviation from the $\Ca^{2/3}$ rule for capillary numbers larger than $0.003$.
-To be able to predict consistently the flow pattern for capillaries in
-different ranges of parameters, simulations have been developed and are typically validated with the
+\cite{kreutzer-pressure-drop} showed a deviation from the $\Ca^{2/3}$ rule and Reynolds number 
+effect on the film thickness for capillary numbers larger than $0.003$.
+To  predict consistently a flow pattern for capillaries in
+different ranges of parameters, simulations have been conducted and are typically validated with the
 small capillary numbers Bretherton problem.
@@
 thickness for rising bubbles in a circular capillary using the finite volume method.
-\citet{kreutzer-pressure-drop} used the finite volume method as well to perform
+\citet{kreutzer-pressure-drop} also used the finite volume method to perform
 simulations of a circular capillary for a number of different
@@
 three-dimensional bubbles in
-polygonal capillaries and showed the derivations of different shapes in the
-slug crossections and menisci appearance.
+polygonal capillaries and showed calculated different shapes in the
+slug cross sections and menisci appearance.
 \citet{heil-bretherton,ingham-plates} studied gas finger propagation in
 a two-dimensional channel for a range of Reynolds and capillary
-numbers using finite element method. \citet{giavedoni-numerical} performed crossvalidation of the
-finite element solution developed by them with previously published results.
-The solution is available for circular and planar cases.
+numbers using finite element method. \citet{giavedoni-numerical} performed cross validation of the
+finite element solution with previously published results.
+The solutions were obtained for circular and planar cases.
 
-While the applicability of other methods is well established for the
+While the applicability of other methods has been demonstrated for the
 simulation of the Bretherton/Taylor problem, this is not the case for
-the lattice Boltzmann method. A thorough parameter study for this approach
-has not yet been done to the best of the authors' knowledge.
+the lattice Boltzmann method. A thorough parametric study for this approach
+has not been done to the best of the authors' knowledge.
 
@@
 in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
-simulations of two-dimensional channel flows for larger capillary numbers, and
+simulations of two-dimensional channel flows for relatively large capillary numbers, and
 the authors found discrepancies with the classical Bretherton theory, which
@@
 
-To resolve the interface on a fine level for capillary numbers less than
-$0.003$ is computationally intensive even in the two-dimensional case
-(the amount of memory necessary to perform the simulation is on the order
-of tens or hundreds of gigabytes).  This work is therefore based on a
-comparison with the already established results, which were themselves
+Resolving the interface on a fine level for capillary numbers smaller than
+$0.003$ is computationally expensive even in the two-dimensional case
+(the amount of memory necessary to perform the simulation is of the order
+of tens or hundreds gigabytes).  This work is therefore based on a
+comparison of the results obtained with the known data
 validated for the Bretherton bubble flow.  In this paper, we present
-recipes for the initialization of the simulations and discuss optimal
+techniques for initialization of the simulations and discuss optimal
 parameter ranges for the binary liquid LB model.
@@
 directions specified by the discrete velocity set.
-The LBE is typically implementated as follows:
+The LBE is typically implemented as follows:
 \begin{equation}
@@
 \end{equation}
-where $F_i$ is the external force population. The binary fluid LB model is
+where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
+relaxation parameter, and $F_i$ is the external force population.
+
+The binary fluid LB model is
 based on a free-energy functional \cite{swift,landau}, and operates with two
-sets of populations: one to track the pressure and velocity fields, and the
-other to represent the phase field.
+sets of populations: one to track the pressure and the velocity fields, and another to represent the
+phase field $\phi$ indicating the gas or liquid.
 
-The model we use in this work is a two-dimensional nine-velocity (D2Q9) model,
+The model we use is a two-dimensional nine-velocity (D2Q9) model,
 with equilibrium populations \cite{pooley-contact}:
 \begin{equation}
+\label{set:equilibrium:binary}
 \begin{aligned}
@@
 where $\Gamma$ is the mobility parameter; the chemical potential
-$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
-tension; $A$ is a parameter of the free-energy model  The bulk pressure
+$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
+tension; $A$ is the parameter of the free-energy model  The bulk pressure
 is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
-Parameters speficic to the D2Q9 grid are the weights
+Parameters specific to the D2Q9 grid are the weights
 $w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
@@
 $w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
-$w^{xy}_{7-8}=-1/4$. The system can be shown to restore the macroscopic
+$w^{xy}_{7-8}=-1/4$. The set of equations (\ref{set:equilibrium:binary}) restores the
+macroscopic
 fluid equations as:
@@
 where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
-$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter. The system allows the separation of the liquid
+$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter.
+
+The system allows the separation of the liquid
 phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
-relaxation time is taken as linearly dependant on the relaxation
+relaxation time is taken as linearly dependent on the relaxation
 times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
@@
 \section{Lattice Boltzmann benchmark}
-To properly simulate and to compare simulation results with the data
+To properly simulate and compare simulation results with the data
 published in the literature, one needs to design a numerical benchmark addressing
@@
  \item The diffusive nature of the interface in the multiphase model should not
-influence the film thickness results. This is obviously achieved only for certain
-resolutions of the interface in comparison with the film thickness. The same applies to spurious
-currents arising from surface tension discretization which
+influence the film thickness results. This is obviously achieved only for certain relation of the
+interface resolution related to the film thickness. The same applies to spurious
+currents arising due to surface tension discretization, which
 can influence mass transfer in the thin film, thereby compromising the
-overall solution. Therefore the diffusive interface should be of nengligible influence on the film
+overall solution. Therefore, the diffusive interface should not noticeably affect the film
 thickness.
@@
 influence the overall solution.
- \item The free-energy model examined in the work is a single density, two
+ \item The free-energy model examined in this work is a single density, two
 viscosities model. The classical Bretherton problem is a free-interface problem between liquid and
 gas
-where the film thickness is established by itself. Thus, the viscosity and density ratios for the
+where the film thickness is established by itself. Thus, the viscosity and the density ratios for
+the
 classical
 Bretherton problem should
-be large. However, in terms of the lattice Boltzmann binary liquid framework only the viscosities
-ratio can be addressed. The results show that the results are consistent with other works and
+be large. However, in terms of the lattice Boltzmann binary liquid framework only the viscosity
+ratio can be addressed. Our computations show that the results are consistent with other works and
 inertia effects can be neglected.
   \item The Bretherton problem addresses a bubble flow driven by a pressure
-difference, not a body force. However, to the best authors knowledge the pressure boundaries
+difference, not a body force. However, to the best authors' knowledge the pressure boundary
 conditions are still not yet developed for the binary liquid model, because it involves the
@@
 This kind of pressure boundary conditions for the binary liquid model is currently
-under development, and preliminary results show that the differences between
-pressure difference flows and flows driven by a body force are small as far as the film
-deposition thickness is of interest.  In this work we limit ourselves to the study
-of body force driven flows, because of their simplicity and better numerical stability. The case of
-body force driven flows implies the periodic boundary conditions. As soon as the periodic boundary
-conditions are applied, not one bubble but whole bubble train is simulated. In this case one needs
-to insure that bubbles do not influence each other making large distance between them.
- \item Most simulations are done for a channel with circular crosssection,
+under development, and preliminary results show that the differences between flows driven by a body
+force and by
+pressure difference are small as far as the film
+ thickness is of interest.  In this work we limit ourselves to the study
+of body force driven flows, because of their simplicity and better numerical stability. This
+case implies the periodic boundary conditions. As soon as the periodic boundary
+conditions are applied, not a single bubble but whole bubble train is simulated. In this case one
+needs
+to ensure that the distance between bubbles is large enough to exclude mutual bubble influence.
+ \item Most simulations are done for a channel with circular cross section,
 which is quite difficult to address in terms of lattice Boltzmann free energy
-framework where one needs to calculate the Laplacian and gradients near boundaries.
+framework, where one needs to calculate the Laplacian and gradients near boundaries.
 Although one can apply certain finite difference stencils
@@
 terms in the continuity and Navier-Stokes equation \cite{halliday-circular}, the lattice Boltzmann
-incorporation of force and mass terms in the process of development and validation. Therefore
+incorporation of force and mass terms is in the process of development and validation. Therefore,
 the benchmark is chosen as a plane-symmetric two-dimensional case.
@@
 is a two-dimensional flow driven by a body force.  An analytical
-solution to this problem is to the best of our knowledge not yet known for the range of capillary
+solution to this problem is not known for a wide range of capillary
 numbers.
@@
 bubble shape
-is different in the front and rear menisci depending on Bond number $Bo$. \citet{wong-films}
+is different in the front and rear menisci depending on Bond number ($Bo=\frac{\Delta \rho g
+l^2}{\gamma}$). \citet{wong-films}
 analytically described the film thickness variation along a bubble varying from $\Ca^{1/2}$ to
@@
 $Ca^{2/3}$. Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
-to be sufficient to give consistent with theory results. The
-channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
-to minimize the influence of one bubble on another,
+to be sufficient to give results consistent with the theory. The
+channel length is taken to be $3$ bubble lengths (or $15$ channel heights)
+to minimize influence of one bubble on another,
 because of the periodicity of boundary conditions.
 Figure \ref{fig:benchmark:sketch} is a sketch of the geometry
-used in the benchmark.
+used for the benchmark.
 \begin{figure}
-%\includegraphics[width=0.97\textwidth]{}
-\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
+\includegraphics[width=0.97\textwidth]{Figures/benchmark.eps}
+\caption{The benchmark sketch. $\delta$ corresponds to film thickness. $\xi$ corresponds to the interface thickness. We take microchannel length to be 15 times larger than height.
 \label{fig:benchmark:sketch}}
@@
 We now introduce the procedure of nondimensionalization, as well as
-the initialization techniques we used to be able to determine parameters necessary for
+the initialization technique we used to determine parameters necessary for
 simulation. The capillary number governs the interface thickness and is defined as:
@@
 \end{equation}
-where $\mu_{\mathrm{liq}}$ is the liquid viscosity, and $\gamma$ is the surface tension. The
+where $\mu_{\mathrm{liq}}$ is the liquid viscosity, and $\gamma$ is the interface tension. The
 bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
@@
 explicitly.
-In the simulation we approximately match and initialize the velocity with the
-Poiseuille profile center velocity, which is given by the equation:
+In our simulations we approximately match and initialize the velocity with the
+Poiseuille profile center velocity, which is given as:
 \begin{equation}
@@
 effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes.
-By taking the Poiseuille profile one can estimate the body force through the
+For the Poiseuille profile one can estimate the body force (pressure gradient) through the
 capillary number and grid resolution:
@@
 \end{equation}
-The body force (\ref{poiseuille:velocity:center}) applied to the problem will cause the bubble
-velocity different from the implied by the capillary number because of the multiphase flow. That's
+The body force (\ref{poiseuille:velocity:center}) causes the difference  in the bubble
+velocity from that implied by the capillary number, because of the bubble presence in a flow. That's
 why to achieve the 
 desired velocity and the
-capillary number the iteration procedure to find the appropriate body force needs to be done.
+capillary number the iteration procedure is needed.
 However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
-starting approximation point.
+starting approximation.
 
-After one calculates the parameters necessary to run a simulation, one needs to
+After one calculates the parameters necessary to run a simulation, it is necessary to
 initialize the macroscopic fields and particle populations.
-The velocity is initialized with $0$ everywhere. Populations are initialized
-with the equilibrium values for the binary liquid model, including all the phase
+The velocity is initialized with zero everywhere. Populations are initialized
+using the equilibrium values for the binary liquid model, including all the phase
 gradients and Laplacians. The liquid phase is initialized with $\phi = 1$ and the
-gas phase is initialized with $\phi = -1$. Though the initial film thickness
-is not involved in the final result for the front meniscus (see Fig.
+gas phase with $\phi = -1$. Though the initial film thickness
+does not affect the final result for the front meniscus (see Fig.
 \ref{fig:different:initialization:widths}) we keep the initialized film
@@
 thickness values \cite{giavedoni-numerical}. This is done to minimize
-the convergence simulation time, as starting far away from the steady-state requies
+the convergence simulation time, because starting far away from the steady-state requires
 additional time for the relaxation.
@@
 430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
-\caption{The phase plots for different bubble widths as
+\caption{The phase plots for different bubble thicknesses as
 $H_{\mathrm{eff}}-12$, $H_{\mathrm{eff}}-16$ and
@@
 Here, two different examples are presented that quantitatively show the
-initialization procedure for quite different capillary numbers. In these
+initialization procedure for different capillary numbers. In these
 examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
@@
 A}{9}}=0.0377$ and the characteristic length of the interface
-$5\sqrt{\frac{k}{A}}=5$ lattice boltzmann units. For the sake of simplicity and
-stability the relaxation times for liquid and gas phases are taken as
-$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which corresponds
-to a gas-liquid kinematic viscosity ratio of $10$.
+$5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units. For the sake of simplicity and
+stability, the relaxation times for liquid and gas phases are taken as
+$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, that correspond
+to the gas liquid kinematic viscosity ratio equal to $10$.
 \begin{description}
  \item[I \bm{$Ca=0.005$}]
-  The film width for such small Capillary numbers \cite{bretherton} is
+  The film thickness for such small Capillary numbers \cite{bretherton} is
 given by the equation:
@@
 Thus, the film
-thickness occupies around $2$ percent of the whole channel height. Let us impose that $8$ lattice
+ occupies about $2$ percent of the whole channel height. Let us impose that $8$ lattice
 units are required to resolve  $2$ percent of the channel. Note that $5$ out of $8$ these lattice
@@
 
-We take the bubble length as $5$ channel widths
-  and the distance between bubbles is $2$ bubble lengths. This gives a
-grid size of $400 \times 6000$. In terms of computer memory
+We take the bubble length equal to $5$ channel widths
+  and the distance between bubbles as $2$ bubble lengths. This gives a
+grid size $400 \times 6000$. In terms of computer memory
 requirements, the calculations with double precision accuracy demand:
@@
 \end{equation}
-Even with this naive assumption for the $62.5$ percent resolution of the interface thickness of the
+Even using this naive assumption that the interface thickness is $62.5$ percent  of the
 film thickness, one can get the memory
@@
   \end{equation}
-Note the velocity is directly
-involved in calculation of the physical time step. Small velocites imply longer running simulations:
+Note that the velocity is directly
+involved in calculation of the physical time step (small velocities imply longer running
+simulations):
 \begin{equation}
@@
 \end{equation}
-where for example the real velocities in the physical world can reach
-approximately $1$ m/s and physical channel height up to $1$ mm. In this case one
+where, for example, the real velocities in the physical world can reach
+approximately $1$ m/s and the real height up to $1$ mm. In this case one
 can easily approximate the physical time step during one lattice Boltzmann iteration:
@@
 which is small. A typical simulation runs approximately
-$100000-300000$ time steps due to computational resources. That gives 
+$100000-300000$ time steps due to computational restrictions. That gives 
 physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
-probably not enough for physical system to come to the pseudo steady state.
+probably not enough for physical system to come to the pseudo-steady state.
 
@@
    The parameters in the previous subsection imply long
-simulations for system to come to equilibrium and this number increase with decrease of the
-capillary number. Bretherton obtained the analytical solution for the range of low capillary numbers
-less than
-  $0.005$. However, a lot of simulations extended it to the range of Capillary numbers greater than
+computations for system to come to equilibrium. The simulation becomes longer with decrease of the
+capillary number. Bretherton obtained the analytical solution for the range of low capillary
+numbers $Ca \leq 0.005$. However, many researchers extended it to the range of Capillary numbers
+greater than
 $0.005$, and to higher Reynolds numbers. One can refer to figures
   \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
-  \citet{giavedoni-numerical,heil-bretherton}, who crossvalidated their
-  simulations with Bretherton and each other. Even though the LBM can simulate
+  \citet{giavedoni-numerical,heil-bretherton}, who cross validated their
+  simulations against the Bretherton data and each other. Even though the LBM can simulate
 flows with small capillary numbers, i.e. less than $0.005$, the simulation time
-is relatively long. Therefore, we focus our work on the simulations of
-\citet{giavedoni-numerical} who crossvalidated
+is relatively long. Therefore, in our work we focus on the range of capillary numbers $Ca$,
+employed by the simulations of
+\citet{giavedoni-numerical} who cross validated
 results for all ranges of capillary numbers.  Our intention is
-  to show proper physics behavior for moderate parameters
+  to show physically justified behavior for moderate parameters,
 which
@@
   Capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
-  number ranges from $0$ to $300$.
+  number from $0$ to $300$.
 
-  The film thickness for such small Capillary numbers is taken from
+  The film thickness for such small Capillary numbers is taken from Fig. 
   \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
@@
 lattice units are related to the interface resolution.  Therefore the channel
-  width is $200$ lattice units. Therefore using the assumptions for the benchmark 
-the grid size is $200 \times 3000$. The velocity can be
+  width is $200$ lattice units. Therefore, 
+the employed grid size is $200 \times 3000$. This velocity can be
 calculated from the Capillary number as:
@@
   \end{equation}
-  Our approach to the initial conditions is  to take values from the previously performed
-numerical simulations and to calculate the grid sizes from them, matching the
-body force gradient through the solution of Poiseuille flow.
+  Our approach to the initial conditions is  to take the flow parameters from the previously
+performed
+numerical simulations and to calculate the grid size from them, matching the
+body force gradient through the solution for Poiseuille flow.
 
@@
 capillary number and the body force. However, in the binary liquid framework
-those correlations are only approximate. The stronggest assumption lies
+those correlations are approximations. The strongest assumption is
 in calculating the body force from the Poiseuille velocity profile.
-In reality, the less viscous bubble will have a slippage
-factor and its velocity is larger than the surrounding liquid velocity. By
-substituting the obtained velocity back to the capillary number equation the
-larger capillary number can be obtained. While we present the initialization
-tips, in practice one needs to take the bubble velocity from the simulations and
-recalculate all the necessary characteristics or in other words adjust the
-forcing to obtain the desired velocity/capillary number.
+In reality, the less viscous bubble moves faster than the surrounding liquid. By
+substituting the obtained velocity in the capillary number equation the
+larger capillary number can be obtained. In practice one needs to take the bubble velocity from the
+simulations and
+recalculate all the necessary flow characteristics, i.e. adjust the
+ applied force to obtain the desired velocity/capillary number.
 
 \subsection{Grid refinement}
-To properly estimate the interface resolution one needs to study the convergence
-of the results as a function of the grid resolution. To do that, all parameters are held fixed,
+To properly estimate the interface resolution one needs to study the convergence as a function of
+the grid resolution. To do that, all parameters are fixed,
 including the bubble velocity, the capillary number, and only the grid
-resolution is changed. Our goal is to determine the ratio of the interface thickness to 
-film thickness where results are not dependent on the grid resolution.
+resolution is varied. Our goal is to determine the ratio of the interface thickness to the 
+film thickness, at which results are not dependent on the grid resolution.
 
-We illustrate the procedure for the capillary number
-$0.05$. As was pointed out earlier, the film thickness in that case is $6$ percent of the effective
+Let us illustrate the procedure for the capillary number
+$0.05$. As it was pointed out above, the film thickness in that case is $6$ percent of the
+effective
 channel $H_{\mathrm{eff}}$.
@@
 If the grid number in the vertical direction is $N_y$ then one has $N_y-1$ regions which represent
+the
 domain. The effective wall location is in the middle between bounce-back and fluid nodes giving
 overall $N_y-2$ representing the channel width. Note that it is a big
-simplification to impose the boundary in the middle beetween the bounce-back
-node and the fluid node which give rise to the viscosity dependent location of
-the wall \cite{ginzburg-multireflection}, however the effective location of the wall for the
+simplification to impose the boundary in the middle between the bounce-back
+node and the fluid node. The location of the wall for simply hydrodynamics flows is
+viscosity dependent \cite{ginzburg-multireflection}.  Note that the effective location of the
+wall for the
 multiphase models to the best authors' knowledge is not yet derived. In the case of $\delta=0.06$
@@
 x } = \mathrm{const}$. We performed a number of simulations for the prescribed grids. The
-simulations in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
+simulation results in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
 velocity $U_{bubble}$ are summarized in Table \ref{table:parameters:grid:refinement}.
 The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
-see the visual results converge for $H_{\mathrm{eff}}\geq 175$. However, one can see
+see that results converge for $H_{\mathrm{eff}}\geq 175$. However, one can see
 that with proper initialization techniques, large enough time and the wall
 wettability, results are different only in the 3rd digit even for
-underresolved film thicknesses. To calculate how well the interface thickness is
+underresolved film thicknesses. To calculate how well the interface is
 resolved, the ratio of the interface thickness to the film thickness is calculated. The
@@
 \ref{table:parameters:grid:refinement} that the bubble velocities are pretty consistent and the
-calculated capillary number corresponding to those values is $0.07$, which is slightly
+calculated capillary number corresponding to these velocities is $0.07$, which is slightly
 larger than the expected capillary number. The corresponding error can be
-attributed to the density and viscosity ratios which are much larger in the
-Bretherton problem and Poiseuille profile body force initialization.
+attributed to the density and the viscosity ratios, which are much larger in the
+Bretherton problem, and Poiseuille profile body force initialization.
 \begin{figure}
@@
 \caption{The parameters and results for studying grid resolution simulations. Domain simulated is
-of size $N_x\, \mathrm{x}\, N_y$. $U_{bubble}$ is center
-bubble velocity measured at the scaled to $H_eff$ $x=14.0$.
+of size $N_x\, \mathrm{x}\, N_y$. $U_{bubble}$ is the velocity of the bubble center measured at the
+scaled to $H_{eff}$ $x=14.0$.
 $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of the film in
@@
 gradient ranging from $-1$ to $1$. The parameters are summarized in Table
-\ref{table:parameters:wall:gradient}. The results are consistent among all the
-wall gradients given that the interface is resolved in the right way. For 
+\ref{table:parameters:wall:gradient}. The results are consistent for all the
+wall gradients given that the interface is properly resolved. For 
 given wall gradients the values of interface thickness values are of $0.2\%$ relative accuracy.  For
-the wall gradients $0.8$
+the given wall gradients $0.8$
 and $1.0$ the
-simulations are unstable. One can see the simulated profiles in Fig.
+simulations are unstable. One can see the calculated profiles in Fig.
 \ref{fig:gradients:profiles}. The latter shows that the negative wall gradient values cause 
-unphysical phase values more than $1.0$ near the wall. One should attribute them to the numerical
-imposement of wall gradients. Therefore, they do not spoil numerical scheme as whole. Note that
+non-physical phase values (above $1.0$ near the wall). One should attribute them to the numerical
+impose of wall gradients through the first-order finite difference scheme. Therefore, they do
+not spoil numerical results. Note that
 negative values of the phase gradient are
@@
 gradients the values near the wall are below $1$, and the gradient profile can
-fuse with the values of the gas phase. This case is attributed when to the slug flow when gas
-touches the wall.
+fuse with the values of the gas phase. This case corresponds to the slug flow when gas is in
+contact with the wall.
 \begin{figure}
@@
 
-Another question which arises is the influence of the wall gradient on the bubble velocity and the
+Another question that arises is the influence of the wall gradient on the bubble velocity and the
 corresponding capillary number. The shear stress controlled by the wall phase gradient changes the
-effective viscosity near the wall, thus changing the bubble velocity as well. If viscosity ratio is
+effective viscosity near the wall and the bubble velocity respectively. If the viscosity ratio is
 high enough then the free surface motion of the bubble should not be dependent on the shear stress
@@
 One can see that if the film thickness is resolved properly the influence of the wall gradient can
-be neglected. The center bubble velocities are presented in Table
+be neglected. The velocities of the bubble center for certain profiles along $x$ direction are
+presented in Table
 \ref{table:parameters:wall:gradient}. 
@@
 %gradients. 
-The average relative error around the mean velocity is $0.1\%$.
+The average relative error for the mean velocity is $0.1\%$.
 % Uncomment if you need this picture
@@
 numbers. Because of limited computational resources, we skip the
-small capillary numbers and compare the film thickness over the capillary number
-range from $0.03$ to $0.8$, which is a computationally reasonable task.  
-To perform numerical simulations we refer
-to the grid refinement section where we established that the interface thickness should be
-at most $50$ percent of the interface thickness.  We therefore choose the grid to be
+small capillary numbers and make calculations for the range of capillary number $0.03-0.8$, which
+is a computationally reasonable task. We choose consistent with the previous sections grid to be
 $202 \times 3001$. $5$ lattice Boltzmann units do not occupy more than $60$
@@
 smaller grids and larger gradients, all the capillary number simulations were
-performed on the chosen grid. To properly initialize the force gradient, the
+performed on the same grid. To properly initialize the force gradient, the
 proportionality law was utilized. The forcing
@@
 \end{equation}
-The film thickness is initialized through the capillary numbers ratio as well:
+The film thickness is initialized through the ratio of capillary numbers as well:
 \begin{equation*}
@@
 \end{equation*}
-The results were obtained after $2\cdot10^5$ steps. Simulation results are presented in Table
+The results obtained after $2\cdot10^5$ steps are presented in Table
 \ref{table:parameters:capillary:number}. One can see that the calculated
 capillary numbers are overpredicted but the obtained thicknesses are overpredicted
-as well due to the improper body force imposement, because of the Poiseuille flow assumption. If one
+as well due to the improper body force setting caused by the Poiseuille flow assumption. If one
 needs to obtain the desired capillary number, the shooting method for the forcing is
 necessary starting as the Poiseuille pressure gradient force as the initial condition. We extracted
-results of 
-\citet{giavedoni-numerical} and \citet{heil-bretherton} and compared them with our results, shown
-in Fig. \ref{fig:capillary:comparison}.
+the data of 
+\citet{giavedoni-numerical} and \citet{heil-bretherton} and compared them with our results (see 
+Fig. \ref{fig:capillary:comparison}).
 \begin{figure}
@@
 Next, we investigated the qualitative variation of the film thickness over the bubble length.
-Though there are simulations for the bubble length for the two-dimensional case, due to
-the presentation of data it is hard to extract numerical values from the plots. One can see bubble lengths
-variations for the two-dimensional simulations for the range of capillary numbers in Fig.
+Though, there are simulations for the bubble length for the two-dimensional case, due to
+the computer resources constraints we are not able to provide simulations with long bubbles. Thus,
+the qualitative comparison for the thickness variation is conducted in Fig.
 \ref{fig:sehgal:bubble:length}.
@@
 the bubble length. Left (courtesy of \citet{sehgal-microchannel}),
-right (present simulations). $x$ is scaled to $H_{\mathrm{eff}}$
-and is taken in the flow direction.\label{fig:sehgal:bubble:length}}
+right (present simulations); $x$ is scaled to $H_{\mathrm{eff}}$
+and increases in the flow direction.\label{fig:sehgal:bubble:length}}
 \end{figure}
@@
 \subsection{The influence of different viscosities}
-The results obtained earlier are taken for the liquid viscosity ratio $10$.
-We performed the same simulations for the liquid viscosity ratio of $20$.
+The results presented above are taken for the liquid viscosity ratio $10$.
+We performed the same simulations for the liquid-gas viscosity ratio of $20$.
 The relaxation parameters were taken as $\tau_{\mathrm{liquid}}=4.5$ and $\tau_{\mathrm{bubble}}=0.7$. We
-did not find any significant differences with the results obtained earlier. The
-comparison for the liquid thicknesses is presented in Fig.
+did not find any significant differences with from the results presented above. The
+comparison for the liquid film thicknesses is presented in Fig.
 \ref{fig:capillary:viscous}. The results validate our assumption that in
 the low capillary flow regime the density ratio does not affect the film thickness,
-and that the gas-liquid viscosity ratio of the order $10$ is enough to obtain results
-consistent with the Bretherton/Taylor problem.
+and that the gas-liquid viscosity ratio of the order $10$ is sufficient to obtain results
+consistent with the literature data.
 \begin{figure}
@@
 Bretherton/Taylor problem using the binary liquid lattice Botlzmann method. To obtain correlations
-we specifically designed the benchmark. The bubble was chosen long enough for the film thickness to
+we specifically designed the benchmark. The bubble was chosen sufficiently long for the film
+thickness to
 stabilize. Also, we used periodic boundary conditions for scheme robustness. Thus, the bubble
-train was simulated instead of one bubble motion. To avoid the influence of other bubbles
-distance between bubbles were chosen as $3$ bubble sizes. The
-results show consistency with the previously published numerical results obtained
-using other methods in terms of capillary number region, shape of the bubble. It comes with surprise
-but with large enough viscosities ratio the results are independent on inertia. This work supports
-this assumption as well. Also, we examined the influence of grid resolution on the results. We found
-that if the phase interface is resolved as $50$ percent of the simulated film thickness, the results
+train was simulated instead of one bubble motion. To avoid the mutual influence of neighboring
+bubbles a
+distance between bubbles was chosen to be $3$ bubble sizes. The computational
+results in terms of capillary number dependence and shape of a bubble show consistency with the
+previously published data. Surprisingly, with large enough viscosity ratio the results are
+independent on inertia. Also, we examined the influence of grid resolution on the results. We found
+that if the phase interface is resolved as $50$ percent of the simulated film thickness, the
+simulations
 are convergent. Though  our results are specific to the binary liquid lattice
 Botlzmann method, the numerical hints and procedures can be used for any
-continuous interface method.  We hope this work will prove to be valuable for
-people simulating microchannel flows.
-\bibliographystyle{unsrtnat}
+continuous interface method. 
+\bibliographystyle{model1-num-names}
 \bibliography{paper}
-
+%\bibliographystyle{unsrtnat}
+%\bibliography{paper}
 \end{document}

### 2011-01-05 Text/paper.tex
Fix typos + style fixes in the abstract and intro.

@@
 \ead{fmostowfi@slb.com}
-\address[google]{Google, Zurich, Switzerland}
+\address[google]{Google Switzerland GmbH, Brandschenkestrasse 110, 8002 Zurich, Switzerland}
 \address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
@@
 numbers
-$\Ca$ (defined in Eq. \ref{capillary:number:definition}), the flow is thoroughly described by
+$\Ca$ (defined in Eq.~\ref{capillary:number:definition}), the flow is thoroughly described by
 Bretherton \cite{bretherton},
 with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
-from the Bretherton correlation for $Ca\geq 0.005$, studied experimentally and numerically. This
-work aims to examine the lattice Boltzmann binary liquid model feasibility to obtain valid results
-for a bubble flow in liquid for a range of $Ca$ number. The binary
+from the Bretherton correlation for $\Ca\geq 0.005$, studied experimentally and numerically. This
+work aims to examine the applicability of the lattice Boltzmann binary liquid model to simulations
+of bubble flow in liquids for a range of $\Ca$ numbers. The binary
 liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
 It was found that the interface thickness needs to be $50$
-percent of the film thickness for simulation to give convergent results. While
-the convergence criteria is fulfilled the results are in a good agreement with the
-numerical simulations of \citet{giavedoni-numerical} in the range of $Ca$ from $0.01$ to $1$. 
+percent of the film thickness for the simulation to give convergent results. When
+the convergence criterion is fulfilled, the results are in a good agreement with the
+numerical simulations of \citet{giavedoni-numerical} in the range of $\Ca$ from $0.01$ to $1$. 
 \end{abstract}
@@
 
-While there are number
+While there are a number
 of works which simulate the Bretherton problem by the boundary value approach
@@
 resolving the parameter range of the binary liquid lattice
-Boltzmann method to validate it in the range of capillary numbers.
+Boltzmann method to validate it for a range of capillary numbers.
 
@@
 multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
-the interface using continuous interface methods where the interfaces spans over several grid nodes.
+the interface using continuous interface methods where the interfaces spans several grid nodes.
 Such a representation
 brings issues of the film thickness resolution versus interface
-resolution. Thus, the diffusive interface should be resolved in such a way as to have
+resolution -- the diffuse interface should be resolved in such a way as to have
 negligible effect on the physics of the film.
@@
 \begin{itemize}
- \item The diffusive nature of the interface in the multiphase model should not
+ \item The diffuse nature of the interface in the multiphase model should not
 influence the film thickness results. This is obviously achieved only for certain relation of the
@@
 can influence mass transfer in the thin film, thereby compromising the
-overall solution. Therefore, the diffusive interface should not noticeably affect the film
+overall solution. Therefore, the diffuse interface should not noticeably affect the film
 thickness.
@@
 conditions are still not yet developed for the binary liquid model, because it involves the
-treatment of associated gradients and laplacians on the boundary.
+treatment of associated gradients and Laplacians on the boundary.
 This kind of pressure boundary conditions for the binary liquid model is currently
@@
 $177 \times 2626$ and the initial film thickness was
-$12$ lattice boltzmann units, which corresponds to the predicted film thickness.
+$12$ lattice Boltzmann units, which corresponds to the predicted film thickness.
 We examined $11$ different values for the wall
@@
 This work presents numerical tips and a benchmark for the
-Bretherton/Taylor problem using the binary liquid lattice Botlzmann method. To obtain correlations
+Bretherton/Taylor problem using the binary liquid lattice Boltzmann method. To obtain correlations
 we specifically designed the benchmark. The bubble was chosen sufficiently long for the film
@@
 are convergent. Though  our results are specific to the binary liquid lattice
-Botlzmann method, the numerical hints and procedures can be used for any
+Boltzmann method, the numerical hints and procedures can be used for any
 continuous interface method. 

### 2011-01-05 Text/paper.tex
Fixes in the remaining part of the text + affiliation change.

@@
 \ead{shurik.kuzmin@gmail.com}
-\author[google]{M.~Januszewski}
+\author[us]{M.~Januszewski}
 \ead{michalj@gmail.com}
@@
 \ead{fmostowfi@slb.com}
-\address[google]{Google Switzerland GmbH, Brandschenkestrasse 110, 8002 Zurich, Switzerland}
+\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
 \address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
@@
 \label{capillary:number:definition}
-Ca=\frac{\mu_{liq} U_{bubble}}{\gamma},
+\Ca=\frac{\mu_\mathrm{liq} U_\mathrm{bubble}}{\gamma},
 \end{equation}
@@
 multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
-the interface using continuous interface methods where the interfaces spans several grid nodes.
+the interface using continuous interface methods where the interface spans over several grid nodes.
 Such a representation
@@
 
-This work examines the binary liquid free-energy LB model \cite{swift}. The
-model simulates two liquids with the assumption of uniform overall
+The binary liquid free-energy LB model \cite{swift} we used
+simulates two liquids with the assumption of uniform overall
 density. While the classical Bretherton problem is stated for gas and liquid, which are of
@@
 neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
-negligible Reynolds number effect on the film thickness for the relatively wide range of Reynolds
+negligible Reynolds number effect on the film thickness for a relatively wide range of Reynolds
 numbers.
@@
 viscosity ratio. Moreover, the results based on the uniform density, as in the case of the LBM
-binary liquid model, are good compared with other simulations
-\cite{giavedoni-numerical,heil-bretherton} taking into the account inertia effects.
+binary liquid model, are in good agreement with other simulations
+\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account.
 
@@
 To  predict consistently a flow pattern for capillaries in
-different ranges of parameters, simulations have been conducted and are typically validated with the
+different ranges of parameters, simulations have been conducted and have been typically validated with the
 small capillary numbers Bretherton problem.
@@
 a two-dimensional channel for a range of Reynolds and capillary
-numbers using finite element method. \citet{giavedoni-numerical} performed cross validation of the
+numbers using the finite element method. \citet{giavedoni-numerical} performed cross validation of the
 finite element solution with previously published results.
@@
 in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
-simulations of two-dimensional channel flows for relatively large capillary numbers, and
+simulations of two-dimensional channel flows for a relatively large capillary numbers, and
 the authors found discrepancies with the classical Bretherton theory, which
@@
 $\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
-tension; $A$ is the parameter of the free-energy model  The bulk pressure
+tension; $A$ is the parameter of the free-energy model.  The bulk pressure
 is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
@@
  \item The diffuse nature of the interface in the multiphase model should not
-influence the film thickness results. This is obviously achieved only for certain relation of the
-interface resolution related to the film thickness. The same applies to spurious
+influence the film thickness results. This is obviously achieved only for certain
+interface resolutions relative to the film thickness. The same applies to spurious
 currents arising due to surface tension discretization, which
@@
 ratio can be addressed. Our computations show that the results are consistent with other works and
-inertia effects can be neglected.
+inertial effects can be neglected.
   \item The Bretherton problem addresses a bubble flow driven by a pressure
@@
 the analysis of error introduced by the boundary conditions becomes an issue. Though the circular
-axis-symmetric case can be simulated on rectangular grid by introducing specific mass and force
+axis-symmetric case can be simulated on a rectangular grid by introducing specific mass and force
 terms in the continuity and Navier-Stokes equation \cite{halliday-circular}, the lattice Boltzmann
@@
 bubble shape
-is different in the front and rear menisci depending on Bond number ($Bo=\frac{\Delta \rho g
+is different in the front and rear menisci depending on Bond number ($\mathrm{Bo}=\frac{\Delta \rho g
 l^2}{\gamma}$). \citet{wong-films}
@@
 $\Ca^{2/3}$. The bubble should be long enough to have the film thickness proportional to
-$Ca^{2/3}$. Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
+$\Ca^{2/3}$. Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
 to be sufficient to give results consistent with the theory. The
@@
 \includegraphics[width=0.97\textwidth]{Figures/benchmark.eps}
-\caption{The benchmark sketch. $\delta$ corresponds to film thickness. $\xi$ corresponds to the interface thickness. We take microchannel length to be 15 times larger than height.
+\caption{The benchmark sketch. $\delta$ corresponds to film thickness. $\xi$ corresponds to the interface thickness. We take the microchannel length to be 15 times larger than its height.
 \label{fig:benchmark:sketch}}
@@
 We now introduce the procedure of nondimensionalization, as well as
-the initialization technique we used to determine parameters necessary for
-simulation. The capillary number governs the interface thickness and is defined as:
-\begin{equation}
-Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
-\end{equation}
-where $\mu_{\mathrm{liq}}$ is the liquid viscosity, and $\gamma$ is the interface tension. The
-bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
+the initialization technique we used to determine the parameters necessary for
+the simulation. The capillary number defined in Eq.~\ref{capillary:number:definition} governs the interface thickness and is.
+The bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
 thickness is examined.
@@
 \end{equation}
-where $N_y$ is the grid number in the vertical direction. As it will be discussed further, the
+where $N_y$ is the grid number in the vertical direction. As will be discussed further, the
 effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes.
@@
 \begin{aligned}
-&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma Ca.
+&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma \Ca.
 \end{aligned}
 \end{equation}
-The body force (\ref{poiseuille:velocity:center}) causes the difference  in the bubble
-velocity from that implied by the capillary number, because of the bubble presence in a flow. That's
-why to achieve the 
-desired velocity and the
-capillary number the iteration procedure is needed.
+The body force (\ref{poiseuille:velocity:center}) causes the bubble
+velocity to differ from that implied by the capillary number because of the bubble presence in a flow. That is
+why the iteration proceduer is needed to achieve the 
+desired velocity and 
+capillary number. 
 However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
@@
 
-After one calculates the parameters necessary to run a simulation, it is necessary to
+After one calculates the parameters necessary to run a simulation, one has to
 initialize the macroscopic fields and particle populations.
@@
 thickness values \cite{giavedoni-numerical}. This is done to minimize
-the convergence simulation time, because starting far away from the steady-state requires
+the time required for the simulation to be convergent, as starting far away from the steady-state requires
 additional time for the relaxation.
@@
 stability, the relaxation times for liquid and gas phases are taken as
-$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, that correspond
+$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which correspond
 to the gas liquid kinematic viscosity ratio equal to $10$.
@@
  \item[I \bm{$Ca=0.005$}]
-  The film thickness for such small Capillary numbers \cite{bretherton} is
+  The film thickness for such small capillary numbers \cite{bretherton} is
 given by the equation:
@@
   and the distance between bubbles as $2$ bubble lengths. This gives a
-grid size $400 \times 6000$. In terms of computer memory
+grid size of $400 \times 6000$. In terms of computer memory
 requirements, the calculations with double precision accuracy demand:
@@
 &2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
-} \times \,\, 9 \text{ number of
+} \times \,\, 9 \text{
 populations}\\
@@
 physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
-probably not enough for physical system to come to the pseudo-steady state.
+probably not enough for the physical system to come to the pseudo-steady state.
 
  \item[II $\bm{Ca=0.05}$] 
-   The parameters in the previous subsection imply long
-computations for system to come to equilibrium. The simulation becomes longer with decrease of the
-capillary number. Bretherton obtained the analytical solution for the range of low capillary
-numbers $Ca \leq 0.005$. However, many researchers extended it to the range of Capillary numbers
+   The parameters in the previous subsection imply that long
+computations are required in order for the system to come to equilibrium. The simulation becomes longer with a decrease of the
+capillary number. Bretherton obtained an analytical solution for the range of low capillary
+numbers $Ca \leq 0.005$. However, many researchers extended it to the range of capillary numbers
 greater than
@@
 flows with small capillary numbers, i.e. less than $0.005$, the simulation time
-is relatively long. Therefore, in our work we focus on the range of capillary numbers $Ca$,
+is relatively long. Therefore, in our work we focus on the range of capillary numbers $\Ca$
 employed by the simulations of
@@
   can be easily and quickly validated on a computer.
-  One can see from Fig. \ref{fig:heil:planar} that the film thickness for the
-  Capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
+  One can see from Fig.~\ref{fig:heil:planar} that the film thickness for the
+  capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
   number from $0$ to $300$.
 
-  The film thickness for such small Capillary numbers is taken from Fig. 
-  \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
+  The film thickness for such small capillary numbers is taken from Fig.~\ref{fig:heil:planar},
+  and for Reynolds number $40$ equals $0.12$. We want to
   resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
-lattice units are related to the interface resolution.  Therefore the channel
-  width is $200$ lattice units. Therefore, 
-the employed grid size is $200 \times 3000$. This velocity can be
-calculated from the Capillary number as:
+lattice units are related to the interface resolution.  Therefore, the channel
+  width is $200$ lattice units and the grid size is $200 \times 3000$. The velocity can be
+calculated from the capillary number as:
   \begin{equation}
@@
   \begin{aligned}
-  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca=3.77\cdot10^{-7}.
+  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma \Ca=3.77\cdot10^{-7}.
   \end{aligned}
@@
 \caption{\citet{giavedoni-numerical} gathered results across the
-literature for different Capillary numbers. Note that results are scaled on a half-width of the
+literature for different capillary numbers. Note that results are scaled to the half-width of the
 channel. \label{fig:giavedoni:planar}}
@@
 \caption{\citet{heil-bretherton} performed simulations for different ranges of
-Capillary numbers and Reynolds numbers. Note that results are scaled on a half-width of the channel.
+capillary numbers and Reynolds numbers. Note that results are scaled to the half-width of the channel.
 \label{fig:heil:planar}}
@@
 In reality, the less viscous bubble moves faster than the surrounding liquid. By
-substituting the obtained velocity in the capillary number equation the
-larger capillary number can be obtained. In practice one needs to take the bubble velocity from the
+substituting the obtained velocity in the capillary number equation, a
+larger capillary number can be obtained. In practice, one needs to take the bubble velocity from the
 simulations and
@@
 To properly estimate the interface resolution one needs to study the convergence as a function of
-the grid resolution. To do that, all parameters are fixed,
-including the bubble velocity, the capillary number, and only the grid
-resolution is varied. Our goal is to determine the ratio of the interface thickness to the 
-film thickness, at which results are not dependent on the grid resolution.
+the grid resolution. To do that the grid resolution is varies while all remaining parameters,
+including the bubble velocity and the capillary number, are fixed.  Our goal is to determine the 
+ratio of the interface thickness to the 
+film thickness at which results are no longer dependent on the grid resolution.
 
@@
 where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
-is related to the effective channel height. That can be cleared out by the following explanations.
-If the grid number in the vertical direction is $N_y$ then one has $N_y-1$ regions which represent
-the
-domain. The effective wall location is in the middle between bounce-back and fluid nodes giving
-overall $N_y-2$ representing the channel width. Note that it is a big
+is related to the effective channel height.
+If the grid size in the vertical direction is $N_y$, then one has $N_y-1$ regions between the grid nodes which represent
+the physical domain. The effective wall location is in the middle between bounce-back and fluid nodes giving
+overall $N_y-2$ nodes representing the fluid. Note that it is a big
 simplification to impose the boundary in the middle between the bounce-back
-node and the fluid node. The location of the wall for simply hydrodynamics flows is
+node and the fluid node. The location of the wall for simple hydrodynamic flows is
 viscosity dependent \cite{ginzburg-multireflection}.  Note that the effective location of the
@@
 The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
-see that results converge for $H_{\mathrm{eff}}\geq 175$. However, one can see
-that with proper initialization techniques, large enough time and the wall
-wettability, results are different only in the 3rd digit even for
+see that results converge for $H_{\mathrm{eff}}\geq 175$, and that with proper
+initialization techniques, large enough time and different wall
+wettabilities, results are different only in the 3rd digit even for
 underresolved film thicknesses. To calculate how well the interface is
@@
 $5\xi/H_{\mathrm{film}}$
- is shown in Table \ref{table:parameters:grid:refinement}. Based on results one can conclude that
+ is shown in Table \ref{table:parameters:grid:refinement}. Based on these results one can conclude that
  the interface needs to be resolved as $40-50$ percent of the
 expected film thickness for simulations to be convergent. We further examine the
-velocities in the center of the bubble to calculate the Capillary number. One can see from Table
+velocities in the center of the bubble to calculate the capillary number. One can see from Table
 \ref{table:parameters:grid:refinement} that the bubble velocities are pretty consistent and the
@@
 larger than the expected capillary number. The corresponding error can be
-attributed to the density and the viscosity ratios, which are much larger in the
-Bretherton problem, and Poiseuille profile body force initialization.
+attributed to the Poiseuille profile body force initialization and the
+ density and viscosity ratios, which are much larger in the
+Bretherton problem.
 \begin{figure}
@@
 channel widths
-$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_{eff}$ and $\delta=0$
-corresponds to the wall location. The profiles were taken at $x=14$ nondimensional coordinate.
-Capillary number is $Ca=0.05$. Capillary number obtained from simulations $Ca=0.07$. One can
+$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_\mathrm{eff}$ and $\delta=0$
+corresponds to the wall location. The profiles were taken at $x=14$ (nondimensional coordinates).
+Capillary number is $\Ca=0.05$. Capillary number obtained from simulations $\Ca=0.07$. One can
 see
@@
 \hline
-$N_x$&$N_y$&$\delta$&$U_{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_{iter}$\\
+$N_x$&$N_y$&$\delta$&$U_\mathrm{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_\mathrm{iter}$\\
 \hline
@@
 \end{tabularx}
-\caption{The parameters and results for studying grid resolution simulations. Domain simulated is
-of size $N_x\, \mathrm{x}\, N_y$. $U_{bubble}$ is the velocity of the bubble center measured at the
-scaled to $H_{eff}$ $x=14.0$.
-$5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of the film in
+\caption{The parameters and results for grid resolution simulations. Simulated domain is
+of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the velocity of the bubble center measured at
+$x=14$.  $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of the film in
 lattice Boltzmann units. $Ca=0.05$ was taken to initialize the simulation.
@@
 non-physical phase values (above $1.0$ near the wall). One should attribute them to the numerical
-impose of wall gradients through the first-order finite difference scheme. Therefore, they do
+adjustment of wall gradients through a first-order finite difference scheme. Therefore, they do
 not spoil numerical results. Note that
@@
 simulations were conducted for $200,000$ iterations. The measured profile was taken at the
-nondimensional coordinate scaled to $H_{\mathrm{eff}}$ as $x=11.42$. $Ca=0.05$ was taken to
+nondimensional coordinate scaled to $H_{\mathrm{eff}}$ as $x=11.42$. $\Ca=0.05$ was taken to
 initialize simulation.
@@
 \end{tabularx}
-\caption{The parameters and results for studying wall gradient simulations. $U_{b}$ stands for the
+\caption{The parameters and results for wall gradient simulations. $U_{b}$ stands for the
 center bubble velocity. $\delta$ is the interface thickness. $\partial_n \phi$ is the phase
 gradient near the wall responsible for hydrophilic and hydrophobic behavior. The results are
-calculated at nondimensional $x=11.42$ after $200000$ time iterations. $Ca=0.05$ was taken to
+calculated at nondimensional $x=11.42$ after $200 000$ time iterations. $\Ca=0.05$ was taken to
 initialize simulation.
@@
 numbers. Because of limited computational resources, we skip the
-small capillary numbers and make calculations for the range of capillary number $0.03-0.8$, which
-is a computationally reasonable task. We choose consistent with the previous sections grid to be
+small capillary numbers and make calculations for the range of capillary numbers $0.03-0.8$, which
+is a computationally reasonable task.  To be consistent with the previous sections, we choose the grid to be
 $202 \times 3001$. $5$ lattice Boltzmann units do not occupy more than $60$
-percent of the effective thickness. Because the simulation get unstable with
+percent of the effective thickness. Because the simulation gets unstable with
 smaller grids and larger gradients, all the capillary number simulations were
@@
 number $0.05$. The forcing for other grids can be obtained using
-simple correlations:
+simple proportionality relationships:
 \begin{equation}
 \begin{aligned}
-&Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
+&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
 &U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\
-&Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
+&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
 &\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{N_y^2},
@@
 involved because all simulations are conducted on the same grid.
-The pressure gradient can be obtained through the capillary numbers
+The pressure gradient can be obtained through the capillary number
 ratio:
 \begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d} x}=6 \cdot 10^{-6}/16 \frac{Ca_{\mathrm{lit}}}{0.05}
+\frac{\mathrm{d}P}{\mathrm{d} x}=6 \cdot 10^{-6}/16 \frac{\Ca_{\mathrm{lit}}}{0.05}
 \end{equation}
@@
 \begin{equation*}
-w=12 \frac{Ca_{\mathrm{lit}}}{0.05}
+w=12 \frac{\Ca_{\mathrm{lit}}}{0.05}
 \end{equation*}
@@
 needs to obtain the desired capillary number, the shooting method for the forcing is
-necessary starting as the Poiseuille pressure gradient force as the initial condition. We extracted
+necessary, starting with the Poiseuille pressure gradient force as the initial condition. We extracted
 the data of 
@@
 \end{tabularx}
-\caption{The parameters and results for studying capillary number region simulations.
-$\delta_{\mathrm{lit}}$ is the film thickness with corresponding $Ca_{\mathrm{lit}}$ taken from
-literature. $\delta_{sim}$ is the simulation film thickness with corresponding $Ca$. $U_b$ is the
-center bubble velocity taken at location $x$. $Ca$ is based on the measured center bubble velocity
+\caption{The parameters and results for capillary number region simulations.
+$\delta_{\mathrm{lit}}$ is the film thickness with corresponding $\Ca_{\mathrm{lit}}$ taken from
+literature. $\delta_\mathrm{sim}$ is the simulation film thickness with corresponding $\Ca$. $U_b$ is the
+center bubble velocity taken at location $x$. $\Ca$ is based on the measured center bubble velocity
 $U_{b}$.
@@
 Next, we investigated the qualitative variation of the film thickness over the bubble length.
-Though, there are simulations for the bubble length for the two-dimensional case, due to
-the computer resources constraints we are not able to provide simulations with long bubbles. Thus,
-the qualitative comparison for the thickness variation is conducted in Fig.
-\ref{fig:sehgal:bubble:length}.
+The comparison for the thickness variation is presented in Fig.~\ref{fig:sehgal:bubble:length}.
 \begin{figure}
@@
 One can see a qualitative agreement, i.e. the thickness increases towards the front meniscus and
-rapidly decreases towards the rear meniscus. This shape sometimes is referred as a "bullet" shape.
+rapidly decreases towards the rear meniscus. This shape is sometimes referred to as a "bullet" shape.
 
@@
 The relaxation parameters were taken as $\tau_{\mathrm{liquid}}=4.5$ and $\tau_{\mathrm{bubble}}=0.7$. We
-did not find any significant differences with from the results presented above. The
+did not find any significant differences from the results presented above. The
 comparison for the liquid film thicknesses is presented in Fig.
@@
 This work presents numerical tips and a benchmark for the
-Bretherton/Taylor problem using the binary liquid lattice Boltzmann method. To obtain correlations
-we specifically designed the benchmark. The bubble was chosen sufficiently long for the film
+Bretherton/Taylor problem using the binary liquid lattice Boltzmann method. The bubble was chosen sufficiently long for the film
 thickness to
-stabilize. Also, we used periodic boundary conditions for scheme robustness. Thus, the bubble
-train was simulated instead of one bubble motion. To avoid the mutual influence of neighboring
-bubbles a
-distance between bubbles was chosen to be $3$ bubble sizes. The computational
-results in terms of capillary number dependence and shape of a bubble show consistency with the
+stabilize, and periodic boundary conditions were used to keep the simulations robust.  A bubble
+train was simulated instead of the motion of a single bubble, and care was taken to minimize
+the mutual influence of neighboring bubbles. The computational
+results in terms of capillary number dependence and shape of the bubbles show consistency with the
 previously published data. Surprisingly, with large enough viscosity ratio the results are
-independent on inertia. Also, we examined the influence of grid resolution on the results. We found
-that if the phase interface is resolved as $50$ percent of the simulated film thickness, the
-simulations
-are convergent. Though  our results are specific to the binary liquid lattice
+independent of any inertial effects.  An examination of the influence of grid resolution on the results allowed
+us to determine that the phase interface should be resolved as at least $50$ percent of the simulated film thickness
+in order for the simulations to be convergent. Though our results are specific to the binary liquid lattice
 Boltzmann method, the numerical hints and procedures can be used for any

### 2011-01-06 Text/paper.tex
Added Mostowfi revisions. Jos revisions are not yet finished. Need to polish figures.

@@
 
-\title{The binary liquid model for gravity-driven microchannel simulations}
+\title{Simulations of gravity-driven flow of binary liquids in microchannels}
 
@@
 \ead{michalj@gmail.com}
-\author[uofa]{J.~Derksen}
-\ead{kuzmin@ualberta.ca}
 \author[schlum]{D.~Eskin}
@@
 \ead{fmostowfi@slb.com}
-\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
+\author[uofa]{J.~Derksen}
+\ead{jos@ualberta.ca}
 \address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
@@
 2V4 Canada}
+\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
 \address[schlum]{DBR Schlumberger Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
@@
 \begin{abstract}
-This work shows that the free-energy binary liquid lattice Boltzmann method is able to
-reproduce the gravity driven microchannel flow of a gas bubble through liquid. In the regime of low
-capillary
-numbers
-$\Ca$ (defined in Eq.~\ref{capillary:number:definition}), the flow is thoroughly described by
-Bretherton \cite{bretherton},
-with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
-from the Bretherton correlation for $\Ca\geq 0.005$, studied experimentally and numerically. This
-work aims to examine the applicability of the lattice Boltzmann binary liquid model to simulations
-of bubble flow in liquids for a range of $\Ca$ numbers. The binary
-liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
-It was found that the interface thickness needs to be $50$
-percent of the film thickness for the simulation to give convergent results. When
-the convergence criterion is fulfilled, the results are in a good agreement with the
-numerical simulations of \citet{giavedoni-numerical} in the range of $\Ca$ from $0.01$ to $1$. 
+%This work shows that the free-energy binary liquid lattice Boltzmann method is able to
+%reproduce the gravity driven microchannel flow of a gas bubble through liquid. In the regime of low
+%capillary
+%numbers
+%$\Ca$ (defined in Eq.~\ref{capillary:number:definition}), the flow is thoroughly described by
+%Bretherton \cite{bretherton},
+%with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
+%from the Bretherton correlation for $\Ca\geq 0.005$, studied experimentally and numerically. This
+%work aims to examine the applicability of the lattice Boltzmann binary liquid model to simulations
+%of bubble flow in liquids for a range of $\Ca$ numbers. The binary
+%liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
+%It was found that the interface thickness needs to be $50$
+%percent of the film thickness for the simulation to give convergent results. When
+%the convergence criterion is fulfilled, the results are in a good agreement with the
+%numerical simulations of \citet{giavedoni-numerical} in the range of $\Ca$ from $0.01$ to $1$. 
+In this work a free-energy binary liquid lattice-Boltzmann scheme is used to simulate
+Taylor/Bretherton flow in a micro-channel where elongated gas bubbles move through a liquid with
+thin liquid films between the bubbles and the channel walls. The numerical scheme has a diffuse
+interface, and a main focus of our work is to assess resolution requirements for correctly resolving
+the liquid film and bubble motion. The simulations are two-dimensional and span a capillary number
+range of $0.03$ to $0.8$ where the capillary number is based on the liquid’s dynamic viscosity, the
+velocity of the bubble, and the interfacial tension. The flow is driven by a body force, and
+periodic boundary conditions apply in the streamwise direction. We obtain grid independent results
+as long as the liquid film thickness is at least two times the width of the diffuse interface, with
+film thicknesses in accordance to literature results. We also show that the results in terms of film
+thicknesses are largely insensitive to the liquid-gas viscosity ratio and wettability parameters. 
 \end{abstract}
@@
 \section{Introduction}
-The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
-microchannels. It was found that the front meniscus film thickness is proportional
+The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
+narrow channels. It was found that the front meniscus film thickness is proportional
 to $\Ca^{2/3}$ for capillary numbers smaller than $0.003$, where the capillary number is defined as:
@@
 where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is the bubble velocity, and
-$\gamma$ is the surface tension between gas and liquid.
+$\gamma$ is the interfacial tension between gas and liquid.
 
-While there are a number
-of works which simulate the Bretherton problem by the boundary value approach
+While there are many works which simulate the Bretherton problem by the boundary value approach
 \cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
@@
 droplet breakup. The continuous interface models are more flexible for those kind of simulations.
-However, if the interface is spread over several grid nodes, the question of
+However, if the interface is smeared our over several grid nodes, the question of
 proper film resolution in comparison with the interface resolution arises.  This work is focused on
-resolving the parameter range of the binary liquid lattice
-Boltzmann method to validate it for a range of capillary numbers.
+exploring the parameter range of the binary liquid lattice
+Boltzmann method for correctly resolving the flow in a range of capillary numbers.
 
@@
 brings issues of the film thickness resolution versus interface
-resolution -- the diffuse interface should be resolved in such a way as to have
+resolution -- the diffuse interface should be dealt with in such a way as to have a
 negligible effect on the physics of the film.
 
-The binary liquid free-energy LB model \cite{swift} we used
+The binary liquid free-energy LB model due to \citet{swift} we used
 simulates two liquids with the assumption of uniform overall
@@
 neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
-negligible Reynolds number effect on the film thickness for a relatively wide range of Reynolds
+negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
 numbers.
@@
 binary liquid model parameters, the viscosity ratio and the grid
-resolution to obtain correct film thicknesses.
+resolution to obtain correct flow physics with an emphasis on film thicknesses.
 
-The paper is organized as follows.  First, we briefly
-explain the binary liquid lattice Boltzmann model and review the literature
-for the Bretherton problem. Then, the parameters involved in the
+The paper is organized as follows.  First, we review the literature
+for the Bretherton problem and briefly
+explain the binary liquid lattice Boltzmann model. Then, the parameters involved in the
 simulation are examined and presented in the results section. The paper is
@@
 moving in a tube filled with liquid. It was found that the deposition film thickness
-is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
+is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later,
 it was realized \cite{wong-films,wong-pressure} that the film thickness
@@
 \cite{kreutzer-pressure-drop} showed a deviation from the $\Ca^{2/3}$ rule and Reynolds number 
-effect on the film thickness for capillary numbers larger than $0.003$.
-To  predict consistently a flow pattern for capillaries in
+effects on the film thickness for capillary numbers larger than $0.003$.
+To consistently  predict a flow pattern for capillaries in
 different ranges of parameters, simulations have been conducted and have been typically validated with the
@@
 three-dimensional bubbles in
-polygonal capillaries and showed calculated different shapes in the
+polygonal capillaries and calculated bubble shapes in the
 slug cross sections and menisci appearance.
-\citet{heil-bretherton,ingham-plates} studied gas finger propagation in
+\citet{heil-bretherton} and \citet{ingham-plates} studied gas finger propagation in
 a two-dimensional channel for a range of Reynolds and capillary
@@
 simulations of two-dimensional channel flows for a relatively large capillary numbers, and
-the authors found discrepancies with the classical Bretherton theory, which
+found discrepancies with the classical Bretherton theory, which
 is limited to the low capillary number regime \cite{giavedoni-numerical}.
@@
 $0.003$ is computationally expensive even in the two-dimensional case
-(the amount of memory necessary to perform the simulation is of the order
+(the amount of memory necessary to perform the simulation would be of the order
 of tens or hundreds gigabytes).  This work is therefore based on a
@@
 &f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
-&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
+&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
 \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
@@
 &\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
--\partial_{\alpha}P_{\alpha \beta} +
+-\partial_{\beta}P_{\alpha \beta} +
 \nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
@@
 \end{equation}
-where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
-$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter.
+where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
+$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
+are the relaxation parameters of density and phase fields.
 
@@
 times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
-$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
-The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k A}{9}}$.
+$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
+to change viscosity from the gas viscosity
+$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
+$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
+accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
+A}{9}}$.
 
@@
  \item The diffuse nature of the interface in the multiphase model should not
-influence the film thickness results. This is obviously achieved only for certain
-interface resolutions relative to the film thickness. The same applies to spurious
+influence the film thickness results.  The same applies to spurious
 currents arising due to surface tension discretization, which
@@
   \item The Bretherton problem addresses a bubble flow driven by a pressure
-difference, not a body force. However, to the best authors' knowledge the pressure boundary
+difference, not a body force. However, to the best of our knowledge the pressure boundary
 conditions are still not yet developed for the binary liquid model, because it involves the
@@
 pressure difference are small as far as the film
- thickness is of interest.  In this work we limit ourselves to the study
+ thickness is concerned.  In this work we limit ourselves to the study
 of body force driven flows, because of their simplicity and better numerical stability. This
-case implies the periodic boundary conditions. As soon as the periodic boundary
-conditions are applied, not a single bubble but whole bubble train is simulated. In this case one
+implies that we can use the periodic boundary conditions in the streamwise direction. As soon as the
+periodic boundary
+conditions are applied, not a single bubble but a bubble train is simulated. In this case one
 needs
@@
  \item Most simulations are done for a channel with circular cross section,
-which is quite difficult to address in terms of lattice Boltzmann free energy
-framework, where one needs to calculate the Laplacian and gradients near boundaries.
+which is quite difficult to address in terms of lattice Boltzmann framework.
 Although one can apply certain finite difference stencils
 \cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
-the analysis of error introduced by the boundary conditions becomes an issue. Though the circular
+the analysis of error introduced by the boundary conditions becomes an issue. Although the circular
 axis-symmetric case can be simulated on a rectangular grid by introducing specific mass and force
@@
 the benchmark is chosen as a plane-symmetric two-dimensional case.  
-
 \end{itemize}
@@
 the initialization technique we used to determine the parameters necessary for
-the simulation. The capillary number defined in Eq.~\ref{capillary:number:definition} governs the interface thickness and is.
-The bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
-thickness is examined.
+the simulation. The capillary number defined in Eq.~\ref{capillary:number:definition} governs the
+interface thickness.
+%The bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
+%thickness is examined.
 While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
@@
 explicitly.
-In our simulations we approximately match and initialize the velocity with the
+In our simulations we approximate $U_{bubble}$ with the
 Poiseuille profile center velocity, which is given as:
@@
 where $N_y$ is the grid number in the vertical direction. As will be discussed further, the
-effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes.
+effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes to
+mimic no-slip conditions at the walls.
 For the Poiseuille profile one can estimate the body force (pressure gradient) through the
@@
 \end{equation}
-The body force (\ref{poiseuille:velocity:center}) causes the bubble
-velocity to differ from that implied by the capillary number because of the bubble presence in a flow. That is
-why the iteration proceduer is needed to achieve the 
+In simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
+velocity to differ from that implied by the Poiseuille relation because of the bubble presence in a
+flow. That is
+why an iteration procedure is needed to achieve the 
 desired velocity and 
@@
 
-After one calculates the parameters necessary to run a simulation, one has to
+After one calculates the parameters to run a simulation, one has to
 initialize the macroscopic fields and particle populations.
@@
 thickness values \cite{giavedoni-numerical}. This is done to minimize
-the time required for the simulation to be convergent, as starting far away from the steady-state requires
-additional time for the relaxation.
+the time  to convergence. %, as starting far away from the steady-state requires
+%additional time for the relaxation.
 \begin{figure}
@@
 430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
-\caption{The phase plots for different bubble thicknesses as
+\caption{Phase plots for different bubble thicknesses as
 $H_{\mathrm{eff}}-12$, $H_{\mathrm{eff}}-16$ and
@@
 system is initialized with different bubble volumes, the bubbles always relax
-to a shape where the film thickness is conserved. For the given parameters the
+to a shape where the film thickness is the same. For the given parameters ($Ca=0.05$) the
 interface thickness rescaled at the effective channel width $H_{\mathrm{eff}}$  is
@@
 
-Here, two different examples are presented that quantitatively show the
+Here, two different examples ($Ca=0.005$ and $Ca=0.05$) are presented that quantitatively show the
 initialization procedure for different capillary numbers. In these
@@
 A}{9}}=0.0377$ and the characteristic length of the interface
-$5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units. For the sake of simplicity and
+$5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units (grid nodes). For the sake of simplicity and
 stability, the relaxation times for liquid and gas phases are taken as
 $\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which correspond
-to the gas liquid kinematic viscosity ratio equal to $10$.
+to the liquid over gas kinematic viscosity ratio equal to $10$.
 \begin{description}
@@
 We take the bubble length equal to $5$ channel widths
-  and the distance between bubbles as $2$ bubble lengths. This gives a
+  and the distance between bubbles as $3$ bubble lengths. This gives a
 grid size of $400 \times 6000$. In terms of computer memory
@@
 \end{equation}
-Even using this naive assumption that the interface thickness is $62.5$ percent  of the
-film thickness, one can get the memory
+Even with an interface thickness of $62.5$ percent  of the
+film thickness, one gets (in 2D) a memory
 requirement of $345 \mathrm{MB}$.
@@
   \begin{equation}
-  U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{slug}}}=0.005
+  U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.005
 \frac{0.0377}{0.6666}=2.82 \cdot10^{-4},
   \end{equation}
-  where $\mu_{\mathrm{slug}}=1/3 (2.5-0.5)=0.6666$.
+  where $\mu_{\mathrm{liq}}=1/3 (2.5-0.5)=0.6666$.
   This velocity implies the following pressure gradient calculated from the
@@
   \end{equation}
-Note that the velocity is directly
-involved in calculation of the physical time step (small velocities imply longer running
+Note that the velocity is proportional to the physical time step (small velocities imply longer
+running
 simulations):
@@
 which is small. A typical simulation runs approximately
-$100000-300000$ time steps due to computational restrictions. That gives 
+$100000-300000$ time steps due to reach steady-state. That gives 
 physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
@@
   \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
-  \citet{giavedoni-numerical,heil-bretherton}, who cross validated their
+  \citet{giavedoni-numerical} and \citet{heil-bretherton}, who cross validated their
   simulations against the Bretherton data and each other. Even though the LBM can simulate
@@
 results for all ranges of capillary numbers.  Our intention is
-  to show physically justified behavior for moderate parameters,
+  to show physically correct behavior for moderate $Ca$,
 which
@@
   One can see from Fig.~\ref{fig:heil:planar} that the film thickness for the
-  capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
-  number from $0$ to $300$.
+  capillary number $0.05$ ranges from $0.12$ to $0.15$ in a wide Reynolds
+  number range (from $0$ to $300$).
 
-  The film thickness for such small capillary numbers is taken from Fig.~\ref{fig:heil:planar},
+  The film thickness for such capillary number is taken from Fig.~\ref{fig:heil:planar},
   and for Reynolds number $40$ equals $0.12$. We want to
-  resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
+  resolve $6\%$ of the whole channel width with $12$ lattice units, where $5$
 lattice units are related to the interface resolution.  Therefore, the channel
@@
   \end{equation}
-  Our approach to the initial conditions is  to take the flow parameters from the previously
-performed
-numerical simulations and to calculate the grid size from them, matching the
-body force gradient through the solution for Poiseuille flow.
+  %Our approach to the initial conditions is  to take the flow parameters from the previously
+%performed
+%numerical simulations and to calculate the grid size from them, matching the
+%body force gradient through the solution for Poiseuille flow.
 
@@
 \caption{\citet{giavedoni-numerical} gathered results across the
-literature for different capillary numbers. Note that results are scaled to the half-width of the
-channel. \label{fig:giavedoni:planar}}
+literature for different capillary numbers (reprinted from \cite{giavedoni-numerical}). Results are
+scaled to the half-width of the
+channel. $h_{\infty}$ is the film thickness of air finger at
+the infinite distance from the front meniscus. \label{fig:giavedoni:planar}}
 \includegraphics[width=0.97\textwidth]{Figures/heil-planar.eps}
 \caption{\citet{heil-bretherton} performed simulations for different ranges of
-capillary numbers and Reynolds numbers. Note that results are scaled to the half-width of the channel.
+capillary numbers and Reynolds numbers (reprinted from \cite{heil-bretherton}). Results
+are scaled to the half-width of the channel. $h_{0}$ is the film thickness of air finger at
+the infinite distance from the front meniscus.
 \label{fig:heil:planar}}
@@
 where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
-is related to the effective channel height.
-If the grid size in the vertical direction is $N_y$, then one has $N_y-1$ regions between the grid nodes which represent
+is the effective channel height.
+If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions between the grid nodes
+which represent
 the physical domain. The effective wall location is in the middle between bounce-back and fluid nodes giving
@@
 simplification to impose the boundary in the middle between the bounce-back
-node and the fluid node. The location of the wall for simple hydrodynamic flows is
+node and the fluid node. The location of the wall is
 viscosity dependent \cite{ginzburg-multireflection}.  Note that the effective location of the
@@
  the interface needs to be resolved as $40-50$ percent of the
-expected film thickness for simulations to be convergent. We further examine the
+expected film thickness for simulations to be grid independent. We further examine the
 velocities in the center of the bubble to calculate the capillary number. One can see from Table
-\ref{table:parameters:grid:refinement} that the bubble velocities are pretty consistent and the
+\ref{table:parameters:grid:refinement} that the bubble velocities are consistent and the
 calculated capillary number corresponding to these velocities is $0.07$, which is slightly
-larger than the expected capillary number. The corresponding error can be
-attributed to the Poiseuille profile body force initialization and the
- density and viscosity ratios, which are much larger in the
-Bretherton problem.
+larger than the capillary number we aimed for. The corresponding difference can be
+attributed to the Poiseuille profile body force initialization.% and the
+ %density and viscosity ratios, which are much larger in the
+%Bretherton problem.
 \begin{figure}
@@
 corresponds to the wall location. The profiles were taken at $x=14$ (nondimensional coordinates).
-Capillary number is $\Ca=0.05$. Capillary number obtained from simulations $\Ca=0.07$. One can
-see
-convergent results even for the underresolved film
-thickness.\label{fig:grid:profiles}}
+Capillary number is $\Ca=0.05$. The capillary number obtained from simulations $\Ca=0.07$. 
+%One can seeconvergent results even for the underresolved film
+%thickness.
+\label{fig:grid:profiles}}
 \end{figure}
@@
 \end{tabularx}
-\caption{The parameters and results for grid resolution simulations. Simulated domain is
+\caption{The parameters and results for grid resolution. The simulated domain is
 of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the velocity of the bubble center measured at
@@
 through the phase gradient \cite{pooley-contact}, $\partial_n \phi$.  We took a large
-enough grid resolution in order for simulations to be consistent -- the grid size was
+enough grid in order for simulations to be consistent -- the grid size was
 $177 \times 2626$ and the initial film thickness was
@@
 We examined $11$ different values for the wall
-gradient ranging from $-1$ to $1$. The parameters are summarized in Table
+gradient ranging from $-1$ to $1$. The parameters and results are summarized in Table
 \ref{table:parameters:wall:gradient}. The results are consistent for all the
@@
 and $1.0$ the
-simulations are unstable. One can see the calculated profiles in Fig.
-\ref{fig:gradients:profiles}. The latter shows that the negative wall gradient values cause 
+simulations are unstable. One can see a few of the calculated phase profiles in Fig.
+\ref{fig:gradients:profiles}. It shows that the negative wall gradient values cause 
 non-physical phase values (above $1.0$ near the wall). One should attribute them to the numerical
-adjustment of wall gradients through a first-order finite difference scheme. Therefore, they do
-not spoil numerical results. Note that
+adjustment of wall gradients through a first-order finite difference scheme. Note that
 negative values of the phase gradient are
@@
 with the bubble which has phase value $-1$.  In case of positive
-gradients the values near the wall are below $1$, and the gradient profile can
-fuse with the values of the gas phase. This case corresponds to the slug flow when gas is in
+gradients, the values near the wall are below $1$, and the gradient profile can
+fuse with the values of the gas phase. This case corresponds to  slug flow when gas is in
 contact with the wall.
@@
 One can see that if the film thickness is resolved properly the influence of the wall gradient can
-be neglected. The velocities of the bubble center for certain profiles along $x$ direction are
+be neglected. The velocities of the bubble center for profiles along $x$ direction are
 presented in Table
@@
 \hline
-$\scriptstyle U_{b}$ &$\scriptstyle 0.0041$& $\scriptstyle 0.0042$& $\scriptstyle 0.0042$
+$\scriptstyle U_{bubble}$ &$\scriptstyle 0.0041$& $\scriptstyle 0.0042$& $\scriptstyle 0.0042$
 &$\scriptstyle 0.0042$ & $\scriptstyle 0.0041$& $\scriptstyle 0.0041$ & $\scriptstyle 0.0042$ &
@@
 \end{tabularx}
-\caption{The parameters and results for wall gradient simulations. $U_{b}$ stands for the
+\caption{The parameters and results for wall gradient effects on the film thickness. $U_{bubble}$
+stands for the
 center bubble velocity. $\delta$ is the interface thickness. $\partial_n \phi$ is the phase
@@
 calculated at nondimensional $x=11.42$ after $200 000$ time iterations. $\Ca=0.05$ was taken to
-initialize simulation.
+initialize the simulations.
 \label{table:parameters:wall:gradient}}
@@
 The purpose of this section is to validate the correlations of
-\citet{giavedoni-numerical}, Fig. \ref{fig:giavedoni:planar}, and
-\citet{heil-bretherton}, Fig. \ref{fig:heil:planar}, for a range of capillary
+\citet{giavedoni-numerical} in Fig. \ref{fig:giavedoni:planar} and
+\citet{heil-bretherton} in Fig. \ref{fig:heil:planar} for a range of capillary
 numbers. Because of limited computational resources, we skip the
@@
 is a computationally reasonable task.  To be consistent with the previous sections, we choose the grid to be
-$202 \times 3001$. $5$ lattice Boltzmann units do not occupy more than $60$
-percent of the effective thickness. Because the simulation gets unstable with
+$202 \times 3001$. Then $5$ lattice Boltzmann units do not occupy more than $60$
+percent of the effective film thickness. Because the simulation gets unstable with
 smaller grids and larger gradients, all the capillary number simulations were
-performed on the same grid. To properly initialize the force gradient, the
+performed on the same grid. To properly initialize the body force, the
 proportionality law was utilized. The forcing
@@
 capillary numbers are overpredicted but the obtained thicknesses are overpredicted
-as well due to the improper body force setting caused by the Poiseuille flow assumption. If one
+as well due to the body force setting caused by the Poiseuille flow assumption. If one
 needs to obtain the desired capillary number, the shooting method for the forcing is

### 2011-01-06 Text/paper.tex
typo fix: our -> out

@@
 droplet breakup. The continuous interface models are more flexible for those kind of simulations.
-However, if the interface is smeared our over several grid nodes, the question of
+However, if the interface is smeared out over several grid nodes, the question of
 proper film resolution in comparison with the interface resolution arises.  This work is focused on

### 2011-01-07 Text/paper.tex
Need only to change figures from desktop computer and submit it.

@@
 \ead{fmostowfi@slb.com}
-\author[uofa]{J.~Derksen}
+\author[uofa]{J.J.~Derksen}
 \ead{jos@ualberta.ca}
@@
 channel widths
-$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_\mathrm{eff}$ and $\delta=0$
+$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_{\mathrm{eff}}$ and $\delta=0$
 corresponds to the wall location. The profiles were taken at $x=14$ (nondimensional coordinates).
@@
 
-Next, we investigated the qualitative variation of the film thickness over the bubble length.
+Next, we investigated the variation of the film thickness over the bubble length.
 The comparison for the thickness variation is presented in Fig.~\ref{fig:sehgal:bubble:length}.
@@
 \includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_length.eps}\\
-\caption{The qualitative comparison for the film thickness across
+\caption{A qualitative comparison for the film thickness across
 the bubble length. Left (courtesy of \citet{sehgal-microchannel}),
@@
 \subsection{The influence of different viscosities}
-The results presented above are taken for the liquid viscosity ratio $10$.
-We performed the same simulations for the liquid-gas viscosity ratio of $20$.
+The results presented above are taken for the liquid over gas kinematic viscosity ratio of $10$.
+We performed the same simulations for the liquid over gas viscosity ratio of $20$.
 The relaxation parameters were taken as $\tau_{\mathrm{liquid}}=4.5$ and $\tau_{\mathrm{bubble}}=0.7$. We
@@
 \ref{fig:capillary:viscous}. The results validate our assumption that in
-the low capillary flow regime the density ratio does not affect the film thickness,
-and that the gas-liquid viscosity ratio of the order $10$ is sufficient to obtain results
+the low capillary flow regime that the sufficiently high gas over liquid viscosity ratio ($\geq 10$) is sufficient to obtain results
 consistent with the literature data.
@@
 \caption{The film thicknesses versus capillary number for
-$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=20$. The
-results show good consistency and mutual agreement.\label{fig:capillary:viscous}}
+$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=20$. \label{fig:capillary:viscous}}
 \end{figure}
@@
 \section{Conclusion}
-This work presents numerical tips and a benchmark for the
+This work presents numerical studies of the
 Bretherton/Taylor problem using the binary liquid lattice Boltzmann method. The bubble was chosen sufficiently long for the film
@@
 independent of any inertial effects.  An examination of the influence of grid resolution on the results allowed
-us to determine that the phase interface should be resolved as at least $50$ percent of the simulated film thickness
-in order for the simulations to be convergent. Though our results are specific to the binary liquid lattice
+us to determine that the phase interface should be resolved as at least $50$ percent of the film thickness
+in order for the simulations to be grid independent. Though our results are specific to the binary liquid lattice
 Boltzmann method, the numerical hints and procedures can be used for any
 continuous interface method.
+\section{Acknowledgement}
+A.~Kuzmin wants to thank Schlumberger for their financial support.
 \bibliographystyle{model1-num-names}

### 2011-01-07 Sailfish/Capillary/capillary.py
All revisions are finished.

@@ -106,7 +106,7 @@ def Analyze_Simulations():
     pylab.ylabel(r'''$\delta$''',fontsize=30)
     
     #labels=[r'''$H_{eff}='''+str(value-2)+r'''$''' for value in ny]
-    pylab.legend(["Giavedoni","Heil","Simulations"],loc=4)
+    pylab.legend(["Giavedoni","Heil","This work"],loc=4)
     #pylab.xlim(xmax=15)
     pylab.savefig("capillaries_comparison.eps",format="EPS",dpi=300)
 
@@ -216,8 +216,8 @@ def Analyze_Velocities():
 
 if __name__=="__main__":
     
-    #Analyze_Simulations()    
+    Analyze_Simulations()    
     #Analyze_Velocities()
     #Run_Simulations()
-    Analyze_Bubble()
+    #Analyze_Bubble()
     pylab.show()

### 2011-01-07 Sailfish/Wall_Gradient/gradient.py

@@ -25,12 +25,14 @@ def Get_Zero(prof):
 
 def Analyze_Simulations():
     print os.getcwd()
+    chosen=[-0.6,0.0,0.6]
     gradient=[-1.0,-0.8,-0.6,-0.4,-0.2,0.0,0.2,0.4,0.6,0.8,1.0]
     str_gradient=['-10','-8','-6','-4','-2','0','2','4','6','8','10']
     style=["bo","rH","g--","k-.","c<-.","y>","bs","g^","rD"]
+    style_chosen=["--","-.",":"]
     fig=pylab.figure()
-    for i in range(0, len(str_gradient)-2):
-        dir_temp="Results/Grad"+str_gradient[i]
+    for i in range(0, len(chosen)):
+        dir_temp="Results/Grad"+str_gradient[gradient.index(chosen[i])]
         os.chdir(dir_temp)
         name="grad180000.npz"
         array=numpy.load(name)
@@ -40,7 +42,7 @@ def Analyze_Simulations():
         prof_domain=prof[1:len(prof)-1]
         x_delta=1.0/len(prof_domain)
         x=numpy.arange(0.5*x_delta,1.0,x_delta)
-        pylab.plot(x,prof_domain,style[i],linewidth=3,markersize=10)
+        pylab.plot(x,prof_domain,style_chosen[i],linewidth=4,markersize=10)
         #pylab.yticks([0,25,50,75,100])
         #pylab.savefig("../../initfinish"+str(width[i])+".eps",format="EPS",dpi=70)
         
@@ -61,15 +63,17 @@ def Analyze_Simulations():
     pylab.xlim(xmax=0.141)
     pylab.xlabel(r'''$\delta$''',fontsize=30)
     pylab.ylabel(r'''$\phi$''',fontsize=30)
-    pylab.legend([r'''$\partial_n\phi=$'''+r'''$'''+str(x)+r'''$''' for x in gradient[:-2]])
+    pylab.legend([r'''$\partial_n\phi=$'''+r'''$'''+str(x)+r'''$''' for x in chosen])
     pylab.xticks(fontsize=20)
     pylab.yticks(fontsize=20)
     pylab.savefig("phase_grad_profiles.eps",format="EPS",dpi=300)
 
 def Analyze_Velocities():
     print os.getcwd()
+    
     gradient=[-1.0,-0.8,-0.6,-0.4,-0.2,0.0,0.2,0.4,0.6,0.8,1.0]
[…]

### 2011-03-22 Review/answers.tex
Started the review corrections. A lot of work.

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{color}
\usepackage{bm}
\usepackage[sort&compress]{natbib}
\usepackage[normalem]{ulem}	% Part of the standard distribution

\begin{document}
\textbf{Question:}
\begin{quotation}
The main quantity that the paper focuses on is the thickness of the film between the bubbles and the
wall. As shown by Figure 8 and as mentioned by the Authors in Section 4, the film thickness varies
along the bubble. However, the Authors do not explain the choice of the location of the measurement
of the film thickness. This choice is of primary importance when comparing results with those of
the
literature and in particular with works on finger propagation in which the film thickness is
measured
”at infinity”.
\end{quotation}

\textbf{Answer:} The following section is added to the text 
\begin{quotation}
We chose to measure the film thickness in the middle of the bubble. The
classical Bretherton formulation is for the measurement of the film thickness at ``infinity''.
However, \citet{cerro-bubble-train} indicate that for the gravity-driven bubble train flow in tubes
it is enough for bubble to have a length of two-three diameters to apply analysis for very long
bubbles. In comparison with the axisymmetric circular geometry, for three-dimensional square shaped
microchannels \citet{heil-threedim} indicate the measurement of the film thickness at $5.5$
channel heights from the front bubble tip. This distance was shown sufficient for the
film thickness to establish itself for the capillary number range $Ca<4$. We chose to measure
the film thickness in the middle of the bubble, which is located at least at the distance of $2.5-3$
channel heights from the front bubble tip.  Moreover, the film thickness examination, see Section
\ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
$0.08\leq Ca \leq 1$, the standard deviation from the bubble middle film thickness is around $7$
 percent for small capillary numbers ($Ca=0.05$) and less than $1$ percent for larger ones
($Ca \geq 0.1$) at the distances ranging from one channel height from the rear tip upto one channel
height to the front tip. This certainly validates the point of choice to measure bubble thickness.
\end{quotation}

\textbf{Question:}
\begin{quotation}
It may be appropriate to introduce Figure 8 on thickness variation before the main results on ﬁlm
thickness.
\end{quotation}

\textbf{Answer:} We agree and moved the section before the results on the capillary number
dependency.

\textbf{Question:}
\begin{quotation}
The lines 54-55 indicating that ”numerical simulations and experimental studies showed [...]
Reynolds
number effects on the ﬁlm thickness for capillary numbers larger than 0.003” is in contradiction
with
the lines 32-33 which state ”negligible Reynolds number effects on the film thickness for a
relatively
wide range of Reynolds numbers.”
\end{quotation}

\textbf{Answer:} The following paragraph is added:
\begin{quotation}
Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
range of Reynolds
numbers. 
{\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
$300$. The indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
significantly the pressure distrubution and the flow field near the front bubble tip. In
the present simulations the largest Reynolds number we achieve is less than $20$ and it allows to
neglect the
inertia effects.}

\citet{kreutzer-pressure-drop} showed a deviation from the $Ca^{2/3}$ rule
{\color{red}\sout{and Reynolds number 
effects on the film thickness}} for capillary numbers larger than $0.003$.
\end{quotation}

\textbf{Question:}
\begin{quotation}
Although inertia has been shown to have a minor effect on ﬁlm thickness, inertia does influence
some flow variables in the Bretherton’s problem (see Ref. 3 for instance). To check whether inertia
is important or not, and to discuss the assumption of uniform density, the Authors should provide
the reader with the Reynolds numbers associated with each simulation. This is also required for the
consistency of Figure 7 in which results are compared with other works of the literature.
\end{quotation}

\textbf{Answer:} Thank you for pointing it out. We included the discussions of the Reynolds
effects in the answer to the question above. As well, we added Reynolds numbers to the Table for
the capillary range regime and the grid depedency study.

\textbf{Question:}
\begin{quotation}
In this work, the diphasic flow is driven by a body force. The body force is therefore the control
parameter in the simulations. The Authors should give more detail about this body force. Indeed,
although the estimation of the body force from the pressure gradient is well described, it is not
clear
where this body force appears in the governing equations (Equations 2, 3 and 4). Does this body
force correspond to the variable $F_i$ in Equation 2? Should not this force or its equivalent appear
in
the Navier-Stokes equations (Equation 4)?
\end{quotation}

\textbf{Answer:} We updated equation (4) to include the force contribution. As well, the definition
of $P_{\alpha\beta}$ is added. $F_i$ is the force population which corresponds to the force
inclusion $\bm{F}$. We added the citation which describes how the force population is formed and
mimics the force behavior.

\textbf{Question:}
\begin{quotation}
I would suggest to calculate the Bond number to check if the shapes of the bubbles and the main
results of the paper are not inﬂuenced.
\end{quotation}

\textbf{Answer:} 
The maximum Bond number $Bo=\frac{\rho g N_y^2}{\gamma}$ in the
current simulation equals to $7$, which is less than the number when the shape of the bubble is
affected \cite{zheng-large-ratio}. However, the point is that the equation of state in the lattice
Boltzmann framework is connected with density. Thus usually either pressure (i.e. density) boundary
conditions are prescribed or the body force. If they are prescribed simultaneously then
we can consider the influence of the Bond number on the simulations. Otherwise prescribing just the
body force is analogous to prescribing the pressure gradient $\frac{\mathrm{d}P}{\mathrm{d}x}$ in
the Navier-Stokes equation. Thus, it's not possbile the Bond number effect. It is because of the
coupled nature of the pressure and density in the LBM framework.

\textbf{Question:}
The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
estima
tions of the body force to apply and of the ﬁlm thickness to choose grid resolution. The
presentation
of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
again described in a diﬀerent way in Section 5-4. Similar equations for the calculation of the
pressure
gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
into one unless they explain their differences. In fact, it would probably be straightforward not to
detail the calculation of estimated variables since they are intermediary variables which allow the
design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
say they control and impose directly the body force without describing all the estimation procedure.

\textbf{Answer:} Thank you for the suggestion. While we agree that there is redundant information
about initialization and this was taken into the account by reducing repetitive text:
\begin{quotation}
\end{quotation}

 However, the
dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
the lattice Boltzmann are refered to the non-dimensionalization procedure. Therefore, we decided to
keep equation 16 as the most informative for the LBM community and to reduce {\color{red} equation
14 and 6}.
\bibliographystyle{unsrtnat}
\bibliography{answers}

\end{document}

### 2011-03-22 Text/paper.tex
Started the review corrections. A lot of work.

@@
 neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
-negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
+{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
+range of Reynolds
 numbers. 
+{\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
+are negligible for $\Ca\leq0.05$ and have moderate changes for $\Ca>0.05$ in the range of Reynolds
+number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
+$300$. The indicate that while the influence of the established film thickness is insignificant
+($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
+significantly the pressure distrubution and the flow field near the front bubble tip. In the present
+simulations the largest Reynolds number we achieve is less than $20$ and it allows to neglect the
+inertia effects.} 
 Therefore, the major governing factor for microchannel flows is not the density ratio, but the
@@
 simulations \cite{giavedoni-numerical} and experimental studies
-\cite{kreutzer-pressure-drop} showed a deviation from the $\Ca^{2/3}$ rule and Reynolds number 
-effects on the film thickness for capillary numbers larger than $0.003$.
+\cite{kreutzer-pressure-drop} showed a deviation from the $\Ca^{2/3}$ rule for capillary numbers
+larger than $0.003$.
 To consistently  predict a flow pattern for capillaries in
@@
 where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
-relaxation parameter, and $F_i$ is the external force population.
+relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
+is connected with the force and is implemented in the current work due to work \cite{guo}.}
 
@@
 &\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
-&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
+&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}= F_{\alpha}
 -\partial_{\beta}P_{\alpha \beta} +
@@
 $M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
-are the relaxation parameters of density and phase fields.
+are the relaxation parameters of density and phase fields, {\color{red}
+$P_{\alpha\beta}=\Bigl(p_0-k\phi \Delta \phi -\frac{k}{2}|\nabla \phi|^2\Bigr)\delta_{\alpha\beta}
++ k \partial_{\alpha} \phi \partial_{\beta} \phi$  \cite{pooley-contact}.}
 
@@
 interface thickness rescaled at the effective channel width $H_{\mathrm{eff}}$  is
-$0.0701$,$0.0697$,$0.0692$ measured at the dimensionless coordinate $x=13$.
+$0.0701$,$0.0697$,$0.0692$ measured at the {\color{red} center of the bubble}.
 \label{fig:different:initialization:widths}}
@@
 
-Here, two different examples ($Ca=0.005$ and $Ca=0.05$) are presented that quantitatively show the
-initialization procedure for different capillary numbers. In these
+{\color{red}
+For demonstration purposes we present the simulation parameters choice for $\Ca=0.05$.
+In these
 examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
@@
 to the liquid over gas kinematic viscosity ratio equal to $10$.
+
+The procedure can be described as follows:
 \begin{description}
- \item[I \bm{$Ca=0.005$}]
+ \item[Capillary number choice] Blah-blah
+ \item[Film thickness] You choose film thickness. 
+ \item[Grid choice]
+ \item[Body force] Blah-blah
+\end{description}
+
 The film thickness for such small capillary numbers \cite{bretherton} is
@@
 
+ The film thickness for such capillary number is taken from Fig.~\ref{fig:heil:planar},
+  and for Reynolds number $40$ equals $0.12$. We want to
+  resolve $6\%$ of the whole channel width with $12$ lattice units, where $5$
+lattice units are related to the interface resolution.  Therefore, the channel
+  width is $200$ lattice units and the grid size is $200 \times 3000$.
+
 We take the bubble length equal to $5$ channel widths
@@
 
- \item[II $\bm{Ca=0.05}$] 
    The parameters in the previous subsection imply that long
-computations are required in order for the system to come to equilibrium. The simulation becomes longer with a decrease of the
+computations are required in order for the system to come to equilibrium. The simulation becomes
+longer with a decrease of the
 capillary number. Bretherton obtained an analytical solution for the range of low capillary
@@
 
-  The film thickness for such capillary number is taken from Fig.~\ref{fig:heil:planar},
-  and for Reynolds number $40$ equals $0.12$. We want to
-  resolve $6\%$ of the whole channel width with $12$ lattice units, where $5$
-lattice units are related to the interface resolution.  Therefore, the channel
-  width is $200$ lattice units and the grid size is $200 \times 3000$. The velocity can be
+  The velocity can be
 calculated from the capillary number as:
@@
   \end{equation}
-  %Our approach to the initial conditions is  to take the flow parameters from the previously
-%performed
-%numerical simulations and to calculate the grid size from them, matching the
-%body force gradient through the solution for Poiseuille flow.
-
-\end{description}
+
+}
+
 \begin{figure}
@@
 To properly estimate the interface resolution one needs to study the convergence as a function of
-the grid resolution. To do that the grid resolution is varies while all remaining parameters,
+the grid resolution. To do that the grid resolution is varied while all remaining parameters,
 including the bubble velocity and the capillary number, are fixed.  Our goal is to determine the 
@@
 the physical domain. The effective wall location is in the middle between bounce-back and fluid nodes giving
-overall $N_y-2$ nodes representing the fluid. Note that it is a big
-simplification to impose the boundary in the middle between the bounce-back
+overall $N_y-2$ nodes representing the fluid. Note that it is a simplification to impose the
+boundary in the middle between the bounce-back
 node and the fluid node. The location of the wall is
-viscosity dependent \cite{ginzburg-multireflection}.  Note that the effective location of the
+viscosity dependent \cite{ginzburg-multireflection}.  The effective location of the
 wall for the
@@
 The bubble is initialized as a rectangular box with coordinates
-$y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
+{\color{red}$y=7\frac{N_y-2}{100}\dots N_y-7\frac{N_y-2}{100}-1$}, $x=\frac{N_x}{3}\dots \frac{2
+N_x}{3}$ and phase
 $\phi_{\mathrm{bubble}}=-1$. All other nodes are initialized with the phase field
@@
 \ref{table:parameters:grid:refinement} that the bubble velocities are consistent and the
-calculated capillary number corresponding to these velocities is $0.07$, which is slightly
-larger than the capillary number we aimed for. The corresponding difference can be
+calculated capillary number corresponding to these velocities is $0.047$, which is {\color{red}
+close to} the capillary number we aimed for. The corresponding difference can be
 attributed to the Poiseuille profile body force initialization.% and the
@@
 corresponds to the wall location. The profiles were taken at $x=14$ (nondimensional coordinates).
-Capillary number is $\Ca=0.05$. The capillary number obtained from simulations $\Ca=0.07$. 
+Capillary number is $\Ca=0.05$. The capillary number obtained from simulations $\Ca=0.047$. 
 %One can seeconvergent results even for the underresolved film
@@
 \begin{table}
-\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|}
-\hline
-$N_x$&$N_y$&$\delta$&$U_\mathrm{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_\mathrm{iter}$\\
-\hline
-$1500$&$102$&$0.0694$&$0.0041$&$0.824$&$200000$\\
-\hline
-$1875$&$127$&$0.0688$&$0.0041$&$0.646$&$250000$\\
-\hline
-$2250$&$152$&$0.0676$&$0.0040$&$0.539$&$300000$\\
-\hline
-$2625$&$177$&$0.0679$&$0.0040$&$0.453$&$350000$\\
-\hline
-$3000$&$202$&$0.0668$&$0.0041$&$0.400$&$400000$\\
-\hline
-$3375$&$227$&$0.0663$&$0.0039$&$0.355$&$450000$\\
+\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|X|}
+\hline
+$N_x$&$N_y$&$\delta$&$U_\mathrm{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_\mathrm{iter}$&{
+\color{red} $Re$}\\
+\hline
+$1500$&$102$&$0.0694$&$0.0028$&$0.824$&$200000$&$0.44$\\
+\hline
+$1875$&$127$&$0.0690$&$0.0028$&$0.646$&$250000$&$0.53$\\
+\hline
+$2250$&$152$&$0.0675$&$0.0027$&$0.539$&$300000$&$0.63$\\
+\hline
+$2625$&$177$&$0.0674$&$0.0027$&$0.453$&$350000$&$0.73$\\
+\hline
+$3000$&$202$&$0.0662$&$0.0027$&$0.400$&$400000$&$0.83$\\
+\hline
+$3375$&$227$&$0.0664$&$0.0026$&$0.355$&$450000$&$0.91$\\
 \hline
@@
 \caption{The parameters and results for grid resolution. The simulated domain is
-of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the velocity of the bubble center measured at
-$x=14$.  $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of the film in
+of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the velocity of the interface at the fron tip of
+the bubble.  $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of
+the film in
 lattice Boltzmann units. $Ca=0.05$ was taken to initialize the simulation.
@@
 gradient near the wall responsible for hydrophilic and hydrophobic behavior. The results are
-calculated at nondimensional $x=11.42$ after $200 000$ time iterations. $\Ca=0.05$ was taken to
+calculated at the middle of the bubble after $200 000$ time iterations. $\Ca=0.05$ was taken to
 initialize the simulations.
@@
 \end{table}
+
+\subsection{The film variation over the bubble}
+\label{section:film:thickness:variation}
+Next, we investigated the variation of the film thickness over the bubble length.
+The comparison for the thickness variation is presented in Fig.~\ref{fig:sehgal:bubble:length}.
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_sehgal.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_length.eps}\\
+\caption{A qualitative comparison for the film thickness across
+the bubble length. Left (courtesy of \citet{sehgal-microchannel}),
+right (present simulations); $x$ is scaled to $H_{\mathrm{eff}}$
+and increases in the flow direction.\label{fig:sehgal:bubble:length}}
+\end{figure}
+One can see a qualitative agreement, i.e. the thickness increases towards the front meniscus and
+rapidly decreases towards the rear meniscus. This shape is sometimes referred as a "bullet"
+shape.
 
@@
 The results obtained after $2\cdot10^5$ steps are presented in Table
-\ref{table:parameters:capillary:number}. One can see that the calculated
+\ref{table:parameters:capillary:number}. 
+{\color{red} We chose to measure the film thickness in the middle of the bubble. The
+classical Bretherton formulation is for the measurement of the film thickness at ``infinity''.
+However, \citet{cerro-bubble-train} indicate that for the gravity-driven bubble train flow in tubes
+it is enough for bubble to have a length of two-three diameters to apply analysis for very long
+bubbles. In comparison with the axisymmetric circular geometry, for three-dimensional square shaped
+microchannels \citet{heil-threedim} indicate the measurement of the film thickness at $5.5$
+channel heights from the front bubble tip. This distance was shown sufficient for the
+film thickness to establish itself for the capillary number range $\Ca<4$.
+\citet{giavedoni-numerical} showed that the film stabilizes at the distances as $2.6-4.0$ diameters
+from the front tip depending on the Reynolds number.  We chose to measure the film thickness in the
+middle of the bubble, which is located at least at the distance of $2.5-3$ channel heights from the
+front bubble tip. Moreover, the film thickness examination, see Section
+\ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
+$0.05\leq \Ca \leq 1$, the standard deviation from the bubble middle film thickness is around $7$
+percent for small capillary numbers ($\Ca=0.05$) and less than $1$ percent for larger ones ($\Ca
+\geq 0.1$) at the distances ranging from one channel height from the rear tip upto one channel
+height to the front tip. This certainly validates the point of choice to measure bubble thickness.}
+One can see that the calculated
 capillary numbers are overpredicted but the obtained thicknesses are overpredicted
@@
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/Capillary/capillaries_comparison.eps}
+\includegraphics[width=0.97\textwidth]{Figures/Capillary/capillaries_comparison_real.eps}
 \caption{A comparison between simulation results and results of
@@
 \hline
-$Ca_{\mathrm{lit}}$&$\delta_{\mathrm{lit}}$&$x$&$\delta$&$U_b$&$Ca$\\
-\hline
-$0.03$&$0.04$&$10.0$&$0.042$&$0.0022$&$0.039$\\
-\hline
-$0.05$&$0.06$&$11.0$&$0.066$&$0.0041$&$0.072$\\
-\hline
-$0.08$&$0.08$&$13.0$&$0.085$&$0.0068$&$0.120$\\
-\hline
-$0.1$&$0.1$&$12.5$&$0.077$&$0.0055$&$0.098$\\
-\hline
-$0.2$&$0.12$&$5.5$&$0.123$&$0.0185$&$0.328$\\
-\hline
-$0.4$&$0.13$&$4.7$&$0.153$&$0.0388$&$0.686$\\
-\hline
-$0.6$&$0.15$&$3.0$&$0.164$&$0.0592$&$1.046$\\
-\hline
-$0.8$&$0.16$&$1.7$&$0.173$&$0.0782$&$1.383$\\
+$Ca_{\mathrm{lit}}$&$\delta_{\mathrm{lit}}$&$\delta$&$U_{bubble}$&$Ca$&{\color{red}$Re$}\\
+\hline
+$0.03$&$0.04$&$0.040$&$0.0014$&$0.026$&$0.449$\\
+\hline
+$0.05$&$0.06$&$0.058$&$0.0027$&$0.047$&$0.820$\\
+\hline
+$0.08$&$0.08$&$0.085$&$0.0045$&$0.080$&$1.378$\\
+\hline
+$0.1$&$0.1$&$0.076$&$0.0037$&$0.065$&$1.126$\\
+\hline
+$0.2$&$0.12$&$0.122$&$0.0125$&$0.222$&$3.807$\\
+\hline
+$0.4$&$0.13$&$0.151$&$0.0271$&$0.479$&$8.222$\\
+\hline
+$0.6$&$0.15$&$0.164$&$0.0416$&$0.736$&$12.617$\\
+\hline
+$0.8$&$0.16$&$0.172$&$0.0559$&$0.989$&$16.960$\\
 \hline
@@
 $\delta_{\mathrm{lit}}$ is the film thickness with corresponding $\Ca_{\mathrm{lit}}$ taken from
-literature. $\delta_\mathrm{sim}$ is the simulation film thickness with corresponding $\Ca$. $U_b$ is the
-center bubble velocity taken at location $x$. $\Ca$ is based on the measured center bubble velocity
-$U_{b}$.
+literature. $\delta_\mathrm{sim}$ is the simulation film thickness with corresponding $\Ca$.
+{\color{red}$U_{bubble}$ is the interface velocity at the center axis.} $\Ca$ is based on
+the measured center bubble velocity
+$U_{bubble}$.
 \label{table:parameters:capillary:number}}
 \end{table}
-
-\subsection{The film variation over the bubble}
-
-Next, we investigated the variation of the film thickness over the bubble length.
-The comparison for the thickness variation is presented in Fig.~\ref{fig:sehgal:bubble:length}.
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_sehgal.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_length.eps}\\
-\caption{A qualitative comparison for the film thickness across
-the bubble length. Left (courtesy of \citet{sehgal-microchannel}),
-right (present simulations); $x$ is scaled to $H_{\mathrm{eff}}$
-and increases in the flow direction.\label{fig:sehgal:bubble:length}}
-\end{figure}
-One can see a qualitative agreement, i.e. the thickness increases towards the front meniscus and
-rapidly decreases towards the rear meniscus. This shape is sometimes referred to as a "bullet" shape.
 

### 2011-03-22 Sailfish/Capillary/capillary.py

@@ -50,7 +50,9 @@ def Analyze_Simulations():
     #fig=pylab.figure()
     widths=[]
     velocities=[]
-    
+    velocities_real=[]
+    re_real=[]
+    widths_real=[]
     #read the giavedoni data (not precise though)
     giavedoni=genfromtxt("planarcasesolution.csv",delimiter=',',dtype=float)[1:]
 
@@ -64,6 +66,8 @@ def Analyze_Simulations():
         name="capillary200000.npz"
         array=numpy.load(name)
         prof=array['phi'][:,exam[i]]
+        dims=array['phi'].shape
+        
         #x=numpy.arange(0.0,float(ny[i]))/ratio
         #pylab.imshow(array['phi'])
     
@@ -71,8 +75,22 @@ def Analyze_Simulations():
             #pylab.plot(prof)
             widths.append(Get_Zero(prof))
             vel=array['v'][0]
+            center=array['phi'][dims[0]/2,:]
+           
+            z1 = numpy.min(numpy.where(center < 0.0))
+            z2 = numpy.max(numpy.where(center < 0.0))
+            if z1==0:
+                z2=numpy.min(numpy.where(center>0.0))+dims[1]
+                z1=numpy.max(numpy.where(center>0.0))
+            print z1,z2
+            
+            prof_real=array['phi'][:,((z1+z2)/2)%dims[1]]
+            widths_real.append(Get_Zero(prof_real))     
+
             vel_prof=vel[:,exam[i]]
             velocities.append(vel_prof[len(vel_prof)/2])
+            velocities_real.append(vel[dims[0]/2,z2%dims[1]])
+            re_real.append(vel[dims[0]/2,z2%dims[1]]*dims[0]/(2.0/3.0))
         
         #pylab.plot(x[0:20],prof[0:20],color[i],linewidth=3)
         #pylab.plot(x,prof,style_diff[i],markersize=10,linewidth=3)
@@ -87,9 +105,16 @@ def Analyze_Simulations():
     
     fig=pylab.figure()
     capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
+    capillaries_real=numpy.array(velocities_real)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
+
     print "Widths=",widths
+    print "Widths_real=",widths_real
     print "Capillaries=",capillaries
[…]

### 2011-03-22 Sailfish/Grid/grid.py

@@ -51,7 +51,17 @@ def Analyze_Simulations():
         #pylab.figure()
         name="grid"+str(200000+i*50000)+".npz"
         array=numpy.load(name)
-        prof=array['phi'][:,int(1400*ratio)]
+        center=array['phi'][ny[i]/2,:]
+        z1 = numpy.min(numpy.where(center < 0.0))
+        z2 = numpy.max(numpy.where(center < 0.0))
+        if z1==0:
+            z2=numpy.min(numpy.where(center>0.0))+nx[i]
+            z1=numpy.max(numpy.where(center>0.0))
+        print z1,z2
+
+        
+        #prof=array['phi'][:,int(1400*ratio)]
+        prof=array['phi'][:,((z1+z2)/2)%nx[i]]
         prof_domain=prof[1:len(prof)-1]
         delta_x=1.0/float(ny[i]-2)
         x_domain=numpy.arange(0.5*delta_x,1.0,delta_x)
@@ -82,7 +92,7 @@ def Analyze_Velocities():
     print os.getcwd()
     ny=[102,127,152,177,202,227]
     nx=[1501,1876,2251,2626,3001,3376]
-
+    
     for i in range(0, len(ny)):
         dir_temp="Results/"+str(ny[i])
         ratio=float(ny[i]-2)/100.0
@@ -91,16 +101,30 @@ def Analyze_Velocities():
         name="grid"+str(200000+50000*i)+".npz"
         array=numpy.load(name)
         ux=array['v'][0]
+        center=array['phi'][ny[i]/2,:]
+        z1 = numpy.min(numpy.where(center < 0.0))
+        z2 = numpy.max(numpy.where(center < 0.0))
+        if z1==0:
+            z2=numpy.min(numpy.where(center>0.0))+nx[i]
+            z1=numpy.max(numpy.where(center>0.0))
+        print z1,z2
+
         #print ux.shape
         prof=ux[:,int(1400*ratio)]
+        prof_real=ux[:,((z1+z2)/2)%nx[i]]
         #pylab.figure()
         #pylab.plot(prof)
         print prof[len(prof)/2]
+        print ux[ny[i]/2,((z1+z2)/2)%nx[i]]
+        print "Interface velocity=",ux[ny[i]/2,z2%nx[i]]        
+        print "Reynolds=", ny[i]*ux[ny[i]/2,z2%nx[i]]/(2.0/3.0)
+        #pylab.figure()
+        #pylab.plot(center)
[…]

### 2011-03-22 Text/paper.tex
Continuing with the revisions.

@@
 
-The procedure can be described as follows:
+Overall, the setting simulations procedure can be described in steps as follows:
 \begin{description}
- \item[Capillary number choice] Blah-blah
- \item[Film thickness] You choose film thickness. 
- \item[Grid choice]
- \item[Body force] Blah-blah
+ \item[Capillary number] One first needs to be supplied with the capillary number for
+simulations. For the demonstration purposes we chose $\Ca=0.05$. 
+ \item[Film thickness] After the capillary number is prescribed, one needs to approximate the film
+thickness $\delta$. It can be done either by taking correlations from numerical simulations by
+\citet{heil-bretherton} and \citet{giavedoni-numerical} or by using the classical Bretherton
+correlation for small cappillary numbers $\delta=0.6687 \Ca^{2/3}$. In the case of $\Ca=0.05$, the 
+film thickness is taken from {\color{green} Fig.~\ref{fig:heil:planar}},
+  and for Reynolds number $40$ equals $0.06$.  
+\item[Grid choice] After pecifying the film thickness, one needs to choose the associate
+number of nodes to resolve the film thickness. Note that for paremeters indicated above the
+interface thickness is spread over approximately $5$ lattice units. The study of the interface
+resolution to the film thickness, Section \ref{section:grid:refinement}, suggests to choose the
+film thickness to be $2-2.5$ times larger than the interface thickness. Therefore, we want to
+  resolve $6\%$ of the whole channel width with $12$ lattice units.  Thus, the channel
+  width is $200$ lattice units. We take the bubble length equal to $5$ channel widths for the film
+thickness to establish
+  and the distance between bubbles as $3$ bubble lengths to avoid mutual influence of bubbles on
+each other. Thus, the whole grid size is $200 \times 3000$.
+ \item[Velocity] The relaxation parameters are taken to be in the stable range for simulations.
+In the present simulations the viscosity ratio is $\dfrac{\mu_{liq}}{\mu_{gas}}=10$ with the
+correspoding relaxation times $\tau_{liq}=2.5$ and $\tau_{gas}=0.7$. Given the viscosity of the
+liquid one can obtain the velocity of the bubble from the capillary number: 
+\begin{equation}
+\begin{aligned}
+&U_{\mathrm{bubble}}=\Ca \frac{\gamma}{\mu_{\mathrm{liq}}}=\Ca \frac{\sqrt{\frac{8 k
+A}{9}}}{\frac{1}{3}(\tau_{\mathrm{liq}}-\frac{1}{2})}\\
+&U_{\mathrm{bubble}}=0.05 \frac{0.0377}{0.6666}=2.82 \cdot10^{-3},
+\end{aligned}
+\end{equation}
+where $\mu_{\mathrm{liq}}=1/3 (2.5-0.5)=0.6666$.
+\item[Body force] 
+  The velocity implies the following pressure gradient calculated from the
+  Poiseuille flow solution:
+  \begin{equation}
+  \begin{aligned}
+  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma \Ca=3.77\cdot10^{-7}.
+  \end{aligned}
+  \end{equation}
+
+After resolving the interface one can  calculate the
+  velocity from the capillary number as:
+  This velocity implies the following pressure gradient calculated from the
+  Poiseuille flow solution, Eq. \ref{poiseuille:velocity:center}: 
+  \begin{equation}
+  \begin{aligned}
+  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
+  \end{aligned}
+  \end{equation}
+Note that the velocity is proportional to the physical time step (small velocities imply longer
+running
+simulations):
+\begin{equation}
+\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
+\end{equation}
+where, for example, the real velocities in the physical world can reach
+approximately $1$ m/s and the real height up to $1$ mm. In this case one
+can easily approximate the physical time step during one lattice Boltzmann iteration:
+\begin{equation}
+\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
+\end{equation}
+which is small. A typical simulation runs approximately
+$100000-300000$ time steps due to reach steady-state. That gives 
+physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
+probably not enough for the physical system to come to the pseudo-steady state.
+
 \end{description}
 
-The film thickness for such small capillary numbers \cite{bretherton} is
-given by the equation:
-  \begin{equation}
-  \frac{h_{\infty}}{H_{\mathrm{eff}}}=0.6687 \Ca^{2/3}=0.0194.
-  \end{equation}
-Thus, the film
- occupies about $2$ percent of the whole channel height. Let us impose that $8$ lattice
-units are required to resolve  $2$ percent of the channel. Note that $5$ out of $8$ these lattice
-units
-($62.5$ percent) are used to resolve the phase interface. Therefore the channel width is $400$
-lattice units.
-
- The film thickness for such capillary number is taken from Fig.~\ref{fig:heil:planar},
-  and for Reynolds number $40$ equals $0.12$. We want to
-  resolve $6\%$ of the whole channel width with $12$ lattice units, where $5$
-lattice units are related to the interface resolution.  Therefore, the channel
-  width is $200$ lattice units and the grid size is $200 \times 3000$.
-
-We take the bubble length equal to $5$ channel widths
-  and the distance between bubbles as $3$ bubble lengths. This gives a
+This gives a
 grid size of $400 \times 6000$. In terms of computer memory
@@
 requirement of $345 \mathrm{MB}$.
-
-After resolving the interface one can  calculate the
-  velocity from the capillary number as:
-  \begin{equation}
-  U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.005
-\frac{0.0377}{0.6666}=2.82 \cdot10^{-4},
-  \end{equation}
-  where $\mu_{\mathrm{liq}}=1/3 (2.5-0.5)=0.6666$.
-  This velocity implies the following pressure gradient calculated from the
-  Poiseuille flow solution, Eq. \ref{poiseuille:velocity:center}: 
-  \begin{equation}
-  \begin{aligned}
-  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
-  \end{aligned}
-  \end{equation}
-Note that the velocity is proportional to the physical time step (small velocities imply longer
-running
-simulations):
-\begin{equation}
-\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
-\end{equation}
-where, for example, the real velocities in the physical world can reach
-approximately $1$ m/s and the real height up to $1$ mm. In this case one
-can easily approximate the physical time step during one lattice Boltzmann iteration:
-\begin{equation}
-\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
-\end{equation}
-which is small. A typical simulation runs approximately
-$100000-300000$ time steps due to reach steady-state. That gives 
-physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
-probably not enough for the physical system to come to the pseudo-steady state.
 
@@
   number range (from $0$ to $300$).
-
-  The velocity can be
-calculated from the capillary number as:
-  \begin{equation}
-  U_{\mathrm{bubble}}=0.05 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.05 \frac{0.037712}{0.6666}=2.82\cdot
-  10^{-3}.
-  \end{equation}
-  The velocity implies the following pressure gradient calculated from the
-  Poiseuille flow solution:
-  \begin{equation}
-  \begin{aligned}
-  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma \Ca=3.77\cdot10^{-7}.
-  \end{aligned}
-  \end{equation}
 
@@
 \subsection{Grid refinement}
+\label{section:grid:refinement}
 To properly estimate the interface resolution one needs to study the convergence as a function of

### 2011-03-25 Text/paper.tex
Adding details for revisions.

@@
 \subsection{The nondimensionalization and initialization procedure}
+\label{section:init}
 We now introduce the procedure of nondimensionalization, as well as
@@
 \item[Body force] 
-  The velocity implies the following pressure gradient calculated from the
-  Poiseuille flow solution:
+Thus, it is desired to obtain the prescribed velocity $U_{\mathrm{bubble}}$ in the simulations. The
+flow in the simulations is conducted by imposing the body force. Thus, the assumption is needed of
+how the body force is connected with the bubble velocity $U_{\mathrm{bubble}}$. We assume that the
+flow is close to the Poiseuille flow:
 \begin{equation}
-  \begin{aligned}
-  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma \Ca=3.77\cdot10^{-7}.
-  \end{aligned}
+\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8\mu}{H_{\mathrm{eff}}^2}
+U_{\mathrm{bubble}}=\frac{8}{N_y^2}\gamma \Ca=3.77\times 10^{-7}.  
 \end{equation}
-
-After resolving the interface one can  calculate the
-  velocity from the capillary number as:
-  This velocity implies the following pressure gradient calculated from the
-  Poiseuille flow solution, Eq. \ref{poiseuille:velocity:center}: 
-  \begin{equation}
-  \begin{aligned}
-  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
-  \end{aligned}
-  \end{equation}
-Note that the velocity is proportional to the physical time step (small velocities imply longer
-running
-simulations):
-\begin{equation}
-\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
-\end{equation}
-where, for example, the real velocities in the physical world can reach
-approximately $1$ m/s and the real height up to $1$ mm. In this case one
-can easily approximate the physical time step during one lattice Boltzmann iteration:
-\begin{equation}
-\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
-\end{equation}
-which is small. A typical simulation runs approximately
-$100000-300000$ time steps due to reach steady-state. That gives 
-physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
-probably not enough for the physical system to come to the pseudo-steady state.
-
+After imposing the body force the simulations can be run. A typical simulation runs approximately
+$100000-300000$ time steps due to reach steady-state.
+%Note that the velocity is proportional to the physical time step (small velocities imply longer
+%running
+%simulations):
+%\begin{equation}
+%\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
+%\end{equation}
+%where, for example, the real velocities in the physical world can reach
+%approximately $1$ m/s and the real height up to $1$ mm. In this case one
+%can easily approximate the physical time step during one lattice Boltzmann iteration:
+%\begin{equation}
+%\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
+%\end{equation}
+%which is small.  That gives 
+%physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
+%probably not enough for the physical system to come to the pseudo-steady state.
 \end{description}
-
-This gives a
-grid size of $400 \times 6000$. In terms of computer memory
-requirements, the calculations with double precision accuracy demand:
-\begin{equation}
-\begin{aligned}
-&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
-} \times \,\, 9 \text{
-populations}\\
-&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
-\end{aligned}
-\end{equation}
-Even with an interface thickness of $62.5$ percent  of the
-film thickness, one gets (in 2D) a memory
-requirement of $345 \mathrm{MB}$.
-
-   The parameters in the previous subsection imply that long
-computations are required in order for the system to come to equilibrium. The simulation becomes
-longer with a decrease of the
-capillary number. Bretherton obtained an analytical solution for the range of low capillary
-numbers $Ca \leq 0.005$. However, many researchers extended it to the range of capillary numbers
+}
+Note that following the prescribed procedure one can see that a computer memory requirements are
+high for small capillary numbers due to large grids to resolve the film thickness. Thus, the
+simulations in the present work cover range of capillary numbers from $0.05$ to $1.0$.
+Although, Bretherton obtained an analytical solution for the range of low capillary
+numbers $Ca \leq 0.005$, many researchers, for example \citet{giavedoni-numerical} and
+\citet{heil-bretherton}, extended it to the range of capillary numbers
 greater than
-$0.005$, and to higher Reynolds numbers. One can refer to figures
-  \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
-  \citet{giavedoni-numerical} and \citet{heil-bretherton}, who cross validated their
-  simulations against the Bretherton data and each other. Even though the LBM can simulate
-flows with small capillary numbers, i.e. less than $0.005$, the simulation time
-is relatively long. Therefore, in our work we focus on the range of capillary numbers $\Ca$
-employed by the simulations of
-\citet{giavedoni-numerical} who cross validated
+$0.005$, and to higher Reynolds numbers. %One can refer to figures
+%  \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
+%  \citet{giavedoni-numerical} and \citet{heil-bretherton}, who cross validated their
+% simulations against the Bretherton data and each other.
+%Even though the LBM can simulate
+%flows with small capillary numbers, i.e. less than $0.005$, the simulation time
+%is relatively long. 
+Therefore, the focus of our work on the capillary range larger $\Ca\geq0.05$ is legitimate as far
+as we base it on comparison with already cross validated
 results for all ranges of capillary numbers.  Our intention is
@@
   can be easily and quickly validated on a computer.
-  One can see from Fig.~\ref{fig:heil:planar} that the film thickness for the
-  capillary number $0.05$ ranges from $0.12$ to $0.15$ in a wide Reynolds
-  number range (from $0$ to $300$).
-
-}
-
+%One can see from Fig.~\ref{fig:heil:planar} that the film thickness for the
+%  capillary number $0.05$ ranges from $0.12$ to $0.15$ in a wide Reynolds
+% number range (from $0$ to $300$).
+%This gives a
+%grid size of $400 \times 6000$. In terms of computer memory
+%requirements, the calculations with double precision accuracy demand:
+%\begin{equation}
+%\begin{aligned}
+%&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
+%} \times \,\, 9 \text{
+%populations}\\
+%&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
+%\end{aligned}
+%\end{equation}
+%Even with an interface thickness of $62.5$ percent  of the
+%film thickness, one gets (in 2D) a memory
+%requirement of $345 \mathrm{MB}$. 
 \begin{figure}
@@
 \end{figure}
+
 Note that we base the initialization techniques on correlations for the
@@
 recalculate all the necessary flow characteristics, i.e. adjust the
- applied force to obtain the desired velocity/capillary number.
-
+ applied force to obtain the desired velocity/capillary number. However, simulations show that the
+Poiseuille flow assumptions work reasonably well.
+
+{\color{red}
+Note that the initialization procedure is a good guide of designing the simulations. For example
+for the grid refinement, see Section \ref{section:grid:refinement}, we keep the same velocity
+$U_{\mathrm{bubble}}$ and $\Ca$. This implies the conservation of the quantity
+$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
+x } = \mathrm{const}$. For the designing capillary range simulations, we use a proportionality law
+to initialize body forces knowing results of just one simulation:
+\begin{equation}
+\begin{aligned}
+&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
+&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} H_{\mathrm{eff}}^2\\
+&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} H_{\mathrm{eff}}^2 \text{ or }\\
+&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{H_{\mathrm{eff}}^2},
+\end{aligned}
+\end{equation}
+where the subscript ,,lit'' stands for the predicted capillary number
+\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
+involved because all simulations are conducted on the same grid.
+}
 \subsection{Grid refinement}
@@
 
-After choosing the reference parameters, the grid refinement procedure
-needs to keep the macroscopic parameters constant.  It is easy to check
-that it yields the following quantity to
-be independent of grid size
-$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
-x } = \mathrm{const}$. We performed a number of simulations for the prescribed grids. The
+{\color{red} After choosing the reference parameters, the grid refinement procedure
+needs to keep the macroscopic parameters constant. One can refer to Section \ref{section:init}.}  
+%It is easy to check
+%that it yields the following quantity to
+%be independent of grid size
+%$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
+%x } = \mathrm{const}$. 
+We performed a number of simulations for the prescribed grids. The
 simulation results in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
@@
 \subsection{Capillary number region}
+\label{section:capillary:range}
 The purpose of this section is to validate the correlations of
@@
 $6 \cdot 10^{-6}/16$ was chosen to obtain the predicted capillary
-number $0.05$. The forcing for other grids can be obtained using
-simple proportionality relationships:
-\begin{equation}
-\begin{aligned}
-&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
-&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\
-&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
-&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{N_y^2},
-\end{aligned}
-\end{equation}
-where the subscript ,,lit'' stands for the predicted capillary number
-\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
-involved because all simulations are conducted on the same grid.
+number $0.05$. 
 The pressure gradient can be obtained through the capillary number

### 2011-03-26 Review/answers.tex
Working with corrections.

@@
 \textbf{Question:}
+\begin{quotation}
 The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
@@
 say they control and impose directly the body force without describing all the estimation procedure.
-
-\textbf{Answer:} Thank you for the suggestion. While we agree that there is redundant information
-about initialization and this was taken into the account by reducing repetitive text:
-\begin{quotation}
 \end{quotation}
 
- However, the
+\textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
+about initialization and this was taken into the account by reducing repetitive text and
+introducing step-by-step explations, see Section about initialization. We also added the
+recommendations for controlling parameters in the same section.  The
 dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
@@
 the lattice Boltzmann are refered to the non-dimensionalization procedure. Therefore, we decided to
-keep equation 16 as the most informative for the LBM community and to reduce {\color{red} equation
-14 and 6}.
+keep suggestions for controlling and initialization of simulations as informative guides for the
+LBM community.
+
+\textbf{Question:}
+\begin{quotation}
+ Moreover, the estimations of the pressure gradient for the initialisation of the scheme are
+obtained
+under the assumption of an axisymmetric Poiseuille ﬂow. However, among the ”Key words”, one can
+read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
+that are done in this work. The Authors should discuss this choice of estimating the pressure
+gradient with the axisymmetric assumption instead of a Hele-Shaw flow (factor 1/12 instead of 1/8
+in Equation 5).
+\end{quotation}
+
+\textbf{Answer:} While the Hele-Shaw flow implies the flow with the infinitisemal distance between
+plates, in our situation the size of the bubble can be compared with the channel height ($5$ times
+larger) and that's why the assumption of the infinitisemal distance between plates is too strict.
+Moreover, after recalculating results with the velocity of the interface the assumption of the
+Poiseulle flow is valid and shows good accuracy.
+
+\textbf{Question:}
+\begin{quotation}
+The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
+journal of applied science. The Authors would improve signiﬁcantly the manuscript by deﬁning more
+rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
+lines 106, 111 and 123 via the parameter k, the weights w and the quantity 8kA/9 respectively.
+A suggestion is to emphasize within an array the links between physical parameters and the model
+eq
+parameters. The variables i, fieq , gi should be deﬁned and the Equations for f and g in Equation
+3 should be described (which equation corresponds to continuity equation...). Equation 2 could
+perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
+physical variable f of Equation 3. The notation i = 1 ÷ 8 is unusual, it would be clearer to use a
+classical notation such as 1 ≤ i ≤ 8. The sound speed c2 should be deﬁned just after its
+introduction in line 107.
+\end{quotation}
+
+\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the physical world
+parameters with the lattice Botlzmann parameters. As well more explanations are put towards the
+interface surface tension description and inclusion it to the model. The reference to the
+continuity and the Navier-Stokes equations are added. The equation (2) is not rewritten in terms of
+generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
+governing equations. The notations are corrected.
+
+\textbf{Question:}
+\begin{quotation}
+The name of the Section ”Lattice Boltzmann benchmark” lets the reader think that this Section
+would describe the numerical procedure (assumptions, choice of the geometry and the grid, boundary
+conditions, control parameters, initialization of the scheme). This is not the case because this
+Section
+develops the challenges and the difficulties of using the Lattice Boltzmann method to bubble ﬂows.
+This Section should be renamed or reorganized.
+\end{quotation}
+
+\textbf{Answer:} We renamed the section to ``Approach to benchmark''.
+
+\textbf{Question:}
+\begin{quotation}
+The Authors should make a distinction between variables in lattice units and dimensional or non
+dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y − 2$ where $N_y$ is
+the
+number of nodes in the vertical direction indicates that $H_{eff}$ has lattice units. However, this
+is in
+contradiction with the use of $H_{eff}$ in Equation 5. The Authors should deﬁne each variable
+without
+any ambiguity on its units. Besides, the unit of $U_{bubble}$ should also be given. These
+improvements
+would ﬁt well in the Subsection named ”The nondimensionalization and initialization procedure”
+which does not in fact introduce much nondimensionalisation procedure.
+\end{quotation}
+
+\textbf{Answer:} We added a decription paragraph in the Subsection ``The nondimensionalization and
+initialization procedure'' about the approach to units. In fact, all equations are the same in the
+lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
+preferred units as soon as capillary number is matched in the LBM and the physical world.
+
+\textbf{Question:}
+\begin{quotation}
+Some notations are unclear. For instance, the ﬁlm thickness is denoted as h∞ in Equation 7 whereas
+it is denoted δ in most of the paper. The notations w or δsim are also found before line 381 or in
+Table 3 respectively. The thickness of the interface is denoted ξ in Figure 1 but the same variable
+is
+denoted 5ξ in Table 1. The speed of the bubble is denoted Ububble in some places and Ub in others.
+The variable Uin Equation 5 does not seem to be deﬁned.
+\end{quotation}
+
+\texbf{Answer:} We thoroughly corrected all the inconsistencies.
+
 \bibliographystyle{unsrtnat}

### 2011-03-26 Text/paper.tex
Working with corrections.

@@
 \end{equation}
-where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
+where $f_i$ is the probability distribution function in the direction $\bm{c_i}$,
+{\color{red} $f_i^{q}$ is the equilibrium probability distribution function}, $\omega$ is the
 relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
@@
 &&&+k\bigl(w_i^{xx} (\partial_x \phi)^2+w_i^{yy} (\partial_y \phi)^2 +w_i^{xy} \partial_x
-\phi \partial_y \phi \bigr), i=1\div8\\
+\phi \partial_y \phi \bigr), 1\leq i \leq 8\\
 &f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
 &g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
-\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), 1\leq i \leq 8 \\
 &g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
@@
 tension; $A$ is the parameter of the free-energy model. The bulk pressure
-is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
+is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$ {\color{red} with
+the sound speed $c_s^2=1/3$}. 
 Parameters specific to the D2Q9 grid are the weights
@@
 \frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
-$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
-the sound speed parameter $c_s^2=1/3$.  The weights related to the
-inclusion of the surface tension coefficient into the equations are as follows:
+$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$.  
+{\color{red}
+Other weights 
+%related to the
+%inclusion of the surface tension coefficient into the equations 
+are as follows:}
 $w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
@@
 $P_{\alpha\beta}=\Bigl(p_0-k\phi \Delta \phi -\frac{k}{2}|\nabla \phi|^2\Bigr)\delta_{\alpha\beta}
-+ k \partial_{\alpha} \phi \partial_{\beta} \phi$  \cite{pooley-contact}.}
-
-The system allows the separation of the liquid
++ k \partial_{\alpha} \phi \partial_{\beta} \phi$  \cite{pooley-contact}.The interface tension value
+in the framework of the binary liquid model is $\gamma=\sqrt{\frac{8 k
+A}{9}}$. The inclusion of the interface tension to the momentum flux tensor is done through the
+coefficients $k$, $A$ and weights $w_i^{\alpha\beta}$.
+
+Note that the first equation of system \ref{standard:implementation} simulates the continuity and
+the Navier-Stokes equations, first two equations in (\ref{binary:fluid:system}). The second equation
+of system \ref{standard:implementation} simulates the phase governing equation, third equation in
+(\ref{binary:fluid:system}).} The system (\ref{binary:fluid:system}) allows the separation of the
+liquid
 phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
@@
 $\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
-accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
-A}{9}}$.
-
-\section{Lattice Boltzmann benchmark}
+accordingly.
+
+{\color{red}
+While the lattice Boltzmann system has parameters as gas and liquid
+viscosities, the surface tension and etc., those parameters are not the representative and
+proportional quantities of the parameters in a physical system. The parameters of the lattice
+Boltzmann are connected with the physical world's parameters only through the non-dimensional
+numbers governing a physics of the problem. In this case it's a capillary number $\Ca$, which is
+obtained from the physical world and then it is matched through the lattice Boltzmann quantities.
+In fact, one should distinguish the lattice Boltzmann system and the physical system. The set of
+the fluid and phase equations \eqref{binary:fluid:system} is valid in the lattice Boltzmann space
+and in the physical domain. Therefore, one can substitute any quantity, i.e.
+$U_{\mathrm{bubble}}$, in the physical units or in the lattice Boltzmann units as soon as the
+capillary number is matched in the both of worlds.}
+
+%\section{Lattice Boltzmann benchmark}
+\section{Approach to benchmark}
 To properly simulate and compare simulation results with the data
@@
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/benchmark.eps}
+\includegraphics[width=0.97\textwidth]{Figures/benchmark_new.eps}
 \caption{The benchmark sketch. $\delta$ corresponds to film thickness. $\xi$ corresponds to the interface thickness. We take the microchannel length to be 15 times larger than its height.
@@
 \begin{aligned}
-U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H_{\mathrm{eff}}^2}{8
+U_{\mathrm{bubble}} \approx \frac{H_{\mathrm{eff}}^2}{8
 \mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
@@
 Note that the initialization procedure is a good guide of designing the simulations. For example
-for the grid refinement, see Section \ref{section:grid:refinement}, we keep the same velocity
+for the grid refinement, Section \ref{section:grid:refinement}, we keep the same velocity
 $U_{\mathrm{bubble}}$ and $\Ca$. This implies the conservation of the quantity
 $H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
-x } = \mathrm{const}$. For the designing capillary range simulations, we use a proportionality law
-to initialize body forces knowing results of just one simulation:
+x } = \mathrm{const}$. For the designing capillary range simulations, Section
+\ref{section:capillary:range}, we use a proportionality law to initialize body forces knowing
+results of just one simulation:
 \begin{equation}
@@
 where the subscript ,,lit'' stands for the predicted capillary number
-\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
-involved because all simulations are conducted on the same grid.
+\cite{giavedoni-numerical,heil-bretherton}. 
 }
@@
 the film in
-lattice Boltzmann units. $Ca=0.05$ was taken to initialize the simulation.
+lattice Boltzmann units. $\Ca=0.05$ was taken to initialize the simulation.
 \label{table:parameters:grid:refinement}}
@@
 \end{tabularx}
-\caption{The parameters and results for wall gradient effects on the film thickness. $U_{bubble}$
-stands for the
-center bubble velocity. $\delta$ is the interface thickness. $\partial_n \phi$ is the phase
+\caption{The parameters and results for wall gradient effects on the film thickness.
+$U_{\mathrm{bubble}}$
+stands for the interface velocity. $\delta$ is the film thickness. $\partial_n \phi$ is the phase
 gradient near the wall responsible for hydrophilic and hydrophobic behavior. The results are
@@
 The pressure gradient can be obtained through the capillary number
-ratio:
+ratio, Section \ref{section:init}:
 \begin{equation}
@@
 \begin{equation*}
-w=12 \frac{\Ca_{\mathrm{lit}}}{0.05}
+12 \frac{\Ca_{\mathrm{lit}}}{0.05}
 \end{equation*}
@@
 \hline
-$Ca_{\mathrm{lit}}$&$\delta_{\mathrm{lit}}$&$\delta$&$U_{bubble}$&$Ca$&{\color{red}$Re$}\\
+$Ca_{\mathrm{lit}}$&$\delta_{\mathrm{lit}}$&$\delta$&$U_{\mathrm{bubble}}$&$Ca$&{\color{red}$Re$}\\
 \hline
@@
 $\delta_{\mathrm{lit}}$ is the film thickness with corresponding $\Ca_{\mathrm{lit}}$ taken from
-literature. $\delta_\mathrm{sim}$ is the simulation film thickness with corresponding $\Ca$.
-{\color{red}$U_{bubble}$ is the interface velocity at the center axis.} $\Ca$ is based on
-the measured center bubble velocity
-$U_{bubble}$.
+literature. $\delta$ is the simulation film thickness with corresponding $\Ca$.
+{\color{red}$U_{\mathrm{bubble}}$ is the interface velocity at the center axis.} $\Ca$ is based on
+the measured bubble tip velocity
+$U_{\mathrm{bubble}}$.
 \label{table:parameters:capillary:number}}

### 2011-03-29 Review/answers.tex
Started producing high quality vector plots.

@@
 
+\textbf{Question:}
+\begin{quotation}
+Three ﬁgures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
+unusual
+for an article which is not a review. I encourage the Authors to remove at least the two ﬁrst ones
+which are not essential.
+\end{quotation}
+
+\textbf{Answer:} We eliminated necessary pictures and references to them.
+
+\textbf{Question:} 
+\begin{quotation}
+In Figure 7, the points of data reported from the numerical work of Ref. 3 exhibit a surprisingly
+noisy trend that corresponds more to an experimental study. The Authors should explain how they
+obtain this data and plot them.
+\end{quotation}
+
+\textbf{Answer:} We added a description to the figure caption explaining data acquisition.
+
+\textbf{Question:} 
+\begin{quotation}
+The plots of Figure 8 are of less quality than the other ﬁgures of the paper.
+\end{quotation}
+
+\textbf{Answer:} Unfortunately, it was not possible to extract data as a picture from the paper by
+\citet{sehgal-microchannel}. Thus, the curves were extracted from the graphs and represented in a
+new figure scaled to the present simulations for comparison reasons.
+
+\textbf{Question:}
+\begin{quotation}
+The work of Yang seems to be close from the present study since it involves the
+solution of Bretherton
+ﬂow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
+differences between their work and this paper.
+\end{quotation}
+
+\textbf{Answer:} We outlined the differences in the introduction section.
+
+\textbf{Question:} It would be interesting if the Authors could mention a study in which a level-set
+method is used in
+order to complete the bibliography.
+
+\textbf{Answer:} We added references to the introduction section.
+
+\textbf{Questions:} 
+\begin{quotation}
+A secondary quantity of interest is the bubble speed or in a non-dimensional
+form the capillary
+number. This quantity is measured at the center of the bubbles. The Authors should describe how
+this quantity as well as the center of the bubble are measured.
+\end{quotation}
+
+\textbf{Answer:} Thank you for pointing that. The classical Bretherton problem
+$U_{\mathrm{bubble}}$ is the bubble velocity, i.e. the interface velocity. We recalculate all the
+results taking into the consideration the interface velocity at the bubble tip. Qualitatively, the
+results are the same. It was found that Poiseuille profile assumption is improved significantly
+after recalculation.
+
+\textbf{Question:} 
+\begin{quotation}
+ If possible, the Authors should present results showing that steady state is reached. It would
+also
+be interesting if the Authors show some patterns of the ﬂow by plotting for instance the vector ﬁeld
+of velocity around the bubbles.
+\end{quotation}
+
+\textbf{Answer:} We added additional section which shows the change of flow pattern.
+
+\textbf{Question:}
+\begin{quotation}
+In the Abstract, the Authors should give the range of real capillary numbers they measured instead
+of the targeted values (Calit ).
+\end{quotation}
+
+\textbf{Answer:} We changed it to the measured capillary numbers.
+
+\textbf{Question:} 
+\begin{quotation}
+The Authors should discuss more their results (Sections 5-4, 5-5, 5-6). Did the Authors measure the
+pressure drops at the front and rear interfaces of the bubbles?
+\end{quotation}
+
 \bibliographystyle{unsrtnat}

### 2011-03-29 Text/paper.tex
Started producing high quality vector plots.

@@
 the liquid film and bubble motion. The simulations are two-dimensional and span a capillary number
-range of $0.03$ to $0.8$ where the capillary number is based on the liquid’s dynamic viscosity, the
+range of $0.05$ to $1.0$ where the capillary number is based on the liquid’s dynamic viscosity, the
 velocity of the bubble, and the interfacial tension. The flow is driven by a body force, and
@@
 found discrepancies with the classical Bretherton theory, which
-is limited to the low capillary number regime \cite{giavedoni-numerical}.
+is limited to the low capillary number regime \cite{giavedoni-numerical}. Authors didn't take into
+the account recent studies extending microchannel simulations for the capillary numbers
+beyond the Bretherton regime. Thus, limiting comparison with other established CFD methods. Author
+used the Shan-Chen model \cite{Shan-chen:extended} to simulate the multiphase behavior. It is also
+the continuous interface model, but the study of grid dependence is absent.
+
+{\color{red}
+Level set method is represented by a number of works \cite{fukugata-levelset,
+lakehal-micro, carlson-comparison}. While the bubble shapes and  were found in a good agreeement
+with the experiments, the range of studied bubble lengths is confined to $2-3$ channels diameters.
+Thus, the bubble radius changes significantly and cannot be used as classical Bretherton problem
+benchmark. However, the works provide insight to the bubble formation and small bubbles flow in the
+microchannels. Also, \citet{carlson-comparison} found a significant influence of the mutual bubbles
+motion on the flow patterns.}
 
@@
 correlation for small cappillary numbers $\delta=0.6687 \Ca^{2/3}$. In the case of $\Ca=0.05$, the 
-film thickness is taken from {\color{green} Fig.~\ref{fig:heil:planar}},
-  and for Reynolds number $40$ equals $0.06$.  
+film thickness is taken from correlations by \citet{heil-bretherton} 
+%{\color{green}
+%Fig.~\ref{fig:heil:planar}}
+, and for Reynolds number $40$ equals $0.06$.  
 \item[Grid choice] After pecifying the film thickness, one needs to choose the associate
@@
 %requirement of $345 \mathrm{MB}$. 
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/giavedoni_planar.eps}
-\caption{\citet{giavedoni-numerical} gathered results across the
-literature for different capillary numbers (reprinted from \cite{giavedoni-numerical}). Results are
-scaled to the half-width of the
-channel. $h_{\infty}$ is the film thickness of air finger at
-the infinite distance from the front meniscus. \label{fig:giavedoni:planar}}
-\includegraphics[width=0.97\textwidth]{Figures/heil-planar.eps}
-\caption{\citet{heil-bretherton} performed simulations for different ranges of
-capillary numbers and Reynolds numbers (reprinted from \cite{heil-bretherton}). Results
-are scaled to the half-width of the channel. $h_{0}$ is the film thickness of air finger at
-the infinite distance from the front meniscus.
-\label{fig:heil:planar}}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=0.97\textwidth]{Figures/giavedoni_planar.eps}
+%\caption{\citet{giavedoni-numerical} gathered results across the
+%literature for different capillary numbers (reprinted from \cite{giavedoni-numerical}). Results are
+%scaled to the half-width of the
+%channel. $h_{\infty}$ is the film thickness of air finger at
+%the infinite distance from the front meniscus. \label{fig:giavedoni:planar}}
+%\includegraphics[width=0.97\textwidth]{Figures/heil-planar.eps}
+%\caption{\citet{heil-bretherton} performed simulations for different ranges of
+%capillary numbers and Reynolds numbers (reprinted from \cite{heil-bretherton}). Results
+%are scaled to the half-width of the channel. $h_{0}$ is the film thickness of air finger at
+%the infinite distance from the front meniscus.
+%\label{fig:heil:planar}}
+%\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_sehgal.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/Bubble/bubble_length.eps}\\
+\includegraphics[width=0.5\textwidth]{Figures/Bubble/bubble_sehgal_new.eps}\hfill
+\includegraphics[width=0.5\textwidth]{Figures/Bubble/bubble_length_new.eps}\\
 \caption{A qualitative comparison for the film thickness across
-the bubble length. Left (courtesy of \citet{sehgal-microchannel}),
-right (present simulations); $x$ is scaled to $H_{\mathrm{eff}}$
-and increases in the flow direction.\label{fig:sehgal:bubble:length}}
+the bubble length. {\color{red} Left} (courtesy of \citet{sehgal-microchannel}),
+{\color{red} Right}  (present simulations); $x$ is scaled to $H_{\mathrm{eff}}$
+and increases in the flow direction.{\color{red} The data from
+\cite{sehgal-microchannel} was digitally acquired by the program ``Engauge Digitizer'' and scaled
+to compare with the present simulations.} \label{fig:sehgal:bubble:length}}
 \end{figure}
@@
 The purpose of this section is to validate the correlations of
-\citet{giavedoni-numerical} in Fig. \ref{fig:giavedoni:planar} and
-\citet{heil-bretherton} in Fig. \ref{fig:heil:planar} for a range of capillary
+\citet{giavedoni-numerical} %in Fig. \ref{fig:giavedoni:planar} 
+and
+\citet{heil-bretherton} %in Fig. \ref{fig:heil:planar} 
+for a range of capillary
 numbers. Because of limited computational resources, we skip the
@@
 reasonable agreement. The plots depict the film thickness as a function of the
-the capillary number.\label{fig:capillary:comparison}}
+the capillary number. {\color{red} Data from \cite{giavedoni-numerical} was
+extracted with the help of ``Engauge Digital Discretizer software'', data from
+\citet{heil-bretherton} was extracted by hands using a set of curves involving interpolation and
+it's the reason for a data to exhibit a noisy trend.} \label{fig:capillary:comparison}}
 \end{figure}

### 2011-03-29 Sailfish/Capillary/capillary.py

@@ -1,7 +1,7 @@
 #!/usr/bin/python
 import os
 import subprocess
-import pylab
+#import pylab
 import numpy
 import math
  
@@ -167,7 +167,8 @@ def Analyze_Bubble():
     good=[0,1,2,3,4,5,6,7]
     good_value=[1,2,4]
     style=["b-.","r--","c."]
-    labels=[r'''$Ca='''+ca+r'''$''' for ca in numpy.array(capillary_theor[good_value],dtype=str)]
+    #labels=[r'''$Ca='''+ca+r'''$''' for ca in numpy.array(capillary_theor[good_value],dtype=str)]
+    labels=[]    
     #color=["b","r","c","y","b","g"]
     #style_diff=["b-","r:","c-.","y--","b^","g<"]
     
@@ -236,7 +237,7 @@ def Analyze_Bubble():
         delta_x=15.0/(len(thicknesses)-1)
         x=delta_x*numpy.arange(0,len(thicknesses))*float(len(thicknesses))/3000.0
         pylab.plot(x,thicknesses,style[counter],linewidth=3)
-
+        labels.append(r'''$Ca='''+str(ux[dims[0]/2,z2%dims[1]]*2.0/3.0/math.sqrt(8.0*0.04*0.04/9.0))[0:4]+r'''$''')
         os.chdir("../..")
     
     pylab.ylim(ymin=0.0,ymax=0.5)
@@ -245,7 +246,7 @@ def Analyze_Bubble():
     leg=pylab.legend(labels)
     legtext = leg.get_texts() # all the text.Text instance in the legend
     for text in legtext:
-        text.set_fontsize(30) 
+        text.set_fontsize(20) 
     #set(ltext, fontsize='large') # the legend text fontsize
 
     fig.subplots_adjust(left=0.15,bottom=0.15)  
@@ -259,8 +260,76 @@ def Analyze_Bubble():
     #pylab.xlim(xmax=15)
     pylab.savefig("bubble_length.eps",format="EPS",dpi=300)
 
+def smoothListGaussian(list,strippedXs=False,degree=2):  
+    window=degree*2-1  
+    weight=numpy.array([1.0]*window)  
+    weightGauss=[]  
+    for i in range(window):  
+        i=i-degree+1  
+        frac=i/float(window)  
+        gauss=1/(numpy.exp((4*(frac))**2))  
+        weightGauss.append(gauss)  
+    weight=numpy.array(weightGauss)*weight  
[…]

### 2011-03-30 Review/answers.tex
Finished revisions. Waiting a reply from Jos.

@@
 \begin{document}
+\section{Reviewer I}
+
+\textbf{Question:}
+\begin{quotation}
+The authors found that quite surprisingly at high viscosity ratio, the results are
+independent of any inertial effects. It is known that Lattice Boltzmann has some problem for high
+viscosity ratio binary fluids. It will be better if the authors provide some details how the high
+viscosity ratio is tackled in their LB code.
+\end{quotation}
+
+\textbf{Answer:} The reviewer is right that accuracy of the method deteriorate with the high
+viscosity ratio. However, the parameters we chose here as $\tau_{\mathrm{liq}}=3.5-7.5$ and with
+grids in the vertical direction as $N_y=202$ is sufficient to suppress effects. We have added all
+the necessary references in the text.
+
+\textbf{Question:}
+\begin{quotation}
+Figure 8 needs to redrawn. Also, there are some inconsistencies in other figures. Authors may
+want to review all the figures and ensure that they have uniform fonts, etc.
+\end{quotation}
+
+\textbf{Answer:} We redrawed all the curves. For the figure 8 the data was extracted from the
+figure and rescaled for the comparison reasons.
+
+\section{Reviewer II}
+
 \textbf{Question:}
@@
 channel heights from the front bubble tip.  Moreover, the film thickness examination, see Section
-\ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
+..., shows that for the capillary range of interest, i.e.
 $0.08\leq Ca \leq 1$, the standard deviation from the bubble middle film thickness is around $7$
@@
 
-\texbf{Answer:} We thoroughly corrected all the inconsistencies.
+\textbf{Answer:} We thoroughly corrected all the inconsistencies.
 
@@
 
-\textbf{Answer:} We added additional section which shows the change of flow pattern.
+\textbf{Answer:} We added an additional paragraph which shows the change of flow pattern for the
+streamlines as it is indicated in literature.
 
@@
 
+\textbf{Answer:} We added additional information about results to the corresponding sections. The
+pressure drop are not measured as they are influenced largely by Reynolds number which is beyond
+the scope of this work.
+
+\textbf{Question:}
+\begin{quotation}
+The Authors mention the shooting method. It is not clear if this method is really used or just
+mentioned as a part of the discussion.
+\end{quotation}
+
+\textbf{Answer:} This is a part of discussion to have a remedy for the velocity. The corresponding
+sentence is added.
+
+\textbf{Question:}
+\begin{quotation}
+It seems more appropriate to introduce the variables of the Bretherton’s problem as well as to plot
+Figure 1 elsewhere than in the Section ”Results”.
+\end{quotation}
+
+\textbf{Answer:} The corresponding figure for the classical Bretherton problem is added. We
+separate it from the numerical benchmark layout.
+
+\textbf{Question:}
+\begin{quotation}
+Minor corrections...
+\end{quotation}
+
+\textbf{Answer:} All the corrections are taken into the account.
+
 \bibliographystyle{unsrtnat}

### 2011-03-30 Text/paper.tex
Finished revisions. Waiting a reply from Jos.

@@
 the liquid film and bubble motion. The simulations are two-dimensional and span a capillary number
-range of $0.05$ to $1.0$ where the capillary number is based on the liquid’s dynamic viscosity, the
+range of $0.05$ to $1.0$ where the capillary number is based on the liquid dynamic viscosity, the
 velocity of the bubble, and the interfacial tension. The flow is driven by a body force, and
 periodic boundary conditions apply in the streamwise direction. We obtain grid independent results
-as long as the liquid film thickness is at least two times the width of the diffuse interface, with
+as long as the liquid film thickness is at least twice the width of the diffuse interface, with
 film thicknesses in accordance to literature results. We also show that the results in terms of film
@@
 The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
-narrow channels. It was found that the front meniscus film thickness is proportional
+narrow channels, Fig. \ref{fig:classical:bretherton}. It was found that the front meniscus film
+thickness is proportional
 to $\Ca^{2/3}$ for capillary numbers smaller than $0.003$, where the capillary number is defined as:
@@
 $\gamma$ is the interfacial tension between gas and liquid.
+\begin{figure}
+\includegraphics[width=\textwidth]{Figures/benchmark_bretherton.eps}
+\caption{The classical Bretherton problem layout. The gas bubble
+propagates with the steady velocity $U_{\mathrm{bubble}}$ through the liquid media and deposits
+the film, where thickness to the height of the channel $H_{\mathrm{eff}}$ ratio is 
+$\delta$.\label{fig:classical:bretherton}} 
+\end{figure}
 
@@
 involving complex geometries, free interface motion, or for problems involving coalescence and/or
-droplet breakup. The continuous interface models are more flexible for those kind of simulations.
+droplet breakup. The continuous interface models are more flexible for those kinds of simulations.
 However, if the interface is smeared out over several grid nodes, the question of
@@
 exploring the parameter range of the binary liquid lattice
-Boltzmann method for correctly resolving the flow in a range of capillary numbers.
-
-The lattice Boltzmann method has emerged as a successful method to simulate
+Boltzmann method to resolve correctly the flow in a range of capillary numbers.
+
+The lattice Boltzmann method (LBM) has emerged as a successful method to simulate
 a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
@@
 inertia effects.} 
-Therefore, the major governing factor for microchannel flows is not the density ratio, but the
+Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
+but the
 viscosity ratio. Moreover, the results based on the uniform density, as in the case of the LBM
@@
 
-The goal of this paper is to understand how to choose the
-binary liquid model parameters, the viscosity ratio and the grid
-resolution to obtain correct flow physics with an emphasis on film thicknesses.
+Resolving the interface on a fine level for capillary numbers smaller than
+$0.003$ is computationally expensive even in the two-dimensional case
+(the amount of memory necessary to perform the simulation would be of the order
+of tens or hundreds gigabytes).  This work is therefore based on a
+comparison of the results obtained with the known data
+validated for the Bretherton bubble flow.  In this paper, we present
+techniques for initialization of the simulations and discuss optimal
+parameter ranges for the binary liquid LB model, as the binary liquid parameters, the viscosity
+ratio and the grid resolution. The emphasis is to obtain correct flow physics, especially in terms
+of film thicknesses.
+
+%The goal of this paper is to understand how to choose the
+%binary liquid model parameters, the viscosity ratio and the grid
+%resolution to obtain correct flow physics with an emphasis on film thicknesses.
 
@@
 Bretherton \cite{bretherton} studied a long bubble
-moving in a tube filled with liquid. It was found that the deposition film thickness
+moving in a tube filled with liquid. It was found that the film thickness
 is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later,
@@
 finite element solution with previously published results.
-The solutions were obtained for circular and planar cases.
+The solutions were obtained for circular and planar geometries.
 
@@
 
-One should acknowledge the works of \citet{pagonabarraga-fingers} on menisci
+One should acknowledge the work of \citet{pagonabarraga-fingers} on menisci
 in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
-simulations of two-dimensional channel flows for a relatively large capillary numbers, and
+simulations of two-dimensional channel flows for relatively large capillary numbers, and
 found discrepancies with the classical Bretherton theory, which
@@
 
-Resolving the interface on a fine level for capillary numbers smaller than
-$0.003$ is computationally expensive even in the two-dimensional case
-(the amount of memory necessary to perform the simulation would be of the order
-of tens or hundreds gigabytes).  This work is therefore based on a
-comparison of the results obtained with the known data
-validated for the Bretherton bubble flow.  In this paper, we present
-techniques for initialization of the simulations and discuss optimal
-parameter ranges for the binary liquid LB model.
-
 \section{Lattice Boltzmann binary liquid model}
@@
 two parts: a local collision step, and a propagation step which transports
-information from one node to another in certain
+information from one node to another along some 
 directions specified by the discrete velocity set.
@@
 p_0 - k \phi \Delta \phi
-+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
++\rho\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\rho \frac{Q_{i\alpha\beta}u_{\alpha } u_ {
 \beta}}{2 c_s^4}\biggr)\\
@@
 &f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
-&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
+&g_i^{eq}&&=w_i\left(\Gamma \mu + \phi\frac{ c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
 \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), 1\leq i \leq 8 \\
@@
 %thickness is examined.
-While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
-or defined through the model parameters, the bubble velocity $U_{\mathrm{bubble}}$ is not given
+{\color{red} Parameters as $\mu_{\mathrm{liq}}$, $\gamma$ are usually taken in the already defined
+stable range. The capillary number $\Ca$ is supplied in order to simulate.} 
+In comparison, the bubble
+velocity $U_{\mathrm{bubble}}$ is not given
 explicitly.
-In our simulations we approximate $U_{bubble}$ with the
+In our simulations we approximate $U_{\mathrm{bubble}}$ with the
 Poiseuille profile center velocity, which is given as:
@@
 \begin{aligned}
-U_{\mathrm{bubble}} \approx \frac{H_{\mathrm{eff}}^2}{8
+&U_{\mathrm{bubble}} \approx \frac{H_{\mathrm{eff}}^2}{8
 \mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
-U_{\mathrm{bubble}}\approx
+&U_{\mathrm{bubble}}\approx
 \frac{{(N_y-2)}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
@@
 \end{equation}
-where $N_y$ is the grid number in the vertical direction. As will be discussed further, the
+where $N_y$ is the number of nodes in the vertical direction. As will be discussed further, the
 effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes to
@@
 That explicitly implies the surface tension to be $\sqrt{\frac{8 k
-A}{9}}=0.0377$ and the characteristic length of the interface
+A}{9}}=0.0377$ and the characteristic width of the interface
 $5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units (grid nodes). For the sake of simplicity and
@@
 %Fig.~\ref{fig:heil:planar}}
-, and for Reynolds number $40$ equals $0.06$.  
-\item[Grid choice] After pecifying the film thickness, one needs to choose the associate
+, and for Reynolds number $40$ equals $0.06-0.07$.  
+\item[Grid choice] After specifying the film thickness, one needs to choose the associate
 number of nodes to resolve the film thickness. Note that for paremeters indicated above the
@@
 film thickness to be $2-2.5$ times larger than the interface thickness. Therefore, we want to
-  resolve $6\%$ of the whole channel width with $12$ lattice units.  Thus, the channel
+  resolve $6\%$ of the whole channel width with $12$ lattice units.  Thus, the physical channel
   width is $200$ lattice units. We take the bubble length equal to $5$ channel widths for the film
@@
   and the distance between bubbles as $3$ bubble lengths to avoid mutual influence of bubbles on
-each other. Thus, the whole grid size is $200 \times 3000$.
+each other. Thus, the whole physical grid size is $200 \times 3000$. Note that in the case of the
+half-way bounce-back walls \cite{yu} which are used in the
+simulations one needs to calculate the film thickness as:
+\begin{equation}
+\delta=(\phi_0-0.5)/(N_y-2),
+\end{equation}
+where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
+is the effective channel height.
+If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions between the grid nodes
+which represent
+the physical domain. The effective wall location is in the middle between bounce-back and fluid
+nodes giving
+overall $N_y-2$ nodes representing the fluid. Note that it is a simplification to impose the
+boundary in the middle between the bounce-back
+node and the fluid node. The location of the wall is
+viscosity dependent \cite{ginzburg-multireflection}.  The effective location of the
+wall for the
+multiphase models to the best authors' knowledge is not yet derived. Thus, the simulation grid size
+is $202\time 3000$. 
  \item[Velocity] The relaxation parameters are taken to be in the stable range for simulations.
@@
 Let us illustrate the procedure for the capillary number
-$0.05$. As it was pointed out above, the film thickness in that case is $6$ percent of the
-effective
-channel $H_{\mathrm{eff}}$.
-Note that in the case of the half-way bounce-back walls \cite{yu} which are used in the
-simulations one needs to calculate the film thickness as:
-\begin{equation}
-\delta=(\phi_0-0.5)/(N_y-2),
-\end{equation}
-where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
-is the effective channel height.
-If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions between the grid nodes
-which represent
-the physical domain. The effective wall location is in the middle between bounce-back and fluid nodes giving
-overall $N_y-2$ nodes representing the fluid. Note that it is a simplification to impose the
-boundary in the middle between the bounce-back
-node and the fluid node. The location of the wall is
-viscosity dependent \cite{ginzburg-multireflection}.  The effective location of the
-wall for the
-multiphase models to the best authors' knowledge is not yet derived. In the case of $\delta=0.06$
-the
-initial grid size is $N_y=102$, which gives the horizontal grid size as $N_x=15(N_y-2)=1500$.
+$0.05$. 
+{\color{red} The procedure of choice and measurement follows the instructions from Section
+\ref{section:init}.
+}
+
+%As it was pointed out above, the film thickness in that case is $6$ percent of the
+%effective
+%channel $H_{\mathrm{eff}}$.
+The simulations start from the initial grid size with $N_y=102$, which gives the horizontal grid
+size as $N_x=15(N_y-2)=1500$, and proceeds with finer grids.
 The bubble is initialized as a rectangular box with coordinates
@@
 We performed a number of simulations for the prescribed grids. The
-simulation results in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
+simulation results in terms of grid dimensions $N_x$, $N_y$, the film thickness $\delta$, center
+bubble
 velocity $U_{bubble}$ are summarized in Table \ref{table:parameters:grid:refinement}.
@@
 shape.
+{\color{red} The shape of the bubble is an important quantity, as the bubble tip shape mainly
+influences the film thickness at infinity \cite{bretherton}. It was indicated by
+\citet{heil-bretherton} that the inertia effects influence insignificantly the film thickness at
+infinity but influence significantly the velocity pattern in front of the bubble. Therefore, the
+bubble shape can not serve as the only characterizing quantity.}
 
@@
 needs to obtain the desired capillary number, the shooting method for the forcing is
-necessary, starting with the Poiseuille pressure gradient force as the initial condition. We extracted
+necessary, starting with the Poiseuille pressure gradient force as the initial condition. 
+{\color{red} Note that in the present work the shooting method is not used but it is presented as
+a possible remedy to exactly match the required capillary number.} 
+We
+extracted
 the data of 
@@
 
-\subsection{The influence of different viscosities}
+{\color{red} For the completeness of the presented data we also indicate the velocity streamlines
+profiles, Fig. \ref{fig:streamlines:pattern}. The pattern shows that there is a transition of
+having vortex in the front of the bubble $\Ca=0.22$ and not having it for larger capillary numbers
+$\Ca=1.00$. \citet{giavedoni-numerical} studied the behavior of the detaching of the vortex for the
+Bretherton problem. They indicate critical capillary numbers as $\Ca=0.73$ and $\Ca=0.89$ for
+detaching and full vortex disappearance. \citet{heil-bretherton} indicates that the behavior of the
+pattern strongly depends on the Reynolds number. The scope of this work is to indicate the
+qualitative capture of the effects. The thorough study of the flow patterns behavior for Reynolds
+number range and for the bubble train is planned for the future. 
+}
+\begin{figure}
+\includegraphics[angle=90,width=\textwidth]{Figures/Capillary/ca022_velocity_new.eps}\\
+\includegraphics[angle=90,width=\textwidth]{Figures/Capillary/ca100_velocity_new.eps}
+\caption{The streamlines plots for $\Ca=0.22$ (top) and $\Ca=1.00$ (bottom). One can see the change
+of pattern as indicated in works of \citet{heil-bretherton} and \citet{giavedoni-numerical}. For
+smaller capillary numbers there exists a vortex in front of the bubble. For the larger once there
+is none. The streamlines are presented in the moving reference of the bubble
+interface.\label{fig:streamlines:pattern}}
+\end{figure}
+
+\subsection{The influence of different viscosities ratio}
 The results presented above are taken for the liquid over gas kinematic viscosity ratio of $10$.
@@
 consistent with the literature data. 
+{\color{red} However, in reality the viscosity ratio affects results as it
+is indicated by \citet{shikazono-square}.}
+
+{\color{red} Also, while it is necessary to increase the viscosities ratio for the binary-liquid
+formulation to be close to the physical Bretherton formulation, one needs to be attentive to the
+numerical accuracy of the LBM formulation. It is known \cite{ginzburg-trt-simple-hydro} that the
+BGK collision operator used in the present scheme has residual error accuracy proportional
+to $\Bigl(\tau-\frac{1}{2}\Bigr)^2$. That means by increasing the viscosity one reduces the spatial
+accuracy of the method. To avoid this one needs to use advanced formulations of the LBM collision
+operator, as two-relaxation-times (TRT) and multiple-relaxation-times (MRT) collision operators
+\cite{ginzburg-main}. Although TRT and MRT collisions operators are studied for simple
+hydrodynamics solutions, their application to the complex fluids flows is sparse and not studied
+well.
+}
 \begin{figure}
@@
 \section{Acknowledgement}
-A.~Kuzmin wants to thank Schlumberger for their financial support.
+A.~Kuzmin wants to thank Schlumberger for their financial support. Authors are grateful to referees
+for their significant contribution to the quality improvement of the manuscript.
 \bibliographystyle{model1-num-names}

### 2011-03-30 Sailfish/Capillary/capillary.py

@@ -1,7 +1,7 @@
 #!/usr/bin/python
 import os
 import subprocess
-#import pylab
+import pylab
 import numpy
 import math
  
@@ -122,7 +122,7 @@ def Analyze_Simulations():
 
     pylab.xlim(0.02,1.5)
     pylab.ylim(ymin=0.01)
-    numpy.savetxt("capillary.dat",zip(capillaries,widths))
+    numpy.savetxt("capillary.dat",zip(capillaries_real,widths_real))
     
     fig.subplots_adjust(left=0.15,bottom=0.15)  
     pylab.xticks(fontsize=20)
@@ -358,7 +358,7 @@ def Analyze_Vectors():
     capillary_str=numpy.array(["3","5","8","10","20","40","60","80"])
  
     good=[0,1,2,3,4,5,6,7]
-    good_value=[4]
+    good_value=[7]
     style=["b-.","r--"]
     #labels=[r'''$Ca='''+ca+r'''$''' for ca in numpy.array(capillary_theor[good_value],dtype=str)]
     labels=[]    
@@ -389,7 +389,7 @@ def Analyze_Vectors():
             z2=numpy.min(numpy.where(center>0.0))+dims[0]
             z1=numpy.max(numpy.where(center>0.0))
         print z1,z2
-
+        print "Capillary=",ux[dims[0]/2,z2%dims[1]]*2.0/3.0/math.sqrt(8.0*0.04*0.04/9.0)
         positive=numpy.where(phase>0.0)
         negative=numpy.where(phase<0.0)
         large=numpy.where(numpy.logical_or(x<20,x>dims[0]-20))
@@ -428,30 +428,55 @@ def Analyze_Vectors():
         #    resources.vfMissingUValueV = uvar._FillValue
         # if hasattr(vvar,"_FillValue"):
         #    resources.vfMissingVValueV = vvar._FillValue
-   
-        resources.tiMainFont    = "Times-Roman"
-        resources.tiXAxisString = "streamlines"
+        
+        
+        #resources.tiMainFont    = "Times-Roman"
+        #resources.tiMainOn=True
+        #resources.tiMainString="Ca=0.22 "
+           
+        #resources.tiXAxisString = "streamlines"
         resources.vpHeightF = 0.25 # Define height, width, and location of plot.
         resources.vpWidthF  = 3*0.25
+        resources.wkPaperSize="A5"
         resources.nglFrame = False
+        resources.vfXArray=numpy.linspace(0.0,15.0,len(ux[1,::50]))
[…]

### 2011-03-30 Sailfish/Capillary_Viscous/capillary.py

@@ -67,16 +67,29 @@ def Analyze_Simulations():
         #pylab.figure()
         name="capillary200000.npz"
         array=numpy.load(name)
-        prof=array['phi'][:,exam[i]]
+        dims=array['phi'].shape
         #x=numpy.arange(0.0,float(ny[i]))/ratio
         #pylab.imshow(array['phi'])
     
         if i in good:
             #pylab.plot(prof)
-            widths.append(Get_Zero(prof))
             vel=array['v'][0]
             vel_prof=vel[:,exam[i]]
-            velocities.append(vel_prof[len(vel_prof)/2])
+            center=array['phi'][dims[0]/2,:]
+             
+            
+            z1 = numpy.min(numpy.where(center < 0.0))
+            z2 = numpy.max(numpy.where(center < 0.0))
+            if z1==0:
+                z2=numpy.min(numpy.where(center>0.0))+dims[1]
+                z1=numpy.max(numpy.where(center>0.0))
+            print z1,z2
+ 
+            prof=array['phi'][:,((z1+z2)/2)%dims[1]]
+            widths.append(Get_Zero(prof))
+
+            
+            velocities.append(vel[dims[0]/2,z2%dims[1]])
         
         #pylab.plot(x[0:20],prof[0:20],color[i],linewidth=3)
         #pylab.plot(x,prof,style_diff[i],markersize=10,linewidth=3)

### 2011-04-04 Review/answers.tex
Made all Jos corrections.

@@
 grids in the vertical direction as $N_y=202$ is sufficient to suppress effects. We have added all
-the necessary references in the text.
+the necessary references in the text, lines $481-491$.
 
@@
 
-\textbf{Answer:} We redrawed all the curves. For the figure 8 the data was extracted from the
+\textbf{Answer:} We redrew all the figures. For the figure 8 the data was extracted from the
 figure and rescaled for the comparison reasons.
@@
 
-\textbf{Answer:} The following section is added to the text 
-\begin{quotation}
-We chose to measure the film thickness in the middle of the bubble. The
-classical Bretherton formulation is for the measurement of the film thickness at ``infinity''.
-However, \citet{cerro-bubble-train} indicate that for the gravity-driven bubble train flow in tubes
-it is enough for bubble to have a length of two-three diameters to apply analysis for very long
-bubbles. In comparison with the axisymmetric circular geometry, for three-dimensional square shaped
-microchannels \citet{heil-threedim} indicate the measurement of the film thickness at $5.5$
-channel heights from the front bubble tip. This distance was shown sufficient for the
-film thickness to establish itself for the capillary number range $Ca<4$. We chose to measure
-the film thickness in the middle of the bubble, which is located at least at the distance of $2.5-3$
-channel heights from the front bubble tip.  Moreover, the film thickness examination, see Section
-..., shows that for the capillary range of interest, i.e.
-$0.08\leq Ca \leq 1$, the standard deviation from the bubble middle film thickness is around $7$
- percent for small capillary numbers ($Ca=0.05$) and less than $1$ percent for larger ones
-($Ca \geq 0.1$) at the distances ranging from one channel height from the rear tip upto one channel
-height to the front tip. This certainly validates the point of choice to measure bubble thickness.
-\end{quotation}
+\textbf{Answer:} The explanations are added in lines 434-453.
 
@@
 Reynolds
-number effects on the ﬁlm thickness for capillary numbers larger than 0.003” is in contradiction
+number effects on the film thickness for capillary numbers larger than 0.003” is in contradiction
 with
@@
 
-\textbf{Answer:} The following paragraph is added:
-\begin{quotation}
-Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
-{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
-range of Reynolds
-numbers. 
-{\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
-are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
-number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
-$300$. The indicate that while the influence of the established film thickness is insignificant
-($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
-significantly the pressure distrubution and the flow field near the front bubble tip. In
-the present simulations the largest Reynolds number we achieve is less than $20$ and it allows to
-neglect the
-inertia effects.}
-
-\citet{kreutzer-pressure-drop} showed a deviation from the $Ca^{2/3}$ rule
-{\color{red}\sout{and Reynolds number 
-effects on the film thickness}} for capillary numbers larger than $0.003$.
-\end{quotation}
+\textbf{Answer:} We elaborate the Reynolds number influence in lines 34-43.
 
@@
 
-\textbf{Answer:} Thank you for pointing it out. We included the discussions of the Reynolds
+\textbf{Answer:} Thank you for pointing this out. We included the discussions of the Reynolds
 effects in the answer to the question above. As well, we added Reynolds numbers to the Table for
@@
 inclusion $\bm{F}$. We added the citation which describes how the force population is formed and
+how it
 mimics the force behavior.
@@
 The maximum Bond number $Bo=\frac{\rho g N_y^2}{\gamma}$ in the
-current simulation equals to $7$, which is less than the number when the shape of the bubble is
-affected \cite{zheng-large-ratio}. However, the point is that the equation of state in the lattice
-Boltzmann framework is connected with density. Thus usually either pressure (i.e. density) boundary
-conditions are prescribed or the body force. If they are prescribed simultaneously then
-we can consider the influence of the Bond number on the simulations. Otherwise prescribing just the
-body force is analogous to prescribing the pressure gradient $\frac{\mathrm{d}P}{\mathrm{d}x}$ in
-the Navier-Stokes equation. Thus, it's not possbile the Bond number effect. It is because of the
-coupled nature of the pressure and density in the LBM framework.
+current simulation equals to $7$, which is less than the number for which the shape of the bubble is
+affected according to \cite{zheng-large-ratio}.
 
@@
 The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
-estima
-tions of the body force to apply and of the ﬁlm thickness to choose grid resolution. The
+estimations of the body force to apply and of the film thickness to choose grid resolution. The
 presentation
@@
 \textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
-about initialization and this was taken into the account by reducing repetitive text and
-introducing step-by-step explations, see Section about initialization. We also added the
+about initialization and this was taken into account by reducing repetitive text and
+introducing step-by-step explations, see Section about initialization (lines 259-305). We also added
+the
 recommendations for controlling parameters in the same section.  The
@@
 Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
-the lattice Boltzmann are refered to the non-dimensionalization procedure. Therefore, we decided to
-keep suggestions for controlling and initialization of simulations as informative guides for the
+the lattice Boltzmann method are refered to the non-dimensionalization procedure. Therefore, we
+decided to
+keep suggestions for controlling and initialization of simulations as informative guidelines for the
 LBM community.
@@
 obtained
-under the assumption of an axisymmetric Poiseuille ﬂow. However, among the ”Key words”, one can
+under the assumption of an axisymmetric Poiseuille flow. However, among the ”Key words”, one can
 read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
@@
 
-\textbf{Answer:} While the Hele-Shaw flow implies the flow with the infinitisemal distance between
-plates, in our situation the size of the bubble can be compared with the channel height ($5$ times
-larger) and that's why the assumption of the infinitisemal distance between plates is too strict.
-Moreover, after recalculating results with the velocity of the interface the assumption of the
-Poiseulle flow is valid and shows good accuracy.
+\textbf{Answer:} The factor $1/8$ is taken for the maximum velocity for the Poiseille flow between
+plates. Our assumption that the bubble travels with the constant velocity which equals to the
+maximum velocity of the Poiseuille profile between plates. Factor $1/12$ comes from the averaging
+of the flow rate. However, bubble moves faster than the liquid, i.e. the average liquid velocity.
+That's why the factor $1/8$ was chosen. The simulations show that factor $1/8$ works reasonably
+well.
 
@@
 The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
-journal of applied science. The Authors would improve signiﬁcantly the manuscript by deﬁning more
+journal of applied science. The Authors would improve signiﬁcantly the manuscript by defining more
 rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
@@
 A suggestion is to emphasize within an array the links between physical parameters and the model
-eq
-parameters. The variables i, fieq , gi should be deﬁned and the Equations for f and g in Equation
+parameters. The variables $i$, $f_i^{eq}$ , $g_i$ should be deﬁned and the Equations for $f$ and $g$
+in Equation
 3 should be described (which equation corresponds to continuity equation...). Equation 2 could
 perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
-physical variable f of Equation 3. The notation i = 1 ÷ 8 is unusual, it would be clearer to use a
-classical notation such as 1 ≤ i ≤ 8. The sound speed c2 should be deﬁned just after its
+physical variable f of Equation 3. The notation $i = 1 \div 8$ is unusual, it would be clearer to
+use a
+classical notation such as $1 \leq i \leq 8$. The sound speed $c_s^2$ should be deﬁned just after
+its
 introduction in line 107.
@@
 \textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the physical world
-parameters with the lattice Botlzmann parameters. As well more explanations are put towards the
-interface surface tension description and inclusion it to the model. The reference to the
-continuity and the Navier-Stokes equations are added. The equation (2) is not rewritten in terms of
+parameters with the lattice Botlzmann parameters (lines $149-160$). As well more explanations are
+put towards the
+interface surface tension description (lines $136-139$). The reference to the
+continuity and the Navier-Stokes equations are added (lines $140-143$). The equation (2) is not
+rewritten in terms of
 generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
-governing equations. The notations are corrected.
+governing equations. The notations have been corrected.
 
@@
 The Authors should make a distinction between variables in lattice units and dimensional or non
-dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y − 2$ where $N_y$ is
+dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y - 2$ where $N_y$ is
 the
@@
 is in
-contradiction with the use of $H_{eff}$ in Equation 5. The Authors should deﬁne each variable
+contradiction with the use of $H_{eff}$ in Equation 5. The Authors should define each variable
 without
@@
 improvements
-would ﬁt well in the Subsection named ”The nondimensionalization and initialization procedure”
+would fit well in the Subsection named ”The nondimensionalization and initialization procedure”
 which does not in fact introduce much nondimensionalisation procedure.
@@
 
-\textbf{Answer:} We added a decription paragraph in the Subsection ``The nondimensionalization and
-initialization procedure'' about the approach to units. In fact, all equations are the same in the
+\textbf{Answer:} We added a decription paragraph about the approach to units (lines 149-160). In
+fact, all equations are the same in the
 lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
-preferred units as soon as capillary number is matched in the LBM and the physical world.
+preferred units as soon as the capillary number is the same in the LBM and the physical world.
 
@@
 \begin{quotation}
-Some notations are unclear. For instance, the ﬁlm thickness is denoted as h∞ in Equation 7 whereas
-it is denoted δ in most of the paper. The notations w or δsim are also found before line 381 or in
-Table 3 respectively. The thickness of the interface is denoted ξ in Figure 1 but the same variable
+Some notations are unclear. For instance, the film thickness is denoted as $h_{\infty}$ in Equation
+7 whereas
+it is denoted $\delta$ in most of the paper. The notations w or $\delta_{sim}$ are also found before
+line 381 or in
+Table 3 respectively. The thickness of the interface is denoted $\xi$ in Figure 1 but the same
+variable
 is
-denoted 5ξ in Table 1. The speed of the bubble is denoted Ububble in some places and Ub in others.
-The variable Uin Equation 5 does not seem to be deﬁned.
+denoted $5\xi$ in Table 1. The speed of the bubble is denoted $U_{\mathrm{bubble}}$ in some places
+and $U_b$ in others.
+The variable $Uin$ Equation 5 does not seem to be deﬁned.
 \end{quotation}
@@
 \begin{quotation}
-Three ﬁgures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
+Three figures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
 unusual
@@
 
-\textbf{Answer:} We eliminated necessary pictures and references to them.
+\textbf{Answer:} We eliminated unnecessary pictures and references to them.
 
@@
 \begin{quotation}
-The plots of Figure 8 are of less quality than the other ﬁgures of the paper.
+The plots of Figure 8 are of less quality than the other figures of the paper.
 \end{quotation}
@@
 solution of Bretherton
-ﬂow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
+flow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
 differences between their work and this paper.
@@
 
-\textbf{Answer:} We outlined the differences in the introduction section.
+\textbf{Answer:} We outlined the differences in the introduction section (lines $92-100$).
 
@@
 
-\textbf{Answer:} We added references to the introduction section.
+\textbf{Answer:} We added references to the introduction section [17-19].
 
@@
 
-\textbf{Answer:} Thank you for pointing that. The classical Bretherton problem
+\textbf{Answer:} Thank you for pointing that out. In the classical Bretherton problem
 $U_{\mathrm{bubble}}$ is the bubble velocity, i.e. the interface velocity. We recalculate all the
 results taking into the consideration the interface velocity at the bubble tip. Qualitatively, the
-results are the same. It was found that Poiseuille profile assumption is improved significantly
-after recalculation.
+results are the same.
 
@@
 also
-be interesting if the Authors show some patterns of the ﬂow by plotting for instance the vector ﬁeld
+be interesting if the Authors show some patterns of the flow by plotting for instance the vector
+ﬁeld
 of velocity around the bubbles.
@@
 
-\textbf{Answer:} We added an additional paragraph which shows the change of flow pattern for the
-streamlines as it is indicated in literature.
+\textbf{Answer:} We added a small steady-state study (lines $334-341$). We added an additional
+paragraph which shows the change of flow pattern for the
+streamlines as it is indicated in literature (lines $463-471$, Figure 8).
 
@@
 In the Abstract, the Authors should give the range of real capillary numbers they measured instead
-of the targeted values (Calit ).
+of the targeted values ($Ca_{lit}$).
 \end{quotation}
@@
 \textbf{Answer:} We added additional information about results to the corresponding sections. The
-pressure drop are not measured as they are influenced largely by Reynolds number which is beyond
-the scope of this work.
+pressure drop is not measured as they are influenced largely by Reynolds number according to
+\citet{heil-bretherton}.
 
@@
 
-\textbf{Answer:} This is a part of discussion to have a remedy for the velocity. The corresponding
-sentence is added.
+\textbf{Answer:} This is a part of discussion. The corresponding
+sentence is added (lines $459-461$).
 
@@
 \textbf{Answer:} The corresponding figure for the classical Bretherton problem is added. We
-separate it from the numerical benchmark layout.
+separated it from the numerical benchmark layout.
 

### 2011-04-04 Text/paper.tex
Made all Jos corrections.

@@
 \cortext[cor1]{Corresponding author}
-\ead{shurik.kuzmin@gmail.com}
+\ead{kuzmin@ualberta.ca}
 \author[us]{M.~Januszewski}
@@
 different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
-neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
+negligible. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
 {\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
@@
 are negligible for $\Ca\leq0.05$ and have moderate changes for $\Ca>0.05$ in the range of Reynolds
-number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
+number from $0$ to $70$. Later on, \citet{heil-bretherton} extended results upto Reynolds number
 $300$. The indicate that while the influence of the established film thickness is insignificant
@@
 significantly the pressure distrubution and the flow field near the front bubble tip. In the present
-simulations the largest Reynolds number we achieve is less than $20$ and it allows to neglect the
+simulations the largest Reynolds number we achieve is less than $20$ and it allows to neglect 
 inertia effects.} 
@@
 but the
-viscosity ratio. Moreover, the results based on the uniform density, as in the case of the LBM
+viscosity ratio. Moreover, the results based on uniform density, as in the case of the LBM
 binary liquid model, are in good agreement with other simulations
@@
 (the amount of memory necessary to perform the simulation would be of the order
-of tens or hundreds gigabytes).  This work is therefore based on a
+of tens or hundreds of gigabytes).  This work is therefore based on a
 comparison of the results obtained with the known data
@@
 techniques for initialization of the simulations and discuss optimal
-parameter ranges for the binary liquid LB model, as the binary liquid parameters, the viscosity
+parameter ranges for the binary liquid LB model, such as the binary liquid parameters, the viscosity
 ratio and the grid resolution. The emphasis is to obtain correct flow physics, especially in terms
@@
 found discrepancies with the classical Bretherton theory, which
-is limited to the low capillary number regime \cite{giavedoni-numerical}. Authors didn't take into
+is limited to the low capillary number regime \cite{giavedoni-numerical}. These authors didn't take
+into
 the account recent studies extending microchannel simulations for the capillary numbers
-beyond the Bretherton regime. Thus, limiting comparison with other established CFD methods. Author
-used the Shan-Chen model \cite{Shan-chen:extended} to simulate the multiphase behavior. It is also
+beyond the Bretherton regime. Thus, the comparison with other established CFD methods is limited.
+\citeauthor{sehgal-microchannel} used the Shan-Chen model \cite{Shan-chen:extended} to simulate the
+multiphase behavior. It is also
 the continuous interface model, but the study of grid dependence is absent.
@@
 Level set method is represented by a number of works \cite{fukugata-levelset,
-lakehal-micro, carlson-comparison}. While the bubble shapes and  were found in a good agreeement
+lakehal-micro, carlson-comparison}. While the bubble shapes  were found in a good agreeement
 with the experiments, the range of studied bubble lengths is confined to $2-3$ channels diameters.
 Thus, the bubble radius changes significantly and cannot be used as classical Bretherton problem
-benchmark. However, the works provide insight to the bubble formation and small bubbles flow in the
-microchannels. Also, \citet{carlson-comparison} found a significant influence of the mutual bubbles
-motion on the flow patterns.}
+benchmark. However, the works provide insight to the bubble formation and to the way small bubbles
+flow in the
+microchannels.}
 
@@
 where $f_i$ is the probability distribution function in the direction $\bm{c_i}$,
-{\color{red} $f_i^{q}$ is the equilibrium probability distribution function}, $\omega$ is the
+{\color{red} $f_i^{eq}$ is the equilibrium probability distribution function}, $\omega$ is the
 relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
@@
 Note that the first equation of system \ref{standard:implementation} simulates the continuity and
-the Navier-Stokes equations, first two equations in (\ref{binary:fluid:system}). The second equation
-of system \ref{standard:implementation} simulates the phase governing equation, third equation in
+the Navier-Stokes equations, i.e. the first two equations in (\ref{binary:fluid:system}). The second
+equation
+of system \ref{standard:implementation} simulates the phase governing equation, i.e. the third
+equation in
 (\ref{binary:fluid:system}).} The system (\ref{binary:fluid:system}) allows the separation of the
@@
 While the lattice Boltzmann system has parameters as gas and liquid
-viscosities, the surface tension and etc., those parameters are not the representative and
+viscosities, the surface tension, etc., those parameters are not the representative and
 proportional quantities of the parameters in a physical system. The parameters of the lattice
 Boltzmann are connected with the physical world's parameters only through the non-dimensional
-numbers governing a physics of the problem. In this case it's a capillary number $\Ca$, which is
-obtained from the physical world and then it is matched through the lattice Boltzmann quantities.
-In fact, one should distinguish the lattice Boltzmann system and the physical system. The set of
+numbers governing the physics of the problem. In this case it is a capillary number $\Ca$ and a
+viscosity ratio $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}$, which are
+obtained from the physical world and then they are matched through the lattice Boltzmann quantities.
+ The set of
 the fluid and phase equations \eqref{binary:fluid:system} is valid in the lattice Boltzmann space
@@
 $U_{\mathrm{bubble}}$, in the physical units or in the lattice Boltzmann units as soon as the
-capillary number is matched in the both of worlds.}
+capillary number is the same in both worlds.}
 
@@
 can influence mass transfer in the thin film, thereby compromising the
-overall solution. Therefore, the diffuse interface should not noticeably affect the film
-thickness.
+overall solution. 
+%Therefore, the diffuse interface should not noticeably affect the film
+%thickness.
  \item The wettability coefficient which in principle can control the
@@
 bubble shape
-is different in the front and rear menisci depending on Bond number ($\mathrm{Bo}=\frac{\Delta \rho g
+is different in the front and rear menisci depending on the Bond number ($\mathrm{Bo}=\frac{\Delta
+\rho g
 l^2}{\gamma}$). \citet{wong-films}
@@
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/benchmark_new.eps}
+\includegraphics[width=\textwidth]{Figures/benchmark_new.eps}
 \caption{The benchmark sketch. $\delta$ corresponds to film thickness. $\xi$ corresponds to the interface thickness. We take the microchannel length to be 15 times larger than its height.
@@
 {\color{red} Parameters as $\mu_{\mathrm{liq}}$, $\gamma$ are usually taken in the already defined
-stable range. The capillary number $\Ca$ is supplied in order to simulate.} 
+stable range. The capillary number $\Ca$ is supplied from the physical world.} 
 In comparison, the bubble
@@
 In our simulations we approximate $U_{\mathrm{bubble}}$ with the
-Poiseuille profile center velocity, which is given as:
+Poiseuille profile maximum velocity, which is given as:
 \begin{equation}
@@
 \end{equation}
-In simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
+In the simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
 velocity to differ from that implied by the Poiseuille relation because of the bubble presence in a
@@
 capillary number. 
-However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
-starting approximation.
+However, the initialization body force as given by (\ref{poiseuille:velocity:center}) is a good
+starting point.
 
@@
 \begin{figure}
-\includegraphics*[bb=-15 360 630 430,width=0.97\textwidth]
+\includegraphics*[bb=-15 360 630 430,width=\textwidth]
 {Figures/Init/initbegin12.eps}
 \includegraphics*[bb=-15 360 630
-430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
+430,width=\textwidth]{Figures/Init/initfinish12.eps}\\
 \includegraphics*[bb=-15 360 630 430,
-width=0.97\textwidth]{Figures/Init/initbegin16.eps}
+width=\textwidth]{Figures/Init/initbegin16.eps}
 \includegraphics*[bb=-15 360 630
-430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
+430,width=\textwidth]{Figures/Init/initfinish16.eps}\\
 \includegraphics*[bb=-15 360 630 430,
-width=0.97\textwidth]{Figures/Init/initbegin20.eps}
+width=\textwidth]{Figures/Init/initbegin20.eps}
 \includegraphics*[bb=-15 360 630
-430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
+430,width=\textwidth]{Figures/Init/initfinish20.eps}\\
 \caption{Phase plots for different bubble thicknesses as
@@
 examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
-That explicitly implies the surface tension to be $\sqrt{\frac{8 k
+That implies the surface tension to be $\sqrt{\frac{8 k
 A}{9}}=0.0377$ and the characteristic width of the interface
@@
 
-Overall, the setting simulations procedure can be described in steps as follows:
+Overall, the setting up the simulations can be described in steps as follows:
 \begin{description}
- \item[Capillary number] One first needs to be supplied with the capillary number for
+ \item[Capillary number] One first needs to set the capillary number for
 simulations. For the demonstration purposes we chose $\Ca=0.05$. 
@@
 %Fig.~\ref{fig:heil:planar}}
-, and for Reynolds number $40$ equals $0.06-0.07$.  
+, and for Reynolds number $40$ is in the range $0.06-0.07$.  
 \item[Grid choice] After specifying the film thickness, one needs to choose the associate
@@
 multiphase models to the best authors' knowledge is not yet derived. Thus, the simulation grid size
-is $202\time 3000$. 
- \item[Velocity] The relaxation parameters are taken to be in the stable range for simulations.
+is $202\times 3000$. 
+ \item[Velocity] The relaxation parameters are taken to be in the stable range for hte simulations.
 In the present simulations the viscosity ratio is $\dfrac{\mu_{liq}}{\mu_{gas}}=10$ with the
@@
 \item[Body force] 
-Thus, it is desired to obtain the prescribed velocity $U_{\mathrm{bubble}}$ in the simulations. The
+It is desired to obtain the prescribed velocity $U_{\mathrm{bubble}}$ in the simulations. The
 flow in the simulations is conducted by imposing the body force. Thus, the assumption is needed of
@@
 \begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8\mu}{H_{\mathrm{eff}}^2}
+\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8\mu_{\mathrm{liq}}}{H_{\mathrm{eff}}^2}
 U_{\mathrm{bubble}}=\frac{8}{N_y^2}\gamma \Ca=3.77\times 10^{-7}.  
@@
 After imposing the body force the simulations can be run. A typical simulation runs approximately
-$100000-300000$ time steps due to reach steady-state.
+$100000-300000$ time steps due to reach steady-state, see Section \ref{sec:steady:state}.
 %Note that the velocity is proportional to the physical time step (small velocities imply longer
@@
 }
-Note that following the prescribed procedure one can see that a computer memory requirements are
+Note that following the prescribed procedure one can see that  computer memory requirements are
 high for small capillary numbers due to large grids to resolve the film thickness. Thus, the
-simulations in the present work cover range of capillary numbers from $0.05$ to $1.0$.
+simulations in the present work cover a range of capillary numbers from $0.05$ to $1.0$.
 Although, Bretherton obtained an analytical solution for the range of low capillary
@@
 results for all ranges of capillary numbers.  Our intention is
-  to show physically correct behavior for moderate $Ca$,
+  to show physically correct behavior for moderate $\Ca$,
 which
@@
 in calculating the body force from the Poiseuille velocity profile.
-In reality, the less viscous bubble moves faster than the surrounding liquid. By
-substituting the obtained velocity in the capillary number equation, a
-larger capillary number can be obtained. In practice, one needs to take the bubble velocity from the
+In reality, the less viscous bubble moves faster than the surrounding liquid.  In practice, one
+needs to take the bubble velocity from the
 simulations and
@@
 $H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
-x } = \mathrm{const}$. For the designing capillary range simulations, Section
+x } = \mathrm{const}$. For designing capillary range simulations, Section
 \ref{section:capillary:range}, we use a proportionality law to initialize body forces knowing
@@
 }
+{\color{red}
+\subsection{Steady state}
+\label{sec:steady:state}
+The simulations were performed in order to determine the number of iterations for the bubble to
+reach the steady state. The grid was chosen as $202 \times 3000$. The body force was $0.375\times
+10^{-6}$. The initial film thickness was taken as $12$ nodes. Table \ref{table:steady:state}
+summarizes the simulation results in terms of the film thickness in the middle of the bubble and the
+velocity of the interface. We can conclude it is enough to have $100000-300000$ iterations for
+bubble to approach the steady state.
+\begin{table}
+\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|}
+\hline
+$N_\mathrm{iter}$&$\delta$&$U_{\mathrm{bubble}}$&{\color{red} $Re$}\\
+\hline
+$10000$&$ 0.0655$&$0.0027$&$0.81$\\
+\hline
+$14000$&$0.0608$&$0.0027$&$0.81$\\
+\hline
+$18000$&$0.0577$&$0.0027$&$0.81$\\
+\hline
+$22000$&$0.0630$&$0.0027$&$0.81$\\
+\hline
+$26000$&$0.0660$&$0.0027$&$0.81$\\
+\hline
+$30000$&$0.0661$&$0.0027$&$0.81$\\
+\hline
+\end{tabularx}
+\caption{The results for steady-state calculations. The simulated domain is
+of size $202 \times 3000$. $U_{\mathrm{bubble}}$ is the velocity of the interface at the fron tip of
+the bubble. $\Ca=0.05$ was taken to initialize the simulation. One can see that in terms of
+velocities and film thicknesses $100000-300000$ iterations are enough.
+\label{table:steady:state}}
+\end{table}
+}
+
 \subsection{Grid refinement}
@@
 $0.05$. 
-{\color{red} The procedure of choice and measurement follows the instructions from Section
+{\color{red} The procedure of parameters choice and film thickness measurements follows the
+instructions from Section
 \ref{section:init}.
@@
 {\color{red} After choosing the reference parameters, the grid refinement procedure
-needs to keep the macroscopic parameters constant. One can refer to Section \ref{section:init}.}  
+needs to keep the macroscopic parameters constant, see Section \ref{section:init}.}  
 %It is easy to check
@@
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
+\includegraphics[width=\textwidth]{Figures/Grid/norm_grid_profs.eps}
 \caption{Grid-refined profiles for the effective
@@
 \begin{table}
-\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|X|}
+\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|}
 \hline
@@
 gradients, the values near the wall are below $1$, and the gradient profile can
-fuse with the values of the gas phase. This case corresponds to  slug flow when gas is in
+merge with the values of the gas phase. Such a situation would correspond to  slug flow when gas is
+in
 contact with the wall.
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
+\includegraphics[width=\textwidth]{Figures/Wall/phase_grad_profiles.eps}
 \caption{Phase profiles for different wall gradients. One can see that if the thickness
@@
 effective viscosity near the wall and the bubble velocity respectively. If the viscosity ratio is
-high enough then the free surface motion of the bubble should not be dependent on the shear stress
+high enough then the free surface motion of the bubble should not depend on the shear stress
 near the wall.
@@
 \citet{heil-bretherton} that the inertia effects influence insignificantly the film thickness at
-infinity but influence significantly the velocity pattern in front of the bubble. Therefore, the
-bubble shape can not serve as the only characterizing quantity.}
+infinity but influence significantly the velocity pattern in front of the bubble. Therefore, to
+properly describe the Bretherton phenomena the
+bubble shape along with velocity pattern need to be examined.}
 
@@
 The pressure gradient can be obtained through the capillary number
-ratio, Section \ref{section:init}:
+ratio, see Section \ref{section:init}:
 \begin{equation}
@@
 However, \citet{cerro-bubble-train} indicate that for the gravity-driven bubble train flow in tubes
-it is enough for bubble to have a length of two-three diameters to apply analysis for very long
+it is enough for bubble to have a length of two-three tube diameters to apply analysis for very long
 bubbles. In comparison with the axisymmetric circular geometry, for three-dimensional square shaped
@@
 from the front tip depending on the Reynolds number.  We chose to measure the film thickness in the
-middle of the bubble, which is located at least at the distance of $2.5-3$ channel heights from the
-front bubble tip. Moreover, the film thickness examination, see Section
+middle of the bubble, which is located at least at a distance of $2.5-3$ channel heights from the 
+bubble tip. Moreover, the film thickness examination, see Section
 \ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
-$0.05\leq \Ca \leq 1$, the standard deviation from the bubble middle film thickness is around $7$
+$0.05\leq \Ca \leq 1$, the standard deviation of the film thickness from the bubble middle film
+thickness is around $7$
 percent for small capillary numbers ($\Ca=0.05$) and less than $1$ percent for larger ones ($\Ca
-\geq 0.1$) at the distances ranging from one channel height from the rear tip upto one channel
-height to the front tip. This certainly validates the point of choice to measure bubble thickness.}
+\geq 0.1$). %at the distances one channel height from the rear tip upto one channel
+%height to the front tip. 
+This certainly validates the point of choice to measure bubble thickness.}
 One can see that the calculated
@@
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/Capillary/capillaries_comparison_real.eps}
+\includegraphics[width=\textwidth]{Figures/Capillary/capillaries_comparison_real.eps}
 \caption{A comparison between simulation results and results of
@@
 the capillary number. {\color{red} Data from \cite{giavedoni-numerical} was
-extracted with the help of ``Engauge Digital Discretizer software'', data from
+extracted with the help of ``Engauge Digitizer'', data from
 \citet{heil-bretherton} was extracted by hands using a set of curves involving interpolation and
@@
 pattern strongly depends on the Reynolds number. The scope of this work is to indicate the
-qualitative capture of the effects. The thorough study of the flow patterns behavior for Reynolds
-number range and for the bubble train is planned for the future. 
+qualitative capture of the effects.  
 }
@@
 smaller capillary numbers there exists a vortex in front of the bubble. For the larger once there
-is none. The streamlines are presented in the moving reference of the bubble
+is none. The streamlines are presented in the reference frame moving with the bubble
 interface.\label{fig:streamlines:pattern}}
@@
 \ref{fig:capillary:viscous}. The results validate our assumption that in
-the low capillary flow regime that the sufficiently high gas over liquid viscosity ratio ($\geq 10$) is sufficient to obtain results
+the low capillary flow regime the sufficiently high gas over liquid viscosity ratio ($\geq 10$)
+is sufficient to obtain results
 consistent with the literature data. 
@@
 BGK collision operator used in the present scheme has residual error accuracy proportional
-to $\Bigl(\tau-\frac{1}{2}\Bigr)^2$. That means by increasing the viscosity one reduces the spatial
+to $\Bigl(\tau-\frac{1}{2}\Bigr)^2$. That means that by increasing the viscosity one reduces the
+spatial
 accuracy of the method. To avoid this one needs to use advanced formulations of the LBM collision
@@
 \cite{ginzburg-main}. Although TRT and MRT collisions operators are studied for simple
-hydrodynamics solutions, their application to the complex fluids flows is sparse and not studied
+hydrodynamics solutions, their application to complex fluid flows is sparse and not studied
 well.
@@
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/Capillary_Viscous/capillaries_viscous.eps}
+\includegraphics[width=\textwidth]{Figures/Capillary_Viscous/capillaries_viscous.eps}
 \caption{The film thicknesses versus capillary number for
@@
 \section{Acknowledgement}
-A.~Kuzmin wants to thank Schlumberger for their financial support. Authors are grateful to referees
+A.~Kuzmin wants to thank Schlumberger for financial support. Authors are grateful to referees
 for their significant contribution to the quality improvement of the manuscript.

### 2011-04-04 Sailfish/Steady/bmc.py

@@ -1,0 +1,59 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo2D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(hy < self.wall_width, hy >= self.lat_ny-self.wall_width)
+        self.set_geo(wall_map, self.NODE_WALL)
+
+    def init_fields(self):
+        hy, hx = np.mgrid[0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(
+                       np.logical_and(hx >= (self.lat_nx-1)/3.0,
+                                      hx <= 2*(self.lat_nx-1)/3.0),
+                       np.logical_and(hy >= width + self.wall_width, hy <= self.lat_ny-width-1-self.wall_width))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_seperation_2d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 1501,
+                    'lat_ny': 102+2, 'grid': 'D2Q9',
+                    'kappa': 0.04, 'Gamma': 1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'bc_wall': 'halfbb',
+                    'periodic_x': True, 'periodic_y': True}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=0.0))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=5))
+
+        lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=opts, defaults=settings)
+
+        self.add_body_force((self.options.force, 0.0), grid=0, accel=False)
+
+        # Use the fluid velocity in the relaxation of the order parameter field,
[…]

### 2011-04-04 Sailfish/Steady/grid.py

@@ -1,0 +1,91 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+ 
+def Run_Simulations():
+    print os.getcwd()
+    ny=[102,127,152,177,202,227]
+    nx=[1501,1876,2251,2626,3001,3376]
+    force_init=6e-6/4;
+    init_width=6
+    #os.mkdir("temp")
+    for i in range(0, 1):
+        dir_temp=str(ny[i])
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','bmc.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        ratio=float(ny[i]-2)/100.0
+        width=int(init_width*ratio)
+        subprocess.call(['./bmc.py','--lat_nx='+str(nx[i]),'--lat_ny='+str(ny[i]),'--bc_wall=halfbb',\
+            '--force='+str(force_init/(ratio*ratio)),'--batch','--every=50000','--max_iters='+str(int(200000*ratio)+1),'--output=grid','--iwidth='+str(width),'--bc_wall_grad_order=1'])
+        
+        os.chdir("..")
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return (zero-0.5)/(len(prof)-2)
+
+    
+def Analyze_Simulations():
+    print os.getcwd()
+        
+    for i in range(0,400000):
+        dir_temp="Proper Grid/"+str(49*i+2)
+        os.chdir(dir_temp)
+        #os.chdir("Force")
+        pylab.figure()
+        
+        
+def Analyze_Simulations():
+    print os.getcwd()
+    
+    
+    lengths=[]
+    thicknesses=[]
+    velocities=[]
+    reynolds=[]
+    for i in range(0,800000,20000):
+        #os.chdir("Force")
+        name="steady"+(6-len(str(i)))*"0"+str(i)+".npz"
+        print name        
+        array=numpy.load(name)
+        phase=array['phi']
+        vel=array['v'][0]
+        dims=phase.shape
+        center=phase[dims[0]/2,:]
+           
+        z1 = numpy.min(numpy.where(center < 0.0))
[…]

### 2011-04-06 Text/paper.tex
Minor style corrections.

@@
 \newcommand{\Ca}{\mathrm{Ca}}
+\renewcommand{\Re}{\mathrm{Re}}
 \newcommand{\todo}[1]{{\color{red}#1}}
@@
 propagates with the steady velocity $U_{\mathrm{bubble}}$ through the liquid media and deposits
-the film, where thickness to the height of the channel $H_{\mathrm{eff}}$ ratio is 
+the film, where the thickness to height of the channel $H_{\mathrm{eff}}$ ratio is 
 $\delta$.\label{fig:classical:bretherton}} 
@@
 \cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
-involving complex geometries, free interface motion, or for problems involving coalescence and/or
+involving complex geometries, free interface motion, or coalescence and/or
 droplet breakup. The continuous interface models are more flexible for those kinds of simulations.
@@
 negligible. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
-{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
+{\color{red} insignificant} Reynolds number ($\Re$) effects on the film thickness for a relatively wide
 range of Reynolds
@@
 {\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
-are negligible for $\Ca\leq0.05$ and have moderate changes for $\Ca>0.05$ in the range of Reynolds
-number from $0$ to $70$. Later on, \citet{heil-bretherton} extended results upto Reynolds number
-$300$. The indicate that while the influence of the established film thickness is insignificant
-($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
-significantly the pressure distrubution and the flow field near the front bubble tip. In the present
-simulations the largest Reynolds number we achieve is less than $20$ and it allows to neglect 
-inertia effects.} 
-Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
+are negligible for $\Ca\leq0.05$ and have moderate impact for $\Ca>0.05$ in the range of Reynolds
+number from $0$ to $70$. Later on, \citet{heil-bretherton} extended these results up to
+$\Re = 300$. This indicates that while the Reynolds number influence on the established film thickness is insignificant
+($7$ percent from the film thickness measured at $\Re=0$), the pressure distribution and the flow field near the front bubble tip
+can be affected by changes in $\Re$.  Since in the present work we only deal with flows with $\Re < 20$, we can safely
+neglect any inertia effects and conclude that the major governing parameter for microchannel Bretherton flows is not the density ratio,
 but the
-viscosity ratio. Moreover, the results based on uniform density, as in the case of the LBM
+viscosity ratio}. Moreover, the results based on uniform density, as in the case of the LBM
 binary liquid model, are in good agreement with other simulations
@@
 One should acknowledge the work of \citet{pagonabarraga-fingers} on menisci
-in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
+in thin films for fingering phenomena, as well as the work of \citet{sehgal-microchannel} who
+performed lattice Boltzmann
 simulations of two-dimensional channel flows for relatively large capillary numbers, and
 found discrepancies with the classical Bretherton theory, which
-is limited to the low capillary number regime \cite{giavedoni-numerical}. These authors didn't take
+is limited to the low capillary number regime \cite{giavedoni-numerical}. These authors did not take
 into
-the account recent studies extending microchannel simulations for the capillary numbers
+account recent studies extending microchannel simulations for the capillary numbers
 beyond the Bretherton regime. Thus, the comparison with other established CFD methods is limited.
 \citeauthor{sehgal-microchannel} used the Shan-Chen model \cite{Shan-chen:extended} to simulate the
-multiphase behavior. It is also
-the continuous interface model, but the study of grid dependence is absent.
+multiphase behavior. The Shan-Chen model is also a
+continuous interface model, but the above-mentioned paper does not contain a study of grid dependence.
 
 {\color{red}
-Level set method is represented by a number of works \cite{fukugata-levelset,
-lakehal-micro, carlson-comparison}. While the bubble shapes  were found in a good agreeement
-with the experiments, the range of studied bubble lengths is confined to $2-3$ channels diameters.
-Thus, the bubble radius changes significantly and cannot be used as classical Bretherton problem
+The level set method is also represented by a number of works \cite{fukugata-levelset,
+lakehal-micro, carlson-comparison}. While the bubble shapes  were found to be in a good agreement
+with the experiments, the range of studied bubble lengths is confined to $2-3$ channel diameters.
+Thus, the bubble radius changes significantly and cannot be used as a classical Bretherton problem
 benchmark. However, the works provide insight to the bubble formation and to the way small bubbles
@@
 {\color{red} $f_i^{eq}$ is the equilibrium probability distribution function}, $\omega$ is the
-relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
-is connected with the force and is implemented in the current work due to work \cite{guo}.}
+relaxation parameter, and $F_i$ is the external force population. {\color{red} The force population
+is represents an texternal physical force and is implemented in the current work using the scheme
+outlined in \citet{guo}.}
 
@@
 {\color{red}
-While the lattice Boltzmann system has parameters as gas and liquid
-viscosities, the surface tension, etc., those parameters are not the representative and
+While the lattice Boltzmann system has parameters such as the surface tension, the gas and liquid
+viscosities, etc., those parameters are not the representative and
 proportional quantities of the parameters in a physical system. The parameters of the lattice
 Boltzmann are connected with the physical world's parameters only through the non-dimensional
-numbers governing the physics of the problem. In this case it is a capillary number $\Ca$ and a
+numbers governing the physics of the problem. In our case, these numbers are the capillary number $\Ca$ and the
 viscosity ratio $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}$, which are
-obtained from the physical world and then they are matched through the lattice Boltzmann quantities.
+obtained from the physical world and then matched through the lattice Boltzmann quantities.
  The set of
@@
 %\section{Lattice Boltzmann benchmark}
-\section{Approach to benchmark}
+\section{Numerical benchmark procedure}
 To properly simulate and compare simulation results with the data
@@
 of body force driven flows, because of their simplicity and better numerical stability. This
-implies that we can use the periodic boundary conditions in the streamwise direction. As soon as the
+implies that we can use periodic boundary conditions in the streamwise direction. As soon as the
 periodic boundary
@@
 where $N_y$ is the number of nodes in the vertical direction. As will be discussed further, the
-effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes to
+effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ due to the use of bounce-back nodes to
 mimic no-slip conditions at the walls.
@@
 In the simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
-velocity to differ from that implied by the Poiseuille relation because of the bubble presence in a
+velocity to differ from that implied by the Poiseuille relation because of the presence of a bubble in the
 flow. That is
-why an iteration procedure is needed to achieve the 
+why an iterative procedure is needed to achieve the 
 desired velocity and 
 capillary number. 
-However, the initialization body force as given by (\ref{poiseuille:velocity:center}) is a good
+However, the initial body force as given by (\ref{poiseuille:velocity:center}) is a good
 starting point.
@@
 number of nodes to resolve the film thickness. Note that for paremeters indicated above the
-interface thickness is spread over approximately $5$ lattice units. The study of the interface
+interface is spread over approximately $5$ lattice units. The study of the interface
 resolution to the film thickness, Section \ref{section:grid:refinement}, suggests to choose the
@@
 is $202\times 3000$. 
- \item[Velocity] The relaxation parameters are taken to be in the stable range for hte simulations.
+ \item[Velocity] The relaxation parameters are taken to be in the stable range for the simulations.
 In the present simulations the viscosity ratio is $\dfrac{\mu_{liq}}{\mu_{gas}}=10$ with the
@@
 velocity of the interface. We can conclude it is enough to have $100000-300000$ iterations for
-bubble to approach the steady state.
+the bubble to approach the steady state.
 \begin{table}
@@
 \hline
-$N_\mathrm{iter}$&$\delta$&$U_{\mathrm{bubble}}$&{\color{red} $Re$}\\
+$N_\mathrm{iter}$&$\delta$&$U_{\mathrm{bubble}}$&{\color{red} $\Re$}\\
 \hline
@@
 film thickness to establish itself for the capillary number range $\Ca<4$.
-\citet{giavedoni-numerical} showed that the film stabilizes at the distances as $2.6-4.0$ diameters
+\citet{giavedoni-numerical} showed that the film stabilizes at the distances of $2.6-4.0$ diameters
 from the front tip depending on the Reynolds number.  We chose to measure the film thickness in the
@@
 extracted with the help of ``Engauge Digitizer'', data from
-\citet{heil-bretherton} was extracted by hands using a set of curves involving interpolation and
-it's the reason for a data to exhibit a noisy trend.} \label{fig:capillary:comparison}}
+\citet{heil-bretherton} was extracted manuallys using a set of curves involving interpolation and
+is the reason while the data exhibits a noisy trend.} \label{fig:capillary:comparison}}
 \end{figure}
@@
 accuracy of the method. To avoid this one needs to use advanced formulations of the LBM collision
-operator, as two-relaxation-times (TRT) and multiple-relaxation-times (MRT) collision operators
+operator, such as two-relaxation-times (TRT) and multiple-relaxation-times (MRT) collision operators
 \cite{ginzburg-main}. Although TRT and MRT collisions operators are studied for simple

### 2011-04-06 Review/answers.tex
Finished all the revisions and submitted it to the journal.

@@
 
-\textbf{Answer:} The reviewer is right that accuracy of the method deteriorate with the high
+\textbf{Answer:} The reviewer is right that accuracy of the method deteriorates with the high
 viscosity ratio. However, the parameters we chose here as $\tau_{\mathrm{liq}}=3.5-7.5$ and with
@@
 \textbf{Answer:} We redrew all the figures. For the figure 8 the data was extracted from the
-figure and rescaled for the comparison reasons.
+figure and rescaled for comparison reasons.
 
@@
 of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
-again described in a diﬀerent way in Section 5-4. Similar equations for the calculation of the
+again described in a different way in Section 5-4. Similar equations for the calculation of the
 pressure
@@
 about initialization and this was taken into account by reducing repetitive text and
-introducing step-by-step explations, see Section about initialization (lines 259-305). We also added
+introducing step-by-step explanations, see Section about initialization (lines $259-305$). We also
+added
 the
@@
 
-\textbf{Answer:} The factor $1/8$ is taken for the maximum velocity for the Poiseille flow between
-plates. Our assumption that the bubble travels with the constant velocity which equals to the
+\textbf{Answer:} The factor $1/8$ in Eq. (5) is taken for the maximum velocity for the Poiseille
+flow between
+plates. Our assumption is that the bubble travels with the constant velocity which equals to the
 maximum velocity of the Poiseuille profile between plates. Factor $1/12$ comes from the averaging
 of the flow rate. However, bubble moves faster than the liquid, i.e. the average liquid velocity.
-That's why the factor $1/8$ was chosen. The simulations show that factor $1/8$ works reasonably
+That is why the factor $1/8$ was chosen. The simulations show that factor $1/8$ works reasonably
 well.
@@
 
-\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the physical world
-parameters with the lattice Botlzmann parameters (lines $149-160$). As well more explanations are
+\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the
+physical parameters with the lattice Botlzmann parameters (lines $149-160$). As well more
+explanations are
 put towards the
@@
 
-\textbf{Answer:} We added a small steady-state study (lines $334-341$). We added an additional
+\textbf{Answer:} We added a steady-state study (lines $334-341$). We added an additional
 paragraph which shows the change of flow pattern for the

### 2011-04-06 Text/paper.tex
Finished all the revisions and submitted it to the journal.

@@
 number from $0$ to $70$. Later on, \citet{heil-bretherton} extended these results up to
-$\Re = 300$. This indicates that while the Reynolds number influence on the established film thickness is insignificant
+$\Re = 300$. They indicate that while the Reynolds number influence on the established film
+thickness is insignificant
 ($7$ percent from the film thickness measured at $\Re=0$), the pressure distribution and the flow field near the front bubble tip
@@
 lakehal-micro, carlson-comparison}. While the bubble shapes  were found to be in a good agreement
-with the experiments, the range of studied bubble lengths is confined to $2-3$ channel diameters.
+with the experiments, the range of studied bubble lengths is limited to $2-3$ channel diameters.
 Thus, the bubble radius changes significantly and cannot be used as a classical Bretherton problem
 benchmark. However, the works provide insight to the bubble formation and to the way small bubbles
-flow in the
-microchannels.}
+flow in microchannels.}
 
@@
 relaxation parameter, and $F_i$ is the external force population. {\color{red} The force population
-is represents an texternal physical force and is implemented in the current work using the scheme
+represents an external physical force and is implemented in the current work using the scheme
 outlined in \citet{guo}.}
@@
 in the framework of the binary liquid model is $\gamma=\sqrt{\frac{8 k
-A}{9}}$. The inclusion of the interface tension to the momentum flux tensor is done through the
+A}{9}}$. The inclusion of the interface tension in the momentum flux tensor is done through the
 coefficients $k$, $A$ and weights $w_i^{\alpha\beta}$.
@@
 proportional quantities of the parameters in a physical system. The parameters of the lattice
-Boltzmann are connected with the physical world's parameters only through the non-dimensional
+Boltzmann are connected with the physical parameters only through the non-dimensional
 numbers governing the physics of the problem. In our case, these numbers are the capillary number $\Ca$ and the
@@
 In these
-examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
+examples we use $k=0.04$ and $A=0.04$ as numerically optimal binary liquid model parameters.
 That implies the surface tension to be $\sqrt{\frac{8 k
 A}{9}}=0.0377$ and the characteristic width of the interface
-$5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units (grid nodes). For the sake of simplicity and
+$5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units (grid spacings). For the sake of simplicity and
 stability, the relaxation times for liquid and gas phases are taken as
@@
 \end{equation}
-where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
+where $\phi_0$ is the grid coordinate where the phase field is $0$, $N_y-2$
 is the effective channel height.
@@
 wall for the
-multiphase models to the best authors' knowledge is not yet derived. Thus, the simulation grid size
+multiphase models to the best of the authors' knowledge is not yet derived. Thus, the simulation
+grid size
 is $202\times 3000$. 
@@
 how the body force is connected with the bubble velocity $U_{\mathrm{bubble}}$. We assume that the
-flow is close to the Poiseuille flow:
+flow is close to planar Poiseuille flow with $U_{\mathrm{bubble}}$ being the maximun in the
+Poiseuille profile:
 \begin{equation}
@@
 After imposing the body force the simulations can be run. A typical simulation runs approximately
-$100000-300000$ time steps due to reach steady-state, see Section \ref{sec:steady:state}.
+$100000-300000$ time steps to reach steady-state, see Section \ref{sec:steady:state}.
 %Note that the velocity is proportional to the physical time step (small velocities imply longer
@@
 summarizes the simulation results in terms of the film thickness in the middle of the bubble and the
-velocity of the interface. We can conclude it is enough to have $100000-300000$ iterations for
+velocity of the interface as a function of time. We can conclude that it is enough to have
+$100000-300000$ iterations for
 the bubble to approach the steady state.
@@
 \hline
-$10000$&$ 0.0655$&$0.0027$&$0.81$\\
-\hline
-$14000$&$0.0608$&$0.0027$&$0.81$\\
-\hline
-$18000$&$0.0577$&$0.0027$&$0.81$\\
-\hline
-$22000$&$0.0630$&$0.0027$&$0.81$\\
-\hline
-$26000$&$0.0660$&$0.0027$&$0.81$\\
-\hline
-$30000$&$0.0661$&$0.0027$&$0.81$\\
+$100000$&$ 0.0655$&$0.0027$&$0.81$\\
+\hline
+$140000$&$0.0608$&$0.0027$&$0.81$\\
+\hline
+$180000$&$0.0577$&$0.0027$&$0.81$\\
+\hline
+$220000$&$0.0630$&$0.0027$&$0.81$\\
+\hline
+$260000$&$0.0660$&$0.0027$&$0.81$\\
+\hline
+$300000$&$0.0661$&$0.0027$&$0.81$\\
 \hline
@@
 influences the film thickness at infinity \cite{bretherton}. It was indicated by
-\citet{heil-bretherton} that the inertia effects influence insignificantly the film thickness at
+\citet{heil-bretherton} that inertia effects influence insignificantly the film thickness at
 infinity but influence significantly the velocity pattern in front of the bubble. Therefore, to
 properly describe the Bretherton phenomena the
-bubble shape along with velocity pattern need to be examined.}
+bubble shape along with velocity pattern needs to be examined.}
 
@@
 However, \citet{cerro-bubble-train} indicate that for the gravity-driven bubble train flow in tubes
-it is enough for bubble to have a length of two-three tube diameters to apply analysis for very long
-bubbles. In comparison with the axisymmetric circular geometry, for three-dimensional square shaped
+it is enough for bubbles to have a length of two-three tube diameters to apply analysis for very
+long
+bubbles. In comparison with the axisymmetric geometry, for three-dimensional square shaped
 microchannels \citet{heil-threedim} indicate the measurement of the film thickness at $5.5$
@@
 profiles, Fig. \ref{fig:streamlines:pattern}. The pattern shows that there is a transition of
-having vortex in the front of the bubble $\Ca=0.22$ and not having it for larger capillary numbers
+having vortex in the front of the bubble for $\Ca=0.22$ and not having it for larger capillary
+numbers
 $\Ca=1.00$. \citet{giavedoni-numerical} studied the behavior of the detaching of the vortex for the
